### 2018-10-29 To_Do.Rmd
First upload

---
title: "To Do"
author: ""
date: "29/10/2018"
output: html_document
---

# Data

## Trends

[ ] Bartomeus PNAS estimates (Nacho) 
[ ] Sheper PNAS estimates (MA)
[ ] UE red list (via @ropensci; MA)
[ ] SPAIN trends (MA: species with Brain size; Nacho or Cristina: trends)

## traits

[ ] LHT Bartomeus PNAS + Shepper PNAS + Traitbase (Var) Nacho
[ ] Brain: MA (Spain + ND + US Ferran + US Parker): MA
[ ] Habitats USA: MA; Others ¿?: MA
[ ] Exotic use ¿? (Nacho)
[ ] Innovation (some in MA bbdd)
[ ] Invasive species in US. (Nacho)

## Phylogeny

[ ] Hedtke et al 2013? + politomies
[ ] Build one as Ferran did?

# Models

- Regresion Trees (see Sol et al 2012 Science)
- Quantile regresions

# Refs:

- Mamifers. Manuela gonzalez??? big-worst
- Birds. Invasion proceses (Sol et al 2012),... 
- Reptiles. Invasion also related with Brain size.

### 2018-11-07 To_Do.Rmd
uploading data

@@
 [ ] UE red list (via @ropensci; MA)
-[ ] SPAIN trends (MA: species with Brain size; Nacho or Cristina: trends)
+[X] SPAIN trends (MA: species with Brain size) 
+[ ] Trends: Nacho or Cristina
 
@@
 [ ] LHT Bartomeus PNAS + Shepper PNAS + Traitbase (Var) Nacho
-[ ] Brain: MA (Spain + ND + US Ferran + US Parker): MA
+[X] Brain: MA (Spain + ND + US Ferran + US Parker): MA
 [ ] Habitats USA: MA; Others ¿?: MA

### 2018-11-07 Scripts/MA_data_cleaning.R

@@ -1,0 +1,21 @@
+brains.it <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/brains.it.csv")
+
+#List of species with brain weight and IT
+
+brains.it$Encephalization<-brains.it$Brain.Weight..mg./brains.it$IT
+brains.t<-subset(brains.it, subset = (is.na(brains.it$Encephalization)==FALSE))
+spain.bees<-subset(brains.t, subset = (brains.t$Country=="Spain"))
+list.of.spanish.bees<-unique(spain.bees$Species)
+list.of.spanish.bees<-as.data.frame(list.of.spanish.bees)
+
+getwd()
+setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/Raw_data")
+write.csv(list.of.spanish.bees, "list.of.spanish.bees.csv")
+setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/")
+
+#Does Queen have bigger brains?------
+boxplot(brains.it$Brain.Weight..mg./brains.it$IT~brains.it$Sex, ylab="Brain/IT")
+
+aggregate(brains.it$Brain.Weight..mg./brains.it$IT~brains.it$Sex, FUN = mean)
+aov.queens<-aov(brains.it$Brain.Weight..mg./brains.it$IT~brains.it$Sex)
+TukeyHSD(aov.queens)

### 2019-01-08 To_Do.Rmd
Corrected Scheper name

--- 2018-10-29 To_Do.Rmd
@@
 [ ] Bartomeus PNAS estimates (Nacho) 
-[ ] Sheper PNAS estimates (MA)
+[ ] Scheper PNAS estimates (MA)
 [ ] UE red list (via @ropensci; MA)
-[ ] SPAIN trends (MA: species with Brain size; Nacho or Cristina: trends)
+[X] SPAIN trends (MA: species with Brain size) 
+[ ] Trends: Nacho or Cristina
 
@@
 
-[ ] LHT Bartomeus PNAS + Shepper PNAS + Traitbase (Var) Nacho
-[ ] Brain: MA (Spain + ND + US Ferran + US Parker): MA
+[ ] LHT Bartomeus PNAS + Scheper PNAS + Traitbase (Var) Nacho
+[X] Brain: MA (Spain + ND + US Ferran + US Parker): MA
 [ ] Habitats USA: MA; Others ¿?: MA

### 2019-01-09 Scripts/MA_data_cleaning.R
Innovation extracted

@@ -1,6 +1,6 @@
 brains.it <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/brains.it.csv")
 
-#List of species with brain weight and IT
+#List of species with brain weight and IT-----
 
 brains.it$Encephalization<-brains.it$Brain.Weight..mg./brains.it$IT
 brains.t<-subset(brains.it, subset = (is.na(brains.it$Encephalization)==FALSE))
@@ -9,11 +9,48 @@ spain.bees<-subset(brains.t, subset = (brains.t$Country=="Spain"))
 list.of.spanish.bees<-unique(spain.bees$Species)
 list.of.spanish.bees<-as.data.frame(list.of.spanish.bees)
 
+list.of.spanish.bees
+list.of.species
+
 getwd()
 setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/Raw_data")
-write.csv(list.of.spanish.bees, "list.of.spanish.bees.csv")
-write.csv(list.of.species, "list.of.species.csv")
+#write.csv(list.of.spanish.bees, "list.of.spanish.bees.csv")
+#write.csv(list.of.species, "list.of.species.csv")
 setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/")
+
+#List of innovators-----
+# load package
+install.packages("googlesheets")
+library(googlesheets)
+gs_ls("Traits")
+be <- gs_title("Traits")
+gs_ws_ls(be)
+Traits <- gs_read(ss=be, ws = "Sheet1", skip=0)
+Traits<- as.data.frame(Traits)
+View(Traits)
+Traits$species
+summary(as.factor(Traits$value), maxsum = 1000)
+list.of.innovators.f<-subset(Traits, subset = (Traits$value == "Brick" | 
+                                                   Traits$value == "Plastic" |
+                                                   Traits$value == "Small mammal house" |
+                                                   Traits$value == "Birdhouse" |
+                                                   Traits$value == "PVC" |
+                                                   Traits$value == "Table" |
+                                                   Traits$value == "Rockwool" |
+                                                   Traits$value == "Plastic table"|
+                                                   Traits$value == "Plastic straw" |
+                                                   Traits$value == "Polycarbonate" |
+                                                   Traits$value == "Paper" |
+                                                   Traits$value == "Metal" |
+                                                   Traits$value == "latch hole" |
+                                                   Traits$value == "Fluff" |
[…]

### 2019-01-09 To_Do.Rmd
Innovation extracted
uploaded scheper trends

@@
 [ ] Bartomeus PNAS estimates (Nacho) 
-[ ] Scheper PNAS estimates (MA)
+[X] Scheper PNAS estimates (MA)
 [ ] UE red list (via @ropensci; MA)
@@
 [ ] Exotic use ¿? (Nacho)
-[ ] Innovation (some in MA bbdd)
+[X] Innovation (some in MA bbdd)
 [ ] Invasive species in US. (Nacho)

### 2019-01-09 Scripts/MA_data_cleaning.R

@@ -50,8 +50,26 @@ list.of.innovators.f<-subset(Traits, subset = (Traits$value == "Brick" |
 list.of.innovators<-unique(list.of.innovators.f$species)
 list.of.innovators
 
+#Import scheper trends----
+
+scheper_trends <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/scheper_trends.csv")
+# 1 means not change along time
+scheper_trends
+
+#Import red list----
+install.packages("rredlist")
+library(rredlist)
 
 
+rl_use_iucn()
+
+
+rl_threats_(name = "Osmia cornuta", id = NULL, region = NULL, key = NULL)
+
+
+#US habitats----
+habitat_preference_US <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/habitat_preference_US.csv")
+habitat_preference_US
 #Does Queen have bigger brains?------
 boxplot(brains.it$Brain.Weight..mg./brains.it$IT~brains.it$Sex, ylab="Brain/IT")
 

### 2019-01-10 To_Do.Rmd
USA preferences done

@@
 [X] Brain: MA (Spain + ND + US Ferran + US Parker): MA
-[ ] Habitats USA: MA; Others ¿?: MA
+[X] Habitats USA: MA 
+    [ ] Others ¿?: MA
 [ ] Exotic use ¿? (Nacho)

### 2019-01-10 Scripts/MA_data_cleaning.R

@@ -1,3 +1,7 @@
+library(reshape2)
+library(googlesheets)
+library(rredlist)
+
 brains.it <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/brains.it.csv")
 
 #List of species with brain weight and IT-----
@@ -9,25 +13,22 @@ spain.bees<-subset(brains.t, subset = (brains.t$Country=="Spain"))
 list.of.spanish.bees<-unique(spain.bees$Species)
 list.of.spanish.bees<-as.data.frame(list.of.spanish.bees)
 
-list.of.spanish.bees
-list.of.species
 
 getwd()
 setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/Raw_data")
 #write.csv(list.of.spanish.bees, "list.of.spanish.bees.csv")
 #write.csv(list.of.species, "list.of.species.csv")
 setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/")
+list.of.spanish.bees
+list.of.species
 
 #List of innovators-----
 # load package
-install.packages("googlesheets")
-library(googlesheets)
 gs_ls("Traits")
 be <- gs_title("Traits")
 gs_ws_ls(be)
 Traits <- gs_read(ss=be, ws = "Sheet1", skip=0)
 Traits<- as.data.frame(Traits)
-View(Traits)
 Traits$species
 summary(as.factor(Traits$value), maxsum = 1000)
 list.of.innovators.f<-subset(Traits, subset = (Traits$value == "Brick" | 
@@ -57,8 +58,6 @@ scheper_trends <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/scheper_tr
 scheper_trends
 
 #Import red list----
-install.packages("rredlist")
-library(rredlist)
 
 
 rl_use_iucn()
@@ -69,6 +68,24 @@ rl_threats_(name = "Osmia cornuta", id = NULL, region = NULL, key = NULL)
 
 #US habitats----
 habitat_preference_US <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/habitat_preference_US.csv")
+habitat_preference_US
+m.habitat_preference_US<-melt(habitat_preference_US)
+m.habitat_preference_US<-as.data.frame(m.habitat_preference_US)
+#We order
+m.habitat_preference_US<-m.habitat_preference_US[order(m.habitat_preference_US$Species),]
[…]

### 2019-01-21 To_Do.Rmd
Added redlist trends

@@
 [X] Scheper PNAS estimates (MA)
-[ ] UE red list (via @ropensci; MA)
+[X] UE red list (via @ropensci; MA)
 [X] SPAIN trends (MA: species with Brain size) 

### 2019-01-21 Scripts/MA_data_cleaning.R

@@ -58,44 +58,64 @@ scheper_trends <- read.csv("~/Desktop/Tesis/R/declines_brain/Raw_data/scheper_tr
 scheper_trends
 
 #Import red list----
-rl_use_iucn
 
+#You need an API token to use this
 rl_use_iucn()
 IUCN_REDLIST_KEY="084775ffa81321631f2583dcd416a81af6fbd73efb60a0eab38e0df10082fe07"
-iucn_redlist_key()
-rl_regions(key = IUCN_REDLIST_KEY, parse = TRUE)
-
-rl_threats(name = 'Emys orbicularis', region = "europe", key = IUCN_REDLIST_KEY)
-A<-rl_threats(name = 'Emys orbicularis', key = IUCN_REDLIST_KEY)
 
 
-rl_threats('Fratercula arctica', key = IUCN_REDLIST_KEY, parse = TRUE)
-rl_threats('Fratercula arctica', key = IUCN_REDLIST_KEY)
-rl_threats('Fratercula arctica', region = 'europe', key = IUCN_REDLIST_KEY)
-rl_threats(id = 12392, key = IUCN_REDLIST_KEY)
-rl_threats(id = 22694927, region = 'europe')
-rl_threats(name = 'Abies numidica')
-rl_threats_('Fratercula arctica', key = IUCN_REDLIST_KEY)
+list.of.species
 
-rl_threats(id = 62290750)
+for (n in 1:nrow(list.of.species)) {
+    tryCatch({temp<-rl_search(as.character(list.of.species[n,1]), key = IUCN_REDLIST_KEY, parse = FALSE)$result[[1]]$category
+    print(temp)}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+}
 
-rl_occ_country('Loxodonta africana', key = IUCN_REDLIST_KEY)
+rl_search(as.character(list.of.species[19,1]), key = IUCN_REDLIST_KEY, parse = FALSE)$result[[1]]$category
+rl_search(as.character(list.of.species[20,1]), key = IUCN_REDLIST_KEY, parse = FALSE)$result[[1]]$category
+rl_search(as.character(list.of.species[26,1]), key = IUCN_REDLIST_KEY, parse = FALSE)$result[[1]]$category
+rl_search(as.character(list.of.species[60,1]), key = IUCN_REDLIST_KEY, parse = FALSE)$result[[1]]$category
[…]

### 2019-01-24 To_Do.Rmd
trends compiled

@@
 
-[ ] Bartomeus PNAS estimates (Nacho) 
+[X] Bartomeus PNAS estimates (Nacho) 
 [X] Scheper PNAS estimates (MA)

### 2019-01-24 Scripts/MA_data_cleaning.R

@@ -14,7 +14,6 @@ list.of.spanish.bees<-unique(spain.bees$Species)
 list.of.spanish.bees<-as.data.frame(list.of.spanish.bees)
 
 
-getwd()
 #setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/Raw_data")
 #write.csv(list.of.spanish.bees, "list.of.spanish.bees.csv")
 #write.csv(list.of.species, "list.of.species.csv")
@@ -111,7 +110,6 @@ category.temp<-c(rl_search(as.character(list.of.species[19,1]), key = IUCN_REDLI
 species.iucn<-as.data.frame(species.temp)          
 category.iucn<-as.data.frame(category.temp)          
 iucn.trends<-(cbind(species.iucn,category.iucn))
-getwd()
 
 # setwd("/Users/Bartomeus_lab/Desktop/Tesis/R/declines_brain/Raw_data")
 # write.csv(iucn.trends, "iucn_trends.csv")
@@ -138,6 +136,70 @@ na.omit(m.habitat_preference_US)
 habitat_preference_US<-na.omit(m.habitat_preference_US)
 
 habitat_preference_US
+
+#Species trends data--------
+scheper_trends <- read.csv("Raw_data/scheper_trends.csv")
+iucn_trends <- read.csv("Raw_data/iucn_trends.csv")
+bartomeus_trends <- read.csv("Raw_data/bartomeus_trends.csv", sep = ";")
+
+colnames(scheper_trends)
+colnames(iucn_trends) <-c("X","Species","category.temp")
+colnames(bartomeus_trends)
+
+colnames(list.of.species)<-c("Species")
+
+merger1<-merge(list.of.species,scheper_trends)
+merger2<-data.frame(merger1$Species, merger1$Change.index)
+colnames(merger2)<-c("Species","Scheper_trends")
+merger3<-merge(list.of.species, iucn_trends)
+merger4<-data.frame(merger3$Species,merger3$category.temp)
+colnames(merger4)<-c("Species","IUCN_trends")
+
+merger5<-merge(list.of.species, bartomeus_trends)
+merger6<-data.frame(merger5$Species, merger5$Status)
+colnames(merger6)<-c("Species","Bartomeus_trends")
+
+merger2$source<-rep("Scheper",nrow(merger2))
+merger4$source<-rep("IUCN",nrow(merger4))
[…]

### 2022-08-24 Summary.Rmd
Preparation of preliminary results

---
output: pdf_document
---

# Brief methological explanation and preliminary results

Here we summarise all the critical steps conducted to analise bee preferences.

## Filtering criteria

1) Filter records above 1985 and 1987 (this is decided based on the GIS data available) for Europe and USA respectively.

2) Filter by unique capture event.

3) Filter by minimum of 3 decimals on coordinates.

4) Filter number of levels per species (minimum sample size, 100 occurrences).

5) Filter by wide geographical distribution.

## Land use classification

**1) Europe**

Link with land European use classes

https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-512.html

```{r Load Europe data, eval=TRUE, include=FALSE}
library(tidyverse)
pref <- read_csv("Data/Europe_data/land_cover_europe.csv.gz") %>% 
    dplyr::select(Species, Cover_names) %>% 
    mutate_if(is.character,as.factor)

```

```{r Table with all levels Europe, echo = FALSE, warning = FALSE, message = FALSE}
library(kableExtra)

pref %>% 
group_by(Cover_names) %>% 
summarise(n_rows = length(Cover_names)) %>% 
filter(!Cover_names=="NODATA") %>% 
rename(Land_use = Cover_names) %>% 
arrange(desc(n_rows)) %>% 
mutate(Cover_names = fct_recode(as.factor(Land_use),
#1 Check with Nacho, dehesas and others 
 Seminatural = "Agro-forestry areas", #1) 2.4.4 
#2 Check with Nacho 
 Urban = "Airports", #2) 1.2.4 
#3 Crops 
 Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
#4 Very different cover type, not included 
 Discard = "Bare rocks", #4) 3.3.2 
#5 Very different cover type, not included 
 Discard = "Beaches, dunes, sands", #5) 3.3.1 
#6 Not included, lack of details 
 Discard = "Burnt areas", #6 3.3.4
#7 Not included, only water surface
 Discard = "Coastal lagoons", #7) 5.2.1 
#8 Straightforward
 Natural = "Coniferous forest", #8) 3.1.2
#9 Excluded, complex to consider because it lacks habitat context
 Urban = "Construction sites", #9) 1.3.3
#10 Based on the images and the description has a lot of vegetation cover
 Seminatural = "Discontinuous urban fabric", #10) 1.1.2
#11 Pollinators here are likely to depend much on the context not on the dump
 Discard = "Dump sites", #11) 1.3.2
#12 Filter out, water body
 Discard = "Estuaries", #12) 5.2.2 
#13 Straightforward
Agricultural = "Fruit trees and berry plantations", #13) 2.2.2
#14 Urban as embedded within the city
 Urban = "Green urban areas", #14) 1.4.1
#15 Highly modified
Urban = "Industrial or commercial units", #15) 1.2.1
#16 By definion, seminatural
Seminatural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #16) 2.4.3
#17 Discard as is not clear the vegetation context
Discard = "Mineral extraction sites", #17) 1.3.1
#18 Straightforward 3.1.3
 Natural = "Mixed forest", #18) 3.1.3
#19 Crystal clear category
 Natural = "Moors and heathland", #19) 3.2.2
#20 Crystal clear category
 Natural = "Natural grasslands", #20) 3.2.1
#21 Filter out
 Discard = "NODATA", #21 
#22 cultivated land
 Agricultural = "Non-irrigated arable land", #22) 2.1.1
#23 As it is a monoculture despite some very
 Agricultural = "Olive groves", #23) 2.2.3
#24 Crystal clear category
 Natural = "Peat bogs", #24) 4.1.2
#25 Crystal clear category
 Seminatural = "Permanently irrigated land", #25) 2.1.2
#26 Check with Nacho
 Urban = "Port areas", #26) 1.2.3
#27 Check with Nacho
 Seminatural = "Road and rail networks and associated land", #27) 1.2.2
#28 Check with Nacho but just only 200 records
 Seminatural = "Salines", #28) 4.2.2 
#29 Flowering plant communities 
 Natural = "Salt marshes", #29) 4.2.1
#30 Sclerophyllous shrubs and low shrubs
 Natural = "Sclerophyllous vegetation", #30) 3.2.3
#31 Discard for now 
 Discard = "Sea and ocean", #31) 5.2.3
#32 Areas with sparse vegetation, covering 10-50% of surface 
 Natural = "Sparsely vegetated areas", #32) 3.3.3
#33 Check with Nacho 
 Seminatural = "Sport and leisure facilities", #33) 1.4.2 
#34 Excluded as it is very general
 Discard = "Water bodies", #34) 5.1.2
#35 Excluded as it is very general
 Discard = "Water courses" #35 5.1.1 
)) %>%  
kbl() %>% 
kable_styling(font_size = 8,latex_options = "hold_position")

```

Check final levels per category:

```{r Show levels Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
pref = pref %>% 
mutate(Cover_names = fct_recode(as.factor(Cover_names),
#1 Check with Nacho, dehesas and others 
Seminatural = "Agro-forestry areas", #1) 2.4.4 
#2 Check with Nacho 
Urban = "Airports", #2) 1.2.4 
#3 Crops 
Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
#4 Very different cover type, not included 
Discard = "Bare rocks", #4) 3.3.2 
#5 Very different cover type, not included 
Discard = "Beaches, dunes, sands", #5) 3.3.1 
#6 Not included, lack of details 
Discard = "Burnt areas", #6 3.3.4
#7 Not included, only water surface
Discard = "Coastal lagoons", #7) 5.2.1 
#8 Straightforward
Natural = "Coniferous forest", #8) 3.1.2
#9 Excluded, complex to consider because it lacks habitat context
Urban = "Construction sites", #9) 1.3.3
#10 Based on the images and the description has a lot of vegetation cover
Urban = "Discontinuous urban fabric", #10) 1.1.2
#11 Pollinators here are likely to depend much on the context not on the dump
Discard = "Dump sites", #11) 1.3.2
#12 Filter out, water body
Discard = "Estuaries", #12) 5.2.2 
#13 Straightforward
Agricultural = "Fruit trees and berry plantations", #13) 2.2.2
#14 Urban as embedded within the city
Urban = "Green urban areas", #14) 1.4.1
#15 Highly modified
Urban = "Industrial or commercial units", #15) 1.2.1
#16 By definion, seminatural
Seminatural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #16) 2.4.3
#17 Discard as is not clear the vegetation context
Discard = "Mineral extraction sites", #17) 1.3.1
#18 Straightforward 3.1.3
Natural = "Mixed forest", #18) 3.1.3
#19 Crystal clear category
Natural = "Moors and heathland", #19) 3.2.2
#20 Crystal clear category
Natural = "Natural grasslands", #20) 3.2.1
#21 Filter out
Discard = "NODATA", #21 
#22 cultivated land
Agricultural = "Non-irrigated arable land", #22) 2.1.1
#23 As it is a monoculture despite some very
Agricultural = "Olive groves", #23) 2.2.3
#24 Crystal clear category
Natural = "Peat bogs", #24) 4.1.2
#25 Crystal clear category
Seminatural = "Permanently irrigated land", #25) 2.1.2
#26 Check with Nacho
Urban = "Port areas", #26) 1.2.3
#27 Check with Nacho
Seminatural = "Road and rail networks and associated land", #27) 1.2.2
#28 Check with Nacho but just only 200 records
Seminatural = "Salines", #28) 4.2.2 
#29 Flowering plant communities 
Natural = "Salt marshes", #29) 4.2.1
#30 Sclerophyllous shrubs and low shrubs
Natural = "Sclerophyllous vegetation", #30) 3.2.3
#31 Discard for now 
Discard = "Sea and ocean", #31) 5.2.3
#32 Areas with sparse vegetation, covering 10-50% of surface 
Natural = "Sparsely vegetated areas", #32) 3.3.3
#33 Check with Nacho 
Seminatural = "Sport and leisure facilities", #33) 1.4.2 
#34 Excluded as it is very general
Discard = "Water bodies", #34) 5.1.2
#35 Excluded as it is very general
Discard = "Water courses" #35 5.1.1 
)) %>% 
filter(!Cover_names == "Discard")

pref %>% 
group_by(Cover_names) %>% 
summarise(Levels = length(Cover_names)) %>% 
kbl() %>% 
kable_styling(latex_options = "hold_position")
```

Then, we calculate preferences and model the data.

```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}

data <- read_csv("Data/Europe_data/data_preference_brain_weight_europe.csv")

model1 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe.csv")

model2 = read_csv("Data/Europe_data/model_output_preference_it_europe.csv")

ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe occurrences")

ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe occurrences")

```

\newpage

**2) USA**

```{r Load USA data, eval=TRUE, include=FALSE}
#Load extracted data
pref <- read_csv("Data/Usa_data/land_cover_usa.csv.gz") %>% 
dplyr::select(species, cover.names) %>% 
mutate_if(is.character,as.factor)

```

```{r Table with all levels USA, echo = FALSE, warning = FALSE, message = FALSE}

pref %>% 
group_by(cover.names) %>% 
summarise(n_rows = length(cover.names)) %>% 
rename(Land_use = cover.names) %>% 
arrange(desc(n_rows)) %>% 
mutate(Cover_names = fct_recode(as.factor(Land_use),
Discard = "Barren Land", #Very different cover type, not included  
Agricultural = "Cultivated Crops", #
Natural = "Deciduous Forest", #
Urban = "Developed, High Intensity", #Greater than 50% of cover  
Urban = "Developed, Medium Intensity", #Greater than 50% of cover  
Seminatural = "Developed, Low Intensity", #Low urban impact 
Seminatural = "Developed, Open Space", #Open areas with vegetation 
Natural = "Emergent Herbaceous Wetlands", #
Natural = "Evergreen Forest", #
Agricultural = "Hay/Pasture", #Highly managed by humans 
Natural = "Herbaceous", #See explanation below 
Natural = "Mixed Forest", #
Discard = "Open Water", #Low vegetation and land, not included 
Natural = "Shrub/Scrub", #Heath classified as natural 
Natural = "Woody Wetlands")) %>% 
kbl() %>% 
kable_styling(latex_options = "hold_position", font_size = 12)

```

Check final levels per category:

```{r Show levels USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}

 pref %>% 
mutate(cover.names = fct_recode(as.factor(cover.names),
      Discard = "Barren Land", #Very different cover type, not included (MA & JB)
      Agricultural = "Cultivated Crops", #(MA & JB)
      Natural = "Deciduous Forest", #(MA & JB)
      Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
      Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
      Seminatural = "Developed, Low Intensity", #Low urban impact (MA & JB)
      Seminatural = "Developed, Open Space", #Open areas with vegetation (MA & JB)
      Natural = "Emergent Herbaceous Wetlands", #(MA & JB)
      Natural = "Evergreen Forest", #(MA & JB)
      Agricultural = "Hay/Pasture", #Highly managed by humans (MA & JB)
      Natural = "Herbaceous", #See explanation below (MA & JB)
      Natural = "Mixed Forest", #(MA & JB)
      Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
      Natural = "Shrub/Scrub", #Heath classified as natural (MA & JB)
      Natural = "Woody Wetlands")) %>% #(MA & JB)
filter(!cover.names == "Discard") %>%  #Discard these categories classified as "Discard"
#Herbaceous and hay/pasture are classified as two different habitats in NLCD. 
#We merged them because herbaceous areas in our sampling region are always 
#for livestock (Koh et al., 2016).
group_by(cover.names) %>% 
summarise(n_rows = length(cover.names)) %>% 
kbl() %>% 
kable_styling(latex_options = "hold_position")

```

Then, we calculate preferences and model the data.

```{r Plotting USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}

data <- read_csv("Data/Usa_data/data_preference_brain_weight_usa.csv")

model1 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa.csv")

model2 = read_csv("Data/Usa_data/model_output_preference_it_usa.csv")

ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA occurrences")

ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA occurrences")

```

Note: The models have been run with non-informative priors from the brms function but some work on priors can be done... I have explored a bit and the results were quite similar overall but reviewers seem to don't like this approach as one of the goals of Bayesian is to use the prior knowledge of our variables to improve our predictions.

**2) Finally, we consider Europe and USA together**

```{r Plotting Europe USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}

data <- read_csv("Data/Europe_USA/data_preference_brain_weight_europe_usa.csv")

model1 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")

model2 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa.csv")

ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe USA occurrences")

ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe USA occurrences")

```

### 2022-08-24 Scripts/3_JBL_Europe_Data_Cleaning.R

@@ -4,7 +4,7 @@
 #1)Filter records above 1985 (this is decided based on the GIS data available)
 #2)Filter by unique capture event
 #3)Filter by minimum of 3 decimals on coordinates
-#4)Filter number of levels per species (minimum N=50)
+#4)Filter number of levels per species (minimum N=100)
 #5)Filter by wide geographical distribution
 ######-
 

### 2022-08-24 Scripts/3_JBL_USA_Data_Cleaning.R

@@ -4,7 +4,7 @@
 #1)Filter records above 1987 (this is decided based on the GIS data available)
 #2)Filter by unique capture event
 #3)Filter by minimum of 3 decimals on coordinates
-#4)Filter number of levels per species (minimum N=50)
+#4)Filter number of levels per species (minimum N=100)
 #5)Filter by wide geographical distribution
 ######-
 

### 2022-08-24 Scripts/6_JBL_Europe_Preferences.R

@@ -21,46 +21,80 @@ check_cover = pref %>%
 
 #Recode levels to calculate preferences
 #Check levels here
-#https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-512.html
 pref = pref %>% 
 mutate(Cover_names = fct_recode(as.factor(Cover_names),
-     Seminatural = "Agro-forestry areas", #1 (2.4.4) Check with Nacho, dehesas and others (MA & JB)
-     Urban = "Airports", #2 Check with Nacho (MA & JB)
-     Agricultural = "Annual crops associated with permanent crops", #3 (2.4.1) Crops (MA & JB)
-     Discard = "Bare rocks", #4 Very different cover type, not included (MA & JB)
-     Discard = "Beaches, dunes, sands", #5 (3.3.1) Very different cover type, not included (MA & JB)
-     Discard = "Burnt areas", #6 Not included, lack of details (MA & JB)
-     Discard = "Coastal lagoons", #7 (5.2.1) Not included, only water surface  (MA & JB)
-     Natural = "Coniferous forest", #8 (MA & JB)
-     Urban = "Construction sites", #9 Excluded, complex to consider because it lacks habitat context (MA & JB)
-     Seminatural = "Discontinuous urban fabric", #10 Based on the images and the description has a lot of vegetation cover (MA & JB)
-     Discard = "Dump sites", #11 Pollinators here are likely to depend much on the context not on the dump (MA & JB)
-     Discard = "Estuaries", #12 Filter out, water body (MA & JB)
-     Agricultural = "Fruit trees and berry plantations", #13 (MA & JB)
-     Urban = "Green urban areas", #14 Urban as embedded within the city (MA & JB)
-     Urban = "Industrial or commercial units", #15 Highly modified (MA & JB)
-     Seminatural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #16 By definion, seminatural (MA & JB)
-     Discard = "Mineral extraction sites", #17 Discard as is not clear the vegetation context (MA & JB)
-     Natural = "Mixed forest", #18 (MA & JB)
-     Natural = "Moors and heathland", #19 (MA & JB)
-     Natural = "Natural grasslands", #20 (MA & JB)
-     Discard = "NODATA", #21 Filter out
[…]

### 2022-08-24 Scripts/8_JBL_Analysis_Europe.R

@@ -14,27 +14,34 @@ A10 = readRDS("Data/Europe_data/phylo_europe.rds")
 
 #Prepare data----
 d = left_join(preferences, brain_weight) %>% 
-    mutate(Species = str_replace_all(Species, " ", "_")) %>% 
+    mutate(Species = str_replace_all(Species, " ", "_")) 
 #Convert to long to model everything at the same time
 long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
 #Analysis preference~brain weight----
-priors_check= get_prior(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
-             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+#priors_check= get_prior(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
+#             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
 
-prior1 <- c(set_prior("normal(0,1)", class = "b", coef = "HabitatNatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatSeminatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatUrban"))
+#prior1 <- c(set_prior("normal(0,1)", class = "b", coef = "HabitatNatural"),
+#            set_prior("normal(0,1)", class = "b", coef = "HabitatSeminatural"),
+#            set_prior("normal(0,1)", class = "b", coef = "HabitatUrban"))
+
+#long_data = long_data %>% 
+#    mutate(Brain.weight = case_when(Brain.weight > 3 ~ 2.5, TRUE ~ as.numeric(Brain.weight)))
 
 model1 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), 
-            data = long_data,prior = prior1, data2 = list(A = A10), family=zero_one_inflated_beta())
+            data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
 
-pp_check(model1,nsamples=100) +ylim(0,5)
+pp_check(model1) +ylim(0,5)
 bayes_R2(model1)
 marginal_effects(model1)
 
 #Plotting preference~brain weight----
 ce1 <- conditional_effects(model1, effects = "Brain.weight:Habitat",points=T) 
+
+#Save data for plotting
[…]

### 2022-08-24 Scripts/8_JBL_Analysis_USA.R

@@ -20,24 +20,49 @@ long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
 
 #Analysis----
-s= get_prior(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
-             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
-
-
-prior1 <- c(set_prior("normal(0,1)", class = "b", coef = "HabitatNatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatSeminatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatUrban"))
-
-
-model = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
-                  data = long_data,prior=prior1, data2 = list(A = A10), family=zero_one_inflated_beta())
+model1 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
+                  data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
 
 
 #Check model fit
-pp_check(model,nsamples=100) + ylim(0,5)
+pp_check(model1,nsamples=100) + ylim(0,5)
 #Check r
-performance::r2(model)
+performance::r2(model1)
 #Check output
-marginal_effects(model)
+marginal_effects(model1)
 
+
+#Plotting preference~brain weight----
+ce1 <- conditional_effects(model1, effects = "Brain.weight:Habitat",points=T) 
+
+#Save data for plotting
+write.csv(long_data, "Data/Usa_data/data_preference_brain_weight_usa.csv")
+write.csv(ce1[[1]], "Data/Usa_data/model_output_preference_brain_weight_usa.csv")
+
+ggplot(ce1[[1]], aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+    geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Brain weight") +
+    ggtitle("USA occurrences")
+
+#Analysis----
+model2 = brm(Preference ~ IT * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
[…]

### 2022-08-24 Scripts/8_JBL_Analysis_all.R

@@ -24,13 +24,6 @@ d = left_join(preferences, brain_weight) %>%
 long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
 #Analysis preference~brain weight----
-priors_check= get_prior(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), warmup = 500, iter = 2000,
-                        data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
-
-prior1 <- c(set_prior("normal(0,1)", class = "b", coef = "HabitatNatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatSeminatural"),
-            set_prior("normal(0,1)", class = "b", coef = "HabitatUrban"))
-
 model1 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), 
              data = long_data,prior = prior1, data2 = list(A = A10), family=zero_one_inflated_beta())
 

### 2022-08-24 Scripts/8_JBL_Analysis_all.R
edits

@@ -25,7 +25,7 @@ long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
 #Analysis preference~brain weight----
 model1 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), 
-             data = long_data,prior = prior1, data2 = list(A = A10), family=zero_one_inflated_beta())
+             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
 
 pp_check(model1,nsamples=100) +ylim(0,5)
 bayes_R2(model1)
@@ -33,6 +33,9 @@ marginal_effects(model1)
 
 #Plotting preference~brain weight----
 ce1 <- conditional_effects(model1, effects = "Brain.weight:Habitat",points=T) 
+write.csv(long_data, "Data/Europe_USA/data_preference_brain_weight_europe_usa.csv")
+write.csv(ce1[[1]], "Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")
+
 
 ggplot(ce1[[1]], aes(x = Brain.weight, y = estimate__, color=Habitat)) +
     geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
@@ -47,12 +50,16 @@ ggplot(ce1[[1]], aes(x = Brain.weight, y = estimate__, color=Habitat)) +
 model2 = brm(Preference ~ IT * Habitat + (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
 
+
 pp_check(model2,nsamples=100) +ylim(0,5)
 bayes_R2(model2)
 marginal_effects(model2)
 
 #Plotting preference~brain weight----
 ce2 <- conditional_effects(model2, effects = "IT:Habitat",points=T) 
+
+#Save data for plotting
+write.csv(ce2[[1]], "Data/Europe_USA/model_output_preference_it_europe_usa.csv")
 
 ggplot(ce2[[1]], aes(x = IT, y = estimate__, color=Habitat)) +
     geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +

### 2022-08-24 Summary.Rmd
Fixing mistake on the summary results

@@
 
+United States land use classification with total number of occurrences within cover type in our dataset
+
 ```{r Table with all levels USA, echo = FALSE, warning = FALSE, message = FALSE}
@@
 kbl() %>% 
-kable_styling(latex_options = "hold_position", font_size = 12)
+kable_styling(latex_options = "hold_position", font_size = 10)
 

### 2022-08-26 Scripts/3_JBL_USA_Data_Cleaning.R
Big commit, organising and checking analysis

@@ -3,7 +3,7 @@
 #We take 5 key decisions in data cleaning
 #1)Filter records above 1987 (this is decided based on the GIS data available)
 #2)Filter by unique capture event
-#3)Filter by minimum of 3 decimals on coordinates
+#3)Filter by minimum of 2 decimals on coordinates #Because MA dataset has only 2
 #4)Filter number of levels per species (minimum N=100)
 #5)Filter by wide geographical distribution
 ######-
@@ -18,6 +18,24 @@ all_east_coast <- data.frame(fread("Data/Usa_data/usa_all_long_lat.csv.gz"))
 names(all_east_coast)[1:(ncol(all_east_coast)-1)] <- names(all_east_coast)[2:ncol(all_east_coast)]
 all_east_coast[, ncol(all_east_coast)] <- NULL
 all <- all_east_coast
+#Read MA tehsis data
+all_east_coast_thesis <- data.frame(fread("Data/Usa_data/usa_all_long_lat_ma_thesis_chapter.csv.gz"))
+#Fix colname position
+names(all_east_coast_thesis)[1:(ncol(all_east_coast_thesis)-1)] <- names(all_east_coast_thesis)[2:ncol(all_east_coast_thesis)]
+all_east_coast_thesis[, ncol(all_east_coast_thesis)] <- NULL
+all_thesis <- all_east_coast_thesis
+
+#Select columns of interest
+all_thesis = all_thesis %>% 
+select(gen_sp, state, lat, long, country) %>% 
+rename(species = gen_sp, Country = country, stateProvince = state)
+
+#Now mege both datasets
+colnames(all_thesis)
+colnames(all)
+
+#Rename to original name
+all = bind_rows(all, all_thesis)
 
 #############-
 #FIRST FILTER----
@@ -63,10 +81,8 @@ count_decimals = function(x) {
 all_unique_event$long_decimals <- as.numeric(all_unique_event$long) %>% count_decimals()
 all_unique_event$lat_decimals <- as.numeric(all_unique_event$lat) %>% count_decimals()
 #Now filter by minimum 3 decimals
-all_unique_event_3_decimals <- all_unique_event %>% filter(long_decimals > 2 & lat_decimals > 2)
+all_unique_event_3_decimals <- all_unique_event %>% filter(long_decimals > 1 & lat_decimals > 1)
 #Seems to work fine
-
-
 
 #############-
 #FOURTH FILTER----

### 2022-08-26 Scripts/6_JBL_Preferences_USA.R

@@ -54,58 +54,59 @@ summarise(n_rows = length(cover.names))
 pref.table = pref %>% 
 count(species, cover.names) %>%
 pivot_wider(names_from = cover.names, values_from = n, values_fill = list(n = 0)) %>% 
-column_to_rownames(var="species") %>% 
-select(Natural,Urban,Agricultural,Seminatural)
+column_to_rownames(var="species") 
 
 #Null model of our pref.table
 n.mod <- nullmodel(pref.table, N=10000, method="r2dtable")
 
 #Calculate preferences
-#Alternative way, probably more elegant
-
 m = pref.table #create matrix to store data
+#check cols
+colnames(pref.table)
 
-#Natural
-pref_nat = pref.table %>% select(Natural)
-n.mod_nat = n.mod
-for(n in 1:1000){
-    n.mod_nat[[n]] = n.mod_nat[[n]][,1]           
-} 
-
-for(k in 1:nrow(pref_nat)){
-    v <- lapply(n.mod_nat, `[[`, k)
-    m[k,1] <- sum(unlist(v) <  pref.table[k,1] ) / length(unlist(v)) 
-    
-}
-#Urban
-pref_urb = pref.table %>% select(Urban)
-n.mod_urb = n.mod
-for(n in 1:1000){
-    n.mod_urb[[n]] = n.mod_urb[[n]][,2]           
-} 
-
-for(k in 1:nrow(pref_urb)){
-    v <- lapply(n.mod_urb, `[[`, k)
-    m[k,2] <- sum(unlist(v) <  pref.table[k,2] ) / length(unlist(v)) 
-    
-}
 #Agricultural
-pref_agr = pref.table %>% select(Agricultural)
-n.mod_agr = n.mod
-for(n in 1:1000){
-    n.mod_agr[[n]] = n.mod_agr[[n]][,3]           
+pref_agr = pref.table %>% dplyr::select(Agricultural)
+n.mod_agr = list()
+for(n in 1:10000){
+    n.mod_agr[[n]] = n.mod[[n]][,1]           
 } 
 
 for(k in 1:nrow(pref_agr)){
     v <- lapply(n.mod_agr, `[[`, k)
+    m[k,1] <- sum(unlist(v) <  pref.table[k,1] ) / length(unlist(v)) 
+    
+}
+
+#Natural
[…]

### 2022-08-26 Summary.Rmd
Update on summary

@@
 ```{r Show levels Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
-pref = pref %>% 
+pref1 = pref %>% 
 mutate(Cover_names = fct_recode(as.factor(Cover_names),
@@
 
-pref %>% 
+pref1 %>% 
 group_by(Cover_names) %>% 
@@
 
-```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}
+```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
 
-data <- read_csv("Data/Europe_data/data_preference_brain_weight_europe.csv")
+long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe.csv")
 
-model1 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe.csv")
+model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe.csv")
 
-model2 = read_csv("Data/Europe_data/model_output_preference_it_europe.csv")
+model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe.csv")
 
-ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+model3 = read_csv("Data/Europe_data/model_output_preference_it_europe.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("Europe")
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Brain weight") +
-ggtitle("Europe occurrences")
+ggtitle("Europe")
 
-ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Intertegular distance") +
-ggtitle("Europe occurrences")
+ggtitle("Europe")
+
+library(patchwork)
+p1 / p2 / p3 & plot_layout(guides = "collect")
+
+```
+
+Now we make more strict land use classifications
+
+```{r Table with restricted levels Europe, echo = FALSE, warning = FALSE, message = FALSE}
+library(kableExtra)
+
+pref %>% 
+group_by(Cover_names) %>% 
+summarise(n_rows = length(Cover_names)) %>% 
+filter(!Cover_names=="NODATA") %>% 
+rename(Land_use = Cover_names) %>% 
+arrange(desc(n_rows)) %>% 
+mutate(Cover_names = fct_recode(as.factor(Land_use),
+#1 Check with Nacho, dehesas and others 
+Discard = "Agro-forestry areas", #1) 2.4.4 
+#2 Check with Nacho 
+Discard = "Airports", #2) 1.2.4 
+#3 Crops 
+Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
+#4 Very different cover type, not included 
+Discard = "Bare rocks", #4) 3.3.2 
+#5 Very different cover type, not included 
+Discard = "Beaches, dunes, sands", #5) 3.3.1 
+#6 Not included, lack of details 
+Discard = "Burnt areas", #6 3.3.4
+#7 Not included, only water surface
+Discard = "Coastal lagoons", #7) 5.2.1 
+#8 Straightforward
+Natural = "Coniferous forest", #8) 3.1.2
+#9 Excluded, complex to consider because it lacks habitat context
+Discard = "Construction sites", #9) 1.3.3
+#10 Based on the images and the description has a lot of vegetation cover
+Discard = "Discontinuous urban fabric", #10) 1.1.2
+#11 Pollinators here are likely to depend much on the context not on the dump
+Discard = "Dump sites", #11) 1.3.2
+#12 Filter out, water body
+Discard = "Estuaries", #12) 5.2.2 
+#13 Straightforward
+Agricultural = "Fruit trees and berry plantations", #13) 2.2.2
+#14 Urban as embedded within the city
+Discard = "Green urban areas", #14) 1.4.1
+#15 Highly modified
+Urban = "Industrial or commercial units", #15) 1.2.1
+#16 By definion, seminatural
+Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #16) 2.4.3
+#17 Discard as is not clear the vegetation context
+Discard = "Mineral extraction sites", #17) 1.3.1
+#18 Straightforward 3.1.3
+Natural = "Mixed forest", #18) 3.1.3
+#19 Crystal clear category
+Discard = "Moors and heathland", #19) 3.2.2
+#20 Crystal clear category
+Discard = "Natural grasslands", #20) 3.2.1
+#21 Filter out
+Discard = "NODATA", #21 
+#22 cultivated land
+Agricultural = "Non-irrigated arable land", #22) 2.1.1
+#23 As it is a monoculture despite some very
+Discard = "Olive groves", #23) 2.2.3
+#24 Crystal clear category
+Discard = "Peat bogs", #24) 4.1.2
+#25 Crystal clear category
+Agricultural = "Permanently irrigated land", #25) 2.1.2
+#26 Check with Nacho
+Discard = "Port areas", #26) 1.2.3
+#27 Check with Nacho
+Discard = "Road and rail networks and associated land", #27) 1.2.2
+#28 Check with Nacho but just only 200 records
+Discard = "Salines", #28) 4.2.2 
+#29 Flowering plant communities 
+Discard = "Salt marshes", #29) 4.2.1
+#30 Sclerophyllous shrubs and low shrubs
+Discard = "Sclerophyllous vegetation", #30) 3.2.3
+#31 Discard for now 
+Discard = "Sea and ocean", #31) 5.2.3
+#32 Areas with sparse vegetation, covering 10-50% of surface 
+Discard = "Sparsely vegetated areas", #32) 3.3.3
+#33 Check with Nacho 
+Discard = "Sport and leisure facilities", #33) 1.4.2 
+#34 Excluded as it is very general
+Discard = "Water bodies", #34) 5.1.2
+#35 Excluded as it is very general
+Discard = "Water courses" #35 5.1.1 
+)) %>%  
+kbl() %>% 
+kable_styling(font_size = 8,latex_options = "hold_position")
+
+```
+
+Check final levels per category:
+
+```{r Show levels restricted land use Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
+pref2 = pref %>% 
+mutate(Cover_names = fct_recode(as.factor(Cover_names),
+#1 Check with Nacho, dehesas and others 
+Discard = "Agro-forestry areas", #1) 2.4.4 
+#2 Check with Nacho 
+Discard = "Airports", #2) 1.2.4 
+#3 Crops 
+Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
+#4 Very different cover type, not included 
+Discard = "Bare rocks", #4) 3.3.2 
+#5 Very different cover type, not included 
+Discard = "Beaches, dunes, sands", #5) 3.3.1 
+#6 Not included, lack of details 
+Discard = "Burnt areas", #6 3.3.4
+#7 Not included, only water surface
+Discard = "Coastal lagoons", #7) 5.2.1 
+#8 Straightforward
+Natural = "Coniferous forest", #8) 3.1.2
+#9 Excluded, complex to consider because it lacks habitat context
+Discard = "Construction sites", #9) 1.3.3
+#10 Based on the images and the description has a lot of vegetation cover
+Discard = "Discontinuous urban fabric", #10) 1.1.2
+#11 Pollinators here are likely to depend much on the context not on the dump
+Discard = "Dump sites", #11) 1.3.2
+#12 Filter out, water body
+Discard = "Estuaries", #12) 5.2.2 
+#13 Straightforward
+Agricultural = "Fruit trees and berry plantations", #13) 2.2.2
+#14 Urban as embedded within the city
+Discard = "Green urban areas", #14) 1.4.1
+#15 Highly modified
+Urban = "Industrial or commercial units", #15) 1.2.1
+#16 By definion, seminatural
+Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #16) 2.4.3
+#17 Discard as is not clear the vegetation context
+Discard = "Mineral extraction sites", #17) 1.3.1
+#18 Straightforward 3.1.3
+Natural = "Mixed forest", #18) 3.1.3
+#19 Crystal clear category
+Discard = "Moors and heathland", #19) 3.2.2
+#20 Crystal clear category
+Discard = "Natural grasslands", #20) 3.2.1
+#21 Filter out
+Discard = "NODATA", #21 
+#22 cultivated land
+Agricultural = "Non-irrigated arable land", #22) 2.1.1
+#23 As it is a monoculture despite some very
+Discard = "Olive groves", #23) 2.2.3
+#24 Crystal clear category
+Discard = "Peat bogs", #24) 4.1.2
+#25 Crystal clear category
+Agricultural = "Permanently irrigated land", #25) 2.1.2
+#26 Check with Nacho
+Discard = "Port areas", #26) 1.2.3
+#27 Check with Nacho
+Discard = "Road and rail networks and associated land", #27) 1.2.2
+#28 Check with Nacho but just only 200 records
+Discard = "Salines", #28) 4.2.2 
+#29 Flowering plant communities 
+Discard = "Salt marshes", #29) 4.2.1
+#30 Sclerophyllous shrubs and low shrubs
+Discard = "Sclerophyllous vegetation", #30) 3.2.3
+#31 Discard for now 
+Discard = "Sea and ocean", #31) 5.2.3
+#32 Areas with sparse vegetation, covering 10-50% of surface 
+Discard = "Sparsely vegetated areas", #32) 3.3.3
+#33 Check with Nacho 
+Discard = "Sport and leisure facilities", #33) 1.4.2 
+#34 Excluded as it is very general
+Discard = "Water bodies", #34) 5.1.2
+#35 Excluded as it is very general
+Discard = "Water courses" #35 5.1.1 
+)) %>% 
+filter(!Cover_names == "Discard")
+
+pref2 %>% 
+group_by(Cover_names) %>% 
+summarise(Levels = length(Cover_names)) %>% 
+kbl() %>% 
+kable_styling(latex_options = "hold_position")
+```
+
+\newpage
+
+And we model the data with this categories:
+
+```{r Plotting Europe subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
+
+long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_subset.csv")
+
+model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe_subset.csv")
+
+model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe_subset.csv")
+
+model3 = read_csv("Data/Europe_data/model_output_preference_it_europe_subset.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("Europe")
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("Europe")
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("Europe")
+
+library(patchwork)
+p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")
 
@@
 summarise(n_rows = length(cover.names)) %>% 
+arrange(desc(n_rows)) %>%     
 kbl() %>% 
@@
 
+\newpage
+
 Then, we calculate preferences and model the data.
 
-```{r Plotting USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}
+```{r Plotting USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
 
-data <- read_csv("Data/Usa_data/data_preference_brain_weight_usa.csv")
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa.csv")
 
-model1 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa.csv")
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa.csv")
 
-model2 = read_csv("Data/Usa_data/model_output_preference_it_usa.csv")
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa.csv")
 
-ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA")
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Brain weight") +
-ggtitle("USA occurrences")
+ggtitle("USA")
 
-ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Intertegular distance") +
-ggtitle("USA occurrences")
+ggtitle("USA")
+
+library(patchwork)
+p1 / p2 / p3 & plot_layout(guides = "collect")
 
@@
 I have explored a bit and the results were quite similar overall but reviewers seem to don't like this approach as one of the goals of Bayesian is to use the prior knowledge of our variables to improve our predictions.
+
+\newpage
+
+Now we make more strict categories of what is Natural, Urban and Agricultural. See below:
+
+```{r Show levels USA subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
+
+pref_subset <- read_csv("Data/Usa_data/land_cover_usa.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor)
+
+pref_subset %>% 
+group_by(cover.names) %>% 
+summarise(n_rows = length(cover.names)) %>% 
+rename(Land_use = cover.names) %>% 
+arrange(desc(n_rows)) %>% 
+mutate(Cover_names = fct_recode(as.factor(Land_use),
+Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+Agricultural = "Cultivated Crops", #(MA & JB)
+Natural = "Deciduous Forest", #(MA & JB)
+Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+Discard = "Developed, Low Intensity", #Low urban impact (MA & JB)
+Discard = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+Discard = "Emergent Herbaceous Wetlands", #(MA & JB)
+Natural = "Evergreen Forest", #(MA & JB)
+Discard = "Hay/Pasture", #Highly managed by humans (MA & JB)
+Discard = "Herbaceous", #See explanation below (MA & JB)
+Natural = "Mixed Forest", #(MA & JB)
+Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
+Natural = "Shrub/Scrub", #Heath classified as natural (MA & JB)
+Natural = "Woody Wetlands")) %>% 
+kbl() %>% 
+kable_styling(latex_options = "hold_position", font_size = 10)
+```
+
+Check number of levels per category:
+
+```{r Show levels USA subset levels, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
+
+ pref_subset %>% 
+mutate(cover.names = fct_recode(as.factor(cover.names),
+Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+Agricultural = "Cultivated Crops", #(MA & JB)
+Natural = "Deciduous Forest", #(MA & JB)
+Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+Discard = "Developed, Low Intensity", #Low urban impact (MA & JB)
+Discard = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+Discard = "Emergent Herbaceous Wetlands", #(MA & JB)
+Natural = "Evergreen Forest", #(MA & JB)
+Discard = "Hay/Pasture", #Highly managed by humans (MA & JB)
+Discard = "Herbaceous", #See explanation below (MA & JB)
+Natural = "Mixed Forest", #(MA & JB)
+Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
+Natural = "Shrub/Scrub", #Heath classified as natural (MA & JB)
+Natural = "Woody Wetlands")) %>% #(MA & JB)
+filter(!cover.names == "Discard") %>% 
+group_by(cover.names) %>% 
+summarise(n_rows = length(cover.names)) %>% 
+arrange(desc(n_rows)) %>%     
+kbl() %>% 
+kable_styling(latex_options = "hold_position")
+
+```
+
+And we model the data with this categories:
+
+```{r Plotting USA subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
+
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset.csv")
+
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_subset.csv")
+
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_subset.csv")
+
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_subset.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA")
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA")
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA")
+
+library(patchwork)
+p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")
+
+```
 

### 2022-08-31 trial.Rmd
organising analysis

---
title: "Untitled"
output: html_document
---

```{r Plotting USA Europe together, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}

library(tidyverse)

long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")

model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")

model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")

model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")

model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa_qualitative.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Load land use filtered data and models
long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_subset_qualitative.csv")

model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa_subset_qualitative.csv")

model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa_subset_qualitative.csv")

model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa_subset_qualitative.csv")

model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa_subset_qualitative.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight /` IT") +
ggtitle("USA levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

```

### 2022-08-31 Scripts/5_2_JBL_USA_Land_Cover_Gbif.R
analyses

@@ -1,0 +1,82 @@
+#Read clean USA data and extract land use from raster file
+#Load libraries
+library(data.table)
+library(tidyverse)
+
+#Load data
+usa <- read.table("Data/Usa_data/all_above_50_usa_gbif.csv.gz",  header=T, quote="\"", sep=",")
+
+#check levels, min 106 max just over 15.0000
+usa %>% 
+    group_by(species) %>% 
+    summarize(no_rows= length(species))
+
+#Extract land use
+#Code from using this code as an example https://www.r-bloggers.com/2014/11/spatial-data-extraction-around-buffered-points-in-r/
+
+year=2011
+
+extract_cover <- function(year,
+                          point_d = "Data/Usa_data/all_above_50_usa_gbif.csv.gz",  
+                          data_dir="Data/Raster_USA",
+                          write_dir="Data/Usa_data"){
+    require(raster)
+    require(rgdal)
+    require(stringr)
+    
+    # load cover data 
+    filename <- paste(data_dir, "/nlcd_",
+                      year,
+                      "_land_cover_l48_20210604.img",
+                      sep="")
+    NLCD <- raster(filename)
+    
+    # load site data
+    Datos <- read.table(point_d,  header=T, quote="\"", sep=",") #Si falla, usar como sep "," o ";"
+    #según tus datos
+    coords <- Datos[, c("longitude", "latitude")]  
+    #convert lat/lon to appropriate projection
+    names(coords) <- c("x", "y")
+    coordinates(coords) <- ~x + y
+    proj4string(coords) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
+    crs_args <- NLCD@crs@projargs
+    Datos_transformed <- spTransform(coords, CRS(crs_args))
+    
+    #extract land cover data for each point, given buffer size
+    Landcover <- raster::extract(NLCD, Datos_transformed)
+    
+    # summarize each site's data by proportion of each cover type
+    summ <- lapply(Landcover, function(x){
+        prop.table(table(x))
+    }
+    )
+    
+    # generate land cover number to name conversions
+    num.codes <- unique(unlist(Landcover))
+    cover.names <- NLCD@data@attributes[[1]]$NLCD.Land.Cover.Class[num.codes + 1]
+    levels(cover.names)[1] <- NA # first level is ""
+    conversions <- data.frame(num.codes, cover.names)
[…]

### 2022-08-31 Scripts/6_2_JBL_Preferences_USA_Gbif.R

@@ -1,0 +1,121 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Usa_data/land_cover_usa_gbif.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, species
+pref %>% 
+    group_by(species) %>% 
+    summarise(n_rows = length(species))
+#Second, cover.names
+S = pref %>% 
+    group_by(cover.names) %>% 
+    summarise(n_rows = length(cover.names))
+
+#Recode levels to calculate preferences
+#Check pdf with landcover types in Data/Usa_data
+#and make decisions based on that
+pref = pref %>% 
+    mutate(cover.names = fct_recode(as.factor(cover.names),
+                                    Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+                                    Agricultural = "Cultivated Crops", #(MA & JB)
+                                    Natural = "Deciduous Forest", #(MA & JB)
+                                    Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Seminatural = "Developed, Low Intensity", #Low urban impact (MA & JB)
+                                    Seminatural = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+                                    Natural = "Emergent Herbaceous Wetlands", #(MA & JB)
+                                    Natural = "Evergreen Forest", #(MA & JB)
+                                    Agricultural = "Hay/Pasture", #Highly managed by humans (MA & JB)
+                                    Natural = "Herbaceous", #See explanation below (MA & JB)
+                                    Natural = "Mixed Forest", #(MA & JB)
+                                    Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
[…]

### 2022-08-31 Scripts/6_2_JBL_Preferences_USA_Subset_Gbif.R

@@ -1,0 +1,109 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Usa_data/land_cover_usa_gbif.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, species
+pref %>% 
+    group_by(species) %>% 
+    summarise(n_rows = length(species))
+#Second, cover.names
+S = pref %>% 
+    group_by(cover.names) %>% 
+    summarise(n_rows = length(cover.names))
+
+#Recode levels to calculate preferences
+#Check pdf with landcover types in Data/Usa_data
+#and make decisions based on that
+pref = pref %>% 
+    mutate(cover.names = fct_recode(as.factor(cover.names),
+                                    Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+                                    Agricultural = "Cultivated Crops", #(MA & JB)
+                                    Natural = "Deciduous Forest", #(MA & JB)
+                                    Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Discard = "Developed, Low Intensity", #Low urban impact (MA & JB)
+                                    Discard = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+                                    Discard = "Emergent Herbaceous Wetlands", #(MA & JB)
+                                    Natural = "Evergreen Forest", #(MA & JB)
+                                    Discard = "Hay/Pasture", #Highly managed by humans (MA & JB)
+                                    Discard = "Herbaceous", #See explanation below (MA & JB)
+                                    Natural = "Mixed Forest", #(MA & JB)
+                                    Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
+                                    Natural = "Shrub/Scrub", #Heath classified as natural (MA & JB)
[…]

### 2022-08-31 Scripts/7_2_JBL_USA_Phylogeny_Gbif.R

@@ -1,35 +1,28 @@
-#Calculate phylogenetic matrix for USA and Europe species
+#Calculate phylogenetic matrix for USA species
 library(visreg) 
 library(data.tree)
 library(ape)
 library(phytools)
-library(tidyverse)
-
 #Load extracted data
-usa <- read_csv("Data/Usa_data/land_cover_usa.csv.gz") %>% 
+pref <- read_csv("Data/Usa_data/land_cover_usa_gbif.csv.gz") %>% 
     dplyr::select(species, cover.names) %>% 
     mutate_if(is.character,as.factor) 
-#Select species
-species_list_usa = usa %>% 
-distinct(species) %>% 
-rename(Species = species)
 
-#Load extracted data
-eur <- read_csv("Data/Europe_data/land_cover_europe.csv.gz") %>% 
-dplyr::select(Species, Cover_names) %>% 
-mutate_if(is.character,as.factor) 
-#Select species
-species_list_eur = eur %>% 
-distinct(Species) 
+#Check levels of species and cover names
+#First, Species
+species_list = pref %>% 
+    distinct(species) 
 
-#Bind lists
-species = bind_rows(species_list_usa , species_list_eur) %>% 
-mutate(species = str_replace_all(Species, " ", "_")) %>% 
-distinct(species) %>% 
-pull()
+#-----
+#Lista de especies
+bee.trees=read.tree(file="Data/phylogeny_genus_level.txt")
+
+#Generate Species vector 
+species = species_list %>%  
+    mutate(species = str_replace_all(species, " ", "_")) %>% 
+    pull()
 
 #Pick tree 1
-bee.trees=read.tree(file="Data/phylogeny_genus_level.txt")
 bee.mcmc=bee.trees[[1]]
 #Make a wasp the outgroup
 bee.mcmc=root(bee.mcmc,outgroup="Tachysphex")
@@ -39,6 +32,7 @@ bee.mcmc=chronos(bee.mcmc)
 
 
 bee.mcmc$tip.label
+species
 
 #Add genus that I don't have 
 bee.tree100=drop.tip(bee.mcmc, tip = c("Xenochilicola", "Geodiscelis", "Xeromelissa", "Chilimelissa",    
@@ -166,7 +160,6 @@ plot(bee.tree100, cex = 0.6)
 
 ##Check for missing species
 setdiff(species,bee.tree100$tip.label)
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_All_Qualitative_Subset.R

@@ -1,4 +1,4 @@
-#Analysis of USA bee preferences
+#Analysis of Europe bee preferences
 
 #Load libraries
 library(tidyverse) 
@@ -8,20 +8,28 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences = read_csv("Data/Usa_data/preferences_usa_subset.csv") 
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative_subset.csv") 
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative_subset.csv") 
+#merge datasets
+preferences = bind_rows(preferences_europe, preferences_usa)
+
 #Load phylogenetic matrix to correct for in analysis
-A10 = readRDS("Data/Usa_data/phylo_usa.rds")
+A10 = readRDS("Data/phylo_all.rds")
 
 #Prepare data----
 d = left_join(preferences, brain_weight) %>% 
     mutate(Species = str_replace_all(Species, " ", "_"))
+
 #Convert to long to model everything at the same time
 long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 
+#Prepare col brain weight/IT
+long_data = long_data %>% 
+    mutate(brain_it = Brain.weight/IT)
 
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
-             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+             data = long_data, data2 = list(A = A10), family=bernoulli())
 
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
@@ -33,18 +41,19 @@ p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
     theme_bw() +
     ylab("Habitat preference") +
     xlab("Residuals") + 
-    ggtitle("USA")
+    ggtitle("USA and Europe")
 
 #Save data and model1 output
-write_csv(long_data , "Data/Usa_data/data_preference_residuals_usa_subset.csv")
-write_csv(ce1[[1]], "Data/Usa_data/model_output_preference_residuals_usa_subset.csv")
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_All_Subset.R

@@ -8,8 +8,8 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") 
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_subset.csv") 
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_subset.csv") 
 #merge datasets
 preferences = bind_rows(preferences_europe, preferences_usa)
 
@@ -21,7 +21,7 @@ d = left_join(preferences, brain_weight) %>%
     mutate(Species = str_replace_all(Species, " ", "_"))
 
 #Convert to long to model everything at the same time
-long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
+long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 
 #Prepare col brain weight/IT
 long_data = long_data %>% 
@@ -34,19 +34,18 @@ model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)),
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
 bayes_R2(model1)
-marginal_effects(model1)
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA and Europe")
+    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Residuals") + 
+    ggtitle("USA and Europe")
 
 #Save data and model1 output
-write_csv(long_data , "Data/Europe_USA/data_preference_residuals_europe_usa.csv")
-write_csv(ce1[[1]], "Data/Europe_USA/model_output_preference_residuals_europe_usa.csv")
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_Europe.R

@@ -19,6 +19,10 @@ mutate(Species = str_replace_all(Species, " ", "_"))
 #Convert to long to model everything at the same time
 long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
+#Prepare col brain weight/IT
+long_data = long_data %>% 
+    mutate(brain_it = Brain.weight/IT)
+
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
@@ -86,3 +90,23 @@ write_csv(ce3[[1]], "Data/Europe_data/model_output_preference_it_europe.csv")
 #Plot all together
 library(patchwork)
 p1 / p2 / p3
+
+
+#Analysis Preference ~ brain_it----
+model4 = brm(Preference ~ brain_it * Habitat + (1|gr(Species, cov = A)), 
+             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+
+ce4 <- conditional_effects(model4, effects = "brain_it:Habitat",points=T) 
+
+p4 = ggplot(ce4[[1]], aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("brain_it") +
+ggtitle("Europe") 
+
+#Save data
+write_csv(ce4[[1]], "Data/Europe_data/model_output_preference_brain_it_europe.csv")
+
+

### 2022-08-31 Scripts/8_JBL_Analysis_Europe_Qualitative.R

@@ -1,0 +1,288 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Europe_data/land_cover_europe.csv.gz") %>% 
+    dplyr::select(Species, Cover_names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, Species
+check_species = pref %>% 
+    group_by(Species) %>% 
+    summarise(n_rows = length(Species))
+#Second, Cover_names
+check_cover = pref %>% 
+    group_by(Cover_names) %>% 
+    summarise(n_rows = length(Cover_names))
+
+#Recode levels to calculate preferences
+#Check levels here
+pref = pref %>% 
+    mutate(Cover_names = fct_recode(as.factor(Cover_names),
+                                    #1 Check with Nacho, dehesas and others 
+                                    Natural = "Agro-forestry areas", #1) 2.4.4 
+                                    #2 Check with Nacho 
+                                    Urban = "Airports", #2) 1.2.4 
+                                    #3 Crops 
+                                    Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
+                                    #4 Very different cover type, not included 
+                                    Discard = "Bare rocks", #4) 3.3.2 
+                                    #5 Very different cover type, not included 
+                                    Natural = "Beaches, dunes, sands", #5) 3.3.1 
+                                    #6 Not included, lack of details 
+                                    Discard = "Burnt areas", #6 3.3.4
+                                    #7 Not included, only water surface
+                                    Natural = "Coastal lagoons", #7) 5.2.1 
+                                    #8 Straightforward
+                                    Natural = "Coniferous forest", #8) 3.1.2
+                                    #9 Excluded, complex to consider because it lacks habitat context
+                                    Urban = "Construction sites", #9) 1.3.3
+                                    #10 Based on the images and the description has a lot of vegetation cover
+                                    Seminatural = "Discontinuous urban fabric", #10) 1.1.2
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_Europe_Qualitative_Subset.R

@@ -1,0 +1,279 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Europe_data/land_cover_europe.csv.gz") %>% 
+    dplyr::select(Species, Cover_names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, Species
+check_species = pref %>% 
+    group_by(Species) %>% 
+    summarise(n_rows = length(Species))
+#Second, Cover_names
+check_cover = pref %>% 
+    group_by(Cover_names) %>% 
+    summarise(n_rows = length(Cover_names))
+
+#Recode levels to calculate preferences
+#Check levels here
+pref = pref %>% 
+    mutate(Cover_names = fct_recode(as.factor(Cover_names),
+                                    #1 Check with Nacho, dehesas and others 
+                                    Discard = "Agro-forestry areas", #1) 2.4.4 
+                                    #2 Check with Nacho 
+                                    Discard = "Airports", #2) 1.2.4 
+                                    #3 Crops 
+                                    Agricultural = "Annual crops associated with permanent crops", #3) 2.4.1 
+                                    #4 Very different cover type, not included 
+                                    Discard = "Bare rocks", #4) 3.3.2 
+                                    #5 Very different cover type, not included 
+                                    Discard = "Beaches, dunes, sands", #5) 3.3.1 
+                                    #6 Not included, lack of details 
+                                    Discard = "Burnt areas", #6 3.3.4
+                                    #7 Not included, only water surface
+                                    Discard = "Coastal lagoons", #7) 5.2.1 
+                                    #8 Straightforward
+                                    Natural = "Coniferous forest", #8) 3.1.2
+                                    #9 Excluded, complex to consider because it lacks habitat context
+                                    Discard = "Construction sites", #9) 1.3.3
+                                    #10 Based on the images and the description has a lot of vegetation cover
+                                    Discard = "Discontinuous urban fabric", #10) 1.1.2
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_Europe_Subset.R

@@ -19,6 +19,10 @@ d = left_join(preferences, brain_weight) %>%
 #Convert to long to model everything at the same time
 long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 
+#Prepare col brain weight/IT
+long_data = long_data %>% 
+    mutate(brain_it = Brain.weight/IT)
+
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
@@ -86,3 +90,20 @@ write_csv(ce3[[1]], "Data/Europe_data/model_output_preference_it_europe_subset.c
 #Plot all together
 library(patchwork)
 p1 / p2 / p3
+
+#Analysis Preference ~ brain_it----
+model4 = brm(Preference ~ brain_it * Habitat + (1|gr(Species, cov = A)), 
+             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+
+ce4 <- conditional_effects(model4, effects = "brain_it:Habitat",points=T) 
+
+p4 = ggplot(ce4[[1]], aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("brain_it") +
+ggtitle("Europe") 
+
+#Save data
+write_csv(ce4[[1]], "Data/Europe_data/model_output_preference_brain_it_europe_subset.csv")

### 2022-08-31 Scripts/8_JBL_Analysis_Qualitative.R

@@ -8,8 +8,8 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") 
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative.csv") 
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative.csv") 
 #merge datasets
 preferences = bind_rows(preferences_europe, preferences_usa)
 
@@ -29,7 +29,7 @@ long_data = long_data %>%
 
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
-             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+             data = long_data, data2 = list(A = A10), family=bernoulli())
 
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
@@ -37,20 +37,20 @@ bayes_R2(model1)
 marginal_effects(model1)
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA and Europe")
+    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Residuals") + 
+    ggtitle("USA and Europe")
 
 #Save data and model1 output
-write_csv(long_data , "Data/Europe_USA/data_preference_residuals_europe_usa.csv")
-write_csv(ce1[[1]], "Data/Europe_USA/model_output_preference_residuals_europe_usa.csv")
+write_csv(long_data , "Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
+write_csv(ce1[[1]], "Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")
 
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Gbif.R

@@ -1,4 +1,4 @@
-#Analysis of Europe bee preferences
+#Analysis of USA bee preferences
 
 #Load libraries
 library(tidyverse) 
@@ -8,13 +8,9 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") 
-#merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
+preferences = read_csv("Data/Usa_data/preferences_usa_gbif.csv") 
 #Load phylogenetic matrix to correct for in analysis
-A10 = readRDS("Data/phylo_all.rds")
+A10 = readRDS("Data/Usa_data/phylo_usa_gbif.rds")
 
 #Prepare data----
 d = left_join(preferences, brain_weight) %>% 
@@ -34,19 +30,18 @@ model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)),
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
 bayes_R2(model1)
-marginal_effects(model1)
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA and Europe")
+    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Residuals") + 
+    ggtitle("USA")
 
 #Save data and model1 output
-write_csv(long_data , "Data/Europe_USA/data_preference_residuals_europe_usa.csv")
-write_csv(ce1[[1]], "Data/Europe_USA/model_output_preference_residuals_europe_usa.csv")
+write_csv(long_data , "Data/Usa_data/data_preference_residuals_usa_gbif.csv")
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Gbif_Thesis_Data.R

@@ -19,6 +19,10 @@ mutate(Species = str_replace_all(Species, " ", "_"))
 #Convert to long to model everything at the same time
 long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
 
+#Prepare col brain weight/IT
+long_data = long_data %>% 
+    mutate(brain_it = Brain.weight/IT)
+
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
@@ -80,3 +84,23 @@ write_csv(ce3[[1]], "Data/Usa_data/model_output_preference_it_usa.csv")
 #Plot all together
 library(patchwork)
 p1 / p2 / p3
+
+#Analysis Preference ~ brain_it----
+model4 = brm(Preference ~ brain_it * Habitat + (1|gr(Species, cov = A)), 
+             data = long_data, data2 = list(A = A10), family=zero_one_inflated_beta())
+
+ce4 <- conditional_effects(model4, effects = "brain_it:Habitat",points=T) 
+
+p4 = ggplot(ce4[[1]], aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("brain_it") +
+ggtitle("USA") 
+
+#Save data
+write_csv(ce4[[1]], "Data/Usa_data/model_output_preference_brain_it_usa.csv")
+
+
+

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Gbif_Thesis_Data_Qualitative.R

@@ -1,0 +1,233 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Usa_data/land_cover_usa.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, species
+pref %>% 
+    group_by(species) %>% 
+    summarise(n_rows = length(species))
+#Second, cover.names
+S = pref %>% 
+    group_by(cover.names) %>% 
+    summarise(n_rows = length(cover.names))
+
+#Recode levels to calculate preferences
+#Check pdf with landcover types in Data/Usa_data
+#and make decisions based on that
+pref = pref %>% 
+    mutate(cover.names = fct_recode(as.factor(cover.names),
+                                    Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+                                    Agricultural = "Cultivated Crops", #(MA & JB)
+                                    Natural = "Deciduous Forest", #(MA & JB)
+                                    Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Seminatural = "Developed, Low Intensity", #Low urban impact (MA & JB)
+                                    Seminatural = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+                                    Natural = "Emergent Herbaceous Wetlands", #(MA & JB)
+                                    Natural = "Evergreen Forest", #(MA & JB)
+                                    Agricultural = "Hay/Pasture", #Highly managed by humans (MA & JB)
+                                    Natural = "Herbaceous", #See explanation below (MA & JB)
+                                    Natural = "Mixed Forest", #(MA & JB)
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Qualitative_Gbif.R

@@ -1,0 +1,232 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Usa_data/land_cover_usa_gbif.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, species
+pref %>% 
+    group_by(species) %>% 
+    summarise(n_rows = length(species))
+#Second, cover.names
+S = pref %>% 
+    group_by(cover.names) %>% 
+    summarise(n_rows = length(cover.names))
+
+#Recode levels to calculate preferences
+#Check pdf with landcover types in Data/Usa_data
+#and make decisions based on that
+pref = pref %>% 
+    mutate(cover.names = fct_recode(as.factor(cover.names),
+                                    Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+                                    Agricultural = "Cultivated Crops", #(MA & JB)
+                                    Natural = "Deciduous Forest", #(MA & JB)
+                                    Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Seminatural = "Developed, Low Intensity", #Low urban impact (MA & JB)
+                                    Seminatural = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+                                    Natural = "Emergent Herbaceous Wetlands", #(MA & JB)
+                                    Natural = "Evergreen Forest", #(MA & JB)
+                                    Agricultural = "Hay/Pasture", #Highly managed by humans (MA & JB)
+                                    Natural = "Herbaceous", #See explanation below (MA & JB)
+                                    Natural = "Mixed Forest", #(MA & JB)
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Subset_Gbif.R

@@ -1,4 +1,4 @@
-#Analysis of Europe bee preferences
+#Analysis of USA bee preferences
 
 #Load libraries
 library(tidyverse) 
@@ -8,20 +8,15 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") 
-#merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
+preferences = read_csv("Data/Usa_data/preferences_usa_subset_gbif.csv") 
 #Load phylogenetic matrix to correct for in analysis
-A10 = readRDS("Data/phylo_all.rds")
+A10 = readRDS("Data/Usa_data/phylo_usa_gbif.rds")
 
 #Prepare data----
 d = left_join(preferences, brain_weight) %>% 
     mutate(Species = str_replace_all(Species, " ", "_"))
-
 #Convert to long to model everything at the same time
-long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
+long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 
 #Prepare col brain weight/IT
 long_data = long_data %>% 
@@ -34,19 +29,18 @@ model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)),
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
 bayes_R2(model1)
-marginal_effects(model1)
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA and Europe")
+    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Residuals") + 
+    ggtitle("USA")
 
 #Save data and model1 output
[…]

### 2022-08-31 Scripts/8_JBL_Analysis_USA_Subset_Gbif_Thesis_Data.R

@@ -1,4 +1,4 @@
-#Analysis of Europe bee preferences
+#Analysis of USA bee preferences
 
 #Load libraries
 library(tidyverse) 
@@ -8,20 +8,15 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") 
-#merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
+preferences = read_csv("Data/Usa_data/preferences_usa_subset.csv") 
 #Load phylogenetic matrix to correct for in analysis
-A10 = readRDS("Data/phylo_all.rds")
+A10 = readRDS("Data/Usa_data/phylo_usa.rds")
 
 #Prepare data----
 d = left_join(preferences, brain_weight) %>% 
     mutate(Species = str_replace_all(Species, " ", "_"))
-
 #Convert to long to model everything at the same time
-long_data = d %>% gather(Habitat, Preference, 2:5, -c(Species))
+long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 
 #Prepare col brain weight/IT
 long_data = long_data %>% 
@@ -34,19 +29,18 @@ model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)),
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
 
 bayes_R2(model1)
-marginal_effects(model1)
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA and Europe")
+    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+    geom_line(aes(color=Habitat)) +
+    theme_bw() +
+    ylab("Habitat preference") +
+    xlab("Residuals") + 
+    ggtitle("USA")
 
 #Save data and model1 output
[…]

### 2022-08-31 Scripts/8_JBL_Preferences_USA_Subset_Gbif_Thesis_Data_Qualitative.R

@@ -1,0 +1,221 @@
+#Calculate habitat preferences
+
+#Load library
+library(tidyverse)
+library(bipartite)
+
+#Load extracted data
+pref <- read_csv("Data/Usa_data/land_cover_usa.csv.gz") %>% 
+    dplyr::select(species, cover.names) %>% 
+    mutate_if(is.character,as.factor) 
+
+#Check levels of species and cover names
+#First, species
+pref %>% 
+    group_by(species) %>% 
+    summarise(n_rows = length(species))
+#Second, cover.names
+S = pref %>% 
+    group_by(cover.names) %>% 
+    summarise(n_rows = length(cover.names))
+
+#Recode levels to calculate preferences
+#Check pdf with landcover types in Data/Usa_data
+#and make decisions based on that
+pref = pref %>% 
+    mutate(cover.names = fct_recode(as.factor(cover.names),
+                                    Discard = "Barren Land", #Very different cover type, not included (MA & JB)
+                                    Agricultural = "Cultivated Crops", #(MA & JB)
+                                    Natural = "Deciduous Forest", #(MA & JB)
+                                    Urban = "Developed, High Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Urban = "Developed, Medium Intensity", #Greater than 50% of cover (MA & JB) 
+                                    Discard = "Developed, Low Intensity", #Low urban impact (MA & JB)
+                                    Discard = "Developed, Open Space", #Open areas with vegetation (MA & JB)
+                                    Discard = "Emergent Herbaceous Wetlands", #(MA & JB)
+                                    Natural = "Evergreen Forest", #(MA & JB)
+                                    Discard = "Hay/Pasture", #Highly managed by humans (MA & JB)
+                                    Discard = "Herbaceous", #See explanation below (MA & JB)
+                                    Natural = "Mixed Forest", #(MA & JB)
+                                    Discard = "Open Water", #Low vegetation and land, not included (MA & JB)
[…]

### 2022-10-19 Manuscript/Ms.Rmd
ms draft

---
title: "Untitled"
output: pdf_document

linkcolor: RoyalBlue
urlcolor: RoyalBlue
fontsize: 12pt

bibliography: 
  - references.bib
---

Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain

# Abstract

Keywords:

\textbf{Author for correspondence:}

# Introduction

# Methods

# Results

# Discussion

### 2022-10-19 Manuscript/Ms.Rmd
MS

@@
 
+The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role in ecosystems for fruit and seed production.
+However, not all pollinators are negatively affected by habitat conversion, but some find in anthropogenic landscapes appropriate resources to survive and reproduce.
+Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities.
+While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
+To address this issue, we measured brains for 145 individuals from 44 species of bees —the most important group of pollinators— from North America and Europe, and compared them with detailed information on habitat preferences and population trends.
+Our analyses revealed that bees that prefer forest habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger brains compared to forest dwellers.
+However, we found no evidence that either larger brains or a preference for urban habitats are associated with positive population trends.
+These results suggest that a large brain can help maintaining bee urban populations, but also highlight that being tolerant to urbanisation is not enough to compensate for the worldwide decline of their populations.
+
 Keywords:
@@
 
+Pollination is a fundamental ecosystem service that connects the earth’s vegetation and human economy.
+Pollination services are performed by both managed and wild pollinators, mostly bees (James & Pitts Singer, 2008; Winfree et al., 2011), both of which might be affected by a range of current and projected human-related environmental changes.
+There are indeed clear shreds of evidence of recent declines in pollinator populations (Bartomeus et al., 2018), and parallel declines in the plants that rely upon them (Potts et al., 2010; Scheper et al., 2014).
+For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 (Powney et al., 2019), including six of all 16 non-parasitic bumblebees (Williams & Osborne, 2009).
+The decline of pollinators currently observed is projected to increase in the future (Dormann et al., 2008).
+Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production (Klein et al., 2007; Ollerton et al., 2011), there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services (Ashman et al., 2004; Klein et al., 2007).
+Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats (Brown & Paxton, 2009; Natural Research Council, 2006).
+The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth (Brown & Paxton, 2009; Natural Research Council, 2006).
+However, not all pollinators species are equally affected by habitat loss and alteration.
+In bees, for example, there are also “winners” —like Bombus terrestris (Rasmont et al., 2008)— that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations (Biesmeijer et al., 2006; Bartomeus et al., 2013; Reemer et al., 2012; Collado et al., 2019).
+Human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants (Cane et al., 2006; Winfree et al., 2007; Collado et al., 2019).
+Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown.
+Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive (Manenti et al., 2013).
+For species that tolerate human-related disturbances, human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants (Cane et al., 2006; Winfree et al., 2007; Collado et al., 2019).
+The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
+Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
+For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders (Sol et al., 2008; Sol et al., 2005; Amiel et al., 2011).
+In vertebrates, relative brain size is a good proxy of problem-solving and learning (e.g.
+see chapter 2 of this thesis for its relationship with learning), which can help individuals to cope with the many challenges on novel environments.
+However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains.
+More important, there was no unequivocal evidence that brain size was related to cognitive performance.
+However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so (Alem et al., 2016).
+There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2).We used previously published data on bee forest and urban preference (Collado et al., 2019) as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
+Species population trends were extracted from Bartomeus et al., 2013 and both datasets were crossed with a newly collected brain weight database.
+We wanted to explore if there is a relationship between brain size, habitat preference, and population trends.
+Hence, our questions were the following:Are brain sizes, as a proxy of behavioural plasticity, related to habitat preferences?Is population stability related to brain size as it has been seen in other taxa?Are habitat preferences related to population stability?
+Or alternatively, are urban dwellers populations increasing?
+
 # Methods
+
+**Animals**
+
+Most bees were opportunistically captured by hand netting in different regions of New York State (USA), and complemented with  bees captured from Spain and the Netherlands for species with a holarctic distribution.
+We captured 145 individuals from 44 species, kept them in cold storage and sent them to the laboratory to be processed.
+Once the bees were in the laboratory, they were anesthetized in cold (Crook, 2013) and decapitated.
+Bees were identified by expert taxonomists (Parker Gambino and Ivo Raemakers).
+The species used for the analysis can be consulted in Table S2.
+We measured the inter-tegular distance, as it is a good proxy of body size (Kendall et al., 2019).
+To prevent degradation, the head was fixed in 4% paraformaldehyde with phosphate buffer saline (PBS).
+Brains were extracted from the head capsule, and tracheae and fat bodies were removed to maximize accuracy in neural tissue weight.
+Brains extracted were placed on a small piece of tared Parafilm® and exceeding fixative solution was dried from the brain using Kimwipes® tissues.
+Finally the brain was weighted in a microbalance.
+Brain weight was used as a proxy of brain size (Sayol et al., 2019 in preparation).
+
+**Data gathering**
+
+The quantitative data of population trends for North American bees were extracted from Bartomeus et al., 2013, where they analyzed trends for 187 species over time, using data from pinned bees from different North American museums and universities.
+We assigned the estimated value of their relative change over time models as values of population trends for the 44 species for which we have brain size measurements (positive for increasing populations and negative for decreasing populations).
+Habitat preferences for USA bee data were extracted from Collado et al., 2019, were habitat preference for forests and urban areas was calculated comparing species occurrence in each habitat to what would be expected if species occur at random using null models.
+This preference index ranges from 0 (avoidance) to 1 (preference).In order to have a representative value of brain weight and body size per species, we used means of the different collected individuals from each species.
+Twenty-four individuals with brain weight values one standard deviation higher or lower than the average were removed (see sup mat analysis) because human errors in their measurement are likely with such delicate tissues.
+Specially, brain weights can give errors due to dehydration or insufficient removal of fat bodies and trachea around the brain.
+For eusocial species, we only used workers.
+We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
+(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p < 0.001, R2= 0.83, n = 163 species).
+Residuals were calculated using our complete brain weights and body size database.
+
+**Data analysis**
+
+We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms (Bürkner, 2017), to assess whether brain sizes, habitat preference and population trends are correlated.
+In all models, we incorporated a phylogenetic covariance matrix.
+The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
+1, Fig.
+S2, Fig.
+S3).
+Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
+Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6‐44, Revell, 2012).
+Because we did not have information on our 3 variables for the 44 species (habitat preference, brain weight and population trends), we used 3 different subsets of our dataset to maximize the sample size in the different models: Relative brain size-habitat preference, population trends-habitat preference and population trends-relative brain size.
 
@@
 # Discussion
+
+Bees with urban habitat preferences tended to have bigger brains related to their body sizes, and those with forest preferences had relatively smaller brains.
+Surprisingly, neither brain sizes nor habitat preference explained population trends, against what we expected.
+Therefore, species with bigger brains appearing on urban habitat can be indicating that while only bees with large cognitive capacities are surviving in these modified environments, they do not have experienced notorious increases in their populations.We hypothesized that bees with bigger brains related to body size should be more flexible in their behaviour because a large brain may confer better cognitive abilities, as it has been observed in some species (Lefebvre et al., 2004) and as we show specifically for bees in relation to learning (Chapter 2 of this thesis).
+Therefore, to exploit complex urban environments, we expect bigger brains to be required.
+Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment (Wheater, 1999).
+Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than the forest dwellers counterpart.
+Some examples of behaviours needed to survive in urban environments (see Table S2 for a full set of examples) are the use of human-made materials to nest (Allasino et al., 2019) or being able to collect pollen and nectar from exotic flowers from gardens (Lowenstein et al., 2014).
+We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them (Smith et al., 2019).
+In accordance, we found smaller brain sizes than expected by their body sizes in bees that preferred forests indicating that an investment in bigger relative brain sizes is not required.
+In addition, other traits have already been described to be filtered in anthropogenic land uses, like early spring or brood parasitism (Harrison et al., 2017), which are present in the forest dwellers.
+We could not explain population trends with brain size as it happens in other taxa.
+For example, in birds, brain size is a good predictor of how a population does in the long-term (Shultz, 2005).
+However, this is not always the case as bigger brains increase the risk of extinction in mammals (González-Voyer et al., 2016).
+Despite brain weight is relatively easy to measure and standardize with the used protocol, we lack information on certain neuropils size, like mushroom bodies (the suspected centers of cognitive processes in bees), which can be variable even within species (Barth & Heisenberg, 1997; Withers et al., 1995; Fahrbachand & Robinson 1995).
+Differences in more subtle brain structures could explain differences in behaviours, and in consequence, shed extra light on species population trends.
+Therefore the isolation and measurement of these neuropils could be interesting because we could get more accurate information.
+On the other hand, also note that population trends data is hard to find or measure, especially for invertebrate taxa.
+Long term data is often lacking and we count only with data indirectly taken from museums.
+This data is also scarce and hard to standardize (Bartomeus et al., 2018), therefore not being optimal for comparative studies of declining or increasing populations (Shultz, 2005).
+Species body size is generally related to population trends.
+In mammals, it is known that bigger body sizes decreases population density and increases the risk of extinction (Damuth, 1981; González-Voyer et al., 2016).
+This relation has also been seen in bees (Scheper et al., 2014).
+Although body size is directly related to absolute brain size, we did not find a strong correlation between body size and population trends in our subset of North American bees as it was found for the complete set in Bartomeus et al., 2013.
+Our subset may be biased towards the most common genera and contain fewer species per genus.
+If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel’s λ > 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
+Population trends or habitat preferences could be explained by the availability of the resources consumed, for example, in the Netherlands, population trends of some bee species are related to the availability of their preferred plants (Scheper et al., 2014).
+A polylectic diet could help species to adapt to new environments, and we know that species able to forage on crop plant families tend to have stable or increasing populations (Scheper et al., 2014, Bartomeus et al.
+2013).
+If we consider habitats being conditioned by their availability of the preferred resources, then we may be able to explain our link between brain size and preference for certain habitats.
+For example, differences in bee community composition and abundance are found for oligolectic (i.e.diet specialist bees) associated to their prefered resources along forested gradients (Grundel et al., 2010).
+Dietary data should be an interesting next step to see if broader (polylectic) or narrower (oligolectic) diet are generally related to habitat preferences and brain sizes.
+In fact, it has been already described for other taxa a relationship between brain size and diet (DeCasien et al., 2017).Finally, the absence of relationship between urban preferences and increasing population trends is interesting.
+Despite these bees have found new niches to occupy, they are not increasing its populations but just persisting.
+The transformation from natural to urban habitats is filtering which species can survive, and hence, biodiversity in urban habitats is usually lower (Collado et al., 2019, Chapter 3 of this thesis).
+However, survivors are not thriving either in this new conditions.
+Some negative effects of urbanization have already been studied in bees, like for example, biased sex ratio, with increasing males in urban areas (Fitch et al., 2019) with unknown future population consequences.
+It should be interesting to fully assess the costs and benefits of living in urban environments, including brain investment and other selected strategies to adapt to the new conditions.

### 2022-10-19 Manuscript/Ms.Rmd
edits

@@
 ---
-title: "Untitled"
+title: "Bee brain size is positively associated with urban lifestyles but not with positive population trends"
 output: pdf_document
 
+link-citations: yes
 linkcolor: RoyalBlue
@@
 Pollination is a fundamental ecosystem service that connects the earth’s vegetation and human economy.
-Pollination services are performed by both managed and wild pollinators, mostly bees (James & Pitts Singer, 2008; Winfree et al., 2011), both of which might be affected by a range of current and projected human-related environmental changes.
-There are indeed clear shreds of evidence of recent declines in pollinator populations (Bartomeus et al., 2018), and parallel declines in the plants that rely upon them (Potts et al., 2010; Scheper et al., 2014).
-For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 (Powney et al., 2019), including six of all 16 non-parasitic bumblebees (Williams & Osborne, 2009).
+Pollination services are performed by both managed and wild pollinators, mostly bees [@james2008; @winfree2011], both of which might be affected by a range of current and projected human-related environmental changes.
+There are indeed clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019], and parallel declines in the plants that rely upon them [@potts2010; @scheper2014].
+For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees (Williams & Osborne, 2009).
 The decline of pollinators currently observed is projected to increase in the future (Dormann et al., 2008).
@@
 
-We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms (Bürkner, 2017), to assess whether brain sizes, habitat preference and population trends are correlated.
+We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms [@burkner2017], to assess whether brain sizes, habitat preference and population trends are correlated.
 In all models, we incorporated a phylogenetic covariance matrix.
@@
 Our subset may be biased towards the most common genera and contain fewer species per genus.
-If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel’s λ > 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
+If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel’s lambda > 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
 Population trends or habitat preferences could be explained by the availability of the resources consumed, for example, in the Netherlands, population trends of some bee species are related to the availability of their preferred plants (Scheper et al., 2014).
@@
 It should be interesting to fully assess the costs and benefits of living in urban environments, including brain investment and other selected strategies to adapt to the new conditions.
+
+# References
+
+::: {#refs}
+:::

### 2022-10-19 Manuscript/Summary/Summary.Rmd
edits

--- 2022-08-26 Summary.Rmd
@@
 
-## Land use classification
+\newpage
 
-**1) Europe**
-
-Link with land European use classes
-
-https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-512.html
+# Europe (4 levels)
 
@@
 
-Then, we calculate preferences and model the data.
+\newpage
 
-```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
-
-long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe.csv")
-
-model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe.csv")
-
-model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe.csv")
-
-model3 = read_csv("Data/Europe_data/model_output_preference_it_europe.csv")
-
-p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("Europe")
-
-p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Brain weight") +
-ggtitle("Europe")
-
-p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Intertegular distance") +
-ggtitle("Europe")
-
-library(patchwork)
-p1 / p2 / p3 & plot_layout(guides = "collect")
-
-```
-
-Now we make more strict land use classifications
+# Europe (3 levels)
 
@@
 
-\newpage
+\clearpage
 
-And we model the data with this categories:
+```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}
 
-```{r Plotting Europe subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
+library(tidyverse)
 
+long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe.csv")
+
+model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe.csv")
+
+model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe.csv")
+
+model3 = read_csv("Data/Europe_data/model_output_preference_it_europe.csv")
+
+model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("Europe 4 levels")+
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("Europe 4 levels")+
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("Europe 4 levels")  +
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("Europe 4 levels")  +
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Read data with land use filtered
 long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_subset.csv")
@@
 
+model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_subset.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("Europe 3 levels")+
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))+
+    theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("Europe 3 levels")+
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))+
+    theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("Europe 3 levels") +
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+    theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("Europe 3 levels") +
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+    theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
+
+```
+
+\newpage
+
+# Europe
+
+Qualitative analysis on the statistical evidences for preference and avoidance by habitat
+
+```{r Plotting Europe qualitative analysis, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="80%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_qualitative.csv")
+
+model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe_qualitative.csv")
+
+model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe_qualitative.csv")
+
+model3 = read_csv("Data/Europe_data/model_output_preference_it_europe_qualitative.csv")
+
+model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_qualitative.csv")
+
 p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
@@
 xlab("Residuals") + 
-ggtitle("Europe")
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
@@
 xlab("Brain weight") +
-ggtitle("Europe")
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
@@
 xlab("Intertegular distance") +
-ggtitle("Europe")
+ggtitle("USA-Europe 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Load land use filtered data and models
+long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_qualitative_subset.csv")
+
+model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe_qualitative_subset.csv")
+
+model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe_qualitative_subset.csv")
+
+model3 = read_csv("Data/Europe_data/model_output_preference_it_europe_qualitative_subset.csv")
+
+model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_qualitative_subset.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA-Europe 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA-Europe 3 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
 
 library(patchwork)
-p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
 
@@
 
-**2) USA**
+# USA (4 levels)
 
@@
 ```
-
-United States land use classification with total number of occurrences within cover type in our dataset
 
@@
 
-Then, we calculate preferences and model the data.
-
-```{r Plotting USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
-
-long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa.csv")
-
-model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa.csv")
-
-model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa.csv")
-
-model3 = read_csv("Data/Usa_data/model_output_preference_it_usa.csv")
-
-p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Residuals") + 
-ggtitle("USA")
-
-p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Brain weight") +
-ggtitle("USA")
-
-p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
-geom_line(aes(color=Habitat)) +
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Intertegular distance") +
-ggtitle("USA")
-
-library(patchwork)
-p1 / p2 / p3 & plot_layout(guides = "collect")
-
-```
-
-Note: The models have been run with non-informative priors from the brms function but some work on priors can be done...
-I have explored a bit and the results were quite similar overall but reviewers seem to don't like this approach as one of the goals of Bayesian is to use the prior knowledge of our variables to improve our predictions.
-
-\newpage
-
-Now we make more strict categories of what is Natural, Urban and Agricultural. See below:
+# USA (3 levels)
 
@@
 
-And we model the data with this categories:
+\newpage
 
-```{r Plotting USA subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}
+# USA (Gbif and MA thesis data)
 
+```{r Plotting USA gbif and MA thesis data, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa.csv")
+
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa.csv")
+
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa.csv")
+
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa.csv")
+
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Load land use filtered data and models
 long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset.csv")
@@
 
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_subset.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA 3 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
+
+```
+
+\newpage
+
+# USA (Gbif and MA thesis data)
+
+Qualitative analysis on the statistical evidences for preference and avoidance by habitat
+
+```{r Plotting USA qualitative, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="80%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_qualitative.csv")
+
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_qualitative.csv")
+
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_qualitative.csv")
+
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_qualitative.csv")
+
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_qualitative.csv")
+
 p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
@@
 xlab("Residuals") + 
-ggtitle("USA")
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
@@
 xlab("Brain weight") +
-ggtitle("USA")
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
@@
 xlab("Intertegular distance") +
-ggtitle("USA")
+ggtitle("USA 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
-library(patchwork)
-p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
 
-```
+#Load land use filtered data and models
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_qualitative_subset.csv")
 
-**2) Finally, we consider Europe and USA together**
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_qualitative_subset.csv")
 
-```{r Plotting Europe USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_qualitative_subset.csv")
 
-data <- read_csv("Data/Europe_USA/data_preference_brain_weight_europe_usa.csv")
+model3 = read_csv( "Data/Usa_data/model_output_preference_it_usa_qualitative_subset.csv")
 
-model1 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_qualitative_subset.csv")
 
-model2 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa.csv")
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
 
-ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Brain weight") +
-ggtitle("Europe USA occurrences")
+ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
 
-ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
 geom_line(aes(color=Habitat)) +
@@
 xlab("Intertegular distance") +
-ggtitle("Europe USA occurrences")
+ggtitle("USA levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
 
 ```
+
+\newpage
+
+# USA (just Gbif without MA thesis data)
+
+```{r Plotting USA gbif, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_gbif.csv")
+
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_gbif.csv")
+
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_gbif.csv")
+
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_gbif.csv")
+
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_gbif.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Load land use filtered data and models
+long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset_gbif.csv")
+
+model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_subset_gbif.csv")
+
+model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_subset_gbif.csv")
+
+model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_subset_gbif.csv")
+
+model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_subset_gbif.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA 3 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
+
+```
+
+\newpage
+
+# USA and Europe together
+
+```{r Plotting USA Europe together, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa.csv")
+
+model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa.csv")
+
+model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")
+
+model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa.csv")
+
+model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA-Europe 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA-Europe 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Load land use filtered data and models
+long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_subset.csv")
+
+model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa_subset.csv")
+
+model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa_subset.csv")
+
+model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa_subset.csv")
+
+model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa_subset.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA-Europe 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA-Europe 3 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
+
+```
+
+\newpage
+
+# USA and Europe together
+
+Qualitative analysis on the statistical evidences for preference and avoidance by habitat
+
+```{r Plotting USA Europe together qualitative, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}
+
+library(tidyverse)
+
+long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
+
+model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")
+
+model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")
+
+model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")
+
+model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa_qualitative.csv")
+
+p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA 4 levels")  + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight / IT") +
+ggtitle("USA 4 levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
+
+#Load land use filtered data and models
+long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_subset_qualitative.csv")
+
+model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa_subset_qualitative.csv")
+
+model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa_subset_qualitative.csv")
+
+model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa_subset_qualitative.csv")
+
+model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa_subset_qualitative.csv")
+
+pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") + 
+ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight") +
+ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Intertegular distance") +
+ggtitle("USA levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
+geom_line(aes(color=Habitat)) +
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Brain weight /` IT") +
+ggtitle("USA levels") + 
+scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
+theme(legend.position = "none")
+
+library(patchwork)
+p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")
+
+```

### 2022-10-20 Manuscript/Ms.Rmd
Changing way of downloading data. occ_search() takes too long, now I use occ_download()

@@
   - references.bib
+
+header-includes:
+- \usepackage{orcidlink}
+
 ---
 
+\textbf{Miguel Á. Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B. Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}
+
+Departamento de xxxxx, Universidad de Sevilla, Seville, Spain
+
 Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
+
+\textbf{Author for correspondence:}
 
@@
 
-Keywords:
-
-\textbf{Author for correspondence:}
+Keywords: XXXX
 
@@
 There are indeed clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019], and parallel declines in the plants that rely upon them [@potts2010; @scheper2014].
-For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees (Williams & Osborne, 2009).
-The decline of pollinators currently observed is projected to increase in the future (Dormann et al., 2008).
-Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production (Klein et al., 2007; Ollerton et al., 2011), there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services (Ashman et al., 2004; Klein et al., 2007).
-Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats (Brown & Paxton, 2009; Natural Research Council, 2006).
-The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth (Brown & Paxton, 2009; Natural Research Council, 2006).
+For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees [@williams2009].
+The decline of pollinators currently observed is projected to increase in the future [@dormann2008].
+Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production [@klein2007; @ollerton2011], there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services [@ashman2004; @klein2007].
+
+Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@brown2009].
+The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth [@brown2009].
 However, not all pollinators species are equally affected by habitat loss and alteration.
-In bees, for example, there are also “winners” —like Bombus terrestris (Rasmont et al., 2008)— that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations (Biesmeijer et al., 2006; Bartomeus et al., 2013; Reemer et al., 2012; Collado et al., 2019).
-Human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants (Cane et al., 2006; Winfree et al., 2007; Collado et al., 2019).
+In bees, for example, there are also “winners” —like Bombus terrestris [@rasmont2008]— that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].
+
+Human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown.
-Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive (Manenti et al., 2013).
-For species that tolerate human-related disturbances, human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants (Cane et al., 2006; Winfree et al., 2007; Collado et al., 2019).
+Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive [@manenti2013].
+For species that tolerate human-related disturbances, human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
 Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders (Sol et al., 2008; Sol et al., 2005; Amiel et al., 2011).
+For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
 In vertebrates, relative brain size is a good proxy of problem-solving and learning (e.g.
@@
 More important, there was no unequivocal evidence that brain size was related to cognitive performance.
-However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so (Alem et al., 2016).
-There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2).We used previously published data on bee forest and urban preference (Collado et al., 2019) as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
-Species population trends were extracted from Bartomeus et al., 2013 and both datasets were crossed with a newly collected brain weight database.
+However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
+There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2).
+We used previously published data on bee forest and urban preference [@collado2019] as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
+Species population trends were extracted from @bartomeus2013 and both datasets were crossed with a newly collected brain weight database.
 We wanted to explore if there is a relationship between brain size, habitat preference, and population trends.

### 2022-10-20 Scripts/1_Alternative_download_occ_search.R

@@ -1,5 +1,10 @@
 #In this script we download data from GBIF----
 
+#Here we download the data directly from gbif without the need of a registered user
+#and the data is directly downloaded into R without logging on gbif
+#HOWEVER, it takes very long!!!
+#I recommend the other approach since is much faster and fully reliable 
+
 #Load libraries
 library(stringr) #data cleaning v.1.4.0
 library(dplyr) #data cleaning v.1.0.7
@@ -8,6 +13,11 @@ library(ggplot2) #plotting v.3.3.5
 library(sp) #manipulate coordinates v.1.4-6
 library(rworldmap) #plot worldmap v.1.3-6
 library(sf) #scale worldmap v 1.0-5
+library(openssl)
+library(usethis)
+
+usethis::edit_r_environ()
+
 
 #Read data
 d <- read.csv("Data/Processing/Especies_para_buscar.csv", row.names = 1)
@@ -26,8 +36,8 @@ d$Species[d$Species=="Anthopora plumipes"] <- "Anthophora plumipes"
 #Filter out Apis mellifera
 d <- d %>% filter(!Species=="Apis mellifera")
 #For doing trials
-#lev_trial <- levels(factor(d$Species))[6:15]
-#d <- d %>% filter(Species %in% lev_trial)
+lev_trial <- levels(factor(d$Species))[1:1]
+d <- d %>% filter(Species %in% lev_trial)
 
 
 ########################---
@@ -83,7 +93,7 @@ dat <-  data.frame(scientificName = NA, decimalLatitude = NA,
 countries <- c("US", "DE", "GB", "LU", "BE", "NL", "DK") 
 
 for(i in gbif_id$key_number){
-    temp <- occ_data(taxonKey= i, 
+    temp <- occ_download(taxonKey= i, 
                        year="1985, 2022",
                        hasCoordinate=TRUE,
                        hasGeospatialIssue=FALSE,
@@ -204,7 +214,7 @@ info <- all_1 %>%
     summarise(no_rows = length(Raw_spp))
 #Seems ok!
 #Save data
-write.csv(all_1, file=gzfile("Data/gbif_data_final.csv.gz"),row.names=FALSE)
+#write.csv(all_1, file=gzfile("Data/gbif_data_final.csv.gz"),row.names=FALSE)
 
 
 

### 2022-10-20 Scripts/1_Main_download_occ_download.R

@@ -1,0 +1,71 @@
+#In this script we download data from GBIF----
+#The download is done with the help of rgbif and the function occ_download
+#Quite fast but we need a gbif account!
+#To reproduce our analysis, 
+#Skip this part and use directly the downloaded data from the DOI
+#https://doi.org/10.15468/dl.re5wwt
+
+#Load libraries
+library(stringr) 
+library(dplyr) 
+library(rgbif) 
+library(sp) 
+library(rworldmap) 
+library(sf) 
+library(openssl)
+library(usethis) 
+
+#Here we need to add our credentials to download the data
+usethis::edit_r_environ()
+
+#Read data
+d <- read.csv("Data/Processing/Especies_para_buscar.csv", row.names = 1)
+#Rename cols
+colnames(d) <- c("Species")
+#Check levels
+levels(unique(factor(d$Species)))
+#Fix one species name
+d$Species[d$Species=="Lasioglossum dialictus spp"] <- "Lasioglossum dialictus" 
+#Delete Species with sp. 
+d <- filter(d, !grepl(" sp.",Species))
+#Some more typos
+d$Species[d$Species=="Agaposemon sericeus"] <- "Agapostemon sericeus"
+d$Species[d$Species=="Rhodantidium sticticum"] <- "Rhodanthidium sticticum"
+d$Species[d$Species=="Anthopora plumipes"] <- "Anthophora plumipes"
+#Filter out Apis mellifera
+d <- d %>% filter(!Species=="Apis mellifera")
+#For doing trials
+#lev_trial <- levels(factor(d$Species))[1:1]
+#d <- d %>% filter(Species %in% lev_trial)
+
+#Save species list
+
+########################---
+#Download data from GBIF----
+########################---
+
+#First create species key (some long numbers that are needed to download the data) 
+#For loop to do it for all species
+i <- NULL
+gbif_id <- list()
+
+for(i in d$Species){
+    
+    j <- gsub(" ", "_", i)
+    gbif_id[[j]] <- name_backbone(name=i, rank = "species")$usageKey
+}
+
+#convert list to dataframe
+gbif_id <- data.frame(unlist(gbif_id))
+#rename col
+colnames(gbif_id) <- "key_number"
+
+#Countries of interest with ISO codes
+countries <- c("US", "DE", "GB", "LU", "BE", "NL", "DK") 
+
+#Download data
[…]

### 2022-10-20 Scripts/1_Process_GBIF_Data.R

@@ -1,0 +1,79 @@
+#In this script we process the data downloaded from GBIF: https://doi.org/10.15468/dl.re5wwt
+#We select cols of interest with georreferenced information
+
+#Notes: 
+#1)this is a very large file that can be downloaded from the DOI above
+#As a consequence, this file is not saved at the repo
+#2) The data has been downloaded by GBIF with the help of occ_dwonload()
+#However, we also have downloaded the data with occ_search()
+#Because, the second option takes very long to download 
+#we have decided that the 1st one is the way to go
+
+#Load libraries
+library(dplyr) 
+library(rworldmap) 
+library(data.table) 
+library(tidyr)
+library(sp)
+library(ggplot2)
+
+#Read data, it takes a while
+d = fread("Data/occurrence.txt")
+#Check colnames
+colnames(d)
+#Select cols of interest
+main_cols = d %>% select("scientificName", "family", "genus", "species","speciesKey",
+             "recordedBy", "recordedByID", "identifiedBy", "year", "month", "day",
+             "decimalLatitude", "decimalLongitude", "stateProvince", 
+             "locality","countryCode", "references", "datasetKey", "institutionCode",
+             "bibliographicCitation")
+
+main_cols_georeferenced = main_cols %>% 
+drop_na("decimalLatitude", "decimalLongitude") %>% 
+filter(year>1984)
+
+
+coords2continent = function(points) 
+{  
+    countriesSP <- getMap(resolution='low')
+    #countriesSP <- getMap(resolution='high') #you could use high res map from rworldxtra if you were concerned about detail
+    
+    # converting points to a SpatialPoints object
+    # setting CRS directly to that from rworldmap
+    pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  
+    
+    
+    # use 'over' to get indices of the Polygons object containing each point 
+    indices = over(pointsSP, countriesSP)
+    
+    #indices$continent   # returns the continent (6 continent model)
+    indices$REGION   # returns the continent (7 continent model)
+    #indices$ADMIN  #returns country name
+    #indices$ISO3 # returns the ISO3 code 
+}
+
+points = cbind(main_cols_georeferenced$decimalLongitude, main_cols_georeferenced$decimalLatitude)
+#Include column of continent on the dataset
+main_cols_georeferenced$Continent <- coords2continent(points)
+
+#Filter out by continent
[…]

### 2022-10-24 Manuscript/Ms.Rmd
Updating manuscript

@@
 - \usepackage{orcidlink}
+- \usepackage{caption,setspace}
+- \captionsetup[figure]{labelformat=empty}
 
@@
 # Methods
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} A) Phylogenetic tree at the genera level for the selected bee species with brain weight and intertegular distance information (N=100).
+Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).
+The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted circumferences at the end of the tip branches.
+Larger circumferences indicate larger brains in proportion to their body and vice versa.
+B) Heatmap indicating the habitat preferences of the different bee species."}
+
+#Phylogenetic tree for USA and Europe
+library(ggtree) 
+library(ggplot2)
+library(stringr)
+library(tidyverse)
+library(tibble)
+
+#Load phylogenetic data from europe
+bee.tree100 = readRDS("../Data/Europe_USA/bee.tree100.rds")
+#Load brain data
+brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+
+#Average preferences between USA and Europe
+preferences = pref %>% 
+group_by(Species) %>%
+summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
+ungroup()
+
+#Delete underscore from tree
+bee.tree100$tip.label = str_replace(bee.tree100$tip.label, "_", " ")
+
+#Generate vector with species
+v = bee.tree100$tip.label
+#Filter species with that are in the phylo
+brain_weight = brain_weight %>% filter(Species %in% v)
+
+#Fill a column based on condition
+brain_weight = brain_weight %>% 
+    mutate(Country = case_when((brain_weight$Species %in% preferences_eu$Species & 
+    brain_weight$Species %in% preferences_usa$Species ~ "USA and Europe"),
+    (brain_weight$Species %in% preferences_usa$Species ~ "USA"),
+    (brain_weight$Species %in% preferences_eu$Species ~ "Europe")))
+
+#Plot tree
+p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country) +  geom_tree(size=0.8)+
+    geom_tippoint(aes(size = residuals, color=Country))+ 
+    theme_tree2()+scale_size_continuous(range = c(-1.2, 3.2),name="Residuals")+
+    scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
+                       labels = c("Europe", "USA", "USA and Europe"),
+                       breaks = c("Europe", "USA", "USA and Europe")) +
+  ggtitle("A") + theme(plot.title = element_text(size=22, face="bold"))
+
+#Convert wide format to long for plotting heatmap
+d=gather(preferences, condition, measurement, Agricultural:Seminatural, factor_key=TRUE)
+
+#Relevel for plotting
+d = d %>% 
+mutate(condition = recode_factor(condition, "Seminatural" = "Semideveloped")) %>% 
+mutate(condition=fct_relevel(condition,c("Natural","Agricultural","Semideveloped", "Urban")))
+
+p2 = ggplot(d, aes(x=condition, y=Species)) + 
+    geom_tile(aes(fill=measurement),color = "white") +
+   # scale_fill_gradient(high="black",  low="white") +
+       scale_fill_viridis_c(name = "Habitat\n preference") + 
+    theme_minimal() +
+    xlab(NULL) + 
+    ylab(NULL) + 
+    theme(axis.text.x = element_text(face="bold", color="black", 
+    size=10,  vjust=1.15),
+    axis.text.y = element_text(face="italic", color="black", size=7, hjust=0),
+    plot.title = element_text(size=22, face="bold")) +
+  ggtitle("B")
+
+#Nice way to insert ordered heatmap
+library(aplot)
+p2 %>% insert_left(p) 
+```
 

### 2022-10-24 Scripts/0_Processing/1_Calculate_residuals.R

@@ -6,6 +6,11 @@ library(tidyverse)
 
 wit.mean = readRDS("Data/Processing/wit.mean.rds")
 
+#IT Xylocopa virginica
+wit.mean %>% 
+mutate(IT = replace(IT, Species == "Xylocopa virginica", 5.518))
+
+
 #Log variables first
 wit.mean = wit.mean %>% 
 mutate(log_brain_weight = log(Brain.weight)) %>% 

### 2022-10-24 Scripts/Collado_analysis/Correlation_preferences_gbif_thesis_data_usa.R

@@ -2,6 +2,7 @@
 #Check correlation between preferences 
 #Load preferences
 library(readr)
+library(dplyr)
 ma_preferences = read_csv("Data/Usa_data/preferences_usa_thesis_ma.csv") 
 gbif_preferences = read_csv("Data/Usa_data/preferences_usa_gbif.csv") 
 #Merge datasets

### 2022-10-24 Scripts/Figure_3_Preference_Models/Figure_3_Plotting_all.R

@@ -1,4 +1,14 @@
 
+
+
+#All----
+
+#Load libraries
+library(readr)
+library(dplyr)
+library(ggplot2)
+library(forcats)
+
 #Load data for plotting
 long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
 
@@ -28,7 +38,7 @@ model3 = model3 %>%
 
 #colorBlindness::displayAllColors(viridis::viridis(4))
 
-p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
 #geom_ribbon(aes(ymin = lower__, ymax = upper__, color=Habitat), linetype = "dashed",alpha = 0)+
@@ -40,7 +50,7 @@ scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))
 # + facet_wrap(~ Habitat)
 
 
-p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+all_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
 #geom_ribbon(aes(ymin = lower__, ymax = upper__, color=Habitat), linetype = "dashed",alpha = 0)+
@@ -52,7 +62,7 @@ scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))
 
 
 
-p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+all_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
 #geom_ribbon(aes(ymin = lower__, ymax = upper__, color=Habitat), linetype = "dashed",alpha = 0)+
[…]

### 2022-10-24 Scripts/0_Processing/1_Calculate_residuals.R
edits

@@ -6,11 +6,6 @@ library(tidyverse)
 
 wit.mean = readRDS("Data/Processing/wit.mean.rds")
 
-#IT Xylocopa virginica
-wit.mean %>% 
-mutate(IT = replace(IT, Species == "Xylocopa virginica", 5.518))
-
-
 #Log variables first
 wit.mean = wit.mean %>% 
 mutate(log_brain_weight = log(Brain.weight)) %>% 

### 2022-10-24 Scripts/Figure_1_Phylo_Tree/Figure_1_Phylo_tree_all.R

@@ -20,6 +20,7 @@ preferences_eu = readr::read_csv("Data/Europe_data/preferences_europe.csv")
 #Bind rows
 pref = bind_rows(preferences_usa, preferences_eu)
 
+
 #Average preferences between USA and Europe
 preferences = pref %>% 
 group_by(Species) %>%

### 2022-10-24 Manuscript/Ms.Rmd
Minor edit caption

@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} A) Phylogenetic tree at the genera level for the selected bee species with brain weight and intertegular distance information (N=100).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
+A) Phylogenetic tree at the genera level.
 Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).

### 2022-10-25 Scripts/Figure_4_Map_Points/Map.Rmd
edits

---
title: ""
output:
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12}

#Plot maps with density of bee occurrences

#Europe----
#Load libraries
library(giscoR) #map of europe
library(ggplot2)
library(dplyr)
library(readr)
library(sf)
library(patchwork)
library(rnaturalearth)

#Load europe map
Europe <- gisco_get_countries(region = "Europe")
#Load countries filtered
euro_map <- st_read("../../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Load data with records filtered
all_long_lat <- read.table("../../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")

#Plot Europe map with counties of interest highlighted
p1 = ggplot(Europe) +
geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white"))

p2 = ggplot(Europe) +
geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 12),
plot.title = element_text(face="bold"),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))

#Insert general plot within specific plot
panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)

#USA-----
#Load libraries

#Load data
usa_states_1 <- read.csv("../../Data/Usa_data/usa_states.csv")
ny_metropolitan_1 <- read.csv("../../Data/Usa_data/ny_metropolitan.csv")
all_long_lat <- read.table("../../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Load maps
canada <- ne_states(country = "canada", returnclass = "sf") %>%
st_as_sf(crs = 4326)

usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
st_as_sf(crs = 4326)

p1 = ggplot(canada) + 
geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white")) +
geom_map(data = ny_metropolitan_1, 
map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
size = 0.1) +
geom_map(data = usa_states_1, map = usa_states_1,
aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
size = 0.1) +
coord_sf(xlim = c(-86, -62), ylim = c(25, 48))

p2 = ggplot(canada) + 
geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray70",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 3, 6)) +
coord_sf(xlim = c(-82, -68), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 12),
plot.title = element_text(face="bold"),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -68, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))

#Insert general plot within specific plot
panel2 = p2 + inset_element(p1, left = 0, bottom = -0.885, right = 0.35, top = 1.2, ignore_tag = TRUE)

library(patchwork)
panel1 + panel2

```

### 2022-10-25 Scripts/Figure_4_Map_Points/Figure_map.R

@@ -1,0 +1,112 @@
+#Plot maps with density of bee occurrences
+
+#Europe----
+#Load libraries
+library(giscoR) #map of europe
+library(ggplot2)
+library(dplyr)
+
+#Load europe map
+Europe <- gisco_get_countries(region = "Europe")
+#Load countries filtered
+euro_map <- st_read("Data/Europe_data/euro_map_1.shp")
+#Load data with records filtered
+all_long_lat <- read.table("Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+#Plot Europe map with counties of interest highlighted
+p1 = ggplot(Europe) +
+geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
+geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
+aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
+coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white"))
+
+
+p2 = ggplot(Europe) +
+geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
+viridis::scale_fill_viridis(name = "log(Count)") +
+coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
+theme(panel.background = element_rect(fill = "aliceblue"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("A") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(), 
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold"),
+plot.title = element_text(face="bold"),
+legend.text = element_text(face = "bold")) +
+scale_y_continuous(breaks = seq(40, 60, by = 6)) +
+scale_x_continuous(breaks = seq(-8, 18, by = 8))
+
[…]

### 2022-10-25 Manuscript/Ms.Rmd
edits

@@
 
-Departamento de xxxxx, Universidad de Sevilla, Seville, Spain
+Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the location of the downloaded GBIF occurrences for the different bee species."}
+
+#Plot maps with density of bee occurrences
+
+#Europe----
+#Load libraries
+library(giscoR) #map of europe
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(sf)
+library(patchwork)
+library(rnaturalearth)
+
+#Load europe map
+Europe <- gisco_get_countries(region = "Europe")
+#Load countries filtered
+euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
+#Load data with records filtered
+all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+#Plot Europe map with counties of interest highlighted
+p1 = ggplot(Europe) +
+geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
+geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
+aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
+coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white"))
+
+p2 = ggplot(Europe) +
+geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8)) +
+coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
+theme(panel.background = element_rect(fill = "aliceblue"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("A") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(), 
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(40, 60, by = 6)) +
+scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))
+
+#Insert general plot within specific plot
+panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
+
+#USA-----
+#Load libraries
+
+#Load data
+usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
+ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
+all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+#Load maps
+canada <- ne_states(country = "canada", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+p1 = ggplot(canada) + 
+geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white")) +
+geom_map(data = ny_metropolitan_1, 
+map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
+size = 0.1) +
+geom_map(data = usa_states_1, map = usa_states_1,
+aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
+size = 0.1) +
+coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
+
+p2 = ggplot(canada) + 
+geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray70",color = "black", alpha = 1,size=0.2) +
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 3, 6)) +
+coord_sf(xlim = c(-82, -68), ylim = c(33, 47))  +
+theme(panel.background = element_rect(fill = "aliceblue"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("B") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(),
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(33, 47, by = 4)) +
+scale_x_continuous(breaks = seq(-82, -68, by = 4)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))
+
+#Insert general plot within specific plot
+panel2 = p2 + inset_element(p1, left = 0, bottom = -0.885, right = 0.35, top = 1.2, ignore_tag = TRUE)
+
+library(patchwork)
+panel1 + panel2
+
+```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
 A) Phylogenetic tree at the genera level.
 Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).
-The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted circumferences at the end of the tip branches.
-Larger circumferences indicate larger brains in proportion to their body and vice versa.
-B) Heatmap indicating the habitat preferences of the different bee species."}
+The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted points at the end of the tip branches.
+Larger points indicate larger brains in proportion to their body size and vice versa.
+B) Heatmap showing the habitat preference for each bee species.
+The columns delimit the habitat type (i.e., natural, agricultural, semideveloped and urban) and the rows the different bee species."}
 
@@
                        breaks = c("Europe", "USA", "USA and Europe")) +
-  ggtitle("A") + theme(plot.title = element_text(size=22, face="bold"))
+  ggtitle("A") + theme(plot.title = element_text(size=18, face="bold"))
 
@@
    # scale_fill_gradient(high="black",  low="white") +
-       scale_fill_viridis_c(name = "Habitat\n preference") + 
+       scale_fill_viridis_c(name = "Habitat\n preference", limits=c(0, 1), breaks=c(0.1, 0.5, 0.9), labels=c("Low","","High")) + 
     theme_minimal() +
@@
     axis.text.y = element_text(face="italic", color="black", size=7, hjust=0),
-    plot.title = element_text(size=22, face="bold")) +
+    plot.title = element_text(size=18, face="bold")) +
   ggtitle("B")
@@
 
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural, semideveloped and urban).
+The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States."}
+
+#Histogram of habitat preferences
+#Load libraries
+library(tidyverse)
+
+#Load preferences
+preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
+preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
+
+#Merge datasets
+preferences = bind_rows(preferences_europe, preferences_usa)
+
+#Convert to long to model everything at the same time
+long_data = preferences %>% gather(Habitat, Preference, 2:5, -c(Species))
+
+#Relevel for plotting
+long_data = long_data %>% 
+    mutate(Habitat = recode_factor(Habitat, "Seminatural" = "Semideveloped")) %>% 
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural","Semideveloped", "Urban")))
+
+labels.minor <- c("Low", "", "High")
+
+#Plot
+ggplot(long_data, aes(x=Preference, color= Habitat)) +
+    geom_histogram(aes(fill= Habitat)) + 
+    xlab("Habitat preference") +
+    ylab("Frequency") +
+    facet_wrap(~ Habitat) +
+    scale_colour_viridis_d(end=0.98)+
+    scale_fill_viridis_d(end=0.98) +
+    theme_bw()+
+    theme(axis.title = element_text(face = "bold", size=14), 
+    strip.text.x = element_blank() , strip.background = element_blank(),
+    panel.border=element_rect(size=0.8))+
+    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5),
+                       labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
+    scale_y_continuous(expand = c(0, 0))+
+    geom_vline(aes(xintercept = 0.05), colour="black", linetype="dashed") +
+    geom_vline(aes(xintercept = 0.95), colour="black", linetype="dashed")
+
+```
+
 # Discussion

### 2022-10-25 Manuscript/Ms.Rmd
Map edits

@@
 p2 = ggplot(Europe) +
-geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
 geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8)) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
 coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "aliceblue"),
+theme(panel.background = element_rect(fill = "white"),
 panel.border = element_rect(colour = "black", fill=NA, size=1)) +
@@
 p2 = ggplot(canada) + 
-geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray70",color = "black", alpha = 1,size=0.2) +
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
 geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 3, 6)) +
-coord_sf(xlim = c(-82, -68), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "aliceblue"),
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
+theme(panel.background = element_rect(fill = "white"),
 panel.border = element_rect(colour = "black", fill=NA, size=1)) +
@@
 scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -68, by = 4)) +
+scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
 guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))
@@
 #Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.885, right = 0.35, top = 1.2, ignore_tag = TRUE)
+panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
 

### 2022-10-26 Manuscript/Ms.Rmd
Reorganising files

@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the location of the downloaded GBIF occurrences for the different bee species."}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the location of the downloaded GBIF occurrences for the different bee species (N = 100).
+A) Selected North European countries that had a high number of records for the selected species.
+B) Selected states from"}
 

### 2022-10-26 Scripts/0_Processing/0_Preprocessing.R

@@ -10,7 +10,7 @@ species.brains = brains.it %>%
 filter(!ID == "F33") %>%  #Dasypoda visnaga's brain was conserved in ethanol, not formol
 rename(Brain.weight = Brain.Weight..mg.) %>% #Rename var
 mutate(Brain.IT = (Brain.weight/IT)) %>% #create new col
-dplyr::select(Species, IT, Brain.weight, Brain.IT) %>% #select cols of interest
+dplyr::select(Species, IT, Brain.weight, Brain.IT, Sex) %>% #select cols of interest
 na.omit() #filter out na's
 
 #Check levels
@@ -107,6 +107,8 @@ species.brains<-subset(species.brains, subset = !((Species == "Dasypoda sp.")))
 species.brains<-subset(species.brains, subset = !((Species == "Coelioxys sp.")))
 species.brains<-subset(species.brains, subset = !((Species == "Andrena (Chlorandrena)")))
 species.brains<-subset(species.brains, subset = !((Species == "Andrena (Chrysandrena)")))
+
+species.brains = species.brains %>%  filter(Sex == "Female") %>% dplyr::select(!Sex)
 
 #Create dataframe with average measurements of IT and brain weight
 weights.mean <- data.frame(aggregate(Brain.weight ~ Species, data = species.brains, FUN = mean))

### 2022-10-26 Scripts/8_Analyses/Qualitative/Europe/1_Analysis_Europe.R
reorganising files

@@ -44,7 +44,9 @@ ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
     theme_bw() +
     ylab("Habitat preference") +
     xlab("Residuals") + 
-    ggtitle("Europe") #+ facet_wrap(~ Habitat)
+    ggtitle("Europe") + facet_wrap(~ Habitat)
+
+
 #Save data and model1 output
 write_csv(long_data , "Data/Europe_data/data_preference_residuals_europe_qualitative.csv")
 write_csv(ce1[[1]], "Data/Europe_data/model_output_preference_residuals_europe_qualitative.csv")

### 2022-10-26 Scripts/8_Analyses/Qualitative/USA/1_Analysis_USA_Gbif_Thesis.R

@@ -23,6 +23,8 @@ long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
 long_data = long_data %>% 
     mutate(brain_it = Brain.weight/IT)
 
+#write_rds(long_data, "Data/Processing/long_data_usa.rds")
+
 #Analysis Preference ~ residuals----
 model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=bernoulli())

### 2022-10-26 Scripts/Figure_1_Phylo_Tree/Figure_1_Phylo_tree_all.R

@@ -54,13 +54,12 @@ p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country)+
                        breaks = c("Europe", "USA", "USA and Europe"))
 
 #Convert wide format to long for plotting heatmap
-d=gather(preferences, condition, measurement, Agricultural:Seminatural, factor_key=TRUE)
+d=gather(preferences, condition, measurement, Agricultural:Urban, factor_key=TRUE)
 
 
 #Relevel for plotting
 d = d %>% 
-mutate(condition = recode_factor(condition, "Seminatural" = "Semi-developed")) %>% 
-mutate(condition=fct_relevel(condition,c("Natural","Agricultural","Semi-developed", "Urban"))) 
+mutate(condition=fct_relevel(condition,c("Natural","Agricultural", "Urban"))) 
 
 
 p2 = ggplot(d, aes(x=condition, y=Species)) + 

### 2022-11-16 Manuscript/Ms.Rmd
reorganising files
edits

--- 2022-10-25 Manuscript/Ms.Rmd
@@
 title: "Bee brain size is positively associated with urban lifestyles but not with positive population trends"
-output: pdf_document
+output:
+  pdf_document:
+    fig_caption: yes
 
@@
 - \captionsetup[figure]{labelformat=empty}
+- \usepackage{float} 
+- \floatplacement{figure}{H} 
+---
 
----
+```{r echo=FALSE,warning=FALSE}
+
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+```
 
@@
 # Introduction
+
+Relevant refs:
+
+https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the location of the downloaded GBIF occurrences for the different bee species."}
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species (N = 85).
+A) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.
+B) East Coast of Unites States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.", fig.pos="p"}
 
@@
 
+\clearpage
+
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
@@
 
-#Phylogenetic tree for USA and Europe
-library(ggtree) 
+library(ggtree) #v3.5.3.002
 library(ggplot2)
@@
 #Plot tree
-p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country) +  geom_tree(size=0.8)+
+p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country)+
     geom_tippoint(aes(size = residuals, color=Country))+
-    theme_tree2()+scale_size_continuous(range = c(-1.2, 3.2),name="Residuals")+
+    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Residuals")+
     scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
@@
 #Convert wide format to long for plotting heatmap
-d=gather(preferences, condition, measurement, Agricultural:Seminatural, factor_key=TRUE)
+d=gather(preferences, condition, measurement, Agricultural:Urban, factor_key=TRUE)
 
@@
 d = d %>% 
-mutate(condition = recode_factor(condition, "Seminatural" = "Semideveloped")) %>% 
-mutate(condition=fct_relevel(condition,c("Natural","Agricultural","Semideveloped", "Urban")))
+mutate(condition=fct_relevel(condition,c("Natural","Agricultural", "Urban")))
 
@@
    # scale_fill_gradient(high="black",  low="white") +
-       scale_fill_viridis_c(name = "Habitat\n preference", limits=c(0, 1), breaks=c(0.1, 0.5, 0.9), labels=c("Low","","High")) + 
+    scale_fill_viridis_c(name = "Habitat\n preference") + 
     theme_minimal() +
@@
     theme(axis.text.x = element_text(face="bold", color="black", 
-    size=10,  vjust=1.15),
-    axis.text.y = element_text(face="italic", color="black", size=7, hjust=0),
+    size=7.5,  vjust=1.15),
+    axis.text.y = element_text(face="italic", color="black", size=7.5, hjust=0),
     plot.title = element_text(size=18, face="bold")) +
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural, semideveloped and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States."}
-
-#Histogram of habitat preferences
-#Load libraries
-library(tidyverse)
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=4, fig.width=10, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
+The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
 
@@
 #Convert to long to model everything at the same time
-long_data = preferences %>% gather(Habitat, Preference, 2:5, -c(Species))
+long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
 
@@
 long_data = long_data %>% 
-    mutate(Habitat = recode_factor(Habitat, "Seminatural" = "Semideveloped")) %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural","Semideveloped", "Urban")))
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
 
 labels.minor <- c("Low", "", "High")
+
+long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),
+                  labels = c("A) Natural", "B) Agricultural", "C) Urban"))
 
@@
     scale_fill_viridis_d(end=0.98) +
-    theme_bw()+
-    theme(axis.title = element_text(face = "bold", size=14), 
-    strip.text.x = element_blank() , strip.background = element_blank(),
-    panel.border=element_rect(size=0.8))+
+    theme_bw() +   
+    theme(axis.title = element_text(face = "bold"),
+    panel.border=element_rect(size=1), 
+    legend.position = "none",
+    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
+    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"))+
     scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5),
                        labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
-    scale_y_continuous(expand = c(0, 0))+
+    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
     geom_vline(aes(xintercept = 0.05), colour="black", linetype="dashed") +
@@
 ```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.}", fig.pos="p"}
+
+#Load libraries
+library(readr)
+library(dplyr)
+library(ggplot2)
+
+#Load data for plotting
+long_data <- read_csv("../Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
+
+model1 = read_csv("../Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")
+
+model2 = read_csv("../Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")
+
+model3 = read_csv("../Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")
+
+#Relevel for plotting
+long_data = long_data %>% 
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
+#Relevel for plotting
+model1 = model1 %>% 
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
+#Relevel for plotting
+model2 = model2 %>% 
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
+#Relevel for plotting
+model3 = model3 %>% 
+    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
+
+#colorBlindness::displayAllColors(viridis::viridis(4))
+
+all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
+geom_line(aes(color=Habitat),size=0.7) +
+geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
+theme_bw() +
+ylab("Habitat preference") +
+xlab("Residuals") +
+scale_colour_viridis_d() +
+scale_fill_viridis_d() +
+scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
+ggtitle("A") +
+theme(plot.title = element_text(face = "bold", size=18))
+
+all_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
+geom_line(aes(color=Habitat),size=0.7) +
+geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
+theme_bw() +
+ylab(NULL) +
+xlab("Brain weight (mg)") +
+scale_colour_viridis_d() +
+scale_fill_viridis_d() +
+scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
+ggtitle("B") +
+theme(plot.title = element_text(face = "bold", size=18))
+
+all_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
+geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
+geom_line(aes(color=Habitat),size=0.7) +
+geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
+theme_bw() +
+ylab(NULL) +
+xlab("Intertegular distance (mm)") +
+scale_colour_viridis_d() +
+scale_fill_viridis_d() +
+scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
+ggtitle("C") +
+theme(plot.title = element_text(face = "bold", size=18))
+
+library(patchwork)
+
+all_1 + (all_2/all_3) + plot_layout(guides = "collect") &
+theme(legend.position = 'bottom', panel.border=element_rect(size=1),
+      axis.title=element_text(face="bold"),
+      legend.box.background = element_rect(colour = "black",size=1))
+
+```
+
+\clearpage
 
@@
 
+\newpage
+
 # References

### 2022-11-16 Scripts/Figure_1_Map_Points/Map.Rmd
edits

--- 2022-10-25 Scripts/Figure_4_Map_Points/Map.Rmd
@@
 p2 = ggplot(Europe) +
-geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
 geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8)) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8),limits=c(0,8)) +
 coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "aliceblue"),
+theme(panel.background = element_rect(fill = "white"),
 panel.border = element_rect(colour = "black", fill=NA, size=1)) +

### 2022-11-16 Scripts/4_Data_For_Analysis.R

@@ -20,9 +20,9 @@ final_data <- rbind(europe, usa)
 colnames(final_data) <- str_to_title(colnames(final_data))
 
 #Save data again
-write.csv(final_data, file=gzfile("Data/final_data.csv.gz"), row.names=FALSE)
+write.csv(final_data, file=gzfile("Data/Europe_USA/final_data.csv.gz"), row.names=FALSE)
 #MA, this would read as follows:
-final_data <- data.frame(fread("Data/final_data.csv.gz"))
+final_data <- data.frame(fread("Data/Europe_USA/final_data.csv.gz"))
 
 #Some quick checks about the data
 #Check number of species

### 2022-11-16 Scripts/Figure_3_Preference_Histogram/Figure_3_Preference_histogram.R

@@ -19,6 +19,9 @@ long_data = long_data %>%
 
 labels.minor <- c("Low", "", "High")
 
+long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),
+                  labels = c("A) Natural", "B) Agricultural", "C) Urban"))
+
 #Plot
 ggplot(long_data, aes(x=Preference, color= Habitat)) +
     geom_histogram(aes(fill= Habitat)) + 
@@ -27,14 +30,15 @@ ggplot(long_data, aes(x=Preference, color= Habitat)) +
     facet_wrap(~ Habitat) +
     scale_colour_viridis_d(end=0.98)+
     scale_fill_viridis_d(end=0.98) +
-    theme_bw()+
-    theme(axis.title = element_text(face = "bold"), 
-    strip.text.x = element_blank() , strip.background = element_blank(),
-    panel.border=element_rect(size=0.8))+
+    jtools::theme_apa() +   
+    theme(axis.title = element_text(face = "bold"),
+    panel.border=element_rect(size=1), 
+    legend.position = "none",
+    strip.text = element_text(face = "bold", size = rel(1),hjust = -0.01),
+    strip.background = element_rect(fill = "white", colour = "black", size = 0,linetype="solid"))+
     scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5),
                        labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
-    scale_y_continuous(expand = c(0, 0))+
+    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
     geom_vline(aes(xintercept = 0.05), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.95), colour="black", linetype="dashed")
-
+    geom_vline(aes(xintercept = 0.95), colour="black", linetype="dashed") 
 

### 2022-11-17 Manuscript/Ms.Rmd
starting with methods

@@
 
+## Overview
+
+Our study uses bee traits measured from North America and Europe to explore how different habitat types influence bee cognitive abilities measured as the deviation of the brain weight relative to their body size.
+To explore this, we first downloaded for our set of species, occurrence data from the Global Biodiversity Information Facility from North America and Europe.
+First, we identified the areas with highest density of occurrences and delimited based on that our study regions (East Coast of Unites States and North Central Europe).
+Second, we filtered the resulting occurrence records (e.g., duplicates and a homogeneous geographical distribution).
+
+## Bee species and measurements
+
+## Species occurrences data
+
+Global Biodiversity Information Facility (GBIF, http://www.gbif.org/)
+
 \clearpage
@@
 scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
 
@@
 scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
 

### 2022-11-17 Scripts/Figure_2_Phylo_Tree/Figure_2_Phylo_tree_all.R

@@ -20,6 +20,8 @@ preferences_eu = readr::read_csv("Data/Europe_data/preferences_europe.csv")
 #Bind rows
 pref = bind_rows(preferences_usa, preferences_eu)
 
+unique(pref$Species)
+
 
 #Average preferences between USA and Europe
 preferences = pref %>% 

### 2022-11-17 Manuscript/Ms.Rmd
methods section

@@
 First, we identified the areas with highest density of occurrences and delimited based on that our study regions (East Coast of Unites States and North Central Europe).
-Second, we filtered the resulting occurrence records (e.g., duplicates and a homogeneous geographical distribution).
+Second, we filtered the resulting occurrence records (e.g., we deleted duplicates and selected species with a homogeneous geographical distribution).
+Third, for the resulting species we extracted
 

### 2022-11-17 Manuscript/Ms.Rmd
overview paragraph on the way

@@
 
-Our study uses bee traits measured from North America and Europe to explore how different habitat types influence bee cognitive abilities measured as the deviation of the brain weight relative to their body size.
+Our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas.
+To investigate this, we have used a dataset of bees captured in North America and Europe with information about brain weight and body size.
+Then, we downloaded for all these species the occurrence information from the Global Biodiversity Information Facility (GBIF) and delimited two areas based on the states or countries with highest density of records.
+After curating the occurrence data to avoid possible biases, we extracted the land use classification
+
+uses bee traits measured from North America and Europe to explore how different habitat types influence bee cognitive abilities measured as the deviation of the brain weight relative to their body size.
 To explore this, we first downloaded for our set of species, occurrence data from the Global Biodiversity Information Facility from North America and Europe.

### 2022-11-17 Manuscript/Ms.Rmd
minor edit

@@
 Then, we downloaded for all these species the occurrence information from the Global Biodiversity Information Facility (GBIF) and delimited two areas based on the states or countries with highest density of records.
-After curating the occurrence data to avoid possible biases, we extracted the land use classification
-
-uses bee traits measured from North America and Europe to explore how different habitat types influence bee cognitive abilities measured as the deviation of the brain weight relative to their body size.
-To explore this, we first downloaded for our set of species, occurrence data from the Global Biodiversity Information Facility from North America and Europe.
-First, we identified the areas with highest density of occurrences and delimited based on that our study regions (East Coast of Unites States and North Central Europe).
-Second, we filtered the resulting occurrence records (e.g., we deleted duplicates and selected species with a homogeneous geographical distribution).
-Third, for the resulting species we extracted
+After curating the occurrence data to avoid possible biases, we extracted the land use classification for each of the records and summarised the most representative classifications into 3 different categories: natural, agricultural and urban.
 

### 2022-11-20 Manuscript/Ms.Rmd
some edits

@@
 While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for 145 individuals from 44 species of bees —the most important group of pollinators— from North America and Europe, and compared them with detailed information on habitat preferences and population trends.
+To address this issue, we measured brains for 145 individuals from 44 species of bees ---the most important group of pollinators--- from North America and Europe, and compared them with detailed information on habitat preferences and population trends.
 Our analyses revealed that bees that prefer forest habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger brains compared to forest dwellers.
@@
 
-https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R
+<https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R>
 
-Pollination is a fundamental ecosystem service that connects the earth’s vegetation and human economy.
+Pollination is a fundamental ecosystem service that connects the earth's vegetation and human economy.
 Pollination services are performed by both managed and wild pollinators, mostly bees [@james2008; @winfree2011], both of which might be affected by a range of current and projected human-related environmental changes.
@@
 However, not all pollinators species are equally affected by habitat loss and alteration.
-In bees, for example, there are also “winners” —like Bombus terrestris [@rasmont2008]— that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].
+In bees, for example, there are also "winners" ---like Bombus terrestris [@rasmont2008]--- that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].
 
-Human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+Human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown.
 Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive [@manenti2013].
-For species that tolerate human-related disturbances, human-dominated habitats —notably cities— offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+For species that tolerate human-related disturbances, human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
@@
 Our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas.
-To investigate this, we have used a dataset of bees captured in North America and Europe with information about brain weight and body size.
-Then, we downloaded for all these species the occurrence information from the Global Biodiversity Information Facility (GBIF) and delimited two areas based on the states or countries with highest density of records.
-After curating the occurrence data to avoid possible biases, we extracted the land use classification for each of the records and summarised the most representative classifications into 3 different categories: natural, agricultural and urban.
+To evaluate this hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size.
+For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and integrated it with an independent data source of occurrences from the East Coast of Unites States.
+Then, we limited the study areas to the states or countries with highest density of records and filtered the data to avoid possible biases.
+For the resulting occurrences, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference with the help of null models.
+Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.
 
-## Bee species and measurements
+## Bee species dataset
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Script to load variables
+library(readr)
+library(dplyr)
+
+raw = read_csv("../Raw_data/all_cleaning.csv")
+
+d = raw %>% 
+group_by(Species) %>% 
+summarise(individuals = n())
+
+```
+
+This dataset contained species captured by hand netting in different areas of the East Coast of Unites States and Europe (i.e., Spain and Netherlands).
+There were in total `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species with brain weight and body size information.
+
+Bees were captured
 
@@
 
-Global Biodiversity Information Facility (GBIF, http://www.gbif.org/)
+Global Biodiversity Information Facility (GBIF, <http://www.gbif.org/>)
 
@@
 We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
-(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p < 0.001, R2= 0.83, n = 163 species).
+(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p \< 0.001, R2= 0.83, n = 163 species).
 Residuals were calculated using our complete brain weights and body size database.
@@
 Our subset may be biased towards the most common genera and contain fewer species per genus.
-If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel’s lambda > 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
+If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel's lambda \> 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
 Population trends or habitat preferences could be explained by the availability of the resources consumed, for example, in the Netherlands, population trends of some bee species are related to the availability of their preferred plants (Scheper et al., 2014).

### 2022-11-20 Scripts/0_Processing/0_Preprocessing.R

@@ -112,6 +112,19 @@ mutate(Sex =replace(Sex, Sex == "Queen?", "Queen"))
 
 #IMPORTANT! select just females
 all_cleaning = all_cleaning %>%  filter(Sex == "Female") %>% dplyr::select(!Sex)
+#Save data to load raw values in methods
+write_csv(all_cleaning, "Raw_data/all_cleaning.csv")
+
+
+#Check unique cases and levels for methods
+
+d = all_cleaning %>% 
+group_by(Species) %>% 
+summarise(individuals = n())
+median(d$individuals)
+sum(d$individuals) #433
+nlevels(factor(unique(d$Species))) #113 species
+
 
 #Create dataframe with average measurements of IT and brain weight
 weights.mean <- data.frame(aggregate(Brain.weight ~ Species, data = all_cleaning, FUN = mean))

### 2022-11-20 Scripts/0_Processing/trial.R

@@ -1,0 +1,10 @@
+all_cleaning = read_csv("Raw_data/all_cleaning.csv")
+
+
+install.packages("taxize")
+
+library(taxize)
+
+specieslist = unique(all_cleaning$Species)
+
+t <- tax_name(query = c(specieslist), get = c("phylum","class", "order", "family", "genus"), db = "ncbi")

### 2022-11-21 Manuscript/Ms.Rmd
uploading data

@@
 For the resulting occurrences, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference with the help of null models.
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference
+by comparing the observed occurrences with the ones of null models.
 Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.
-
-## Bee species dataset
 
@@
 
-This dataset contained species captured by hand netting in different areas of the East Coast of Unites States and Europe (i.e., Spain and Netherlands).
-There were in total `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species with brain weight and body size information.
+## Bee species dataset
 
-Bees were captured
+Our dataset consist in bee specimens captured by hand netting in different areas of the East Coast of the United States and Europe (i.e., Spain and Netherlands).
+In total, it contains `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species with brain weight and body size information.
+Of these, 93 species with their respective traits were previously published in @sayol2020.
+We considered exclusively female individuals as most bees exhibit sexual dimorphism [@shreeves2008] that likely imply intraspecific differences in brain size between male and female individuals.
+
+## Brain and body size measurements
 

### 2022-11-21 Scripts/0_Processing/Taxonomic_rank.R

@@ -1,0 +1,50 @@
+
+#Here I load the filtered data and check the families and genera of the different species
+
+#Load library
+library(taxize)
+library(readr)
+library(dplyr)
+library(stringr)
+
+#Read data
+all_cleaning = read_csv("Raw_data/all_cleaning.csv")
+
+#Create spp list
+specieslist = unique(all_cleaning$Species)
+
+#Download data
+d <- tax_name(query = c(specieslist), get = c("order", "family", "genus"), db = "ncbi")
+
+#Missing values
+d_missing = d %>% 
+filter(is.na(order))
+
+#Non missing
+d_non_missing = d %>% 
+filter(!is.na(order)) 
+
+#Try to add missing variables with mutate
+d_missing = d_missing %>% 
+mutate(order = "Hymenoptera") %>% 
+mutate(family = case_when(query=="Osmia lateralis" ~ "Megachilidae",
+                          query=="Panurgus venustus" ~ "Andrenidae",
+                          query=="Andrena varians" ~ "Andrenidae",
+                          query=="Andrena rhyssonota" ~ "Andrenidae",
+                          query=="Anthophora dispar" ~ "Apidae",
+                          query=="Eucera rufa" ~ "Apidae",
+                          query=="Chelostoma philadephi" ~ "Megachilidae",
+                          query=="Lasioglossum dialictus" ~ "Halictidae")) %>% 
+mutate(genus = word(query, 1))
+
+#Bind rows again
+all = bind_rows(d_non_missing, d_missing)
+#Save data
+write_csv(all, "Data/Processing/taxonomic_rank.csv")
+
+#Check total number of families and genus
+
+nlevels(factor(unique(all$family))) #6
+nlevels(factor(unique(all$genus))) #30
+                    
+ 

### 2022-11-21 Scripts/0_Processing/trial.R

@@ -1,10 +1,0 @@
-all_cleaning = read_csv("Raw_data/all_cleaning.csv")
-
-
-install.packages("taxize")
-
-library(taxize)
-
-specieslist = unique(all_cleaning$Species)
-
-t <- tax_name(query = c(specieslist), get = c("phylum","class", "order", "family", "genus"), db = "ncbi")

### 2022-11-21 Manuscript/Ms.Rmd
edits

@@
 
-#Script to load variables
 library(readr)
@@
 
+#Load final dataset 
 raw = read_csv("../Raw_data/all_cleaning.csv")
@@
 
+#Load taxonomic rank script
+tax = read_csv("../Data/Processing/taxonomic_rank.csv")
+
 ```
@@
 
-Our dataset consist in bee specimens captured by hand netting in different areas of the East Coast of the United States and Europe (i.e., Spain and Netherlands).
-In total, it contains `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species with brain weight and body size information.
-Of these, 93 species with their respective traits were previously published in @sayol2020.
-We considered exclusively female individuals as most bees exhibit sexual dimorphism [@shreeves2008] that likely imply intraspecific differences in brain size between male and female individuals.
+Our dataset contains measurements of brain weight and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
+In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
+We have considered exclusively female specimens as most bees exhibit sexual dimorphism [@shreeves2008] that likely imply intraspecific differences in brain size between male and female individuals.
 

### 2022-11-21 Manuscript/Ms.Rmd
edits

@@
 Our dataset contains measurements of brain weight and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
-In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
+In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
 We have considered exclusively female specimens as most bees exhibit sexual dimorphism [@shreeves2008] that likely imply intraspecific differences in brain size between male and female individuals.

### 2022-11-22 Manuscript/Ms.Rmd
methods update

@@
 
+This may be also relevant:
+
+Although recent studies challenge the use of absolute brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein), many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008).
+
 Pollination is a fundamental ecosystem service that connects the earth's vegetation and human economy.
@@
 
-## Overview
+## (a) Overview
 
@@
 
-## Bee species dataset
+## (b) Bee species dataset
 
-Our dataset contains measurements of brain weight and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
+Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
 In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We have considered exclusively female specimens as most bees exhibit sexual dimorphism [@shreeves2008] that likely imply intraspecific differences in brain size between male and female individuals.
+We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between sexes that can be correlated with the size and structure of the brain [@streinzer2013; @roselino2015].
+In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
+Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
 
-## Brain and body size measurements
+## (c) Brain and body size measurements
 
-## Species occurrences data
+We used as a measurement of brain size the weight of fixed brains.
+To weight the brains, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
+Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
+Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies.
+By doing this, we ensured that our brain measurements mainly accounted for neural tissue.
+We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
+Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
+
+We considered as a proxy of body size the intertegular span (ITS) that represents the distance between the base of the wings (tegulae) on the bee thorax.
+ITS has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
+All ITS measurements were conducted with a stereomicroscope (magnification 16 x –80 x ) with a calibrated ocular micrometre (resolution down to 0.02 mm).
+
+## (d) Species occurrences data
 

### 2022-11-23 Manuscript/Ms.Rmd
Edits

@@
 In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between sexes that can be correlated with the size and structure of the brain [@streinzer2013; @roselino2015].
+We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between female and male individuals [@shreeves2008] that often correlate with the size and structure of the brain [@streinzer2013; @roselino2015].
 In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
@@
 
-## (d) Species occurrences data
+## (d) Occurrence data
 
-Global Biodiversity Information Facility (GBIF, <http://www.gbif.org/>)
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Calculate study site areas
+#Load libraries
+library(sf)
+library(sfheaders)
+
+#EUROPE:
+#Read data
+euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
+#Add area col
+euro_map$area <- st_area(euro_map)
+#Create dataframe with the correct units
+d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
+mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
+#Calculate area
+eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")
+
+#USA:
+#Read data
+usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
+#Not an sf object, convert into multypolygon
+usa = sfheaders::sf_multipolygon(
+  obj = usa_states_1
+  , multipolygon_id = "group"
+  , x = "long"
+  , y = "lat"
+) %>% st_set_crs(4326)
+#Calculate area
+usa$area <- st_area(usa)
+#Create dataframe with the correct units
+d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
+mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
+#Sum areas
+usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")
+
+```
+
+We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
+The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
+The occurrence data can be downloaded from https://doi.org/10.15468/dl.nenuem where just records for the countries of interest were retrieved.
+Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
+For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$.
+These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+Similarly, we selected countries located on the north and center of Europe (Figure 1B) that represented a total area of `r eukm2` km$^{2}$.
+These countries were: England, Belgium, Netherlands, Germany and Denmark.
+
+External data information [...] Origin?
 
@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species (N = 85).
-A) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.
-B) East Coast of Unites States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.", fig.pos="p"}
+A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.", fig.pos="p"}
 
@@
 xlab("Longitude") +
-ggtitle("A") +
+ggtitle("B") +
 theme(legend.position = c(0, 1),legend.justification = c(0, 1),
@@
 xlab("Longitude") +
-ggtitle("B") +
+ggtitle("A") +
 theme(legend.position = c(0, 1),legend.justification = c(0, 1),
@@
 library(patchwork)
-panel1 + panel2
+panel2  + panel1
 
@@
 \clearpage
+
+## (e) Construction of the phylogenetic tree
+
+**Data gathering**
+
+The quantitative data of population trends for North American bees were extracted from Bartomeus et al., 2013, where they analyzed trends for 187 species over time, using data from pinned bees from different North American museums and universities.
+We assigned the estimated value of their relative change over time models as values of population trends for the 44 species for which we have brain size measurements (positive for increasing populations and negative for decreasing populations).
+Habitat preferences for USA bee data were extracted from Collado et al., 2019, were habitat preference for forests and urban areas was calculated comparing species occurrence in each habitat to what would be expected if species occur at random using null models.
+This preference index ranges from 0 (avoidance) to 1 (preference).In order to have a representative value of brain weight and body size per species, we used means of the different collected individuals from each species.
+Twenty-four individuals with brain weight values one standard deviation higher or lower than the average were removed (see sup mat analysis) because human errors in their measurement are likely with such delicate tissues.
+Specially, brain weights can give errors due to dehydration or insufficient removal of fat bodies and trachea around the brain.
+For eusocial species, we only used workers.
+We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
+(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p \< 0.001, R2= 0.83, n = 163 species).
+Residuals were calculated using our complete brain weights and body size database.
+
+**Data analysis**
+
+We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms [@burkner2017], to assess whether brain sizes, habitat preference and population trends are correlated.
+In all models, we incorporated a phylogenetic covariance matrix.
+The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
+1, Fig.
+S2, Fig.
+S3).
+Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
+Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6‐44, Revell, 2012).
+Because we did not have information on our 3 variables for the 44 species (habitat preference, brain weight and population trends), we used 3 different subsets of our dataset to maximize the sample size in the different models: Relative brain size-habitat preference, population trends-habitat preference and population trends-relative brain size.
+
+# Results
 
@@
 
-**Animals**
-
-Most bees were opportunistically captured by hand netting in different regions of New York State (USA), and complemented with bees captured from Spain and the Netherlands for species with a holarctic distribution.
-We captured 145 individuals from 44 species, kept them in cold storage and sent them to the laboratory to be processed.
-Once the bees were in the laboratory, they were anesthetized in cold (Crook, 2013) and decapitated.
-Bees were identified by expert taxonomists (Parker Gambino and Ivo Raemakers).
-The species used for the analysis can be consulted in Table S2.
-We measured the inter-tegular distance, as it is a good proxy of body size (Kendall et al., 2019).
-To prevent degradation, the head was fixed in 4% paraformaldehyde with phosphate buffer saline (PBS).
-Brains were extracted from the head capsule, and tracheae and fat bodies were removed to maximize accuracy in neural tissue weight.
-Brains extracted were placed on a small piece of tared Parafilm® and exceeding fixative solution was dried from the brain using Kimwipes® tissues.
-Finally the brain was weighted in a microbalance.
-Brain weight was used as a proxy of brain size (Sayol et al., 2019 in preparation).
-
-**Data gathering**
-
-The quantitative data of population trends for North American bees were extracted from Bartomeus et al., 2013, where they analyzed trends for 187 species over time, using data from pinned bees from different North American museums and universities.
-We assigned the estimated value of their relative change over time models as values of population trends for the 44 species for which we have brain size measurements (positive for increasing populations and negative for decreasing populations).
-Habitat preferences for USA bee data were extracted from Collado et al., 2019, were habitat preference for forests and urban areas was calculated comparing species occurrence in each habitat to what would be expected if species occur at random using null models.
-This preference index ranges from 0 (avoidance) to 1 (preference).In order to have a representative value of brain weight and body size per species, we used means of the different collected individuals from each species.
-Twenty-four individuals with brain weight values one standard deviation higher or lower than the average were removed (see sup mat analysis) because human errors in their measurement are likely with such delicate tissues.
-Specially, brain weights can give errors due to dehydration or insufficient removal of fat bodies and trachea around the brain.
-For eusocial species, we only used workers.
-We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
-(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p \< 0.001, R2= 0.83, n = 163 species).
-Residuals were calculated using our complete brain weights and body size database.
-
-**Data analysis**
-
-We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms [@burkner2017], to assess whether brain sizes, habitat preference and population trends are correlated.
-In all models, we incorporated a phylogenetic covariance matrix.
-The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
-1, Fig.
-S2, Fig.
-S3).
-Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
-Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6‐44, Revell, 2012).
-Because we did not have information on our 3 variables for the 44 species (habitat preference, brain weight and population trends), we used 3 different subsets of our dataset to maximize the sample size in the different models: Relative brain size-habitat preference, population trends-habitat preference and population trends-relative brain size.
-
-# Results
+\clearpage
 

### 2022-11-23 Scripts/1_Download_GBIF_data/1_Main_download_occ_download.R

@@ -3,7 +3,7 @@
 #Quite fast but we need a gbif account!
 #To reproduce our analysis, 
 #Skip this part and use directly the downloaded data from the DOI
-#https://doi.org/10.15468/dl.re5wwt
+#https://doi.org/10.15468/dl.nenuem
 
 #Load libraries
 library(stringr) 

### 2022-11-23 Scripts/3_Filter_data/Calculate_areas.R

@@ -1,0 +1,35 @@
+#Calculate study site areas
+
+#Load libraries
+library(sf)
+library(sfheaders)
+
+#EUROPE:
+#Read data
+euro_map <- st_read("Data/Europe_data/euro_map_1.shp")
+#Add area col
+euro_map$area <- st_area(euro_map)
+#Create dataframe with the correct units
+d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
+mutate(area_NAME1 = units::set_units(area_NAME, value = "km^2"))
+#Calculate area
+sum(d1$area_NAME1)
+
+#USA:
+#Read data
+usa_states_1 <- read.csv("Data/Usa_data/usa_states.csv")
+#Not an sf object, convert into multypolygon
+usa = sfheaders::sf_multipolygon(
+  obj = usa_states_1
+  , multipolygon_id = "group"
+  , x = "long"
+  , y = "lat"
+) %>% st_set_crs(4326)
+#Calculate area
+usa$area <- st_area(usa)
+#Create dataframe with the correct units
+d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
+mutate(area_NAME1 = units::set_units(area_NAME, value = "km^2"))
+#Sum areas
+sum(d2$area_NAME1)
+usakm2 = round(sum(d2$areakm2),0)

### 2022-11-23 Scripts/Figure_1_Map_Points/Figure_map.R

@@ -42,7 +42,8 @@ axis.title = element_text(face="bold"),
 plot.title = element_text(face="bold"),
 legend.text = element_text(face = "bold")) +
 scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8))
+scale_x_continuous(breaks = seq(-8, 18, by = 8))+
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15) 
 
 #Insert general plot within specific plot
 panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE) 
@@ -99,9 +100,8 @@ axis.title = element_text(face="bold"),
 plot.title = element_text(face="bold"),
 legend.text = element_text(face = "bold")) +
 scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -68, by = 4))
-
-
+scale_x_continuous(breaks = seq(-82, -68, by = 4)) + 
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
 
 
 #Insert general plot within specific plot

### 2022-11-23 Scripts/5_Add_land_cover/USA_Land_Cover.R
Updating analysis of USA

@@ -2,7 +2,7 @@
 #Load libraries
 library(data.table)
 library(tidyverse)
-library(raster)
+library(terra)
 #Load data
 data <- read.table("Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",") 
 
@@ -12,29 +12,16 @@ group_by(species) %>%
 summarize(no_rows= length(species))
 
 #Extract land use
-#Adapted code from this example https://www.r-bloggers.com/2014/11/spatial-data-extraction-around-buffered-points-in-r/
+#Adapted code from  https://stackoverflow.com/a/74418321/13362117
 #Extract land use ----
 #Raster data downloaded from: https://www.mrlc.gov/data/nlcd-2006-land-cover-conus
-NLCD <- raster("Data/Raster_USA/nlcd_2006_land_cover_l48_20210604.img")
-coords <- data[, c("long", "lat")]  
-#convert lat/lon to appropriate projection
-names(coords) <- c("x", "y")
-coordinates(coords) <- ~x + y
-proj4string(coords) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
-crs_args <- NLCD@crs@projargs
-Datos_transformed <- spTransform(coords, CRS(crs_args))
-#extract land cover data for each point
-data$num.codes <- raster::extract(NLCD, Datos_transformed)
+NLCD <- rast("Data/Raster_USA/nlcd_2006_land_cover_l48_20210604.img")
 
-#Generate land cover number to name conversions
-num.codes <- unique(unlist(data$num.codes))
-cover.names <- NLCD@data@attributes[[1]]$NLCD.Land.Cover.Class[num.codes + 1]
-levels(cover.names)[1] <- NA # first level is ""
-conversions <- data.frame(num.codes, cover.names)
-conversions <- na.omit(conversions)
-conversions <- conversions[order(conversions$num.codes),]
-#left_join
-data = left_join(data,conversions)
+#Create vector with coordinates and correct projection
+coord <- vect(data, c("long", "lat"), crs="+proj=longlat") |> project(NLCD)    
+#Add cover names to data
+cover.names <- extract(NLCD, coord, ID=FALSE)
+data$cover.names <- cover.names$`NLCD Land Cover Class`
 #Select cols of interest
 data = data %>% dplyr::select(species, cover.names)
 #Save data

### 2022-11-23 Scripts/8_Analyses/Qualitative/All/2_Analysis_all_Subset.R

@@ -1,112 +1,0 @@
-#Analysis of Europe bee preferences
-
-#Load libraries
-library(tidyverse) 
-library(brms)
-
-#Load data ----
-#Load brain data
-brain_weight = read_csv("Data/brain_weight_data.csv")
-#Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative_subset.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative_subset.csv") 
-#merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
-#Load phylogenetic matrix to correct for in analysis
-A10 = readRDS("Data/phylo_all.rds")
-
-#Prepare data----
-d = left_join(preferences, brain_weight) %>% 
-    mutate(Species = str_replace_all(Species, " ", "_"))
-
-#Convert to long to model everything at the same time
-long_data = d %>% gather(Habitat, Preference, 2:4, -c(Species))
-
-#Prepare col brain weight/IT
-long_data = long_data %>% 
-    mutate(brain_it = Brain.weight/IT)
-
-#Analysis Preference ~ residuals----
-model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
-             data = long_data, data2 = list(A = A10), family=bernoulli())
-
-ce1 <- conditional_effects(model1, effects = "residuals:Habitat",points=T) 
-
-bayes_R2(model1)
-
-p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
-    geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
-    geom_line(aes(color=Habitat)) +
-    theme_bw() +
-    ylab("Habitat preference") +
-    xlab("Residuals") + 
-    ggtitle("USA and Europe")
-
-#Save data and model1 output
-write_csv(long_data , "Data/Europe_USA/data_preference_residuals_europe_usa_subset_qualitative.csv")
-write_csv(ce1[[1]], "Data/Europe_USA/model_output_preference_residuals_europe_usa_subset_qualitative.csv")
-
-#Analysis Preference ~ brain weight----
-model2 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)), 
[…]

### 2022-11-24 Manuscript/Ms.Rmd
Updating analysis of USA
large update

@@
 To evaluate this hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size.
-For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and integrated it with an independent data source of occurrences from the East Coast of Unites States.
-Then, we limited the study areas to the states or countries with highest density of records and filtered the data to avoid possible biases.
-For the resulting occurrences, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference
-by comparing the observed occurrences with the ones of null models.
+For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and selected as study area the states or countries with highest density of records.
+After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models.
 Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.
@@
 The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
-The occurrence data can be downloaded from https://doi.org/10.15468/dl.nenuem where just records for the countries of interest were retrieved.
 Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
@@
 These countries were: England, Belgium, Netherlands, Germany and Denmark.
+All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
 
-External data information [...] Origin?
+To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals on latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 that allowed us to center the land use classification information for North America and Europe on 2006 (i.e, +/- 16 year difference with the extremes).
+
+## (e) Land use classification
 
@@
 
-## (e) Construction of the phylogenetic tree
+## (f) Construction of the phylogenetic tree
 

### 2022-11-24 Scripts/3_Filter_data/Europe_Data_Cleaning.R

@@ -1,10 +1,10 @@
 ######-
 #In this script we prepare Europe data for analysis
 #We take 5 key decisions in data cleaning
-#1)Filter records above 1985 (this is decided based on the GIS data available)
+#1)Filter records above 1990 (this is decided based on the GIS data available)
 #2)Filter by unique capture event (locality/coordinate)
 #3)Filter by minimum of 3 decimals on coordinates
-#4)Filter number of levels per species (minimum N=100)
+#4)Filter number of levels per species (minimum N=50)
 #5)Filter by wide geographical distribution
 ######-
 
@@ -30,13 +30,13 @@ mutate(lat = gsub(")", "", lat))
 #FIRST FILTER----
 #############-
 #First filter all records above 1985 
-all_above_1985 <- all %>% filter(year>1985)
+all_above_1990 <- all %>% filter(year>1990)
 
 #############-
 #SECOND FILTER----
 #############-
 #Filter by unique coordinate and locality 
-coordinate <- all_above_1985 %>% distinct(scientificName, year, month,day, long, lat,  .keep_all = T)
+coordinate <- all_above_1990 %>% distinct(scientificName, year, month,day, long, lat,  .keep_all = T)
 all_unique_event <- coordinate %>% distinct(scientificName, year, month,day,locality,  .keep_all = T)
 
 #############-
@@ -155,7 +155,7 @@ s <- data.frame(all_above_50 %>%
 #SAVE DATA----
 #############-
 
-#58 records
+#56 levels
 unique(as.factor(all_above_50$species))
 
 #Save

### 2022-11-24 Scripts/3_Filter_data/USA_Data_Cleaning.R

@@ -1,15 +1,12 @@
 ######-
 #In this script we prepare USA data for analysis
 #We take 5 key decisions in data cleaning
-#1)Filter records above 1987 (this is decided based on the GIS data available)
+#1)Filter records above 1990 (this is decided based on the GIS data available)
 #2)Filter by unique capture event
 #3)Filter by minimum of 2 decimals on coordinates #Because MA dataset has only 2
-#4)Filter number of levels per species (minimum N=100)
+#4)Filter number of levels per species (minimum N=50)
 #5)Filter by wide geographical distribution
 ######-
-
-#Note we also add the data used in the thesis of MA Collado.
-#This data is processed in the folder Collado_data
 
 #Load libraries
 library(data.table)
@@ -25,14 +22,14 @@ all <- all_east_coast
 #############-
 #FIRST FILTER----
 #############-
-#First filter all records above 1987
-all_above_1987 <- all %>% filter(year>1987)
+#First filter all records above 1990
+all_above_1990 <- all %>% filter(year>1990)
 
 #############-
 #SECOND FILTER----
 #############-
 #Filter by unique coordinate and locality 
-coordinate <- all_above_1987 %>% distinct(species, year, month,day, long, lat,  .keep_all = T)
+coordinate <- all_above_1990 %>% distinct(species, year, month,day, long, lat,  .keep_all = T)
 all_unique_event <- coordinate %>% distinct(species, year, month,day,locality,  .keep_all = T)
 
 #############-
@@ -144,6 +141,9 @@ colnames(all_above_50)
 #############-
 #SAVE DATA----
 #############-
+
+#37 levels
+unique(as.factor(all_above_50$species))
 
 #Save
 write.csv(all_above_50, file=gzfile("Data/Usa_data/all_above_50_usa.csv.gz"),row.names=FALSE)

### 2022-11-24 Scripts/5_Add_land_cover/USA_Land_Cover.R

@@ -2,7 +2,7 @@
 #Load libraries
 library(data.table)
 library(tidyverse)
-library(terra)
+library(raster)
 #Load data
 data <- read.table("Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",") 
 
@@ -11,19 +11,29 @@ data %>%
 group_by(species) %>% 
 summarize(no_rows= length(species))
 
-#Extract land use
-#Adapted code from  https://stackoverflow.com/a/74418321/13362117
-#Extract land use ----
 #Raster data downloaded from: https://www.mrlc.gov/data/nlcd-2006-land-cover-conus
-NLCD <- rast("Data/Raster_USA/nlcd_2006_land_cover_l48_20210604.img")
+NLCD <- raster("Data/Raster_USA/nlcd_2006_land_cover_l48_20210604.img")
 
-#Create vector with coordinates and correct projection
-coord <- vect(data, c("long", "lat"), crs="+proj=longlat") |> project(NLCD)    
-#Add cover names to data
-cover.names <- extract(NLCD, coord, ID=FALSE)
-data$cover.names <- cover.names$`NLCD Land Cover Class`
+coords <- data[, c("long", "lat")]  
+#convert lat/lon to appropriate projection
+names(coords) <- c("x", "y")
+coordinates(coords) <- ~x + y
+proj4string(coords) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
+crs_args <- NLCD@crs@projargs
+Datos_transformed <- spTransform(coords, CRS(crs_args))
+#extract land cover data for each point
+data$num.codes <- raster::extract(NLCD, Datos_transformed)
+
+#Generate land cover number to name conversions
+num.codes <- unique(unlist(data$num.codes))
+cover.names <- NLCD@data@attributes[[1]]$NLCD.Land.Cover.Class[num.codes + 1]
+levels(cover.names)[1] <- NA # first level is ""
+conversions <- data.frame(num.codes, cover.names)
+conversions <- na.omit(conversions)
+conversions <- conversions[order(conversions$num.codes),]
+#left_join
+data = left_join(data,conversions)
 #Select cols of interest
 data = data %>% dplyr::select(species, cover.names)
 #Save data
 write_csv(data, file=gzfile("Data/Usa_data/land_cover_usa.csv.gz"))
-

### 2022-11-24 Manuscript/Ms.Rmd
Ms update

@@
 All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
+To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large  differences with the extracted land cover classifications for United States and Europe (2006).
 
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals on latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 that allowed us to center the land use classification information for North America and Europe on 2006 (i.e, +/- 16 year difference with the extremes).
-
-## (e) Land use classification
+## (e) Land cover classification
 

### 2022-11-24 Manuscript/Ms.Rmd
minor update on methods

@@
 
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Read occurrence records
+#Load libraries
+library(data.table)
+#Load clean data with bee occurrences
+eu_occurrences <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+eu_occurrences = prettyNum(nrow(eu_occurrences), big.mark=",")
+
+usa_occurrences <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+usa_occurrences = prettyNum(nrow(usa_occurrences), big.mark=",")
+
+```
+
 We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
@@
 All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large  differences with the extracted land cover classifications for United States and Europe (2006).
-
-## (e) Land cover classification
-
-\clearpage
+To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006 for United States and Europe.
+After pruning the data, our final dataset comprised `r usa_occurrences` occurrence records for United States and `r eu_occurrences` records for Europe.
 
@@
 
+## (e) Land cover classification
+
+\clearpage
+
 ## (f) Construction of the phylogenetic tree
+
+## (g) Analysis
 

### 2022-11-25 Manuscript/Ms.Rmd
Update on methods

--- 2022-11-24 Manuscript/Ms.Rmd
@@
 
+\clearpage
+
 # Introduction
@@
 
+\clearpage
+
 # Methods
 
-## (a) Overview
+## Overview
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
 
@@
 
-## (b) Bee species dataset
+## (a) Bee species dataset
 
@@
 
-## (c) Brain and body size measurements
+## (b) Brain and body size measurements
 
@@
 
-## (d) Occurrence data
+## (c) Occurrence data
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+```{r Loading study site areas, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
 
@@
 
+```{r Species and occurrences info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Read occurrence records and extract relevant info
+#Load libraries
+library(data.table)
+#Load clean data with bee occurrences
+europe <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+usa <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+#Europe
+eu_occurrences = prettyNum(nrow(europe), big.mark=",")
+records_eu = europe %>% 
+group_by(species) %>% 
+summarize(no_rows= length(species))
+
+#USA
+usa_occurrences = prettyNum(nrow(usa), big.mark=",")
+records_usa = usa %>% 
+group_by(species) %>% 
+summarize(no_rows= length(species))
+
+#Number of species
+usa_spp = nrow(records_usa)
+eu_spp = nrow(records_eu)
+spp_number = nrow(records_eu) + nrow(records_usa)
+#Median number of records per dataset
+eu_median = median(records_eu$no_rows)
+usa_median = median(records_usa$no_rows)
+
+```
+
 We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
@@
 All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large  differences with the extracted land cover classifications for United States and Europe (2006).
-
-## (e) Land cover classification
-
-\clearpage
+To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006.
+After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for North America (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).
 
@@
 A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.", fig.pos="p"}
+B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
 
@@
 
+## (d) Land cover classification
+
+We assigned a habitat type for all GBIF occurrences by merging the information from land cover inventories with the different georeferenced records.
+The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
+After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 to read and obtain the cover classification, respectively.
+Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
+The original extracted cover classes and our classification into these three different categories are shown in Table S1 for United States and Table S2 for Europe.
+
+## (e) Construction of the phylogenetic tree
+
+## (f) Analysis
+
 \clearpage
-
-## (f) Construction of the phylogenetic tree
 
@@
 
-The quantitative data of population trends for North American bees were extracted from Bartomeus et al., 2013, where they analyzed trends for 187 species over time, using data from pinned bees from different North American museums and universities.
-We assigned the estimated value of their relative change over time models as values of population trends for the 44 species for which we have brain size measurements (positive for increasing populations and negative for decreasing populations).
 Habitat preferences for USA bee data were extracted from Collado et al., 2019, were habitat preference for forests and urban areas was calculated comparing species occurrence in each habitat to what would be expected if species occur at random using null models.
@@
 :::
+
+\clearpage
+
+# Supplementary material
+
+```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
+
+#Create table of occurrences per habitat type (USA)
+#Load libraries
+library(tidyverse)
+library(kableExtra)
+
+#Load extracted data
+pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
+dplyr::select(species, cover.names) %>% 
+mutate_if(is.character,as.factor)
+
+#Create table with original and new cover categories
+pref %>% 
+rename(Cover_names = cover.names) %>% 
+select(Cover_names) %>% 
+group_by(Cover_names) %>% 
+summarise(Occurrences = length(Cover_names)) %>% 
+mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
+       Discarded = "Barren Land", 
+       Agricultural = "Cultivated Crops", 
+       Natural = "Deciduous Forest", 
+       Urban = "Developed, High Intensity", 
+       Urban = "Developed, Medium Intensity",
+       Urban = "Developed, Low Intensity",
+       Urban = "Developed, Open Space", 
+       Natural = "Emergent Herbaceous Wetlands", 
+       Natural = "Evergreen Forest", 
+       Agricultural = "Hay/Pasture", 
+       Natural = "Herbaceous", 
+       Natural = "Mixed Forest", 
+       Discarded = "Open Water", 
+       Natural = "Shrub/Scrub", 
+       Natural = "Woody Wetlands")) %>% 
+arrange(desc(Occurrences)) %>%  
+kbl() %>% 
+kable_styling(font_size = 12,latex_options = "hold_position")
+
+```
+
+\clearpage
+
+```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
+
+#Create table of occurrences per habitat type (USA)
+#Load libraries
+library(tidyverse)
+library(kableExtra)
+
+#Load extracted data
+pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
+    dplyr::select(Species, Cover_names) %>% 
+    mutate_if(is.character,as.factor)
+
+#Create table with original and new cover categories
+pref %>% 
+select(Cover_names) %>% 
+group_by(Cover_names) %>% 
+summarise(Occurrences = length(Cover_names)) %>% 
+mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
+      Urban = "Airports", #1
+      Discarded = "Bare rocks", #2
+      Natural = "Beaches, dunes, sands", #3
+      Natural = "Broad-leaved forest", #4
+      Natural = "Coastal lagoons", #5
+      Agricultural = "Complex cultivation patterns", #6
+      Natural = "Coniferous forest", #7
+      Urban = "Construction sites", #8
+      Urban = "Continuous urban fabric", #9
+      Urban = "Discontinuous urban fabric", #10
+      Discarded = "Dump sites", #11
+      Discarded = "Estuaries", #12
+      Agricultural = "Fruit trees and berry plantations", #13
+      Urban = "Green urban areas", #14)
+      Urban = "Industrial or commercial units", #15
+      Natural = "Inland marshes", #16
+      Natural = "Intertidal flats", #17
+      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
+      Discarded = "Mineral extraction sites", #19
+      Natural = "Mixed forest", #20
+      Natural = "Moors and heathland", #21
+      Agricultural = "Natural grasslands", #22
+      Agricultural = "Non-irrigated arable land", #23
+      Agricultural = "Pastures", #24
+      Natural = "Peat bogs", #25
+      Urban = "Port areas", #26) 
+      Discarded = "Road and rail networks and associated land", #27
+      Natural = "Salt marshes", #28
+      Discarded = "Sea and ocean", #29
+      Natural = "Sparsely vegetated areas", #30
+      Urban = "Sport and leisure facilities", #31
+      Natural = "Transitional woodland-shrub", #32
+      Agricultural = "Vineyards", #33
+      Discarded = "Water bodies", #34
+      Discarded = "Water courses" #35  
+)) %>% 
+mutate(Cover_names = fct_recode(as.factor(Cover_names),
+      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
+arrange(desc(Occurrences)) %>%  
+kbl() %>% 
+kable_styling(font_size = 12,latex_options = "hold_position")
+
+```

### 2022-11-26 Manuscript/Ms.Rmd
ms

@@
 ---
+
+\captionsetup[table]{labelformat=empty}
 
@@
 
-We used as a measurement of brain size the weight of fixed brains.
+We used as a measurement of brain size the average weight of fixed brains per species.
 To weight the brains, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
@@
 
-We considered as a proxy of body size the intertegular span (ITS) that represents the distance between the base of the wings (tegulae) on the bee thorax.
-ITS has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
+We considered as a proxy of body size the average intertegular span (ITS) per species.
+ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
 All ITS measurements were conducted with a stereomicroscope (magnification 16 x –80 x ) with a calibrated ocular micrometre (resolution down to 0.02 mm).
@@
 
-We assigned a habitat type for all GBIF occurrences by merging the information from land cover inventories with the different georeferenced records.
+We assigned a habitat type for all GBIF occurrences by merging land cover information with the different georeferenced records.
 The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
-After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 to read and obtain the cover classification, respectively.
+After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 to read and obtain the cover classification of the different georeferenced records, respectively.
 Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
@@
 arrange(desc(Occurrences)) %>%  
-kbl() %>% 
-kable_styling(font_size = 12,latex_options = "hold_position")
+kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.}",
+booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
+kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
+row_spec(0,bold=TRUE)
 
@@
 arrange(desc(Occurrences)) %>%  
-kbl() %>% 
-kable_styling(font_size = 12,latex_options = "hold_position")
+kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.}",booktabs = T,
+align = c("l", "c", "l"), linesep = "") %>% 
+kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
+row_spec(0,bold=TRUE)
 

### 2022-11-26 Manuscript/Ms.Rmd
mathods update

@@
 
-We used as a measurement of brain size the average weight of fixed brains per species.
-To weight the brains, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
+Brain size was measured as the weight of fixed brain for each specimen.
+First, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
 Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
@@
 We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
-Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
+Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis® and calculated the average weight for the species with more than one specimen.
 
@@
 The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
-After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 to read and obtain the cover classification of the different georeferenced records, respectively.
+After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 [@REF] to read and obtain the cover classification of the different georeferenced records, respectively.
 Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
@@
 
+## (f) Analysis
+
+### Habitat preference
+
+Habitat preference was estimated by comparing species occurrences per habitat with the ones expected at random.
+For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@REF].
+We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@REF].
+In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
+We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones resulting in percentiles that describe species habitat preference.
+We considered that a bee species exhibited statistical preference when the percentiles were under the 5th percentile (low preference) or above the 95th percentile (high preference).
+
+### Modelling
+
+Habitat preference -Null models -Bipartite
+
+Bayesian:
+
+1st residuals from log brain weight log ITS We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
+(see Wurm & Fisicaro, 2014)
+
+2nd
+
+All analyses were undertaken in R (version 4.0.5) [@REF].
+
 ## (e) Construction of the phylogenetic tree
 
-## (f) Analysis
+We have constructed a phylogenetic tree for our set of species with the help
+
+for the phylogenetic relatedness of the species in our analyses by constructing phylogenetic trees at the genera level. For this,
+
+From MA:
+
+The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
+1, Fig.
+S2, Fig.
+S3).
+Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
+Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6-44, Revell, 2012).
+
+From Liam:
+
+We explored the influence of phylogenetic relatedness in predicting dry weight for bees only because a well-resolved hoverfly phylogeny was not available.
+We constructed an applicable phylogeny for our dataset using a bee genera backbone tree (Hedtke, Patiny, & Danforth, 2013).
+We removed nonrepresented genera using the ape package (version 5.1, Paradis, Claude, & Strimmer, 2004).
+Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6-44, Revell, 2012).
+This excluded a total of three species whose genera weren't included in Hedtke et al.
+(2013)’s phylogeny: Flavipanurgus venustus, Protomeliturga turnerea, and Tetrapedia diversipes.
+As the infrageneric polytomies add an artificial element to the phylogeny, we made the explicit assumption that phylogenetic patterns in body size were assessed at and above the genus level.
 
@@
 This preference index ranges from 0 (avoidance) to 1 (preference).In order to have a representative value of brain weight and body size per species, we used means of the different collected individuals from each species.
-Twenty-four individuals with brain weight values one standard deviation higher or lower than the average were removed (see sup mat analysis) because human errors in their measurement are likely with such delicate tissues.
-Specially, brain weights can give errors due to dehydration or insufficient removal of fat bodies and trachea around the brain.
-For eusocial species, we only used workers.
-We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
-(see Wurm & Fisicaro, 2014) (LM estimate ± SE = 1.88 ± 0.06, p \< 0.001, R2= 0.83, n = 163 species).
-Residuals were calculated using our complete brain weights and body size database.
 

### 2022-11-27 Manuscript/Ms.Rmd
ms update

@@
 The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
-After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.5-21 [@REF] to read and obtain the cover classification of the different georeferenced records, respectively.
+After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
 Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
@@
 
-Habitat preference was estimated by comparing species occurrences per habitat with the ones expected at random.
-For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@REF].
-We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@REF].
+Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
+For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
+We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
 In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones resulting in percentiles that describe species habitat preference.
-We considered that a bee species exhibited statistical preference when the percentiles were under the 5th percentile (low preference) or above the 95th percentile (high preference).
+We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones.
+We considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference).
+
+### Encephalisation
+
+We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
+For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version xxxx where brain weight was the response variable and intertegular span the predictor.
+
+brain weight was the response variable and intertegular span the predictor.
 

### 2022-11-27 Manuscript/Ms.Rmd
methods update

@@
 
+## (e) Construction of the phylogenetic tree
+
+We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013].
+
 ## (f) Analysis
 
-### Habitat preference
+**(i) Habitat preference**
 
@@
 
-### Encephalisation
+**(ii) Encephalisation**
 
 We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
-For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version xxxx where brain weight was the response variable and intertegular span the predictor.
+For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
+To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
+The model was run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017].
+After conducting the posterior predictive checks with the function *pp_check* from *brms*, we extracted the residuals that represented species encephalisation.
 
-brain weight was the response variable and intertegular span the predictor.
-
-### Modelling
+**(iii) Habitat preference ~ encephalisation**
 

### 2022-11-28 Manuscript/Ms.Rmd
Methods done (for now!)

@@
 
+All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
+
 ```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
@@
 In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between female and male individuals [@shreeves2008] that often correlate with the size and structure of the brain [@streinzer2013; @roselino2015].
-In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
+We only considered female specimens, because females are involved in a larger number of tasks than males and could experience greater selective pressures from the environment.
+In addition, these differences between female and male individuals often correlate with the size and structure of the brain [@streinzer2013; @roselino2015].
+Because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
 Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
@@
 ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
-All ITS measurements were conducted with a stereomicroscope (magnification 16 x –80 x ) with a calibrated ocular micrometre (resolution down to 0.02 mm).
+All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
 
@@
 
+#Load now dataset with full set of species
+all_spp = readRDS("../Data/Processing/wit.mean.rds")
+spp_number_nofilter = nrow(all_spp) #total number of species
+
 ```
@@
 To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006.
-After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for North America (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).
+After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species (N = 85).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
 A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
@@
 We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013].
+For this, we first read the backbone phylogeny with the function *read.tree* and drop the missing genera with function *drop.tip* from the package *ape* version 5.6-2 [@paradis2019].
+Then, we selected as an outgroup a wasp genera (*Tachysphex*) and recalculated distances to the root with the function *chronos* from *ape*.
+Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package version 1.2-0 [@revell2012].
+From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
 
 ## (f) Analysis
-
-**(i) Habitat preference**
 
@@
 
-**(ii) Encephalisation**
-
 We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
@@
 To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
-The model was run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017].
-After conducting the posterior predictive checks with the function *pp_check* from *brms*, we extracted the residuals that represented species encephalisation.
+Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
+After evaluating model fit, we extracted the residuals that represented the a quantitative measurement of encephalisation.
 
-**(iii) Habitat preference ~ encephalisation**
+To evaluate if the degree of encephalisation changes by habitat type, we modeled with Bayesian generalised linear models their association.
+For this, we first joined the resulting habitat preference and encephalisation datasets by species and extracted those with statistical habitat avoidance or preference on any of these three habitats.
+Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and encephalisation the predictor.
+Again, we included the phylogenetic covariance matrix as random factor.
+Moreover, we also explored how habitat preference changed by brain weight and intertegular span independently.
+For this, we conducted two analogous models with these two different predictors.
+Finally, we also investigated the different trends for United States and Europe separately.
 
-Habitat preference -Null models -Bipartite
-
-Bayesian:
-
-1st residuals from log brain weight log ITS We estimated relative brain size using residuals from a log-log correlation of brain weight against body size to remove body size allometric relationships.
-(see Wurm & Fisicaro, 2014)
-
-2nd
-
-All analyses were undertaken in R (version 4.0.5) [@REF].
-
-## (e) Construction of the phylogenetic tree
-
-We have constructed a phylogenetic tree for our set of species with the help
-
-for the phylogenetic relatedness of the species in our analyses by constructing phylogenetic trees at the genera level. For this,
-
-From MA:
-
-The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
-1, Fig.
-S2, Fig.
-S3).
-Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
-Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6-44, Revell, 2012).
-
-From Liam:
-
-We explored the influence of phylogenetic relatedness in predicting dry weight for bees only because a well-resolved hoverfly phylogeny was not available.
-We constructed an applicable phylogeny for our dataset using a bee genera backbone tree (Hedtke, Patiny, & Danforth, 2013).
-We removed nonrepresented genera using the ape package (version 5.1, Paradis, Claude, & Strimmer, 2004).
-Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6-44, Revell, 2012).
-This excluded a total of three species whose genera weren't included in Hedtke et al.
-(2013)’s phylogeny: Flavipanurgus venustus, Protomeliturga turnerea, and Tetrapedia diversipes.
-As the infrageneric polytomies add an artificial element to the phylogeny, we made the explicit assumption that phylogenetic patterns in body size were assessed at and above the genus level.
-
-\clearpage
-
-**Data gathering**
-
-Habitat preferences for USA bee data were extracted from Collado et al., 2019, were habitat preference for forests and urban areas was calculated comparing species occurrence in each habitat to what would be expected if species occur at random using null models.
-This preference index ranges from 0 (avoidance) to 1 (preference).In order to have a representative value of brain weight and body size per species, we used means of the different collected individuals from each species.
-
-**Data analysis**
-
-We used Bayesian phylogenetic generalised linear mixed models (BPGMM), as implemented in the package brms [@burkner2017], to assess whether brain sizes, habitat preference and population trends are correlated.
-In all models, we incorporated a phylogenetic covariance matrix.
-The phylogeny used were maximum-likelihood phylogenetic trees of the superfamily Apoidea at the genera level modified from (Hedtke et al., 2013, Fig.
-1, Fig.
-S2, Fig.
-S3).
-Due to the absence of infrageneric phylogenies for all our genera, we simulated infrageneric polytomies within our phylogeny.
-Species tips were added to the phylogenetic tree genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package (version 0.6‐44, Revell, 2012).
-Because we did not have information on our 3 variables for the 44 species (habitat preference, brain weight and population trends), we used 3 different subsets of our dataset to maximize the sample size in the different models: Relative brain size-habitat preference, population trends-habitat preference and population trends-relative brain size.
+All our models were run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017].
+Finally, the different posterior predictive checks were conducted with the function *pp_check* also from the *brms* package.
 

### 2022-12-01 Manuscript/Ms.Rmd
Random changes in methods

@@
 In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens, because females are involved in a larger number of tasks than males and could experience greater selective pressures from the environment.
-In addition, these differences between female and male individuals often correlate with the size and structure of the brain [@streinzer2013; @roselino2015].
-Because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
+We only considered female specimens because: (i) females are involved in a larger number of tasks than males and are likely to experience greater selective pressures from the environment; and (ii), there are evident behavioral and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
+In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
 Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
@@
 Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
+We selected the boundaries from the maps packages *spData* version 2.2.0 [@spdata] and *giscoR* vesion 0.3.2 [@giscoR], and extracted the occurrences with the help of the function *st_intersection* from the package *sf* version 1.0-7 [@pebesma2018].
 For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$.
@@
 We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones.
-We considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference).
+Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference) in comparison with the simulated ones.
 
@@
 Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
-After evaluating model fit, we extracted the residuals that represented the a quantitative measurement of encephalisation.
+Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation.
+High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
 

### 2022-12-02 Manuscript/Ms.Rmd
edit

@@
 
+We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
+For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
+To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
+Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
+Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation.
+High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
+
 Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
@@
 Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference) in comparison with the simulated ones.
-
-We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
-For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
-To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
-Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
-Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation.
-High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
 
@@
     geom_tippoint(aes(size = residuals, color=Country))+
-    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Residuals")+
+    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Encephalisation")+
     scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),

### 2022-12-02 Scripts/0_Processing/0_Preprocessing.R

@@ -125,6 +125,43 @@ median(d$individuals)
 sum(d$individuals) #433
 nlevels(factor(unique(d$Species))) #113 species
 
+#Check for possible outliers
+#1st Brain weight
+#Extract levels
+l = levels(factor(all_cleaning$Species))
+#Create function to plot everything
+s = function(v) {
+ggplot(all_cleaning %>% filter(Species %in% v), aes(Species, Brain.weight)) +
+geom_boxplot() +
+theme(axis.text.x = element_text(angle = 70, vjust = 1, hjust=1, size=4))
+}
+#Plot 1st 50 records
+s(l[1:50])
+s(l[50:100])
+s(l[100:113])
+
+#Is quite easy to have outliers with brain weights (Weight them is highly ticky and values are ver sensitive to mistakes)
+#For this, we select only values between 
+all_cleaning = all_cleaning %>%
+group_by(Species) %>% 
+mutate(Max_val = quantile(Brain.weight, 0.75) + 1.5 *IQR(Brain.weight)) %>% 
+mutate(Min_val = quantile(Brain.weight, 0.25) -  1.5 * IQR(Brain.weight)) %>% 
+filter(Brain.weight <= Max_val) %>% 
+filter(Brain.weight >= Min_val) %>% 
+ungroup()
+
+s(l[1:50])
+s(l[50:100])
+s(l[100:113])
+
+#Just for safety do the same with IT's
+all_cleaning = all_cleaning %>%
+group_by(Species) %>% 
+mutate(Max_val = quantile(IT, 0.75) + 1.5 *IQR(IT)) %>% 
+mutate(Min_val = quantile(IT, 0.25) -  1.5 * IQR(IT)) %>% 
+filter(IT <= Max_val) %>% 
+filter(IT >= Min_val) %>% 
+ungroup()
 
 #Create dataframe with average measurements of IT and brain weight
 weights.mean <- data.frame(aggregate(Brain.weight ~ Species, data = all_cleaning, FUN = mean))

### 2022-12-02 Scripts/0_Processing/1_Phylo_for_residuals.R

@@ -19,7 +19,7 @@ species = wit.mean %>%
     pull()
 
 #Pick tree 1
-bee.trees=read.tree(file="Data/phylogeny_genus_level.txt")
+bee.trees=read.tree(file="Data/Processing/phylogeny_genus_level.txt")
 bee.mcmc=bee.trees[[1]]
 #Make a wasp the outgroup
 bee.mcmc=root(bee.mcmc,outgroup="Tachysphex")

### 2022-12-02 Scripts/0_Processing/2_Calculate_residuals.R

@@ -14,7 +14,7 @@ mutate(log_it = log(IT))
 ggplot(wit.mean, aes(x= IT, y = Brain.weight)) + geom_point()
 ggplot(wit.mean, aes(x= log_it, y = log_brain_weight)) + geom_point()
 
-#Read with phylogenetic distance to add in the modelling process
+#Read with phylogenetic distance to add it to the modelling process
 phylo =readRDS("Data/Processing/phylo_raw.rds")
 
 #Add under score to unify species names
@@ -27,6 +27,15 @@ library(brms)
 m1 <-  brm(log_brain_weight ~ log_it + (1|gr(Species, cov = A)) , data = wit.mean,
                  data2 = list(A = phylo),family=gaussian(), iter=4000, 
                  warmup = 1000)
+
+
+bayes_R2(m1)
+
+ce1 <- conditional_effects(m1, effects = "log_it", points=T) 
+
+#Save data
+write_csv(ce1[[1]] , "Data/Processing/residuals_model.csv")
+write_csv(wit.mean , "Data/Processing/residuals_data.csv")
 
 #Model diagnostics
 pp_check(m1)

### 2022-12-02 Scripts/0_Processing/3_Plot_encephalisation.R

@@ -1,0 +1,32 @@
+
+
+
+#Plot encephalization
+#Load libraries
+library(stringr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+
+#Read data
+m = read_csv("Data/Processing/residuals_model.csv")
+r = read_csv("Data/Processing/residuals_data.csv")
+d = read_csv("Data/Processing/taxonomic_rank.csv") %>% 
+mutate(Species = str_replace_all(query, " ", "_")) %>% 
+select(!c(query, db))
+
+#Add taxonomic rank to dataset
+r = left_join(r,d)
+
+#Add ggplot
+ggplot(m, aes(x = log_it, y = estimate__)) +
+geom_line() +
+geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily)) + 
+theme_bw() +
+ylab("log(Brain size)") +
+xlab("log(Body size)") #+
+#theme(legend.box.background = element_rect(colour = "black"))
+
+
+
+

### 2022-12-02 Scripts/0_Processing/Taxonomic_rank.R

@@ -14,7 +14,7 @@ all_cleaning = read_csv("Raw_data/all_cleaning.csv")
 specieslist = unique(all_cleaning$Species)
 
 #Download data
-d <- tax_name(query = c(specieslist), get = c("order", "family", "genus"), db = "ncbi")
+d <- tax_name(query = c(specieslist), get = c("order", "family", "subfamily", "genus"), db = "ncbi")
 
 #Missing values
 d_missing = d %>% 
@@ -35,7 +35,15 @@ mutate(family = case_when(query=="Osmia lateralis" ~ "Megachilidae",
                           query=="Eucera rufa" ~ "Apidae",
                           query=="Chelostoma philadephi" ~ "Megachilidae",
                           query=="Lasioglossum dialictus" ~ "Halictidae")) %>% 
-mutate(genus = word(query, 1))
+mutate(genus = word(query, 1)) %>% 
+mutate(subfamily = case_when(query=="Osmia lateralis" ~ "Megachilinae",
+                          query=="Panurgus venustus" ~ "Panurginae",
+                          query=="Andrena varians" ~ "Andreninae",
+                          query=="Andrena rhyssonota" ~ "Andreninae",
+                          query=="Anthophora dispar" ~ "Apinae",
+                          query=="Eucera rufa" ~ "Apinae",
+                          query=="Chelostoma philadephi" ~ "Megachilinae",
+                          query=="Lasioglossum dialictus" ~ "Halictinae"))
 
 #Bind rows again
 all = bind_rows(d_non_missing, d_missing)
@@ -46,5 +54,6 @@ write_csv(all, "Data/Processing/taxonomic_rank.csv")
 
 nlevels(factor(unique(all$family))) #6
 nlevels(factor(unique(all$genus))) #30
-                    
+nlevels(factor(unique(all$subfamily))) #11
+  
  

### 2022-12-02 Manuscript/Ms.Rmd
ms

@@
 ```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S1.}", fig.pos="p"}
+
+#Plot encephalization
+#Load libraries
+library(stringr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+
+#Read data
+m = read_csv("../Data/Processing/residuals_model.csv")
+r = read_csv("../Data/Processing/residuals_data.csv")
+d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
+mutate(Species = str_replace_all(query, " ", "_")) %>% 
+select(!c(query, db))
+
+#Add taxonomic rank to dataset
+r = left_join(r,d)
+
+#Add ggplot
+ggplot(m, aes(x = log_it, y = estimate__)) +
+geom_line(size=1) +
+geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
+theme_bw() +
+ylab("log(Brain size)") +
+xlab("log(Body size)") +
+theme(legend.box.background = element_rect(colour = "black"))
+
+```

### 2022-12-05 Manuscript/Ms.Rmd
edits

--- 2022-12-02 Manuscript/Ms.Rmd
@@
 
-Keywords: XXXX
+Keywords: encephalisation, brain-body mass ratio, habitat preference, Apoidea, GBIF
 
@@
 ylab("Habitat preference") +
-xlab("Residuals") +
+xlab("Encephalisation") +
 scale_colour_viridis_d() +
@@
 ```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.}", fig.pos="p", out.width="80%"}
+
+#Plot encephalization
+#Load libraries
+library(stringr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(elementalist) #devtools::install_github("teunbrand/elementalist")
+
+#Read data
+m = read_csv("../Data/Processing/residuals_model.csv")
+r = read_csv("../Data/Processing/residuals_data.csv")
+d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
+mutate(Species = str_replace_all(query, " ", "_")) %>% 
+select(!c(query, db))
+
+#Add taxonomic rank to dataset
+r = left_join(r,d)
+
+#Select just species that fulfilled our criteria of filtering
+#Load preferences
+preferences_europe = read_csv("../Data/Europe_data/preferences_europe_qualitative.csv") 
+preferences_usa = read_csv("../Data/Usa_data/preferences_usa_qualitative.csv")
+
+sp = c(preferences_europe$Species, preferences_usa$Species)
+
+r = r %>% 
+mutate(Species = str_replace_all(Species, "_", " ")) %>% 
+filter(Species %in% sp)
+
+#Add ggplot
+ggplot(m, aes(x = log_it, y = estimate__)) +
+geom_line(size=1) +
+geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
+theme_bw() +
+ylab("log(Brain size)") +
+xlab("log(Body size)") +
+theme(legend.position = c(.95, .60),
+legend.justification = c("right", "top"),
+legend.box.just = "right",
+legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
+legend.key = element_rect(fill = "grey85")) +
+guides(colour=guide_legend(title="Subfamily")) +
+theme(axis.title = element_text(face = "bold"))
+
+```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
+
+plot_all = readRDS("../Data/Processing/plot_all.rds")
+
+plot_all
+
+```

### 2022-12-05 Scripts/0_Processing/0_Preprocessing.R

@@ -117,7 +117,6 @@ write_csv(all_cleaning, "Raw_data/all_cleaning.csv")
 
 
 #Check unique cases and levels for methods
-
 d = all_cleaning %>% 
 group_by(Species) %>% 
 summarise(individuals = n())
@@ -155,13 +154,13 @@ s(l[50:100])
 s(l[100:113])
 
 #Just for safety do the same with IT's
-all_cleaning = all_cleaning %>%
-group_by(Species) %>% 
-mutate(Max_val = quantile(IT, 0.75) + 1.5 *IQR(IT)) %>% 
-mutate(Min_val = quantile(IT, 0.25) -  1.5 * IQR(IT)) %>% 
-filter(IT <= Max_val) %>% 
-filter(IT >= Min_val) %>% 
-ungroup()
+#all_cleaning = all_cleaning %>%
+#group_by(Species) %>% 
+#mutate(Max_val = quantile(IT, 0.75) + 1.5 *IQR(IT)) %>% 
+#mutate(Min_val = quantile(IT, 0.25) -  1.5 * IQR(IT)) %>% 
+#filter(IT <= Max_val) %>% 
+#filter(IT >= Min_val) %>% 
+#ungroup()
 
 #Create dataframe with average measurements of IT and brain weight
 weights.mean <- data.frame(aggregate(Brain.weight ~ Species, data = all_cleaning, FUN = mean))

### 2022-12-05 Scripts/0_Processing/3_Plot_encephalisation.R

@@ -1,32 +1,51 @@
 
+#In this script we plot encephalisation 
+#(log(brain)~log(body))
+#Although more species were used
+#I am going to show only the ones from
+#our final analyses (after filtering)
 
 
-#Plot encephalization
-#Load libraries
+#Load libraries----
 library(stringr)
 library(readr)
 library(ggplot2)
 library(dplyr)
+library(elementalist) #devtools::install_github("teunbrand/elementalist")
 
-#Read data
+#Read data----
 m = read_csv("Data/Processing/residuals_model.csv")
 r = read_csv("Data/Processing/residuals_data.csv")
 d = read_csv("Data/Processing/taxonomic_rank.csv") %>% 
 mutate(Species = str_replace_all(query, " ", "_")) %>% 
 select(!c(query, db))
-
 #Add taxonomic rank to dataset
 r = left_join(r,d)
 
-#Add ggplot
+#Select just species that fulfilled our criteria of filtering
+#Load preferences
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative.csv") 
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative.csv") 
+
+sp = c(preferences_europe$Species, preferences_usa$Species)
+
+r = r %>% 
+mutate(Species = str_replace_all(Species, "_", " ")) %>% 
+filter(Species %in% sp)
+
+
+#Plot----
 ggplot(m, aes(x = log_it, y = estimate__)) +
 geom_line() +
 geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily)) + 
 theme_bw() +
 ylab("log(Brain size)") +
-xlab("log(Body size)") #+
-#theme(legend.box.background = element_rect(colour = "black"))
-
+xlab("log(Body size)") +
+theme(legend.position = c(.95, .75),
+legend.justification = c("right", "top"),
+legend.box.just = "right",
+legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey"),
+legend.key = element_rect(fill = "grey"))
 
 
 

### 2022-12-05 Scripts/0_Processing/Plot_ecephalisation_subfamily.R

@@ -1,0 +1,38 @@
+
+
+#Load brain data
+brain_weight = read_csv("Data/Processing/brain_weight_data.csv")
+#Load preferences
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative.csv") 
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative.csv") 
+
+#Extract sepcecies from preferences
+sp = c(preferences_europe$Species, preferences_usa$Species)
+#Select final species from the trait datatset
+brain_weight = brain_weight %>% 
+filter(Species %in% sp)
+
+#Add subfamily to species dataset
+d = read_csv("Data/Processing/taxonomic_rank.csv") %>% 
+mutate(Species = query) %>% 
+select(!c(query, db))
+#Add taxonomic rank to dataset
+brain_weight = left_join(brain_weight,d)
+
+brain_subfamily = brain_weight %>% 
+group_by(subfamily) %>% 
+summarise(mean_sub = mean(residuals),
+          sd_sub = sd(residuals),
+          n_lev = n()) %>% 
+mutate(se_sub = sd_sub / sqrt(n_lev),
+         lower_ci = mean_sub - qt(1 - (0.05 / 2), n_lev - 1) * se_sub,
+         upper_ci = mean_sub + qt(1 - (0.05 / 2), n_lev - 1) * se_sub)
+
+
+ggplot(brain_subfamily, aes(mean_sub, subfamily)) +
+geom_point() +
+geom_errorbar(aes(xmin = lower_ci,
+xmax = upper_ci))
+  
+  
+  

### 2022-12-05 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -285,7 +285,7 @@ legend2 <- get_legend(
 # now add the title
 title2 <- ggdraw() + 
   draw_label(
-paste("United States (N =", format(nrow(usa), big.mark = ",") ,")"),    fontface = 'bold',
+paste("United States occurrences (N =", format(nrow(usa), big.mark = ",") ,")"),    fontface = 'bold',
     x = 0,
     hjust = 0
   ) +
@@ -302,4 +302,6 @@ panel2 = plot_grid(
   rel_heights = c(0.1, 0.3)
 )
 
-plot_grid(panel1, panel2, legend2, ncol=1, rel_heights = c(1, 1, 0.4))
+plot_all = plot_grid(panel2, panel1, legend2, ncol=1, rel_heights = c(1, 1, 0.4))
+
+saveRDS(plot_all, "Data/Processing/plot_all.rds")

### 2022-12-06 Manuscript/Ms.Rmd
analyses update

@@
 
-labels.minor <- c("Low", "", "High")
+labels.minor <- c("Avoidance", "", "Preference")
 
@@
     strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"))+
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5),
+    scale_x_continuous(breaks = c(0.2,0.5,0.8),
                        labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
     scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
-    geom_vline(aes(xintercept = 0.05), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.95), colour="black", linetype="dashed")
+    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
+    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed")
 

### 2022-12-06 Scripts/0_Processing/0_Preprocessing.R

@@ -1,172 +1,0 @@
-#Preprocessing----
-
-#In this script we fix typos and rename species when neccesary to their synonyms
-
-#Load libraries
-library(readr)
-library(dplyr)
-library(stringr)
-
-#First we clean USA records again (there were some wrong measurements in the old dataset)
-#Read both datasets
-#read usa
-usa_brain = read_csv("Raw_data/DB_brain_US.csv") %>% 
-  dplyr::select(!c(head.weight.g, no.optic.lobes.weight.mg)) %>% 
-  mutate(species = str_replace(species, "_", " "))
-#read raw dataset sent by Collado
-brains.it <- read.csv("Raw_data/brains.it.csv", dec = ",") %>% 
-  dplyr::select(!c(Place, No.optic.lobes.weight, 
-            Notes, Head.weight..g., Plant, Date)) %>% 
-  rename(Body.weight = Weigth..g.) %>% 
-  rename(Brain.weight = Brain.Weight..mg.)
-
-#Filter out USA brains from sayol, these ones are not corrected!
-#brains.it = brains.it %>% 
-#filter(!Country == "USA")
-
-spp = usa_brain$species
-#Do it just for USA spp 
-brains.it.usa = brains.it %>% filter(Country=="USA")
-brains.it.eur = brains.it %>% filter(!Country=="USA")
-
-#Filter out Sayol species
-brains.it.usa = brains.it.usa %>% 
-filter(!Species %in% spp)
-
-#Bind again datasets
-brains.it = bind_rows(brains.it.eur, brains.it.usa)
-
-#Prepare data----
-#Important:
-#The column of interest is brain fix (mg)
-#Heads are extracted and fixed and then the brain is removed and weighted
-#Fresh brains can be converted with a conversion factor to fix! 
-#brain.fix.mg = brain.fresh.mg x 1.2262. #Check sayol etal
-#Convert fresh to fix when value is na (I guess that almost always when fresh is measured)
-#The trend is the same but looks a bit nicer without adding the brains with this conversion
-
-usa_brain = usa_brain %>% 
-#mutate(brain.fix.mg = case_when(is.na(brain.fix.mg) == T ~ brain.fresh.mg * 1.2262,
-#       is.na(brain.fix.mg) == F ~ brain.fix.mg)) %>% 
-rename(Genus = genus) %>% 
[…]

### 2022-12-06 Scripts/0_Processing/0_Preprocessing_new.R

@@ -1,5 +1,7 @@
 #Preprocessing----
 #In this script we fix typos and rename species when neccesary to their synonyms
+#The data source is Sayol et al., but we also have added extra species from Collado thesis
+#Sayol http://dx.doi.org/10.1098/rspb.2020.0762
 
 #Load libraries----
 library(readr)
@@ -8,12 +10,12 @@ library(stringr)
 library(ggplot2)
 
 #Read datasets
-#1st Sayol http://dx.doi.org/10.1098/rspb.2020.0762
+#1st Sayol 
 sayol = read_csv("Raw_data/BeeTraits_individual.csv") %>% 
 dplyr::select(!c(head.weight.g, no.optic.lobes.weight.mg,body.mass.g)) %>% 
 mutate(species = str_replace(species, "_", " "))
 #snd Collado (is similar but Collado has some extra species)
-#Note Sayol data is clean but Collado still has typos and errors in measurements
+#Note Sayol data is clean but Collado still needs some processing
 collado <- read.csv("Raw_data/brains.it.csv", dec = ",") %>% 
 dplyr::select(!c(Place, No.optic.lobes.weight, 
 Notes, Head.weight..g., Plant, Date, Weigth..g., ID, Genus)) %>% 
@@ -48,8 +50,7 @@ spp = s$Species
 #Subset collado
 collado_filtered = collado %>% 
 filter(!Species == "Dasypoda visnaga") %>%  #Dasypoda visnaga's brain was conserved in ethanol, not formol
-filter(!Species %in% spp)
-
+filter(!Species %in% spp) 
 #Bind both datasets
 all = bind_rows(sayol_main, collado_filtered) %>% 
 na.omit() #filter out na's
@@ -108,29 +109,19 @@ s(l[1:50])
 s(l[50:100])
 s(l[100:130])
 
-#Is quite easy to have outliers with brain weights (Weight them is highly ticky and values are ver sensitive to mistakes)
-#For this, we select only values between 
+#Is quite easy to have outliers with brain weights 
+#For this, we filter outliers
 all_cleaning = all_cleaning %>%
 group_by(Species) %>% 
 mutate(Max_val = quantile(Brain.weight, 0.75) + 1.5 *IQR(Brain.weight)) %>% 
 mutate(Min_val = quantile(Brain.weight, 0.25) -  1.5 * IQR(Brain.weight)) %>% 
[…]

### 2022-12-06 Scripts/5_Preferences/Preferences_Europe.R

@@ -126,9 +126,9 @@ write_csv(m1, "Data/Europe_data/preferences_europe.csv")
 
 #Prepare qualitative data for analysis
 #Convert to qualitative
-m[m >= 0.95] <- 1
-m[m <= 0.05] <- 0
-m[m > 0.05 & m < 0.95 ] <- NA
+m[m >= 0.80] <- 1
+m[m <= 0.20] <- 0
+m[m > 0.20 & m < 0.80 ] <- NA
 
 m= rownames_to_column(m, var = "Species")
 

### 2022-12-06 Scripts/5_Preferences/Preferences_USA.R

@@ -110,11 +110,11 @@ write_csv(m1, "Data/Usa_data/preferences_usa.csv")
 
 #Prepare qualitative data for analysis
 #Convert to qualitative
-m[m >= 0.95] <- 1
-m[m <= 0.05] <- 0
-m[m > 0.05 & m < 0.95 ] <- NA
+m[m >= 0.80] <- 1
+m[m <= 0.20] <- 0
+m[m > 0.20 & m < 0.80 ] <- NA
 
-m= rownames_to_column(m, var = "Species")
+m = rownames_to_column(m, var = "Species")
 
 #Save data
 write_csv(m, "Data/Usa_data/preferences_usa_qualitative.csv")

### 2022-12-06 Scripts/7_Analyses/Analysis_all.R

@@ -8,14 +8,13 @@ library(brms)
 #Load brain data
 brain_weight = read_csv("Data/Processing/brain_weight_data.csv")
 #Load preferences
-preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative.csv") 
-preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative.csv") 
+preferences_europe = read_csv("Data/Europe_data/preferences_europe_qualitative.csv") %>% 
+mutate(country = "europe")
+preferences_usa = read_csv("Data/Usa_data/preferences_usa_qualitative.csv") %>% 
+mutate(country = "usa")
 
 #There are duplicated species between usa and europe
-#Here is the list:
-f = c("Andrena wilkella", "Anthidium manicatum","Anthidium oblongatum",    
-      "Halictus confusus", "Halictus rubicundus", "Lasioglossum leucozonium",
-      "Megachile rotundata")
+f <- intersect(preferences_europe$Species, preferences_usa$Species)
 #Filter them out from usa and just select europe which had higher number of records
 preferences_usa = preferences_usa %>% 
 filter(!Species %in% f)
@@ -38,10 +37,11 @@ long_data = long_data %>%
     mutate(brain_it = Brain.weight/IT)
 
 #Analysis Preference ~ residuals----
-model1 = brm(Preference ~ residuals * Habitat + (1|gr(Species, cov = A)), 
+model1 = brm(Preference ~ residuals * Habitat +  (1|gr(Species, cov = A)), 
              data = long_data, data2 = list(A = A10), family=bernoulli())
 
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat", points=T) 
+pp_check(model1)
 
 bayes_R2(model1)
 

### 2022-12-06 Manuscript/Ms.Rmd
figure 3 update

@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=4, fig.width=10, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
 The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
+
+#Histogram of habitat preferences
+#Load libraries
+library(tidyverse)
 
@@
 
-labels.minor <- c("Avoidance", "", "Preference")
+labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
 
-long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),
-                  labels = c("A) Natural", "B) Agricultural", "C) Urban"))
+long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
 
-#Plot
-ggplot(long_data, aes(x=Preference, color= Habitat)) +
-    geom_histogram(aes(fill= Habitat)) + 
+long_data = long_data %>% 
+mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
+
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
+
+ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
+    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
     xlab("Habitat preference") +
@@
     facet_wrap(~ Habitat) +
-    scale_colour_viridis_d(end=0.98)+
-    scale_fill_viridis_d(end=0.98) +
+    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
     theme_bw() +   
-    theme(axis.title = element_text(face = "bold"),
+    theme(axis.title = element_text(face = "bold", size = 14),
     panel.border=element_rect(size=1),
-    legend.position = "none",
     strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
-    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"))+
-    scale_x_continuous(breaks = c(0.2,0.5,0.8),
-                       labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
+    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
+    panel.grid.minor=element_blank())+
     scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
     geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed")
+    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
+    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
+                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
+  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="80%"}
 

### 2022-12-06 Manuscript/trial.Rmd
figure 3 update

--- 2022-12-06 Manuscript/Ms.Rmd
@@
 ---
-title: "Bee brain size is positively associated with urban lifestyles but not with positive population trends"
+title: "Untitled"
 output:
-  pdf_document:
-    fig_caption: yes
-
-link-citations: yes
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
-fontsize: 12pt
-
-bibliography: 
-  - references.bib
-
-header-includes:
-- \usepackage{orcidlink}
-- \usepackage{caption,setspace}
-- \captionsetup[figure]{labelformat=empty}
-- \usepackage{float} 
-- \floatplacement{figure}{H} 
+  pdf_document: default
+  html_document: default
+date: "2022-12-06"
 ---
-
-\captionsetup[table]{labelformat=empty}
-
-```{r echo=FALSE,warning=FALSE}
-
-knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
-```
-
-\textbf{Miguel Á. Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B. Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}
-
-Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
-
-Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
-
-\textbf{Author for correspondence:}
-
-# Abstract
-
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role in ecosystems for fruit and seed production.
-However, not all pollinators are negatively affected by habitat conversion, but some find in anthropogenic landscapes appropriate resources to survive and reproduce.
-Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities.
-While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for 145 individuals from 44 species of bees ---the most important group of pollinators--- from North America and Europe, and compared them with detailed information on habitat preferences and population trends.
-Our analyses revealed that bees that prefer forest habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger brains compared to forest dwellers.
-However, we found no evidence that either larger brains or a preference for urban habitats are associated with positive population trends.
-These results suggest that a large brain can help maintaining bee urban populations, but also highlight that being tolerant to urbanisation is not enough to compensate for the worldwide decline of their populations.
-
-Keywords: encephalisation, brain-body mass ratio, habitat preference, Apoidea, GBIF
-
-\clearpage
-
-# Introduction
-
-Relevant refs:
-
-<https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R>
-
-This may be also relevant:
-
-Although recent studies challenge the use of absolute brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein), many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008).
-
-Pollination is a fundamental ecosystem service that connects the earth's vegetation and human economy.
-Pollination services are performed by both managed and wild pollinators, mostly bees [@james2008; @winfree2011], both of which might be affected by a range of current and projected human-related environmental changes.
-There are indeed clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019], and parallel declines in the plants that rely upon them [@potts2010; @scheper2014].
-For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees [@williams2009].
-The decline of pollinators currently observed is projected to increase in the future [@dormann2008].
-Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production [@klein2007; @ollerton2011], there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services [@ashman2004; @klein2007].
-
-Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@brown2009].
-The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth [@brown2009].
-However, not all pollinators species are equally affected by habitat loss and alteration.
-In bees, for example, there are also "winners" ---like Bombus terrestris [@rasmont2008]--- that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].
-
-Human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown.
-Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive [@manenti2013].
-For species that tolerate human-related disturbances, human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
-Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
-In vertebrates, relative brain size is a good proxy of problem-solving and learning (e.g.
-see chapter 2 of this thesis for its relationship with learning), which can help individuals to cope with the many challenges on novel environments.
-However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains.
-More important, there was no unequivocal evidence that brain size was related to cognitive performance.
-However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
-There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2).
-We used previously published data on bee forest and urban preference [@collado2019] as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
-Species population trends were extracted from @bartomeus2013 and both datasets were crossed with a newly collected brain weight database.
-We wanted to explore if there is a relationship between brain size, habitat preference, and population trends.
-Hence, our questions were the following:Are brain sizes, as a proxy of behavioural plasticity, related to habitat preferences?Is population stability related to brain size as it has been seen in other taxa?Are habitat preferences related to population stability?
-Or alternatively, are urban dwellers populations increasing?
-
-\clearpage
-
-# Methods
-
-## Overview
-
-Our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas.
-To evaluate this hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size.
-For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and selected as study area the states or countries with highest density of records.
-After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models.
-Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.
-
-All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
-
-```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-library(readr)
-library(dplyr)
-
-#Load final dataset 
-raw = read_csv("../Raw_data/all_cleaning.csv")
-
-d = raw %>% 
-group_by(Species) %>% 
-summarise(individuals = n())
-
-#Load taxonomic rank script
-tax = read_csv("../Data/Processing/taxonomic_rank.csv")
-
-```
-
-## (a) Bee species dataset
-
-Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
-In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because: (i) females are involved in a larger number of tasks than males and are likely to experience greater selective pressures from the environment; and (ii), there are evident behavioral and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
-In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
-Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
-
-## (b) Brain and body size measurements
-
-Brain size was measured as the weight of fixed brain for each specimen.
-First, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
-Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
-Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies.
-By doing this, we ensured that our brain measurements mainly accounted for neural tissue.
-We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
-Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis® and calculated the average weight for the species with more than one specimen.
-
-We considered as a proxy of body size the average intertegular span (ITS) per species.
-ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
-All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
-
-## (c) Occurrence data
-
-```{r Loading study site areas, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Calculate study site areas
-#Load libraries
-library(sf)
-library(sfheaders)
-
-#EUROPE:
-#Read data
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Add area col
-euro_map$area <- st_area(euro_map)
-#Create dataframe with the correct units
-d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Calculate area
-eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")
-
-#USA:
-#Read data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-#Not an sf object, convert into multypolygon
-usa = sfheaders::sf_multipolygon(
-  obj = usa_states_1
-  , multipolygon_id = "group"
-  , x = "long"
-  , y = "lat"
-) %>% st_set_crs(4326)
-#Calculate area
-usa$area <- st_area(usa)
-#Create dataframe with the correct units
-d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Sum areas
-usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")
-
-```
-
-```{r Species and occurrences info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read occurrence records and extract relevant info
-#Load libraries
-library(data.table)
-#Load clean data with bee occurrences
-europe <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-usa <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Europe
-eu_occurrences = prettyNum(nrow(europe), big.mark=",")
-records_eu = europe %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#USA
-usa_occurrences = prettyNum(nrow(usa), big.mark=",")
-records_usa = usa %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#Number of species
-usa_spp = nrow(records_usa)
-eu_spp = nrow(records_eu)
-spp_number = nrow(records_eu) + nrow(records_usa)
-#Median number of records per dataset
-eu_median = median(records_eu$no_rows)
-usa_median = median(records_usa$no_rows)
-
-#Load now dataset with full set of species
-all_spp = readRDS("../Data/Processing/wit.mean.rds")
-spp_number_nofilter = nrow(all_spp) #total number of species
-
-```
-
-We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
-The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
-Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
-We selected the boundaries from the maps packages *spData* version 2.2.0 [@spdata] and *giscoR* vesion 0.3.2 [@giscoR], and extracted the occurrences with the help of the function *st_intersection* from the package *sf* version 1.0-7 [@pebesma2018].
-For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$.
-These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-Similarly, we selected countries located on the north and center of Europe (Figure 1B) that represented a total area of `r eukm2` km$^{2}$.
-These countries were: England, Belgium, Netherlands, Germany and Denmark.
-All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006.
-After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
-
-## (d) Land cover classification
-
-We assigned a habitat type for all GBIF occurrences by merging land cover information with the different georeferenced records.
-The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
-After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
-Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
-The original extracted cover classes and our classification into these three different categories are shown in Table S1 for United States and Table S2 for Europe.
-
-## (e) Construction of the phylogenetic tree
-
-We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013].
-For this, we first read the backbone phylogeny with the function *read.tree* and drop the missing genera with function *drop.tip* from the package *ape* version 5.6-2 [@paradis2019].
-Then, we selected as an outgroup a wasp genera (*Tachysphex*) and recalculated distances to the root with the function *chronos* from *ape*.
-Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package version 1.2-0 [@revell2012].
-From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
-
-## (f) Analysis
-
-We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
-For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
-To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
-Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
-Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation.
-High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
-
-Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
-For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
-We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
-In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones.
-Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference) in comparison with the simulated ones.
-
-To evaluate if the degree of encephalisation changes by habitat type, we modeled with Bayesian generalised linear models their association.
-For this, we first joined the resulting habitat preference and encephalisation datasets by species and extracted those with statistical habitat avoidance or preference on any of these three habitats.
-Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and encephalisation the predictor.
-Again, we included the phylogenetic covariance matrix as random factor.
-Moreover, we also explored how habitat preference changed by brain weight and intertegular span independently.
-For this, we conducted two analogous models with these two different predictors.
-Finally, we also investigated the different trends for United States and Europe separately.
-
-All our models were run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017].
-Finally, the different posterior predictive checks were conducted with the function *pp_check* also from the *brms* package.
-
-# Results
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
-A) Phylogenetic tree at the genera level.
-Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).
-The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted points at the end of the tip branches.
-Larger points indicate larger brains in proportion to their body size and vice versa.
-B) Heatmap showing the habitat preference for each bee species.
-The columns delimit the habitat type (i.e., natural, agricultural, semideveloped and urban) and the rows the different bee species."}
-
-library(ggtree) #v3.5.3.002
-library(ggplot2)
-library(stringr)
-library(tidyverse)
-library(tibble)
-
-#Load phylogenetic data from europe
-bee.tree100 = readRDS("../Data/Europe_USA/bee.tree100.rds")
-#Load brain data
-brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-
-#Average preferences between USA and Europe
-preferences = pref %>% 
-group_by(Species) %>%
-summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-ungroup()
-
-#Delete underscore from tree
-bee.tree100$tip.label = str_replace(bee.tree100$tip.label, "_", " ")
-
-#Generate vector with species
-v = bee.tree100$tip.label
-#Filter species with that are in the phylo
-brain_weight = brain_weight %>% filter(Species %in% v)
-
-#Fill a column based on condition
-brain_weight = brain_weight %>% 
-    mutate(Country = case_when((brain_weight$Species %in% preferences_eu$Species & 
-    brain_weight$Species %in% preferences_usa$Species ~ "USA and Europe"),
-    (brain_weight$Species %in% preferences_usa$Species ~ "USA"),
-    (brain_weight$Species %in% preferences_eu$Species ~ "Europe")))
-
-#Plot tree
-p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country)+
-    geom_tippoint(aes(size = residuals, color=Country))+
-    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Encephalisation")+
-    scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
-                       labels = c("Europe", "USA", "USA and Europe"),
-                       breaks = c("Europe", "USA", "USA and Europe")) +
-  ggtitle("A") + theme(plot.title = element_text(size=18, face="bold"))
-
-#Convert wide format to long for plotting heatmap
-d=gather(preferences, condition, measurement, Agricultural:Urban, factor_key=TRUE)
-
-#Relevel for plotting
-d = d %>% 
-mutate(condition=fct_relevel(condition,c("Natural","Agricultural", "Urban")))
-
-p2 = ggplot(d, aes(x=condition, y=Species)) + 
-    geom_tile(aes(fill=measurement),color = "white") +
-   # scale_fill_gradient(high="black",  low="white") +
-    scale_fill_viridis_c(name = "Habitat\n preference") + 
-    theme_minimal() +
-    xlab(NULL) + 
-    ylab(NULL) + 
-    theme(axis.text.x = element_text(face="bold", color="black", 
-    size=7.5,  vjust=1.15),
-    axis.text.y = element_text(face="italic", color="black", size=7.5, hjust=0),
-    plot.title = element_text(size=18, face="bold")) +
-   ggtitle("B")
-
-#Nice way to insert ordered heatmap
-library(aplot)
-p2 %>% insert_left(p) 
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=4, fig.width=4, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
 The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
@@
 
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Avoidance", "No preference", "Preference"))
 
@@
     theme_bw() +   
-    theme(axis.title = element_text(face = "bold", size = 14),
+    theme(axis.title = element_text(face = "bold"),
     panel.border=element_rect(size=1),
@@
     scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
-                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
-  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
+                       labels = c(0,0.2,0.4,0.6,0.8,1))
 
 ```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.}", fig.pos="p"}
-
-#Load libraries
-library(readr)
-library(dplyr)
-library(ggplot2)
-
-#Load data for plotting
-long_data <- read_csv("../Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
-
-model1 = read_csv("../Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")
-
-model2 = read_csv("../Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")
-
-model3 = read_csv("../Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")
-
-#Relevel for plotting
-long_data = long_data %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model1 = model1 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model2 = model2 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model3 = model3 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-#colorBlindness::displayAllColors(viridis::viridis(4))
-
-all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab("Habitat preference") +
-xlab("Encephalisation") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("A") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-all_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab(NULL) +
-xlab("Brain weight (mg)") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("B") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-all_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab(NULL) +
-xlab("Intertegular distance (mm)") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("C") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-library(patchwork)
-
-all_1 + (all_2/all_3) + plot_layout(guides = "collect") &
-theme(legend.position = 'bottom', panel.border=element_rect(size=1),
-      axis.title=element_text(face="bold"),
-      legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-\clearpage
-
-# Discussion
-
-Bees with urban habitat preferences tended to have bigger brains related to their body sizes, and those with forest preferences had relatively smaller brains.
-Surprisingly, neither brain sizes nor habitat preference explained population trends, against what we expected.
-Therefore, species with bigger brains appearing on urban habitat can be indicating that while only bees with large cognitive capacities are surviving in these modified environments, they do not have experienced notorious increases in their populations.We hypothesized that bees with bigger brains related to body size should be more flexible in their behaviour because a large brain may confer better cognitive abilities, as it has been observed in some species (Lefebvre et al., 2004) and as we show specifically for bees in relation to learning (Chapter 2 of this thesis).
-Therefore, to exploit complex urban environments, we expect bigger brains to be required.
-Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment (Wheater, 1999).
-Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than the forest dwellers counterpart.
-Some examples of behaviours needed to survive in urban environments (see Table S2 for a full set of examples) are the use of human-made materials to nest (Allasino et al., 2019) or being able to collect pollen and nectar from exotic flowers from gardens (Lowenstein et al., 2014).
-We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them (Smith et al., 2019).
-In accordance, we found smaller brain sizes than expected by their body sizes in bees that preferred forests indicating that an investment in bigger relative brain sizes is not required.
-In addition, other traits have already been described to be filtered in anthropogenic land uses, like early spring or brood parasitism (Harrison et al., 2017), which are present in the forest dwellers.
-We could not explain population trends with brain size as it happens in other taxa.
-For example, in birds, brain size is a good predictor of how a population does in the long-term (Shultz, 2005).
-However, this is not always the case as bigger brains increase the risk of extinction in mammals (González-Voyer et al., 2016).
-Despite brain weight is relatively easy to measure and standardize with the used protocol, we lack information on certain neuropils size, like mushroom bodies (the suspected centers of cognitive processes in bees), which can be variable even within species (Barth & Heisenberg, 1997; Withers et al., 1995; Fahrbachand & Robinson 1995).
-Differences in more subtle brain structures could explain differences in behaviours, and in consequence, shed extra light on species population trends.
-Therefore the isolation and measurement of these neuropils could be interesting because we could get more accurate information.
-On the other hand, also note that population trends data is hard to find or measure, especially for invertebrate taxa.
-Long term data is often lacking and we count only with data indirectly taken from museums.
-This data is also scarce and hard to standardize (Bartomeus et al., 2018), therefore not being optimal for comparative studies of declining or increasing populations (Shultz, 2005).
-Species body size is generally related to population trends.
-In mammals, it is known that bigger body sizes decreases population density and increases the risk of extinction (Damuth, 1981; González-Voyer et al., 2016).
-This relation has also been seen in bees (Scheper et al., 2014).
-Although body size is directly related to absolute brain size, we did not find a strong correlation between body size and population trends in our subset of North American bees as it was found for the complete set in Bartomeus et al., 2013.
-Our subset may be biased towards the most common genera and contain fewer species per genus.
-If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel's lambda \> 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
-Population trends or habitat preferences could be explained by the availability of the resources consumed, for example, in the Netherlands, population trends of some bee species are related to the availability of their preferred plants (Scheper et al., 2014).
-A polylectic diet could help species to adapt to new environments, and we know that species able to forage on crop plant families tend to have stable or increasing populations (Scheper et al., 2014, Bartomeus et al.
-2013).
-If we consider habitats being conditioned by their availability of the preferred resources, then we may be able to explain our link between brain size and preference for certain habitats.
-For example, differences in bee community composition and abundance are found for oligolectic (i.e.diet specialist bees) associated to their prefered resources along forested gradients (Grundel et al., 2010).
-Dietary data should be an interesting next step to see if broader (polylectic) or narrower (oligolectic) diet are generally related to habitat preferences and brain sizes.
-In fact, it has been already described for other taxa a relationship between brain size and diet (DeCasien et al., 2017).Finally, the absence of relationship between urban preferences and increasing population trends is interesting.
-Despite these bees have found new niches to occupy, they are not increasing its populations but just persisting.
-The transformation from natural to urban habitats is filtering which species can survive, and hence, biodiversity in urban habitats is usually lower (Collado et al., 2019, Chapter 3 of this thesis).
-However, survivors are not thriving either in this new conditions.
-Some negative effects of urbanization have already been studied in bees, like for example, biased sex ratio, with increasing males in urban areas (Fitch et al., 2019) with unknown future population consequences.
-It should be interesting to fully assess the costs and benefits of living in urban environments, including brain investment and other selected strategies to adapt to the new conditions.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-\clearpage
-
-# Supplementary material
-
-```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
-dplyr::select(species, cover.names) %>% 
-mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-rename(Cover_names = cover.names) %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-       Discarded = "Barren Land", 
-       Agricultural = "Cultivated Crops", 
-       Natural = "Deciduous Forest", 
-       Urban = "Developed, High Intensity", 
-       Urban = "Developed, Medium Intensity",
-       Urban = "Developed, Low Intensity",
-       Urban = "Developed, Open Space", 
-       Natural = "Emergent Herbaceous Wetlands", 
-       Natural = "Evergreen Forest", 
-       Agricultural = "Hay/Pasture", 
-       Natural = "Herbaceous", 
-       Natural = "Mixed Forest", 
-       Discarded = "Open Water", 
-       Natural = "Shrub/Scrub", 
-       Natural = "Woody Wetlands")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.}",
-booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-\clearpage
-
-```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
-    dplyr::select(Species, Cover_names) %>% 
-    mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-      Urban = "Airports", #1
-      Discarded = "Bare rocks", #2
-      Natural = "Beaches, dunes, sands", #3
-      Natural = "Broad-leaved forest", #4
-      Natural = "Coastal lagoons", #5
-      Agricultural = "Complex cultivation patterns", #6
-      Natural = "Coniferous forest", #7
-      Urban = "Construction sites", #8
-      Urban = "Continuous urban fabric", #9
-      Urban = "Discontinuous urban fabric", #10
-      Discarded = "Dump sites", #11
-      Discarded = "Estuaries", #12
-      Agricultural = "Fruit trees and berry plantations", #13
-      Urban = "Green urban areas", #14)
-      Urban = "Industrial or commercial units", #15
-      Natural = "Inland marshes", #16
-      Natural = "Intertidal flats", #17
-      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
-      Discarded = "Mineral extraction sites", #19
-      Natural = "Mixed forest", #20
-      Natural = "Moors and heathland", #21
-      Agricultural = "Natural grasslands", #22
-      Agricultural = "Non-irrigated arable land", #23
-      Agricultural = "Pastures", #24
-      Natural = "Peat bogs", #25
-      Urban = "Port areas", #26) 
-      Discarded = "Road and rail networks and associated land", #27
-      Natural = "Salt marshes", #28
-      Discarded = "Sea and ocean", #29
-      Natural = "Sparsely vegetated areas", #30
-      Urban = "Sport and leisure facilities", #31
-      Natural = "Transitional woodland-shrub", #32
-      Agricultural = "Vineyards", #33
-      Discarded = "Water bodies", #34
-      Discarded = "Water courses" #35  
-)) %>% 
-mutate(Cover_names = fct_recode(as.factor(Cover_names),
-      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.}",booktabs = T,
-align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.}", fig.pos="p", out.width="80%"}
-
-#Plot encephalization
-#Load libraries
-library(stringr)
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(elementalist) #devtools::install_github("teunbrand/elementalist")
-
-#Read data
-m = read_csv("../Data/Processing/residuals_model.csv")
-r = read_csv("../Data/Processing/residuals_data.csv")
-d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
-mutate(Species = str_replace_all(query, " ", "_")) %>% 
-select(!c(query, db))
-
-#Add taxonomic rank to dataset
-r = left_join(r,d)
-
-#Select just species that fulfilled our criteria of filtering
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe_qualitative.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa_qualitative.csv")
-
-sp = c(preferences_europe$Species, preferences_usa$Species)
-
-r = r %>% 
-mutate(Species = str_replace_all(Species, "_", " ")) %>% 
-filter(Species %in% sp)
-
-#Add ggplot
-ggplot(m, aes(x = log_it, y = estimate__)) +
-geom_line(size=1) +
-geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
-theme_bw() +
-ylab("log(Brain size)") +
-xlab("log(Body size)") +
-theme(legend.position = c(.95, .60),
-legend.justification = c("right", "top"),
-legend.box.just = "right",
-legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
-legend.key = element_rect(fill = "grey85")) +
-guides(colour=guide_legend(title="Subfamily")) +
-theme(axis.title = element_text(face = "bold"))
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="80%"}
-
-plot_all = readRDS("../Data/Processing/plot_all.rds")
-
-plot_all
-
-```

### 2022-12-06 Scripts/Figure_3_Preference_Histogram/Figure_3_Preference_histogram.R

@@ -17,28 +17,38 @@ long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
 long_data = long_data %>% 
     mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
 
-labels.minor <- c("Low", "", "High")
+labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
 
 long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),
                   labels = c("A) Natural", "B) Agricultural", "C) Urban"))
 
+long_data = long_data %>% 
+mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", 
+                                     Preference >= 0.8 ~ "B",
+                                     T ~ "C"))
+
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"),
+                            labels = c("Avoidance", "No preference", "Preference"))
+
+
 #Plot
-ggplot(long_data, aes(x=Preference, color= Habitat)) +
-    geom_histogram(aes(fill= Habitat)) + 
+ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
+    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
     xlab("Habitat preference") +
     ylab("Frequency") +
     facet_wrap(~ Habitat) +
-    scale_colour_viridis_d(end=0.98)+
-    scale_fill_viridis_d(end=0.98) +
+    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
     jtools::theme_apa() +   
     theme(axis.title = element_text(face = "bold"),
     panel.border=element_rect(size=1), 
-    legend.position = "none",
     strip.text = element_text(face = "bold", size = rel(1),hjust = -0.01),
     strip.background = element_rect(fill = "white", colour = "black", size = 0,linetype="solid"))+
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5),
-                       labels = paste0(c(0,0.5,1), "\n", labels.minor)) +
     scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
[…]

### 2022-12-06 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -239,9 +239,7 @@ theme(legend.position = 'bottom', panel.border=element_rect(size=1),
 library(cowplot)
 
 eu <- read.table("Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-nrow(usa)
 usa <- read.table("Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-nrow(usa)
 
 #Europe
 #Plot row 

### 2022-12-07 Manuscript/Ms.Rmd
phylo signal and starting results

@@
 
-\textbf{Miguel Á. Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B. Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}
+\textbf{Miguel Á.
+Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B.
+Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}
 
@@
 Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
+
+Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
 
@@
 All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
+All data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
 
@@
 We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
-Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis® and calculated the average weight for the species with more than one specimen.
+Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
+Finally, we calculated the average weight for the species with more than one specimen.
+Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.
 
@@
 eu_spp = nrow(records_eu)
-spp_number = nrow(records_eu) + nrow(records_usa)
+spp_number = nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
+spp_shared = nrow(bind_rows(records_eu, records_usa)) - nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
+
 #Median number of records per dataset
@@
 These countries were: England, Belgium, Netherlands, Germany and Denmark.
-All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved.
+All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf where just the countries of interest were retrieved.
 To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006.
-After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).
+After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`) wit a total of `r spp_shared` shared species between both regions.
 
@@
 We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones.
-Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference) in comparison with the simulated ones.
+Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 20th percentile (low preference) or above the 80th percentile (high preference) in comparison with the simulated ones.
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100).
+```{r Trait info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+library(dplyr)
+#Load brain data
+brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+#Select unique species
+sp = pref %>% distinct(Species) %>% pull()
+
+#Filter species that are in the phylo
+brain_weight = brain_weight %>% filter(Species %in% sp)
+largest_encephalisation = brain_weight %>% filter(residuals == max(residuals)) %>% select(Species) %>% pull()
+lowest_encephalisation = brain_weight %>% filter(residuals == min(residuals)) %>% select(Species) %>% pull()
+
+#Read model output
+r2 = read_rds("../Data/Processing/bayessianr2.rds")
+r2 = round(r2[1],2)
+
+```
+
+For the `r nlevels(factor(unique(d$Species)))` bee species, brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`).
+However, there were considerable variability in brain size relative to the body across species (Figure S1).
+Considering this association, the species with the largest and smallest brains relative to the body were the American bee *`r largest_encephalisation`* and the European bee *`r lowest_encephalisation`*, respectively.
+
+Mention how this is constrained by phylogenetic relationship but there is still large variability.
+
+Histogram
+
+Model
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
 A) Phylogenetic tree at the genera level.
@@
     geom_tippoint(aes(size = residuals, color=Country))+
-    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Encephalisation")+
+    theme_tree2()+scale_size_continuous(range = c(-1, 4),name="Relative\nbrain size")+
     scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
@@
    # scale_fill_gradient(high="black",  low="white") +
-    scale_fill_viridis_c(name = "Habitat\n preference") + 
+    scale_fill_viridis_c(name = "Habitat\npreference") + 
     theme_minimal() +
@@
 ylab("Habitat preference") +
-xlab("Encephalisation") +
+xlab("Relative brain size") +
 scale_colour_viridis_d() +
@@
 
-#Select just species that fulfilled our criteria of filtering
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe_qualitative.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa_qualitative.csv")
-
-sp = c(preferences_europe$Species, preferences_usa$Species)
-
-r = r %>% 
-mutate(Species = str_replace_all(Species, "_", " ")) %>% 
-filter(Species %in% sp)
-
 #Add ggplot
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="80%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
 

### 2022-12-07 Scripts/0_Processing/0_Preprocessing_new.R

@@ -27,8 +27,6 @@ rename(Brain.weight = Brain.Weight..mg.)
 #Heads are extracted and fixed and then the brain is removed and weighted
 #Fresh brains can be converted with a conversion factor to fix! 
 #brain.fix.mg = brain.fresh.mg x 1.2262. #Check sayol etal
-#Convert fresh to fix when value is na (I guess that almost always when fresh is measured)
-#The trend is the same but looks a bit nicer without adding the brains with this conversion
 sayol = sayol %>% 
 #mutate(brain.fix.mg = case_when(is.na(brain.fix.mg) == T ~ brain.fresh.mg * 1.2262,
 #is.na(brain.fix.mg) == F ~ brain.fix.mg)) %>% 
@@ -95,7 +93,7 @@ sum(d$individuals) #433
 nlevels(factor(unique(d$Species))) #113 species
 
 #Check for possible outliers
-#1st Brain weight
+#Brain weight
 #Extract levels
 l = levels(factor(all_cleaning$Species))
 #Create function to plot everything
@@ -110,7 +108,7 @@ s(l[50:100])
 s(l[100:130])
 
 #Is quite easy to have outliers with brain weights 
-#For this, we filter outliers
+#For this, we filter the ones out of 1.5 times IQR
 all_cleaning = all_cleaning %>%
 group_by(Species) %>% 
 mutate(Max_val = quantile(Brain.weight, 0.75) + 1.5 *IQR(Brain.weight)) %>% 
@@ -119,16 +117,15 @@ filter(Brain.weight <= Max_val) %>% #Species with one unique value are not consi
 filter(Brain.weight >= Min_val) %>% #Species with one unique value are not considered!
 ungroup()
 
+#Check again filtered plots
 s(l[1:50])
 s(l[50:100])
 s(l[100:130])
-
 
 #Create dataframe with average measurements of IT and brain weight
 weights.mean <- data.frame(aggregate(Brain.weight ~ Species, data = all_cleaning, FUN = mean))
 IT.mean <- data.frame(aggregate(IT ~ Species, data = all_cleaning, FUN = mean))
 wit.mean <- merge(weights.mean, IT.mean)
-
 #Check number of levels
 #levels(factor(wit.mean$Species))
 

### 2022-12-07 Scripts/0_Processing/2_Calculate_residuals.R

@@ -29,13 +29,14 @@ m1 <-  brm(log_brain_weight ~ log_it + (1|gr(Species, cov = A)) , data = wit.mea
                  warmup = 1000)
 
 
-bayes_R2(m1)
+r2 = bayes_R2(m1)
 
 ce1 <- conditional_effects(m1, effects = "log_it", points=T) 
 
 #Save data
 write_csv(ce1[[1]] , "Data/Processing/residuals_model.csv")
 write_csv(wit.mean , "Data/Processing/residuals_data.csv")
+write_rds(r2 , "Data/Processing/bayessianr2.rds")
 
 #Model diagnostics
 pp_check(m1)
@@ -54,6 +55,16 @@ ggplot(res_df,
 
 #Extract residuals (estimates) 
 wit.mean$residuals = residuals(m1)[,1]
+
+#visual comparison between fitted and observed values
+#residuals(m1)
+#r =  wit.mean$log_brain_weight - fitted(m1)[,1]
+#r= tibble(a = head(fitted(m1)[,1]), b = head(wit.mean$log_brain_weight), c = head(wit.mean$log_it))
+#ggplot(ce1[[1]], aes(x = log_it, y = estimate__)) +
+#geom_line() +
+#geom_point(data = r[4,], aes(c,a), color="blue") +
+#geom_point(data = r[4,], aes(c,b))
+#wit.mean$residuals[4]
 
 #Check species
 wit.mean$Species

### 2022-12-07 Scripts/1_Download_GBIF_data/1_Main_download_occ_download.R

@@ -3,7 +3,7 @@
 #Quite fast but we need a gbif account!
 #To reproduce our analysis, 
 #Skip this part and use directly the downloaded data from the DOI
-#https://doi.org/10.15468/dl.nenuem
+#https://doi.org/10.15468/dl.5s5kuf
 
 #Load libraries
 library(stringr) 
@@ -52,5 +52,5 @@ countries <- c("US", "DE", "GB", "LU", "BE", "NL", "DK")
 occ_download(pred_in("taxonKey",c(gbif_id$key_number)),  pred_in("country", c("US", "DE", "GB", "LU", "BE", "NL", "DK")))
 
 #The data is downloaded through my gbif account
-#Here is the final data: https://doi.org/10.15468/dl.nenuem
-#Last accessed 11/11/2022 
+#Here is the final data: https://doi.org/10.15468/dl.5s5kuf
+#Last accessed 05/12/2022 

### 2022-12-07 Scripts/1_Download_GBIF_data/2_Process_downloaded_data.R

@@ -1,4 +1,4 @@
-#In this script we process the data downloaded from GBIF: https://doi.org/10.15468/dl.re5wwt
+#In this script we process the data downloaded from GBIF: https://doi.org/10.15468/dl.5s5kuf
 #We select cols of interest with georreferenced information
 
 #Notes: 

### 2022-12-07 Scripts/6_Phylo_for_analyses/Phylo_signal_raltive_brain_size.R

@@ -1,0 +1,22 @@
+
+#Here we calculate the phylogenetic signal of relative brain size
+#Load libraries
+library(ape)
+library(phytools)
+#Read tarit data and phylogenetic signal
+trait = read_csv("Data/Processing/brain_weight_data.csv") %>% 
+mutate(Species = str_replace_all(Species, " ", "_"))
+phylo = readRDS("Data/Processing/phylo_raw.rds")
+
+#Prepare names for analysis
+d = setNames(trait1[,"residuals"], trait$Species)
+
+#Convert to phylo tree
+p_tree = nj(phylo)
+#Check tip labels format
+p_tree$tip.label
+
+#Run analysis
+output = phylosig(p_tree, d, method="lambda",test=TRUE)
+#Check output
+output

### 2022-12-07 Scripts/Figure_2_Phylo_Tree/Figure_2_Phylo_tree_all.R

@@ -35,7 +35,7 @@ bee.tree100$tip.label = str_replace(bee.tree100$tip.label, "_", " ")
 
 #Generate vector with species
 v = bee.tree100$tip.label
-#Filter species with that are in the phylo
+#Filter species that are in the phylo
 brain_weight = brain_weight %>% filter(Species %in% v)
 
 

### 2022-12-07 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -115,11 +115,12 @@ nrow(europe)
 eu_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
-#geom_ribbon(aes(ymin = lower__, ymax = upper__, color=Habitat), linetype = "dashed",alpha = 0)+
+geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
 ylab("Habitat preference") +
 xlab("Residuals") +
 scale_colour_viridis_d() +
+scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
 theme(legend.position="none", panel.border=element_rect(size=1),
 axis.title=element_text(face="bold"),
@@ -128,11 +129,12 @@ legend.box.background = element_rect(colour = "black",size=1))
 eu_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
-#geom_ribbon(aes(ymin = lower__, ymax = upper__, color=Habitat), linetype = "dashed",alpha = 0)+
+geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
 ylab(NULL) +
 xlab("Brain weight (mg)") +
 scale_colour_viridis_d() +
+scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))+
 theme(legend.position="none", panel.border=element_rect(size=1),
 axis.title=element_text(face="bold"),
@@ -142,11 +144,12 @@ legend.box.background = element_rect(colour = "black",size=1))
 eu_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
[…]

### 2022-12-24 Manuscript/Ms.Rmd
adding a relevant ref
organising ideas for results

@@
 
-\textbf{Miguel Á.
-Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B.
-Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}
+\orcidlink{0000-0002-4216-317X} \textbf{Jose B.
+Lanuza$^{1}$} \textbf{Miguel Á.
+Collado$^{2}$}  \orcidlink{0000-0002-0287-409X} \textbf{Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+
+Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
 
@@
 
-Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
-
 Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
 
-\textbf{Author for correspondence:}
+\textbf{Author for correspondence: barragansljose@gmail.com}
 
@@
 <https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R>
+
+https://www.biorxiv.org/content/10.1101/2022.12.09.519739v1
 
@@
 
+#Read phylo signal
+phylo = read_rds("../Data/Processing/phylo_signal.rds")
+lambda = round(phylo$lambda, 2)
 ```
 
-For the `r nlevels(factor(unique(d$Species)))` bee species, brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`).
-However, there were considerable variability in brain size relative to the body across species (Figure S1).
-Considering this association, the species with the largest and smallest brains relative to the body were the American bee *`r largest_encephalisation`* and the European bee *`r lowest_encephalisation`*, respectively.
+Our dataset showed that brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`).
+As expected, this allometric relationship was constrained by the evolutive history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
+However, we found considerable variability in relative brain size (Figure S1; Figure 2A) within and across the different taxonomic groups (i.e., genera and subfamily level).
 
-Mention how this is constrained by phylogenetic relationship but there is still large variability.
+Explain now habitat preference trends by species (Figure 2B)
 
-Histogram
+Explain overall habitat preference patterns (Figure 3)
 
-Model
+Explain main models and model by continent (Figure 4 and Figure S2)
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
 

### 2022-12-24 Scripts/6_Phylo_for_analyses/Phylo_signal_raltive_brain_size.R

@@ -3,13 +3,20 @@
 #Load libraries
 library(ape)
 library(phytools)
+library(dplyr)
+library(readr)
+library(stringr)
+library(data.table)
+
 #Read tarit data and phylogenetic signal
 trait = read_csv("Data/Processing/brain_weight_data.csv") %>% 
 mutate(Species = str_replace_all(Species, " ", "_"))
 phylo = readRDS("Data/Processing/phylo_raw.rds")
 
+
+trait1 = unlist(trait$residuals)
 #Prepare names for analysis
-d = setNames(trait1[,"residuals"], trait$Species)
+d = setNames(trait1, trait$Species)
 
 #Convert to phylo tree
 p_tree = nj(phylo)
@@ -20,3 +27,6 @@ p_tree$tip.label
 output = phylosig(p_tree, d, method="lambda",test=TRUE)
 #Check output
 output
+#Save output
+saveRDS(output, "Data/Processing/phylo_signal.rds")
+

### 2022-12-24 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -40,7 +40,7 @@ geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
 ylab("Habitat preference") +
-xlab("Residuals") +
+xlab("Relative brain size") +
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) 
@@ -118,7 +118,7 @@ geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
 ylab("Habitat preference") +
-xlab("Residuals") +
+xlab("Relative brain size") +
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
@@ -196,7 +196,7 @@ geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
 ylab("Habitat preference") +
-xlab("Residuals") +
+xlab("Relative brain size") +
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))+

### 2022-12-27 Manuscript/Ms.Rmd
Update on ms, methods and results

@@
 
-Keywords: encephalisation, brain-body mass ratio, habitat preference, Apoidea, GBIF
+Keywords: relative brain size, brain-body mass ratio, habitat preference, Apoidea, GBIF
 
@@
 To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models.
-Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.
+Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
 
@@
 From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
+Finally, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats (see next section) for our set of species with the help of the function *phylosig* from the *phytools* package.
 
@@
 
-We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span.
+We estimated brain size relative to the body modeling the log-transformed association between brain weight and intertegular span.
 For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
@@
 Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
-Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation.
+Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the relative brain size.
 High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
@@
 In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones.
-Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 20th percentile (low preference) or above the 80th percentile (high preference) in comparison with the simulated ones.
+We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
+Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were below 20th percentile (i.e., low preference) or above the 80th percentile (i.e., high preference) in comparison with the values obtained from the simulations.
 
-To evaluate if the degree of encephalisation changes by habitat type, we modeled with Bayesian generalised linear models their association.
-For this, we first joined the resulting habitat preference and encephalisation datasets by species and extracted those with statistical habitat avoidance or preference on any of these three habitats.
-Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and encephalisation the predictor.
+To evaluate how the association between habitat preference and relative brain size changed by habitat type, we modeled with Bayesian generalised linear models their association.
+For this, we first joined the resulting habitat preference and relative brain size datasets by species and extracted those with statistical habitat preference on any of these three habitats.
+Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
 Again, we included the phylogenetic covariance matrix as random factor.
@@
 lambda = round(phylo$lambda, 2)
+phylo_nat = read_rds("../Data/Processing/phylo_signal_natural.rds")
+lambda_nat = round(phylo_nat$lambda, 2)
+p_nat = round(phylo_nat$P,2)
+
+#Read summary of habitat preference
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+#Average preferences between USA and Europe
+preferences = pref %>% 
+group_by(Species) %>%
+summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
+ungroup()
+#Agricultural
+pref_agr = pref %>% 
+dplyr::select(Agricultural) %>%  
+filter(Agricultural > 0.80) %>% 
+nrow()
+#Natural
+pref_nat = pref %>% 
+dplyr::select(Natural) %>%  
+filter(Natural > 0.80) %>% 
+nrow()
+#Urban
+pref_urb1 = pref %>% 
+dplyr::select(Urban) %>%  
+filter(Urban < 0.20) %>% 
+nrow()
+#Urban
+pref_urb2 = pref %>% 
+dplyr::select(Urban) %>%  
+filter(Urban > 0.80) %>% 
+nrow()
+
+#Read R2 MAIN MODEL
+r2_main = readRDS("../Data/Processing/r2_main.rds")
+r2_main = round(r2_main[[1]],2)
+#Read R2 MODEL BRAIN WEIGHT
+r2_brain = readRDS("../Data/Processing/r2_brain.rds")
+r2_brain = round(r2_brain[[1]],2)
+#Read R2 MODEL ITS
+r2_its = readRDS("../Data/Processing/r2_its.rds")
+r2_its = round(r2_its[[1]],2)
+
 ```
@@
 As expected, this allometric relationship was constrained by the evolutive history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
-However, we found considerable variability in relative brain size (Figure S1; Figure 2A) within and across the different taxonomic groups (i.e., genera and subfamily level).
+However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 2A).
+Similarly, habitat preference varied substantially across species (Figure 2B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+In general, most bee species preferred one or two habitat types but more rarely occurred indistinctively in the three habitats (Figure 2B).
+The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 2B and Figure 3).
+In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 2B and Figure 3).
+The resulting distribution of habitat preferences for each of the different habitats followed a zero-one inflated beta distribution (Figure 3).
+In other words, there were high frequencies of habitat preferences equal to 0 or 1 but low frequencies of intermediate values between 0 and 1.
 
-Explain now habitat preference trends by species (Figure 2B)
-
-Explain overall habitat preference patterns (Figure 3)
-
-Explain main models and model by continent (Figure 4 and Figure S2)
+Our models evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`).
+However, we found differences when we explored this association for each of the habitat types (Figure 4A).
+Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 4A).
+Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 4A).
+Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 4B and 4C).
+Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 4B and 4C).
+These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S2).
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.}", fig.pos="p"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.} Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
+The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p"}
 
@@
 
+set.seed(3)
+
 all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
+geom_jitter(width = 0.05, height = 0,data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
 geom_line(aes(color=Habitat),size=0.7) +
@@
 arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.}",
+kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.} Original downloaded cover classes for United States with the respective assigned cover classes and number of bee occurrences found on each of them.",
 booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
@@
 arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.}",booktabs = T,
+kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.} Original downloaded cover classes for Europe with the respective assigned cover classes and number of bee occurrences found on each of them.",booktabs = T,
 align = c("l", "c", "l"), linesep = "") %>% 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.}", fig.pos="p", out.width="80%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116).
+The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2.}", fig.pos="p", out.width="100%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2}.
+Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
+The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}
 

### 2022-12-27 Scripts/6_Phylo_for_analyses/Phylo_signal_raltive_brain_size.R

@@ -1,5 +1,5 @@
 
-#Here we calculate the phylogenetic signal of relative brain size
+#Here we calculate the phylogenetic signal of relative brain size----
 #Load libraries
 library(ape)
 library(phytools)
@@ -30,3 +30,41 @@ output
 #Save output
 saveRDS(output, "Data/Processing/phylo_signal.rds")
 
+#Here we calculate the phylogenetic signal of habitat preference----
+#Load preferences
+preferences_europe = read_csv("Data/Europe_data/preferences_europe.csv") %>% 
+mutate(country = "europe")
+preferences_usa = read_csv("Data/Usa_data/preferences_usa.csv") %>% 
+mutate(country = "usa")
+phylo = readRDS("Data/Europe_USA/phylo_all.rds")
+
+#There are duplicated species between usa and europe
+f <- intersect(preferences_europe$Species, preferences_usa$Species)
+#Filter them out from usa and just select europe which had higher number of records
+preferences_usa = preferences_usa %>% 
+filter(!Species %in% f)
+
+#Merge datasets
+preferences = bind_rows(preferences_europe, preferences_usa)%>% 
+mutate(Species = str_replace_all(Species, " ", "_"))
+trait1 = unlist(preferences$Urban)
+trait2 = unlist(preferences$Natural)
+trait3 = unlist(preferences$Agricultural)
+#Prepare names for analysis
+d1 = setNames(trait1, preferences$Species)
+d2 = setNames(trait2, preferences$Species)
+d3 = setNames(trait3, preferences$Species)
+#Convert to phylo tree
+p_tree = nj(phylo)
+#Check tip labels format
+p_tree$tip.label
+
+#Run analysis
+output1 = phylosig(p_tree, d1, method="lambda",test=TRUE)
+output1
+output2 = phylosig(p_tree, d2, method="lambda",test=TRUE)
+output2
+output3 = phylosig(p_tree, d3, method="lambda",test=TRUE)
+output3
+#Save output
+saveRDS(output2, "Data/Processing/phylo_signal_natural.rds")

### 2022-12-27 Scripts/7_Analyses/Analysis_all.R

@@ -43,7 +43,8 @@ model1 = brm(Preference ~ residuals * Habitat +  (1|gr(Species, cov = A)),
 ce1 <- conditional_effects(model1, effects = "residuals:Habitat", points=T) 
 pp_check(model1)
 
-bayes_R2(model1)
+r2_main = bayes_R2(model1)
+saveRDS(r2_main, "Data/processing/r2_main.rds")
 
 p1 = ggplot(ce1[[1]], aes(x = residuals, y = estimate__, color=Habitat)) +
     geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
@@ -63,6 +64,9 @@ model2 = brm(Preference ~ Brain.weight * Habitat + (1|gr(Species, cov = A)),
 
 pp_check(model2)
 bayes_R2(model2)
+
+r2_brain = bayes_R2(model2)
+saveRDS(r2_brain, "Data/processing/r2_brain.rds")
 
 ce2 <- conditional_effects(model2, effects = "Brain.weight:Habitat",points=T) 
 
@@ -84,6 +88,8 @@ model3 = brm(Preference ~ IT * Habitat + (1|gr(Species, cov = A)),
 
 pp_check(model3)
 bayes_R2(model3)
+r2_its = bayes_R2(model3)
+saveRDS(r2_its, "Data/processing/r2_its.rds")
 
 ce3 <- conditional_effects(model3, effects = "IT:Habitat",points=T) 
 

### 2022-12-27 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -119,12 +119,14 @@ geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linety
 theme_bw() +
 ylab("Habitat preference") +
 xlab("Relative brain size") +
+ggtitle("D") +  
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
 theme(legend.position="none", panel.border=element_rect(size=1),
 axis.title=element_text(face="bold"),
-legend.box.background = element_rect(colour = "black",size=1))
+legend.box.background = element_rect(colour = "black",size=1),
+plot.title = element_text(face="bold"))
 
 eu_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
 geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
@@ -133,12 +135,14 @@ geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linety
 theme_bw() +
 ylab(NULL) +
 xlab("Brain weight (mg)") +
+ggtitle("E") +  
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))+
 theme(legend.position="none", panel.border=element_rect(size=1),
 axis.title=element_text(face="bold"),
-legend.box.background = element_rect(colour = "black",size=1))
+legend.box.background = element_rect(colour = "black",size=1),
+plot.title = element_text(face="bold"))
 
 
 eu_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
@@ -148,12 +152,14 @@ geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linety
 theme_bw() +
 ylab(NULL) +
 xlab("Intertegular distance (mm)") +
+ggtitle("F") +  
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
 scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High"))+
 theme(legend.position="none", panel.border=element_rect(size=1),
 axis.title=element_text(face="bold"),
-legend.box.background = element_rect(colour = "black",size=1))
[…]

### 2023-02-02 Manuscript/Ms.Rmd
Re-write intro and dicusion.

@@
 ---
-title: "Bee brain size is positively associated with urban lifestyles but not with positive population trends"
+title: "Urban bees have larger brains"
 output:
@@
 \orcidlink{0000-0002-4216-317X} \textbf{Jose B.
-Lanuza$^{1}$} \textbf{Miguel Á.
-Collado$^{2}$}  \orcidlink{0000-0002-0287-409X} \textbf{Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+Lanuza$^{1}$} \textbf{, Miguel Á.
+Collado$^{2}$}  \orcidlink{0000-0002-0287-409X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
 
-Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
+$^{1}$Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
 
-Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
+$^{2}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
 
-Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
+$^{3}$Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
 
@@
 While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for 145 individuals from 44 species of bees ---the most important group of pollinators--- from North America and Europe, and compared them with detailed information on habitat preferences and population trends.
-Our analyses revealed that bees that prefer forest habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger brains compared to forest dwellers.
-However, we found no evidence that either larger brains or a preference for urban habitats are associated with positive population trends.
-These results suggest that a large brain can help maintaining bee urban populations, but also highlight that being tolerant to urbanisation is not enough to compensate for the worldwide decline of their populations.
+To address this issue, we measured brains for XXX individuals from XX species of bees, the most important group of pollinators, and compared them with detailed information on habitat preferences.
+Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
+In addition, urban bees also tend to be larger, and hence, show larger absolute brain sizes.
+These results suggest that a large brain can help maintaining bee urban populations.
 
-Keywords: relative brain size, brain-body mass ratio, habitat preference, Apoidea, GBIF
+Keywords: relative brain size, habitat preference, Apoidea, Urbanization.
 
@@
 
+[REFEREBCES TO ADD: https://link.springer.com/article/10.1007/s10980-017-0592-x; https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/1365-2435.12486; https://link.springer.com/article/10.1007/s10531-020-02003-8; https://pubmed.ncbi.nlm.nih.gov/28976620/ ; https://www.biorxiv.org/content/10.1101/2022.12.09.519739v1 (this one for big body size in cities in disc.
+also see https://onlinelibrary.wiley.com/doi/10.1111/eva.13087); https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R (maybe in the intro); https://www.frontiersin.org/articles/10.3389/fevo.2020.00058/full]()
+
 # Introduction
 
-Relevant refs:
+Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [Ollerton](https://www.annualreviews.org/doi/abs/10.1146/annurev-ecolsys-110316-022919).
+Regrettably, there are clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
+Although the reason why many pollinators are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
+The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011, @brown2009].
+However, not all pollinator species are equally affected by habitat loss and alteration.
+In bees, for example, there are also "winners" ---like *Bombus terrestris* [@rasmont2008]--- that are able to thrive in human-altered environments, as well as species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; @collado2019].
 
-<https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R>
+Human-dominated habitats ---notably cities--- drastically modify the original conditions were pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+The question is then why only some species are able to enjoy the benefits of an urban life.
+In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
+Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
+For example, reptiles, amphibians and birds that thrieved after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
+[maybe also a sentence on urban dwelers having larger brains?
+Or if there are good references, replace the concept of invaders, by the one on urban species]() However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently [citar a chitka]().
+We now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
+Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not beeing explored.
 
-https://www.biorxiv.org/content/10.1101/2022.12.09.519739v1
-
-This may be also relevant:
-
-Although recent studies challenge the use of absolute brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein), many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008).
-
-Pollination is a fundamental ecosystem service that connects the earth's vegetation and human economy.
-Pollination services are performed by both managed and wild pollinators, mostly bees [@james2008; @winfree2011], both of which might be affected by a range of current and projected human-related environmental changes.
-There are indeed clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019], and parallel declines in the plants that rely upon them [@potts2010; @scheper2014].
-For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees [@williams2009].
-The decline of pollinators currently observed is projected to increase in the future [@dormann2008].
-Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production [@klein2007; @ollerton2011], there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services [@ashman2004; @klein2007].
-
-Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@brown2009].
-The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth [@brown2009].
-However, not all pollinators species are equally affected by habitat loss and alteration.
-In bees, for example, there are also "winners" ---like Bombus terrestris [@rasmont2008]--- that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].
-
-Human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown.
-Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive [@manenti2013].
-For species that tolerate human-related disturbances, human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
-Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
-In vertebrates, relative brain size is a good proxy of problem-solving and learning (e.g.
-see chapter 2 of this thesis for its relationship with learning), which can help individuals to cope with the many challenges on novel environments.
-However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains.
-More important, there was no unequivocal evidence that brain size was related to cognitive performance.
-However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
-There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2).
-We used previously published data on bee forest and urban preference [@collado2019] as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
-Species population trends were extracted from @bartomeus2013 and both datasets were crossed with a newly collected brain weight database.
-We wanted to explore if there is a relationship between brain size, habitat preference, and population trends.
-Hence, our questions were the following:Are brain sizes, as a proxy of behavioural plasticity, related to habitat preferences?Is population stability related to brain size as it has been seen in other taxa?Are habitat preferences related to population stability?
-Or alternatively, are urban dwellers populations increasing?
+Here, we characterize habitat preference for XXX bee species [@collado2019] and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
+In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
+Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [Collado et al 2021 PSOS](cite).
+Hence, our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas applies also to insects.
 
@@
 
-Our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas.
-To evaluate this hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size.
-For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and selected as study area the states or countries with highest density of records.
-After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models.
+To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [Sayol et al.](citar).
+For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF).
+After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories: (1) natural, (2) agricultural and (3) urban.
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [collado et al 2019]().
 Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
@@
 
-Brain size was measured as the weight of fixed brain for each specimen.
-First, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013].
+Brain size was measured as the weight of the fixed brain for each specimen.
+First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anesthetized in cold [@crook2013].
 Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
@@
 All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf where just the countries of interest were retrieved.
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006.
+To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a distribution larger than the area selected; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications, which is from 2006.
 After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`) wit a total of `r spp_shared` shared species between both regions.
@@
 
-We assigned a habitat type for all GBIF occurrences by merging land cover information with the different georeferenced records.
+We assigned a habitat type to each GBIF occurrence by merging land cover information with the different georeferenced records.
 The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
@@
 
-We estimated brain size relative to the body modeling the log-transformed association between brain weight and intertegular span.
+We estimated brain size relative to body size modeling the log-transformed association between brain weight and intertegular span.
 For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
@@
 
-Bees with urban habitat preferences tended to have bigger brains related to their body sizes, and those with forest preferences had relatively smaller brains.
-Surprisingly, neither brain sizes nor habitat preference explained population trends, against what we expected.
-Therefore, species with bigger brains appearing on urban habitat can be indicating that while only bees with large cognitive capacities are surviving in these modified environments, they do not have experienced notorious increases in their populations.We hypothesized that bees with bigger brains related to body size should be more flexible in their behaviour because a large brain may confer better cognitive abilities, as it has been observed in some species (Lefebvre et al., 2004) and as we show specifically for bees in relation to learning (Chapter 2 of this thesis).
-Therefore, to exploit complex urban environments, we expect bigger brains to be required.
+Bees with a preference for inhabiting urban habitats tended to have bigger brains related to their body sizes, and those with forest or agricultural  preferences had relatively smaller brains.
+This pattern is highly suggestive, because investing in brain tissue is expensive (REF), and hence it only payoff in unpredictable habitats where a flexible behavior is needed.
+When the environment is stable, small relative brains are cost-effective to exploit predictable resources, but in highly unpredictable environments, a large brain can confer the needed flexibility to exploit unpredictable resources.
+Indeed, specialist bees also tend to have larger brains [@Sayol](cite apropietely), as they need to locate and memorize patchy resources.
+This same complex learning and navigation skills may be also needed in urban environments.
 Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment (Wheater, 1999).
-Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than the forest dwellers counterpart.
-Some examples of behaviours needed to survive in urban environments (see Table S2 for a full set of examples) are the use of human-made materials to nest (Allasino et al., 2019) or being able to collect pollen and nectar from exotic flowers from gardens (Lowenstein et al., 2014).
+Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than forest or agricultural bees.
+
+Some examples of behaviours needed to survive in urban environments are the use of human-made materials to nest (Allasino et al., 2019) or being able to collect pollen and nectar from exotic flowers from gardens (Lowenstein et al., 2014).
 We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them (Smith et al., 2019).
-In accordance, we found smaller brain sizes than expected by their body sizes in bees that preferred forests indicating that an investment in bigger relative brain sizes is not required.
-In addition, other traits have already been described to be filtered in anthropogenic land uses, like early spring or brood parasitism (Harrison et al., 2017), which are present in the forest dwellers.
-We could not explain population trends with brain size as it happens in other taxa.
-For example, in birds, brain size is a good predictor of how a population does in the long-term (Shultz, 2005).
-However, this is not always the case as bigger brains increase the risk of extinction in mammals (González-Voyer et al., 2016).
-Despite brain weight is relatively easy to measure and standardize with the used protocol, we lack information on certain neuropils size, like mushroom bodies (the suspected centers of cognitive processes in bees), which can be variable even within species (Barth & Heisenberg, 1997; Withers et al., 1995; Fahrbachand & Robinson 1995).
-Differences in more subtle brain structures could explain differences in behaviours, and in consequence, shed extra light on species population trends.
-Therefore the isolation and measurement of these neuropils could be interesting because we could get more accurate information.
-On the other hand, also note that population trends data is hard to find or measure, especially for invertebrate taxa.
-Long term data is often lacking and we count only with data indirectly taken from museums.
-This data is also scarce and hard to standardize (Bartomeus et al., 2018), therefore not being optimal for comparative studies of declining or increasing populations (Shultz, 2005).
-Species body size is generally related to population trends.
-In mammals, it is known that bigger body sizes decreases population density and increases the risk of extinction (Damuth, 1981; González-Voyer et al., 2016).
-This relation has also been seen in bees (Scheper et al., 2014).
-Although body size is directly related to absolute brain size, we did not find a strong correlation between body size and population trends in our subset of North American bees as it was found for the complete set in Bartomeus et al., 2013.
-Our subset may be biased towards the most common genera and contain fewer species per genus.
-If the body size-decline relationship occurs mainly within genus, and not across genus, as indicated by the strong phylogenetic signal of body size (Pagel's lambda \> 0.9), we may lack the power to detect this variation within genus in our restricted analysis.
-Population trends or habitat preferences could be explained by the availability of the resources consumed, for example, in the Netherlands, population trends of some bee species are related to the availability of their preferred plants (Scheper et al., 2014).
-A polylectic diet could help species to adapt to new environments, and we know that species able to forage on crop plant families tend to have stable or increasing populations (Scheper et al., 2014, Bartomeus et al.
-2013).
-If we consider habitats being conditioned by their availability of the preferred resources, then we may be able to explain our link between brain size and preference for certain habitats.
-For example, differences in bee community composition and abundance are found for oligolectic (i.e.diet specialist bees) associated to their prefered resources along forested gradients (Grundel et al., 2010).
-Dietary data should be an interesting next step to see if broader (polylectic) or narrower (oligolectic) diet are generally related to habitat preferences and brain sizes.
-In fact, it has been already described for other taxa a relationship between brain size and diet (DeCasien et al., 2017).Finally, the absence of relationship between urban preferences and increasing population trends is interesting.
-Despite these bees have found new niches to occupy, they are not increasing its populations but just persisting.
-The transformation from natural to urban habitats is filtering which species can survive, and hence, biodiversity in urban habitats is usually lower (Collado et al., 2019, Chapter 3 of this thesis).
-However, survivors are not thriving either in this new conditions.
-Some negative effects of urbanization have already been studied in bees, like for example, biased sex ratio, with increasing males in urban areas (Fitch et al., 2019) with unknown future population consequences.
-It should be interesting to fully assess the costs and benefits of living in urban environments, including brain investment and other selected strategies to adapt to the new conditions.
+In addition, other traits have already been described to be filtered out in cities, such as smaller body sizes (REF), early spring species or brood parasites (Harrison et al., 2017).
+We also found small species to be less frequent in urban habitats, maybe as a consequence of the patchy resource distribution, which need from larger body sizes, which confer larger foraging ranges, to be accessed [Kendal et al](cite the 2022 on foraging-body size relationship).
+
+Despite brain weight is relatively easy to measure and standardize with the used protocol, recent studies challenge the use of brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein).
+However, many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008), and brain size is correlated with learning abilities in bees (Collado et al 2021).
+Finer scale measures such as neuropils size or detailed mushroom bodies measurements (the suspected centers of cognitive processes in bees; Barth & Heisenberg, 1997; Withers et al., 1995; Fahrbachand & Robinson 1995) could further explain differences in behaviours, and in consequence, shed extra light on species habitat associations.
+In any case, raw measures of brain size have the advantage to allow for larger comparisons across species.
+
+In conclusion, the fact that urban bees tend to have larger relative brain sizes highlights the importance of cognitive process for understanding the dynamics of insect populations, and stress the importance of considering behavioural responses beyond vertabrate taxa.
 

### 2023-02-04 Manuscript/Ms.Rmd
adding refs and moving figure 2 and 3 to supp mat

@@
 
-\orcidlink{0000-0002-4216-317X} \textbf{Jose B.
-Lanuza$^{1}$} \textbf{, Miguel Á.
-Collado$^{2}$}  \orcidlink{0000-0002-0287-409X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+\textbf{Jose B.
+Lanuza$^{1}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
+Collado$^{2}$}  \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
 
@@
 \textbf{Author for correspondence: barragansljose@gmail.com}
+
+[Not a great fan of the title, maybe something like this?](J)
+[Bees with larger brains relative to their body are more common in urban environments](J)
 
@@
 
-# Introduction
+# 1. Introduction
 
-Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [Ollerton](https://www.annualreviews.org/doi/abs/10.1146/annurev-ecolsys-110316-022919).
+Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
 Regrettably, there are clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
-Although the reason why many pollinators are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
+Although the reason why many pollinators are declining is insufficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
 The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011, @brown2009].
@@
 Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that thrieved after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
+For example, reptiles, amphibians and birds that thrived after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
 [maybe also a sentence on urban dwelers having larger brains?
-Or if there are good references, replace the concept of invaders, by the one on urban species]() However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently [citar a chitka]().
+Or if there are good references, replace the concept of invaders, by the one on urban species]() However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently [citar a chitka]()[Que refs?](J).
 We now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
-Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not beeing explored.
+Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not being explored.
 
@@
 In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
-Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [Collado et al 2021 PSOS](cite).
+Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [@collado2021].
 Hence, our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas applies also to insects.
@@
 
-# Methods
+# 2. Methods
 
@@
 
-To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [Sayol et al.](citar).
+[yo no pondria refs en el parrafo de overview](J)
+
+To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [@sayol2020].
 For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF).
 After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [collado et al 2019]().
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [@collado2019].
 Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
@@
 We selected the boundaries from the maps packages *spData* version 2.2.0 [@spdata] and *giscoR* vesion 0.3.2 [@giscoR], and extracted the occurrences with the help of the function *st_intersection* from the package *sf* version 1.0-7 [@pebesma2018].
-For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$.
+For North America, we selected states located on the East Coast of United States (Figure S1A) that covered an approximate area of `r usakm2` km$^{2}$.
 These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-Similarly, we selected countries located on the north and center of Europe (Figure 1B) that represented a total area of `r eukm2` km$^{2}$.
+Similarly, we selected countries located on the north and center of Europe (Figure S1B) that represented a total area of `r eukm2` km$^{2}$.
 These countries were: England, Belgium, Netherlands, Germany and Denmark.
@@
 After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`) wit a total of `r spp_shared` shared species between both regions.
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
 
@@
 
-# Results
+# 3. Results
 
@@
 As expected, this allometric relationship was constrained by the evolutive history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
-However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 2A).
-Similarly, habitat preference varied substantially across species (Figure 2B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species preferred one or two habitat types but more rarely occurred indistinctively in the three habitats (Figure 2B).
-The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 2B and Figure 3).
-In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 2B and Figure 3).
-The resulting distribution of habitat preferences for each of the different habitats followed a zero-one inflated beta distribution (Figure 3).
+However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S2; Figure 1A).
+Similarly, habitat preference varied substantially across species (Figure 1B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+In general, most bee species preferred one or two habitat types but more rarely occurred indistinctively in the three habitats (Figure 1B).
+The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S3).
+In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
+The resulting distribution of habitat preferences for each of the different habitats followed a zero-one inflated beta distribution (Figure S3).
 In other words, there were high frequencies of habitat preferences equal to 0 or 1 but low frequencies of intermediate values between 0 and 1.
@@
 Our models evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`).
-However, we found differences when we explored this association for each of the habitat types (Figure 4A).
-Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 4A).
-Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 4A).
-Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 4B and 4C).
-Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 4B and 4C).
-These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S2).
+However, we found differences when we explored this association for each of the habitat types (Figure 2A).
+Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
+Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 2A).
+Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
+Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 2B and 2C).
+These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S4).
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
 A) Phylogenetic tree at the genera level.
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
-
-#Histogram of habitat preferences
-#Load libraries
-library(tidyverse)
-
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
-
-#Merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
-#Convert to long to model everything at the same time
-long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
-
-#Relevel for plotting
-long_data = long_data %>% 
-mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
-
-long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
-
-long_data = long_data %>% 
-mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
-
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
-
-ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
-    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
-    xlab("Habitat preference") +
-    ylab("Frequency") +
-    facet_wrap(~ Habitat) +
-    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    theme_bw() +   
-    theme(axis.title = element_text(face = "bold", size = 14),
-    panel.border=element_rect(size=1),
-    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
-    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
-    panel.grid.minor=element_blank())+
-    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
-    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
-                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
-  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.} Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 2.} Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
 The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p"}
@@
 
-# Discussion
+# 4. Discussion
 
@@
 When the environment is stable, small relative brains are cost-effective to exploit predictable resources, but in highly unpredictable environments, a large brain can confer the needed flexibility to exploit unpredictable resources.
-Indeed, specialist bees also tend to have larger brains [@Sayol](cite apropietely), as they need to locate and memorize patchy resources.
+Indeed, specialist bees also tend to have larger brains [@sayol2020], as they need to locate and memorize patchy resources.
 This same complex learning and navigation skills may be also needed in urban environments.
@@
 In addition, other traits have already been described to be filtered out in cities, such as smaller body sizes (REF), early spring species or brood parasites (Harrison et al., 2017).
-We also found small species to be less frequent in urban habitats, maybe as a consequence of the patchy resource distribution, which need from larger body sizes, which confer larger foraging ranges, to be accessed [Kendal et al](cite the 2022 on foraging-body size relationship).
+We also found small species to be less frequent in urban habitats, maybe as a consequence of the patchy resource distribution, which need from larger body sizes, which confer larger foraging ranges, to be accessed [@kendall2019].
 
@@
 
-In conclusion, the fact that urban bees tend to have larger relative brain sizes highlights the importance of cognitive process for understanding the dynamics of insect populations, and stress the importance of considering behavioural responses beyond vertabrate taxa.
+In conclusion, the fact that urban bees tend to have larger relative brain sizes highlights the importance of the cognitive process for understanding the dynamics of insect populations, and stress the importance of considering behavioural responses beyond vertabrate taxa.
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
+A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
+
+#Plot maps with density of bee occurrences
+
+#Europe----
+#Load libraries
+library(giscoR) #map of europe
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(sf)
+library(patchwork)
+library(rnaturalearth)
+
+#Load europe map
+Europe <- gisco_get_countries(region = "Europe")
+#Load countries filtered
+euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
+#Load data with records filtered
+all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+#Plot Europe map with counties of interest highlighted
+p1 = ggplot(Europe) +
+geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
+geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
+aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
+coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white"))
+
+p2 = ggplot(Europe) +
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("B") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(), 
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(40, 60, by = 6)) +
+scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
+
+#Insert general plot within specific plot
+panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
+
+#USA-----
+#Load libraries
+
+#Load data
+usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
+ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
+all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+#Load maps
+canada <- ne_states(country = "canada", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+p1 = ggplot(canada) + 
+geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white")) +
+geom_map(data = ny_metropolitan_1, 
+map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
+size = 0.1) +
+geom_map(data = usa_states_1, map = usa_states_1,
+aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
+size = 0.1) +
+coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
+
+p2 = ggplot(canada) + 
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("A") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(),
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(33, 47, by = 4)) +
+scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
+
+#Insert general plot within specific plot
+panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
+
+library(patchwork)
+panel2  + panel1
+
+```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.} Association between the log-transformed values of brain and body size (N = 116).
 The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2}.
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
+The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
+
+#Histogram of habitat preferences
+#Load libraries
+library(tidyverse)
+
+#Load preferences
+preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
+preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
+
+#Merge datasets
+preferences = bind_rows(preferences_europe, preferences_usa)
+
+#Convert to long to model everything at the same time
+long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
+
+#Relevel for plotting
+long_data = long_data %>% 
+mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
+
+labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
+
+long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
+
+long_data = long_data %>% 
+mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
+
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
+
+ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
+    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
+    xlab("Habitat preference") +
+    ylab("Frequency") +
+    facet_wrap(~ Habitat) +
+    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    theme_bw() +   
+    theme(axis.title = element_text(face = "bold", size = 14),
+    panel.border=element_rect(size=1),
+    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
+    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
+    panel.grid.minor=element_blank())+
+    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
+    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
+    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
+    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
+                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
+  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
+
+```
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.
 Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).

### 2023-02-05 Manuscript/Ms.Rmd
Preparing the MS for biology letters, formatting on length, number of figures and adding pertinent refs

@@
 ---
-title: "Urban bees have larger brains"
+title: "Bees need larger brains to thrive in urban environmnets"
 output:
@@
 - \floatplacement{figure}{H}
+
+csl: BiologyLetters.csl
+
 ---
@@
 
-[Not a great fan of the title, maybe something like this?](J)
-[Bees with larger brains relative to their body are more common in urban environments](J)
+```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+library(readr)
+library(dplyr)
+
+#Load final dataset 
+raw = read_csv("../Raw_data/all_cleaning.csv")
+
+d = raw %>% 
+group_by(Species) %>% 
+summarise(individuals = n())
+
+#Load taxonomic rank script
+tax = read_csv("../Data/Processing/taxonomic_rank.csv")
+
+```
 
@@
 While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for XXX individuals from XX species of bees, the most important group of pollinators, and compared them with detailed information on habitat preferences.
+To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and compared them with detailed information on habitat preferences.
 Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
@@
 
-Keywords: relative brain size, habitat preference, Apoidea, Urbanization.
+Keywords: relative brain size, habitat preference, Apoidea, urbanization.
 
@@
 
-Here, we characterize habitat preference for XXX bee species [@collado2019] and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
+Here, we characterize habitat preference for `r nlevels(factor(unique(d$Species)))` bee species and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
 In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
@@
 
-## Overview
-
-[yo no pondria refs en el parrafo de overview](J)
-
-To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [@sayol2020].
-For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF).
-After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [@collado2019].
-Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
-
-All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
-All data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
-
-```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-library(readr)
-library(dplyr)
-
-#Load final dataset 
-raw = read_csv("../Raw_data/all_cleaning.csv")
-
-d = raw %>% 
-group_by(Species) %>% 
-summarise(individuals = n())
-
-#Load taxonomic rank script
-tax = read_csv("../Data/Processing/taxonomic_rank.csv")
-
-```
-
 ## (a) Bee species dataset
@@
 Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
+Detail information of how brain and body size were measured can be found in supplementary material.
 
-## (b) Brain and body size measurements
-
-Brain size was measured as the weight of the fixed brain for each specimen.
-First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anesthetized in cold [@crook2013].
-Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
-Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies.
-By doing this, we ensured that our brain measurements mainly accounted for neural tissue.
-We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
-Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
-Finally, we calculated the average weight for the species with more than one specimen.
-Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.
-
-We considered as a proxy of body size the average intertegular span (ITS) per species.
-ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019].
-All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
-
-## (c) Occurrence data
+## (b) Occurrence data
 
@@
 Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
-We selected the boundaries from the maps packages *spData* version 2.2.0 [@spdata] and *giscoR* vesion 0.3.2 [@giscoR], and extracted the occurrences with the help of the function *st_intersection* from the package *sf* version 1.0-7 [@pebesma2018].
 For North America, we selected states located on the East Coast of United States (Figure S1A) that covered an approximate area of `r usakm2` km$^{2}$.
-These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
 Similarly, we selected countries located on the north and center of Europe (Figure S1B) that represented a total area of `r eukm2` km$^{2}$.
-These countries were: England, Belgium, Netherlands, Germany and Denmark.
-All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf where just the countries of interest were retrieved.
 To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a distribution larger than the area selected; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications, which is from 2006.
-After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`) wit a total of `r spp_shared` shared species between both regions.
 
-## (d) Land cover classification
+## (c) Land cover classification
 
@@
 
-## (e) Construction of the phylogenetic tree
+## (d) Construction of the phylogenetic tree
 
@@
 
-## (f) Analysis
+## (e) Analysis
 
@@
 For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
-To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor.
+To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor (see supplementary text X for phylogenetic tree construction).
 Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
@@
 Finally, the different posterior predictive checks were conducted with the function *pp_check* also from the *brms* package.
+
+All our analyses were undertaken in R version 4.0.5 [@r2021], and all data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
 
@@
 
-Our models evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`).
-However, we found differences when we explored this association for each of the habitat types (Figure 2A).
-Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
-Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 2A).
-Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
-Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 2B and 2C).
-These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S4).
-
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
@@
 \clearpage
+
+Our models evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`).
+However, we found differences when we explored this association for each of the habitat types (Figure 2A).
+Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
+Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 2A).
+Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
+Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 2B and 2C).
+These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S4).
 
@@
 This same complex learning and navigation skills may be also needed in urban environments.
-Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment (Wheater, 1999).
+Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment [@wheater2002].
 Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than forest or agricultural bees.
 
-Some examples of behaviours needed to survive in urban environments are the use of human-made materials to nest (Allasino et al., 2019) or being able to collect pollen and nectar from exotic flowers from gardens (Lowenstein et al., 2014).
-We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them (Smith et al., 2019).
-In addition, other traits have already been described to be filtered out in cities, such as smaller body sizes (REF), early spring species or brood parasites (Harrison et al., 2017).
+Some examples of behaviours needed to survive in urban environments are the use of human-made materials to nest [@allasino2019] or being able to collect pollen and nectar from exotic flowers from gardens [@lowenstein2019].
+We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them [@smith2019].
+In addition, other traits have already been described to be filtered out in cities, such as smaller body sizes (REF), early spring species or brood parasites [@harrison2019].
 We also found small species to be less frequent in urban habitats, maybe as a consequence of the patchy resource distribution, which need from larger body sizes, which confer larger foraging ranges, to be accessed [@kendall2019].
 
-Despite brain weight is relatively easy to measure and standardize with the used protocol, recent studies challenge the use of brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein).
-However, many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008), and brain size is correlated with learning abilities in bees (Collado et al 2021).
-Finer scale measures such as neuropils size or detailed mushroom bodies measurements (the suspected centers of cognitive processes in bees; Barth & Heisenberg, 1997; Withers et al., 1995; Fahrbachand & Robinson 1995) could further explain differences in behaviours, and in consequence, shed extra light on species habitat associations.
+Despite brain weight is relatively easy to measure and standardize with the used protocol, recent studies challenge the use of brain size as an estimator for behavioral complexity and computational power [@chittka2009\; and references therein].
+However, many studies have shown correlations between brain size and performance [@deaner2007; @gronenberg2010; @sol2008], and brain size is correlated with learning abilities in bees [@collado2021].
+Finer scale measures such as neuropils size or detailed mushroom bodies measurements [the suspected centers of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] could further explain differences in behaviours, and in consequence, shed extra light on species habitat associations.
 In any case, raw measures of brain size have the advantage to allow for larger comparisons across species.
@@
 In conclusion, the fact that urban bees tend to have larger relative brain sizes highlights the importance of the cognitive process for understanding the dynamics of insect populations, and stress the importance of considering behavioural responses beyond vertabrate taxa.
+
+# Data accessibility
+
+All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf.
+In addition, bee trait data and code can be found in the online repositories of Zenodo XXXXXXXKXX and Github https://github.com/ibartomeus/declines_brain.git.
+
+# Authors' contributions
+
+JBL, MAC, FS, IB...
+
+# Funding
+
+# Acknowledgements
+
+# Conflict of interest declaration
+
+We declare we have no competing interests.
 
@@
 # Supplementary material
+
+## Brain and body size measurements
+
+Brain size was measured as the weight of the fixed brain for each specimen.
+First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anesthetized in cold.
+Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
+Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies.
+By doing this, we ensured that our brain measurements mainly accounted for neural tissue.
+We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
+Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
+Finally, we calculated the average weight for the species with more than one specimen.
+Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.
+
+We considered as a proxy of body size the average intertegular span (ITS) per species.
+ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight (Cane et al., 2006; Kendall et al., 2019).
+All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
+
+Cane, J.
+H., Minckley, R.
+L., Kervin, L.
+J., Roulston, T.
+A.
+H., & Williams, N.
+M.
+(2006).
+Complex responses within a desert bee guild (Hymenoptera: Apiformes) to urban habitat fragmentation.
+_Ecological applications_, _16_(2), 632-644.
+
+Kendall, L.
+K., Rader, R., Gagic, V., Cariveau, D.
+P., Albrecht, M., Baldock, K.
+C., ...
+& Bartomeus, I.
+(2019).
+Pollinator size and its consequences: Robust estimates of body size in pollinating insects.
+_Ecology and Evolution_, _9_(4), 1702-1714.
 
@@
 ```
+
+\clearpage
 
@@
 
+\clearpage
+
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.} Association between the log-transformed values of brain and body size (N = 116).
@@
 
+\clearpage
+
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.

### 2023-02-05 Manuscript/Word.Rmd
Preparing the MS for biology letters, formatting on length, number of figures and adding pertinent refs

--- 2023-02-04 Manuscript/Ms.Rmd
@@
 ---
-title: "Urban bees have larger brains"
+title: "Bees with larger brains thrive in urban environmnets"
 output:
+  word_document: default
   pdf_document:
     fig_caption: yes
-
 link-citations: yes
@@
 fontsize: 12pt
-
-bibliography: 
-  - references.bib
-
+bibliography: references.bib
 header-includes:
@@
 ```
-
-\textbf{Jose B.
-Lanuza$^{1}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{2}$}  \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
-
-$^{1}$Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
-
-$^{2}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
-
-$^{3}$Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
-
-\textbf{Author for correspondence: barragansljose@gmail.com}
-
-[Not a great fan of the title, maybe something like this?](J)
-[Bees with larger brains relative to their body are more common in urban environments](J)
-
-# Abstract
-
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role in ecosystems for fruit and seed production.
-However, not all pollinators are negatively affected by habitat conversion, but some find in anthropogenic landscapes appropriate resources to survive and reproduce.
-Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities.
-While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for XXX individuals from XX species of bees, the most important group of pollinators, and compared them with detailed information on habitat preferences.
-Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
-In addition, urban bees also tend to be larger, and hence, show larger absolute brain sizes.
-These results suggest that a large brain can help maintaining bee urban populations.
-
-Keywords: relative brain size, habitat preference, Apoidea, Urbanization.
-
-\clearpage
-
-[REFEREBCES TO ADD: https://link.springer.com/article/10.1007/s10980-017-0592-x; https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/1365-2435.12486; https://link.springer.com/article/10.1007/s10531-020-02003-8; https://pubmed.ncbi.nlm.nih.gov/28976620/ ; https://www.biorxiv.org/content/10.1101/2022.12.09.519739v1 (this one for big body size in cities in disc.
-also see https://onlinelibrary.wiley.com/doi/10.1111/eva.13087); https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R (maybe in the intro); https://www.frontiersin.org/articles/10.3389/fevo.2020.00058/full]()
-
-# 1. Introduction
-
-Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
-Regrettably, there are clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
-Although the reason why many pollinators are declining is insufficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
-The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011, @brown2009].
-However, not all pollinator species are equally affected by habitat loss and alteration.
-In bees, for example, there are also "winners" ---like *Bombus terrestris* [@rasmont2008]--- that are able to thrive in human-altered environments, as well as species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; @collado2019].
-
-Human-dominated habitats ---notably cities--- drastically modify the original conditions were pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to enjoy the benefits of an urban life.
-In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
-Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that thrived after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
-[maybe also a sentence on urban dwelers having larger brains?
-Or if there are good references, replace the concept of invaders, by the one on urban species]() However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently [citar a chitka]()[Que refs?](J).
-We now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
-Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not being explored.
-
-Here, we characterize habitat preference for XXX bee species [@collado2019] and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
-In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
-Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [@collado2021].
-Hence, our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas applies also to insects.
-
-\clearpage
-
-# 2. Methods
-
-## Overview
-
-[yo no pondria refs en el parrafo de overview](J)
-
-To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [@sayol2020].
-For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF).
-After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories: (1) natural, (2) agricultural and (3) urban.
-To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [@collado2019].
-Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
-
-All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
-All data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
 
@@
 
+# 1. Introduction
+
+Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
+Regrettably, there are clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
+Although the reason why many pollinators are declining is insufficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
+The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011, @brown2009].
+However, not all pollinator species are equally affected by habitat loss and alteration.
+In bees, for example, there are also "winners" ---like *Bombus terrestris* [@rasmont2008]--- that are able to thrive in human-altered environments, as well as species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; @collado2019].
+
+Human-dominated habitats ---notably cities--- drastically modify the original conditions were pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+The question is then why only some species are able to enjoy the benefits of an urban life.
+In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
+Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
+For example, reptiles, amphibians and birds that thrived after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
+However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently.
+We now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
+Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not being explored.
+
+Here, we characterize habitat preference for `r nlevels(factor(unique(d$Species)))` bee species and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
+In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
+Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [@collado2021].
+Hence, our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas applies also to insects.
+
+\clearpage
+
+# 2. Methods
+
+## Overview
+
+To evaluate our main hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size [@sayol2020].
+For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF).
+After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories: (1) natural, (2) agricultural and (3) urban.
+To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models assuming random distributions across habitats [@collado2019].
+Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and relative brain size changed by habitat type.
+
+All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].
+All data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
+
 ## (a) Bee species dataset
@@
 
-## (d) Land cover classification
+## (c) Land cover classification
 
@@
 
-## (e) Construction of the phylogenetic tree
-
-We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013].
-For this, we first read the backbone phylogeny with the function *read.tree* and drop the missing genera with function *drop.tip* from the package *ape* version 5.6-2 [@paradis2019].
-Then, we selected as an outgroup a wasp genera (*Tachysphex*) and recalculated distances to the root with the function *chronos* from *ape*.
-Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package version 1.2-0 [@revell2012].
-From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
-Finally, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats (see next section) for our set of species with the help of the function *phylosig* from the *phytools* package.
-
-## (f) Analysis
+## (d) Analysis
 
@@
 :::
-
-\clearpage
-
-# Supplementary material
-
-```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
-dplyr::select(species, cover.names) %>% 
-mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-rename(Cover_names = cover.names) %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-       Discarded = "Barren Land", 
-       Agricultural = "Cultivated Crops", 
-       Natural = "Deciduous Forest", 
-       Urban = "Developed, High Intensity", 
-       Urban = "Developed, Medium Intensity",
-       Urban = "Developed, Low Intensity",
-       Urban = "Developed, Open Space", 
-       Natural = "Emergent Herbaceous Wetlands", 
-       Natural = "Evergreen Forest", 
-       Agricultural = "Hay/Pasture", 
-       Natural = "Herbaceous", 
-       Natural = "Mixed Forest", 
-       Discarded = "Open Water", 
-       Natural = "Shrub/Scrub", 
-       Natural = "Woody Wetlands")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.} Original downloaded cover classes for United States with the respective assigned cover classes and number of bee occurrences found on each of them.",
-booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-\clearpage
-
-```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
-    dplyr::select(Species, Cover_names) %>% 
-    mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-      Urban = "Airports", #1
-      Discarded = "Bare rocks", #2
-      Natural = "Beaches, dunes, sands", #3
-      Natural = "Broad-leaved forest", #4
-      Natural = "Coastal lagoons", #5
-      Agricultural = "Complex cultivation patterns", #6
-      Natural = "Coniferous forest", #7
-      Urban = "Construction sites", #8
-      Urban = "Continuous urban fabric", #9
-      Urban = "Discontinuous urban fabric", #10
-      Discarded = "Dump sites", #11
-      Discarded = "Estuaries", #12
-      Agricultural = "Fruit trees and berry plantations", #13
-      Urban = "Green urban areas", #14)
-      Urban = "Industrial or commercial units", #15
-      Natural = "Inland marshes", #16
-      Natural = "Intertidal flats", #17
-      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
-      Discarded = "Mineral extraction sites", #19
-      Natural = "Mixed forest", #20
-      Natural = "Moors and heathland", #21
-      Agricultural = "Natural grasslands", #22
-      Agricultural = "Non-irrigated arable land", #23
-      Agricultural = "Pastures", #24
-      Natural = "Peat bogs", #25
-      Urban = "Port areas", #26) 
-      Discarded = "Road and rail networks and associated land", #27
-      Natural = "Salt marshes", #28
-      Discarded = "Sea and ocean", #29
-      Natural = "Sparsely vegetated areas", #30
-      Urban = "Sport and leisure facilities", #31
-      Natural = "Transitional woodland-shrub", #32
-      Agricultural = "Vineyards", #33
-      Discarded = "Water bodies", #34
-      Discarded = "Water courses" #35  
-)) %>% 
-mutate(Cover_names = fct_recode(as.factor(Cover_names),
-      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.} Original downloaded cover classes for Europe with the respective assigned cover classes and number of bee occurrences found on each of them.",booktabs = T,
-align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.} Association between the log-transformed values of brain and body size (N = 116).
-The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
-
-#Plot encephalization
-#Load libraries
-library(stringr)
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(elementalist) #devtools::install_github("teunbrand/elementalist")
-
-#Read data
-m = read_csv("../Data/Processing/residuals_model.csv")
-r = read_csv("../Data/Processing/residuals_data.csv")
-d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
-mutate(Species = str_replace_all(query, " ", "_")) %>% 
-select(!c(query, db))
-
-#Add taxonomic rank to dataset
-r = left_join(r,d)
-
-#Add ggplot
-ggplot(m, aes(x = log_it, y = estimate__)) +
-geom_line(size=1) +
-geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
-theme_bw() +
-ylab("log(Brain size)") +
-xlab("log(Body size)") +
-theme(legend.position = c(.95, .60),
-legend.justification = c("right", "top"),
-legend.box.just = "right",
-legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
-legend.key = element_rect(fill = "grey85")) +
-guides(colour=guide_legend(title="Subfamily")) +
-theme(axis.title = element_text(face = "bold"))
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
-
-#Histogram of habitat preferences
-#Load libraries
-library(tidyverse)
-
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
-
-#Merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
-#Convert to long to model everything at the same time
-long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
-
-#Relevel for plotting
-long_data = long_data %>% 
-mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
-
-long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
-
-long_data = long_data %>% 
-mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
-
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
-
-ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
-    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
-    xlab("Habitat preference") +
-    ylab("Frequency") +
-    facet_wrap(~ Habitat) +
-    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    theme_bw() +   
-    theme(axis.title = element_text(face = "bold", size = 14),
-    panel.border=element_rect(size=1),
-    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
-    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
-    panel.grid.minor=element_blank())+
-    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
-    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
-                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
-  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.
-Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
-The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}
-
-plot_all = readRDS("../Data/Processing/plot_all.rds")
-
-plot_all
-
-```

### 2023-02-05 Scripts/Google_drive.R

@@ -16,13 +16,13 @@ library(trackdown)
 
 #update_file
 update_file(
-  file = "Manuscript/Lanuza_et_al_2022.Rmd", 
-  gfile = "Reproductive_traits"
+  file = "Manuscript/Ms.Rmd", 
+  gfile = "Bee_brains"
 )
 
 
 #download file 
 download_file(
-  file = "Manuscript/Lanuza_et_al_2022.Rmd", 
-  gfile = "Reproductive_traits"
+  file = "Manuscript/Ms.Rmd", 
+  gfile = "Bee_brains"
 )

### 2023-03-17 Manuscript/Ms.Rmd
Bee brain

@@
 ---
-title: "Bees need larger brains to thrive in urban environmnets"
+title: "Bees need larger brains to thrive in urban environments"
 output:
@@
 Lanuza$^{1}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{2}$}  \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+Collado$^{1,2}$}  \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{3}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
 
@@
 
-$^{3}$Centre de Recerca Ecològica i Aplicacions Forestals: Bellaterra, Catalonia, Spain
+$^{3}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
 
@@
 
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role in ecosystems for fruit and seed production.
-However, not all pollinators are negatively affected by habitat conversion, but some find in anthropogenic landscapes appropriate resources to survive and reproduce.
+The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role as plant pollinators in natural and agricultural systems.
+However, not all pollinators are negatively affected by habitat conversion, as some find in anthropogenic landscapes appropriate resources to survive and reproduce.
 Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities.
-While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and compared them with detailed information on habitat preferences.
+This theory -known as the cognitive buffer hypothesis- has received ample support in studies of vertebrates.
+However, it is less obvious whether the tiny brain of insects can similarly buffer individuals against the many challenges of anthropogenic habitats.
+To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and evaluated the association of their brain size with detailed information on habitat preference.
 Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
-In addition, urban bees also tend to be larger, and hence, show larger absolute brain sizes.
-These results suggest that a large brain can help maintaining bee urban populations.
+In addition, urban bees also tend to have larger body sizes, and hence, show larger absolute brain sizes.
+Our results provide the first empirical support of the cognitive buffer hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
 
@@
 
-[REFEREBCES TO ADD: https://link.springer.com/article/10.1007/s10980-017-0592-x; https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/1365-2435.12486; https://link.springer.com/article/10.1007/s10531-020-02003-8; https://pubmed.ncbi.nlm.nih.gov/28976620/ ; https://www.biorxiv.org/content/10.1101/2022.12.09.519739v1 (this one for big body size in cities in disc.
-also see https://onlinelibrary.wiley.com/doi/10.1111/eva.13087); https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R (maybe in the intro); https://www.frontiersin.org/articles/10.3389/fevo.2020.00058/full]()
-
 # 1. Introduction
@@
 Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
-Regrettably, there are clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
-Although the reason why many pollinators are declining is insufficiently understood, a factor that is considered central is the loss and alteration of their habitats [@winfree2011].
-The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011, @brown2009].
-However, not all pollinator species are equally affected by habitat loss and alteration.
-In bees, for example, there are also "winners" ---like *Bombus terrestris* [@rasmont2008]--- that are able to thrive in human-altered environments, as well as species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; @collado2019].
+Regrettably, there is increasing evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
+One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
+Anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011; @brown2009 ; @harrison2019].
+Yet, not all pollinator species are negatively affected by land use change.
+For instance, some bee species are able to tolerate human-altered environments  [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive such as *Bombus terrestris* [@rasmont2008].
+However, we still lack generalizable information about the functional and ecological consequences of inhabiting these novel environments [@buchholz2020].
 
-Human-dominated habitats ---notably cities--- drastically modify the original conditions were pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to enjoy the benefits of an urban life.
-In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information.
-Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular.
-For example, reptiles, amphibians and birds that thrived after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011].
-[maybe also a sentence on urban dwelers having larger brains?
-Or if there are good references, replace the concept of invaders, by the one on urban species]() However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains until recently [citar a chitka]()[Que refs?](J).
-We now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016].
-Which role plays bees cognitive abilities to adapt to anthropocentric habitats has so far not being explored.
+Human-dominated habitats ---notably cities--- drastically modify the original conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+The question is then why only some species are able to tolerate urban environments.
+The “cognitive buffer hypothesis” provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on cognition [@sol2009], that is, the processes involved in gathering, storing and reacting to environmental information [@shettleworth2009].
+While the “cognitive buffer hypothesis” receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
 
-Here, we characterize habitat preference for `r nlevels(factor(unique(d$Species)))` bee species and relate this to their brain size, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats.
-In vertebrates, relative brain size is a good proxy of problem-solving and learning (REF!!), which can help individuals to cope with the many challenges on novel environments.
-Despite there was no unequivocal evidence that insect brain size is related to cognitive performance, learning abilities across bee species are reflected in differences in their brain size [@collado2021].
-Hence, our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas applies also to insects.
+Insects have been long associated with the most basic types of learning due to their miniature brain [@giurfa2015; @perry2017].
+However, numerous studies have shown that bees and other insects have sophisticated cognition [@chittka2009; @avargues2011; @sheehan2011].
+For instance, some social insects can learn by observing other individuals how to use tools to solve complex tasks [@alem2016].
+In addition, there is also evidence for substantial variation across species in brain size—both in absolute terms and relative to body size— with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
+Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
+
+Here, we report the first test of the “cognitive-buffer hypothesis“ in insects.
+Our test is based on a unique database of brain measures for `r nlevels(factor(unique(d$Species)))` European and North American bee species.
+By means of detailed information of geographic distribution (GBIF; http://www.gbif.org/), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
 
@@
 
+```{r Trait info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+library(dplyr)
+#Load brain data
+brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+#Select unique species
+sp = pref %>% distinct(Species) %>% pull()
+
+#Filter species that are in the phylo
+brain_weight = brain_weight %>% filter(Species %in% sp)
+largest_encephalisation = brain_weight %>% filter(residuals == max(residuals)) %>% select(Species) %>% pull()
+lowest_encephalisation = brain_weight %>% filter(residuals == min(residuals)) %>% select(Species) %>% pull()
+
+#Read model output
+r2 = read_rds("../Data/Processing/bayessianr2.rds")
+r2 = round(r2[1],2)
+
+#Read phylo signal
+phylo = read_rds("../Data/Processing/phylo_signal.rds")
+lambda = round(phylo$lambda, 2)
+phylo_nat = read_rds("../Data/Processing/phylo_signal_natural.rds")
+lambda_nat = round(phylo_nat$lambda, 2)
+p_nat = round(phylo_nat$P,2)
+
+#Read summary of habitat preference
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+#Average preferences between USA and Europe
+preferences = pref %>% 
+group_by(Species) %>%
+summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
+ungroup()
+#Agricultural
+pref_agr = pref %>% 
+dplyr::select(Agricultural) %>%  
+filter(Agricultural > 0.80) %>% 
+nrow()
+#Natural
+pref_nat = pref %>% 
+dplyr::select(Natural) %>%  
+filter(Natural > 0.80) %>% 
+nrow()
+#Urban
+pref_urb1 = pref %>% 
+dplyr::select(Urban) %>%  
+filter(Urban < 0.20) %>% 
+nrow()
+#Urban
+pref_urb2 = pref %>% 
+dplyr::select(Urban) %>%  
+filter(Urban > 0.80) %>% 
+nrow()
+
+#Read R2 MAIN MODEL
+r2_main = readRDS("../Data/Processing/r2_main.rds")
+r2_main = round(r2_main[[1]],2)
+#Read R2 MODEL BRAIN WEIGHT
+r2_brain = readRDS("../Data/Processing/r2_brain.rds")
+r2_brain = round(r2_brain[[1]],2)
+#Read R2 MODEL ITS
+r2_its = readRDS("../Data/Processing/r2_its.rds")
+r2_its = round(r2_its[[1]],2)
+
+```
+
 ## (a) Bee species dataset
 
-Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands).
-In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because: (i) females are involved in a larger number of tasks than males and are likely to experience greater selective pressures from the environment; and (ii), there are evident behavioral and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
-In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers.
-Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.
-Detail information of how brain and body size were measured can be found in supplementary material.
+Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe (Spain and the Netherlands).
+The dataset includes information of `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
+We only considered female specimens because: (i) females are involved in a greater diversity of tasks than males, and hence are expected to experience greater selective pressures from the environment; and (ii), there are evident behavioural and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
+Brain size was measured as the weight of fixed brains and body size as intertegular span (see supplementary text).
+Because brain size scales allometrically with body size, we considered for our analysis the size of the brain relative to the body.
+
+We estimated brain size relative to body size through modelling the log-transformed association between brain weight and intertegular span.
+For this, we specified a Bayesian linear model with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
+To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as a random factor.
+Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the relative brain size.
+High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
+We found that brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`) and that this relationship was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
+However, there were considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 1A)
 
@@
 The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
-Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species.
-For North America, we selected states located on the East Coast of United States (Figure S1A) that covered an approximate area of `r usakm2` km$^{2}$.
-Similarly, we selected countries located on the north and center of Europe (Figure S1B) that represented a total area of `r eukm2` km$^{2}$.
-To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a distribution larger than the area selected; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications, which is from 2006.
+We selected the states or countries with the highest density of records for our set of species.
+For North America, we selected states located on the East Coast of the United States (Figure S2A), covering an approximate area of `r usakm2` km$^{2}$.
+For Europe, we selected countries located on the north and centre of the continent (Figure S2B), representing a total area of `r eukm2` km$^{2}$.
+To further reduce biases in the data, we only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
+In addition, we optimised the match between species occurrence and the land cover data by only using georeferenced records obtained between 1990 and 2022 with a minimum of two decimals of latitude/longitude coordinates.
 
@@
 
-We assigned a habitat type to each GBIF occurrence by merging land cover information with the different georeferenced records.
+We assigned a habitat type to each GBIF occurrence by merging land cover information with the georeferenced records of species occurrences.
 The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
-After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
-Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban.
-The original extracted cover classes and our classification into these three different categories are shown in Table S1 for United States and Table S2 for Europe.
+After downloading these inventories as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
+To simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
 
@@
 From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
-Finally, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats (see next section) for our set of species with the help of the function *phylosig* from the *phytools* package.
 
@@
 
-We estimated brain size relative to body size modeling the log-transformed association between brain weight and intertegular span.
-For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
-To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor (see supplementary text X for phylogenetic tree construction).
-Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`).
-Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the relative brain size.
-High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
+Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
+For this, we generated 10,000 randomised matrices based on the occurrence matrix (i.e., species in rows and habitats in columns) with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
+We implemented this function with the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
+This maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
+We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
+Lastly, we considered that a bee species exhibited a “high preference” for any of the three studied habitats when the number of  occurrences observed in this habitat exceeded  the 80th percentile of the values obtained from the simulations.
+On the contrary, the species was considered to exhibit a “low preference” for the habitat when the observed occurrences were below the 20th percentile.
+To better understand if our findings are affected by the evolutionary histories of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
 
-Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
-For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
-We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
-In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
-Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were below 20th percentile (i.e., low preference) or above the 80th percentile (i.e., high preference) in comparison with the values obtained from the simulations.
-
-To evaluate how the association between habitat preference and relative brain size changed by habitat type, we modeled with Bayesian generalised linear models their association.
+To evaluate how the association between habitat preference and relative brain size changed by habitat type, we modelled with Bayesian generalised linear models their association.
 For this, we first joined the resulting habitat preference and relative brain size datasets by species and extracted those with statistical habitat preference on any of these three habitats.
 Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
-Again, we included the phylogenetic covariance matrix as random factor.
+Again, we included the phylogenetic covariance matrix as a random factor.
 Moreover, we also explored how habitat preference changed by brain weight and intertegular span independently.
 For this, we conducted two analogous models with these two different predictors.
-Finally, we also investigated the different trends for United States and Europe separately.
+In addition, we also investigated the different trends for the United States and Europe separately.
 
-All our models were run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017].
-Finally, the different posterior predictive checks were conducted with the function *pp_check* also from the *brms* package.
-
+All our models were run with 4,000 iterations with previous 1,000 warm up iterations, usingnon-informative or weakly informative priors [@burkner2017].
+Further, the different posterior predictive checks were conducted with the function *pp_check*, also from the *brms* package.
 All our analyses were undertaken in R version 4.0.5 [@r2021], and all data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
@@
 
-```{r Trait info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr)
-#Load brain data
-brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Select unique species
-sp = pref %>% distinct(Species) %>% pull()
-
-#Filter species that are in the phylo
-brain_weight = brain_weight %>% filter(Species %in% sp)
-largest_encephalisation = brain_weight %>% filter(residuals == max(residuals)) %>% select(Species) %>% pull()
-lowest_encephalisation = brain_weight %>% filter(residuals == min(residuals)) %>% select(Species) %>% pull()
-
-#Read model output
-r2 = read_rds("../Data/Processing/bayessianr2.rds")
-r2 = round(r2[1],2)
-
-#Read phylo signal
-phylo = read_rds("../Data/Processing/phylo_signal.rds")
-lambda = round(phylo$lambda, 2)
-phylo_nat = read_rds("../Data/Processing/phylo_signal_natural.rds")
-lambda_nat = round(phylo_nat$lambda, 2)
-p_nat = round(phylo_nat$P,2)
-
-#Read summary of habitat preference
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Average preferences between USA and Europe
-preferences = pref %>% 
-group_by(Species) %>%
-summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-ungroup()
-#Agricultural
-pref_agr = pref %>% 
-dplyr::select(Agricultural) %>%  
-filter(Agricultural > 0.80) %>% 
-nrow()
-#Natural
-pref_nat = pref %>% 
-dplyr::select(Natural) %>%  
-filter(Natural > 0.80) %>% 
-nrow()
-#Urban
-pref_urb1 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban < 0.20) %>% 
-nrow()
-#Urban
-pref_urb2 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban > 0.80) %>% 
-nrow()
-
-#Read R2 MAIN MODEL
-r2_main = readRDS("../Data/Processing/r2_main.rds")
-r2_main = round(r2_main[[1]],2)
-#Read R2 MODEL BRAIN WEIGHT
-r2_brain = readRDS("../Data/Processing/r2_brain.rds")
-r2_brain = round(r2_brain[[1]],2)
-#Read R2 MODEL ITS
-r2_its = readRDS("../Data/Processing/r2_its.rds")
-r2_its = round(r2_its[[1]],2)
-
-```
-
-Our dataset showed that brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`).
-As expected, this allometric relationship was constrained by the evolutive history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
-However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S2; Figure 1A).
-Similarly, habitat preference varied substantially across species (Figure 1B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species preferred one or two habitat types but more rarely occurred indistinctively in the three habitats (Figure 1B).
+Habitat preference varied substantially across species (Figure 1B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three habitats (Figure 1B).
 The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S3).
 In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
-The resulting distribution of habitat preferences for each of the different habitats followed a zero-one inflated beta distribution (Figure S3).
-In other words, there were high frequencies of habitat preferences equal to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3) but note that for our analyses we only modelled the extremes of the distribution (i.e., species with statistical avoidance or preference).
+Hence, there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
 
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
-A) Phylogenetic tree at the genera level.
+A) Phylogenetic tree at the genus level.
 Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).
@@
 
-Our models evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`).
-However, we found differences when we explored this association for each of the habitat types (Figure 2A).
-Specifically, we found that bees with larger brains relative to their body showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
-Contrary, bees with smaller brains relative to their body showed higher preference in natural and agricultural habitats that bees with larger relative brain sizes (Figure 2A).
+Although our model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found evident differences when we explored this association by habitat type (Figure 2A).
+Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
+Contrarily, bees with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
 Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
-Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural one (Figures 2B and 2C).
+Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
 These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S4).
@@
 
-Bees with a preference for inhabiting urban habitats tended to have bigger brains related to their body sizes, and those with forest or agricultural  preferences had relatively smaller brains.
-This pattern is highly suggestive, because investing in brain tissue is expensive (REF), and hence it only payoff in unpredictable habitats where a flexible behavior is needed.
-When the environment is stable, small relative brains are cost-effective to exploit predictable resources, but in highly unpredictable environments, a large brain can confer the needed flexibility to exploit unpredictable resources.
-Indeed, specialist bees also tend to have larger brains [@sayol2020], as they need to locate and memorize patchy resources.
-This same complex learning and navigation skills may be also needed in urban environments.
-Urban habitats are typically harsher, with less and/or new resources to explore, also containing new unknown threats and a highly changing environment [@wheater2002].
-Hence, urban dwellers may need to have better cognitive abilities and flexibility in their behavior than forest or agricultural bees.
+Bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body, and those with forest or agricultural preferences had relatively smaller brains.
+These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that novel environments should select for larger brains.
+In urban environments, individuals are frequently exposed to a variety of new challenges such as novel resources or human disturbances that can change fast in time and space [@wheater2002].
+In such scenarios, a large brain could confer the behavioural flexibility needed to efficiently exploit novel resources or avoid novel predators.
+Our findings are also consistent with previous evidence found in vertebrates that having a large brain size can allow survival in urban environments [@santini2019; @sayol2020b], and suggest that, in bees, urban dwellers may need to have better cognitive abilities and flexibility in their behaviour than forest or agricultural bees.
 
-Some examples of behaviours needed to survive in urban environments are the use of human-made materials to nest [@allasino2019] or being able to collect pollen and nectar from exotic flowers from gardens [@lowenstein2019].
-We should not expect forested habitats to require such a plastic behaviour because the resources found in these habitats are homogeneous, stable, and bees are already pre-adapted to use them [@smith2019].
-In addition, other traits have already been described to be filtered out in cities, such as smaller body sizes (REF), early spring species or brood parasites [@harrison2019].
-We also found small species to be less frequent in urban habitats, maybe as a consequence of the patchy resource distribution, which need from larger body sizes, which confer larger foraging ranges, to be accessed [@kendall2019].
+The exact mechanism that allows large-brained bee species to prefer urban habitats is still unclear.
+However,  there are some examples of novel behaviours that might be needed to survive in urban environments, such as the use of human-made materials to nest [@allasino2019] or being able to collect pollen and nectar from exotic flowers [@lowenstein2019].
+Despite the lack of broad consensus of the functional consequences of urbanisation [@buchholz2020], there is support to the idea that cities select against small body sizes [@ahrne2009; @theodorou2021].
+We found support for this functional consequence with occurrence data from two different continents and over 100 bee species.
+This may be a consequence of the patchy resource distribution as the foraging distance in bees is positively associated with body size [@greenleaf2007; @kendall2019].
+Further, the patchiness of urban resources might also favour species that are already adapted to look for target resources (i.e.
+specialist bees), which tend to have relatively larger brains [@sayol2020].
+However, this hypothesis would need to be properly tested for bees, as there are shreds of evidence that associate increasing urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022].
 
-Despite brain weight is relatively easy to measure and standardize with the used protocol, recent studies challenge the use of brain size as an estimator for behavioral complexity and computational power [@chittka2009\; and references therein].
-However, many studies have shown correlations between brain size and performance [@deaner2007; @gronenberg2010; @sol2008], and brain size is correlated with learning abilities in bees [@collado2021].
-Finer scale measures such as neuropils size or detailed mushroom bodies measurements [the suspected centers of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] could further explain differences in behaviours, and in consequence, shed extra light on species habitat associations.
-In any case, raw measures of brain size have the advantage to allow for larger comparisons across species.
-
-In conclusion, the fact that urban bees tend to have larger relative brain sizes highlights the importance of the cognitive process for understanding the dynamics of insect populations, and stress the importance of considering behavioural responses beyond vertabrate taxa.
+Our analyses focused on brain size mainly for reasons of data availability.
+The use of brain size is based on the existence of previous evidence that enlarged brains enhance cognitive performance in bees [@collado2021] and is justified because the brain is less subject
+to measurement error or context-dependent biases than are experimental measures of cognition [@kamil1987].
+However, the use of brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
+Thus, finer brain measurements like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] are likely to help improve our understanding of bee neurobiology.
+Overall, our findings highlight the importance of the behavioural responses for understanding the dynamics of insect populations in altered environments, and stresses the need to avoid viewing them as passive agents of external factors.
 
@@
 
-JBL, MAC, FS, IB...
+MAC, IB and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC and IB. JBL and IB wrote the manuscript with contributions from all authors.
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116).
+The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
+
+#Plot encephalization
+#Load libraries
+library(stringr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(elementalist) #devtools::install_github("teunbrand/elementalist")
+
+#Read data
+m = read_csv("../Data/Processing/residuals_model.csv")
+r = read_csv("../Data/Processing/residuals_data.csv")
+d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
+mutate(Species = str_replace_all(query, " ", "_")) %>% 
+select(!c(query, db))
+
+#Add taxonomic rank to dataset
+r = left_join(r,d)
+
+#Add ggplot
+ggplot(m, aes(x = log_it, y = estimate__)) +
+geom_line(size=1) +
+geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
+theme_bw() +
+ylab("log(Brain size)") +
+xlab("log(Body size)") +
+theme(legend.position = c(.95, .60),
+legend.justification = c("right", "top"),
+legend.box.just = "right",
+legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
+legend.key = element_rect(fill = "grey85")) +
+guides(colour=guide_legend(title="Subfamily")) +
+theme(axis.title = element_text(face = "bold"))
+
+```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
 A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S2.} Association between the log-transformed values of brain and body size (N = 116).
-The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
-
-#Plot encephalization
-#Load libraries
-library(stringr)
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(elementalist) #devtools::install_github("teunbrand/elementalist")
-
-#Read data
-m = read_csv("../Data/Processing/residuals_model.csv")
-r = read_csv("../Data/Processing/residuals_data.csv")
-d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
-mutate(Species = str_replace_all(query, " ", "_")) %>% 
-select(!c(query, db))
-
-#Add taxonomic rank to dataset
-r = left_join(r,d)
-
-#Add ggplot
-ggplot(m, aes(x = log_it, y = estimate__)) +
-geom_line(size=1) +
-geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
-theme_bw() +
-ylab("log(Brain size)") +
-xlab("log(Body size)") +
-theme(legend.position = c(.95, .60),
-legend.justification = c("right", "top"),
-legend.box.just = "right",
-legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
-legend.key = element_rect(fill = "grey85")) +
-guides(colour=guide_legend(title="Subfamily")) +
-theme(axis.title = element_text(face = "bold"))
-
-```
-
-\clearpage
-
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).

### 2023-04-17 Manuscript/Ms.Rmd
Update on ms from google doc

@@
 However, it is less obvious whether the tiny brain of insects can similarly buffer individuals against the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and evaluated the association of their brain size with detailed information on habitat preference.
-Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
+To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and evaluated the association of their brain sizes with detailed information on habitat preferences.
+Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size than those who prefer urban habitats.
 In addition, urban bees also tend to have larger body sizes, and hence, show larger absolute brain sizes.
@@
 
-Keywords: relative brain size, habitat preference, Apoidea, urbanization.
+Keywords: relative brain size, habitat preference, Apoidea, urbanisation.
+[a]
 
@@
 While the “cognitive buffer hypothesis” receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
+[b]
 
@@
 However, numerous studies have shown that bees and other insects have sophisticated cognition [@chittka2009; @avargues2011; @sheehan2011].
-For instance, some social insects can learn by observing other individuals how to use tools to solve complex tasks [@alem2016].
+For instance, some social insects can learn by observing other individuals how to use tools to solve complex tasks [@alem2016][c].
 In addition, there is also evidence for substantial variation across species in brain size—both in absolute terms and relative to body size— with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
@@
 We only considered female specimens because: (i) females are involved in a greater diversity of tasks than males, and hence are expected to experience greater selective pressures from the environment; and (ii), there are evident behavioural and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains and body size as intertegular span (see supplementary text).
+Brain size was measured as the weight of fixed brains[d] and body size as intertegular[e] span (see supplementary text).
 Because brain size scales allometrically with body size, we considered for our analysis the size of the brain relative to the body.
@@
 In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
-The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3) but note that for our analyses we only modelled the extremes of the distribution (i.e., species with statistical avoidance or preference).
-Hence, there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+Hence, in our analyses we only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
 
@@
 
-Although our model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found evident differences when we explored this association by habitat type (Figure 2A).
+Although our model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found differences when we explored this association by habitat type (Figure 2A).
 Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
 Contrarily, bees with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
-Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
+Although the models of the association between absolute brain size and intertegular span with habitat preference captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found marked differences between habitat types (Figures 2B and 2C).
 Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
@@
 However,  there are some examples of novel behaviours that might be needed to survive in urban environments, such as the use of human-made materials to nest [@allasino2019] or being able to collect pollen and nectar from exotic flowers [@lowenstein2019].
-Despite the lack of broad consensus of the functional consequences of urbanisation [@buchholz2020], there is support to the idea that cities select against small body sizes [@ahrne2009; @theodorou2021].
-We found support for this functional consequence with occurrence data from two different continents and over 100 bee species.
+In addition, despite the lack of broad consensus of the functional consequences of urbanisation @buchholz2020], there is support to the idea that cities select against small body sizes [@ahrne2009; @theodorou2021].
+We also found larger bees in urban areas using data from two different continents and over 100 bee species.
 This may be a consequence of the patchy resource distribution as the foraging distance in bees is positively associated with body size [@greenleaf2007; @kendall2019].
@@
 specialist bees), which tend to have relatively larger brains [@sayol2020].
-However, this hypothesis would need to be properly tested for bees, as there are shreds of evidence that associate increasing urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022].
+However, this hypothesis requires that the needed resources are present.
+In any case, this idea would need to be properly tested for bees, as there are also shreds of evidence that associate increasing urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022].
 
@@
 However, the use of brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
-Thus, finer brain measurements like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] are likely to help improve our understanding of bee neurobiology.
-Overall, our findings highlight the importance of the behavioural responses for understanding the dynamics of insect populations in altered environments, and stresses the need to avoid viewing them as passive agents of external factors.
+Thus, finer brain measurements like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] are likely to help improve our understanding of bee cognitive abilities in relationship to urbanization.
+Overall, our findings highlight the importance of behavioural responses for understanding the dynamics of insect populations in altered environments, and stresses the need to avoid viewing them as passive agents of external factors.
 
@@
 
-# Funding
+# Funding[f][g][h]
 
@@
 ```
+
+[a]Quizá completar el cupo de keywords? propongo Urban habitats, pollinators, cognitive buffer hypothesis.
+[b]Tenemos alguna forma de dar pruebas de esto?
+Sé que demostrar la no-existencia de algo es muy difícil, pero se me queda un poco cojo afirmar esto sin nada más.
+No he encontrado nada.
+Supongo que si hay problemas con esto los revisores lo indicarán.
+[c]El artículo de los bombus jugando llegó al mainstream, quizá sería buena estrategia citarlo, además así tenemos una cita de algo actual!
+
+https://www.sciencedirect.com/science/article/pii/S0003347222002366?via%3Dihub
+[d]cita sayol aqui.
+[e]citar Kendall aqui
+[f]Qué pondrías aquí?
+[g]puedes rescatar los proyectos del mineco que agradece MA en sus papers? Survive_change, creo. Nada más por mi parte.
+[h]@xmiguelangelcolladox@gmail.com podrías rescatar el funding que creas más oportuno?

### 2023-06-12 Manuscript/Ms.Rmd
text

--- 2023-03-17 Manuscript/Ms.Rmd
@@
 output:
+  word_document: default
   pdf_document:
     fig_caption: yes
-
 link-citations: yes
@@
 fontsize: 12pt
-
-bibliography: 
-  - references.bib
-
+bibliography: references.bib
+always_allow_html: true
 header-includes:
@@
 - \floatplacement{figure}{H}
-
 csl: BiologyLetters.csl
-
 ---
@@
 summarise(individuals = n())
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+s = unique(pref$Species)
+
+raw = raw %>% filter(Species %in% s)
 
@@
 
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role as plant pollinators in natural and agricultural systems.
-However, not all pollinators are negatively affected by habitat conversion, as some find in anthropogenic landscapes appropriate resources to survive and reproduce.
-Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities.
-This theory -known as the cognitive buffer hypothesis- has received ample support in studies of vertebrates.
-However, it is less obvious whether the tiny brain of insects can similarly buffer individuals against the many challenges of anthropogenic habitats.
-To address this issue, we measured brains for `r sum(d$individuals)` individuals from `r nlevels(factor(unique(d$Species)))` species of bees, the most important group of pollinators, and evaluated the association of their brain size with detailed information on habitat preference.
-Our analyses revealed that bees that prefer forested or agricultural habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger relative brain size.
-In addition, urban bees also tend to have larger body sizes, and hence, show larger absolute brain sizes.
-Our results provide the first empirical support of the cognitive buffer hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
+The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role as plant pollinators.
+However, not all pollinators are negatively affected by habitat conversion, as certain species find in anthropogenic landscapes appropriate resources to persist and proliferate.
+The reason why some species thrive in anthropogenic environments while most find them inhospitable remains poorly understood.
+The cognitive buffer hypothesis, widely supported in vertebrates but untested in insects, offers a potential explanation.
+This theory  suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
+To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` species of bees, and evaluated the association of their brain sizes with detailed information on habitat preferences.
+Our analyses revealed that bees that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.
+Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
+Our results provide the first empirical support to this hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
 
-Keywords: relative brain size, habitat preference, Apoidea, urbanization.
+Keywords: relative brain size, habitat preference, Apoidea, urbanisation, pollinators
 
@@
 One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction [@winfree2011; @brown2009 ; @harrison2019].
+Anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction by maldaptation [@winfree2011; @brown2009 ; @harrison2019].
 Yet, not all pollinator species are negatively affected by land use change.
-For instance, some bee species are able to tolerate human-altered environments  [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive such as *Bombus terrestris* [@rasmont2008].
-However, we still lack generalizable information about the functional and ecological consequences of inhabiting these novel environments [@buchholz2020].
+For instance, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [e.g., _Bombus terrestris_\; @rasmont2008; @theodorou2020].
 
@@
 Insects have been long associated with the most basic types of learning due to their miniature brain [@giurfa2015; @perry2017].
-However, numerous studies have shown that bees and other insects have sophisticated cognition [@chittka2009; @avargues2011; @sheehan2011].
-For instance, some social insects can learn by observing other individuals how to use tools to solve complex tasks [@alem2016].
+However, numerous recent studies have shown that bees and other insects have sophisticated cognition that goes beyond simple associative learning or conditioning [@chittka2009; @avargues2011; @sheehan2011].
+Some of these more complex forms of cognition involve the use of simple tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
 In addition, there is also evidence for substantial variation across species in brain size—both in absolute terms and relative to body size— with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
@@
 Our test is based on a unique database of brain measures for `r nlevels(factor(unique(d$Species)))` European and North American bee species.
-By means of detailed information of geographic distribution (GBIF; http://www.gbif.org/), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
+By means of detailed georeferenced information of species occurrences (GBIF; http://www.gbif.org/), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
 
@@
 Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe (Spain and the Netherlands).
-The dataset includes information of `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because: (i) females are involved in a greater diversity of tasks than males, and hence are expected to experience greater selective pressures from the environment; and (ii), there are evident behavioural and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains and body size as intertegular span (see supplementary text).
-Because brain size scales allometrically with body size, we considered for our analysis the size of the brain relative to the body.
+The dataset includes information of `r nrow(raw)` female individuals from `r nlevels(factor(unique(raw$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
+We only considered female specimens because: (i) they are involved in a greater diversity of tasks than males, and hence are expected to experience greater environmental pressures; and (ii), have brains that are structurally and functionally different from those of males [@streinzer2013; @roselino2015].
+Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019\; see supplementary text].
+Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
 
-We estimated brain size relative to body size through modelling the log-transformed association between brain weight and intertegular span.
-For this, we specified a Bayesian linear model with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor.
-To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as a random factor.
-Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the relative brain size.
-High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.
-We found that brain size scaled allometrically with body size (Bayesian R$^2$ = `r r2`) and that this relationship was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
-However, there were considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 1A)
+Following @sayol2020, we estimated relative brain size as the residuals of a log-log phylogenetic linear model of brain mass against body size, using the Bayesian approximation implemented in the function *brm* from the package *brms* version 2.18 [@burkner2017].
+The phylogeny was built with the help of a previously published genus-level phylogeny and it was processed with the help of the packages ape 5.6-2 [@paradis2019], phytools 1.2-0 [@revell2012] and MCMCglmm version 2.33 [@hadfield2010].
+High values of relative brain size indicate larger brains than expected for their body size and low values denote smaller brains than expected.
+Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
+However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 1A).
 
@@
 After downloading these inventories as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
-To simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
+To simplify the interpretation and conduct a joint analysis for both regions, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
 
-## (d) Construction of the phylogenetic tree
+## (d) Analysis
 
-We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013].
-For this, we first read the backbone phylogeny with the function *read.tree* and drop the missing genera with function *drop.tip* from the package *ape* version 5.6-2 [@paradis2019].
-Then, we selected as an outgroup a wasp genera (*Tachysphex*) and recalculated distances to the root with the function *chronos* from *ape*.
-Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package version 1.2-0 [@revell2012].
-From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].
-
-## (e) Analysis
-
-Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random.
+Habitat preferences were estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
 For this, we generated 10,000 randomised matrices based on the occurrence matrix (i.e., species in rows and habitats in columns) with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
-We implemented this function with the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
+We used the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
 This maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
@@
 Lastly, we considered that a bee species exhibited a “high preference” for any of the three studied habitats when the number of  occurrences observed in this habitat exceeded  the 80th percentile of the values obtained from the simulations.
-On the contrary, the species was considered to exhibit a “low preference” for the habitat when the observed occurrences were below the 20th percentile.
+On the contrary, the species was considered to exhibit “low preference or avoidance” for the habitat when the observed occurrences were below the 20th percentile.
 To better understand if our findings are affected by the evolutionary histories of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
@@
 For this, we first joined the resulting habitat preference and relative brain size datasets by species and extracted those with statistical habitat preference on any of these three habitats.
+The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+Hence, in our analyses we only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
 Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
@@
 
-All our models were run with 4,000 iterations with previous 1,000 warm up iterations, usingnon-informative or weakly informative priors [@burkner2017].
+All our models were run with 4,000 iterations with previous 1,000 warm up iterations, using non-informative or weakly informative priors [@burkner2017].
 Further, the different posterior predictive checks were conducted with the function *pp_check*, also from the *brms* package.
@@
 The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S3).
-In contrast, most species avoided or had low preference for urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
-The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3) but note that for our analyses we only modelled the extremes of the distribution (i.e., species with statistical avoidance or preference).
-Hence, there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+In contrast, most species avoided urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
 
@@
 
-Although our model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found evident differences when we explored this association by habitat type (Figure 2A).
+Although the model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found evident trends when we explored this association by habitat type (Figure 2A).
 Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
 Contrarily, bees with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
-Although the models of the association between brain size and intertegular span captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively), we found more marked differences in this case between habitat types (Figures 2B and 2C).
+The models of the association between absolute brain size and intertegular span with habitat preference captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
+However, we found again marked differences between habitat types (Figures 2B and 2C).
 Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
@@
 
-Bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body, and those with forest or agricultural preferences had relatively smaller brains.
-These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that novel environments should select for larger brains.
+Using data from two different continents and 89 bee species, we find that bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
+These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to persist and thrive in novel environments.
 In urban environments, individuals are frequently exposed to a variety of new challenges such as novel resources or human disturbances that can change fast in time and space [@wheater2002].
-In such scenarios, a large brain could confer the behavioural flexibility needed to efficiently exploit novel resources or avoid novel predators.
-Our findings are also consistent with previous evidence found in vertebrates that having a large brain size can allow survival in urban environments [@santini2019; @sayol2020b], and suggest that, in bees, urban dwellers may need to have better cognitive abilities and flexibility in their behaviour than forest or agricultural bees.
+In such scenarios, a large brain could confer the cognitive flexibility needed to efficiently exploit novel resources while avoiding risks.
+While the exact mechanisms are unclear, we expect cognitive flexibility to be essential in a variety of contexts such as the use of human-made materials for the nest [@allasino2019] or the need to collect pollen and nectar from exotic flowers [@lowenstein2019].
 
-The exact mechanism that allows large-brained bee species to prefer urban habitats is still unclear.
-However,  there are some examples of novel behaviours that might be needed to survive in urban environments, such as the use of human-made materials to nest [@allasino2019] or being able to collect pollen and nectar from exotic flowers [@lowenstein2019].
-Despite the lack of broad consensus of the functional consequences of urbanisation [@buchholz2020], there is support to the idea that cities select against small body sizes [@ahrne2009; @theodorou2021].
-We found support for this functional consequence with occurrence data from two different continents and over 100 bee species.
-This may be a consequence of the patchy resource distribution as the foraging distance in bees is positively associated with body size [@greenleaf2007; @kendall2019].
-Further, the patchiness of urban resources might also favour species that are already adapted to look for target resources (i.e.
-specialist bees), which tend to have relatively larger brains [@sayol2020].
-However, this hypothesis would need to be properly tested for bees, as there are shreds of evidence that associate increasing urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022].
+Also in line with the cognitive buffer hypothesis [@ducatez2020], we find that bees that proliferate in urban habitats have bigger bodies and larger absolute brains.
+Larger species obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
+The patchiness of urban resources seems indeed to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
+There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion.
+Further understanding all these processes represent an important avenue of future research.
 
+Overall, our findings support and expand previous evidence found in vertebrates that having a large brain size can allow survival in novel environments [@santini2019; @sayol2020b], highlighting that a cognitive buffer is possible even with tiny brains.
+The use of the entire brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
 Our analyses focused on brain size mainly for reasons of data availability.
-The use of brain size is based on the existence of previous evidence that enlarged brains enhance cognitive performance in bees [@collado2021] and is justified because the brain is less subject
-to measurement error or context-dependent biases than are experimental measures of cognition [@kamil1987].
-However, the use of brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
-Thus, finer brain measurements like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] are likely to help improve our understanding of bee neurobiology.
-Overall, our findings highlight the importance of the behavioural responses for understanding the dynamics of insect populations in altered environments, and stresses the need to avoid viewing them as passive agents of external factors.
+The use of brain size is justified by the existence of previous evidence that enlarged brains enhance cognitive performance in bees [@collado2021].
+Moreover, brain size is less subject to measurement error or context-dependent biases in comparison with other experimental measures of cognition [@chittka2009; @lihoreau2012].
+Although the use in the future of finer measures like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] is likely to help improve our understanding of bee cognitive abilities.
+Our findings highlight the importance of behavioural responses for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
 
@@
 # Funding
+
+Ministerio de Economía y Competitividad, Gobierno de España, Grant/Award Number: CGL2013-47448-P
 
@@
 Brain size was measured as the weight of the fixed brain for each specimen.
-First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anesthetized in cold.
+First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anaesthetised in cold.
 Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.

### 2023-06-19 Manuscript/Ms.Rmd
Manuscript updated to its latest version before submission

@@
 Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
-Our results provide the first empirical support to this hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
+Our results provide the first empirical support for the cognitive buffer hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
 
@@
 One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing their risk of extinction by maldaptation [@winfree2011; @brown2009 ; @harrison2019].
+Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maldaptation [@winfree2011; @brown2009 ; @harrison2019].
 Yet, not all pollinator species are negatively affected by land use change.
@@
 
-## (a) Bee species dataset
+## (a) Brain measurements
 
@@
 
-## (b) Occurrence data
+## (b) Habitat preferences
 
@@
 
-We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
+We downloaded occurrence information for all the measured species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
 The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
@@
 
-## (c) Land cover classification
-
 We assigned a habitat type to each GBIF occurrence by merging land cover information with the georeferenced records of species occurrences.
@@
 To simplify the interpretation and conduct a joint analysis for both regions, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
-
-## (d) Analysis
+With this information we can build an occurrence matrix with species in rows, habitats in columns, and cells depicting the number of occurrences per species-habitat combination.
 
 Habitat preferences were estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
-For this, we generated 10,000 randomised matrices based on the occurrence matrix (i.e., species in rows and habitats in columns) with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
+For this, we generated 10,000 randomised matrices based on the occurrence matrix with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
 We used the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
@@
 
+## (c) Analysis
+
 To evaluate how the association between habitat preference and relative brain size changed by habitat type, we modelled with Bayesian generalised linear models their association.
-For this, we first joined the resulting habitat preference and relative brain size datasets by species and extracted those with statistical habitat preference on any of these three habitats.
+For this, we first joined the resulting habitat preference and relative brain size datasets by species.
 The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
-Hence, in our analyses we only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
-Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
+Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
+Because we assessed habitat preference as binary (“avoiding” or “preferring”), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
 Again, we included the phylogenetic covariance matrix as a random factor.
@@
 
-Habitat preference varied substantially across species (Figure 1B) but showed none (i.e., preference for agricultural and rural habitats) or moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+Habitat preference varied substantially across species (Figure 1B) and showed moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
 In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three habitats (Figure 1B).
@@
 B) Heatmap showing the habitat preference for each bee species.
-The columns delimit the habitat type (i.e., natural, agricultural, semideveloped and urban) and the rows the different bee species."}
+The columns delimit the habitat type (i.e., natural, agricultural and urban) and the rows the different bee species."}
 
@@
 
-Although the model evaluating how relative brain size was associated with habitat preference explained little of the overall variance (Bayesian R$^2$ = `r r2_main`), we found evident trends when we explored this association by habitat type (Figure 2A).
+We found that bee species relative brain size is associated with the habitat type preference (Figure 2A; Bayesian R$^2$ = `r r2_main`).
 Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
 Contrarily, bees with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
-The models of the association between absolute brain size and intertegular span with habitat preference captured also little of the overall variance (Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
-However, we found again marked differences between habitat types (Figures 2B and 2C).
+The models of the association between absolute brain size and intertegular span with habitat preference also showed a marked differences between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
 Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
@@
 
-Using data from two different continents and 89 bee species, we find that bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
+Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
 These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to persist and thrive in novel environments.
@@
 Also in line with the cognitive buffer hypothesis [@ducatez2020], we find that bees that proliferate in urban habitats have bigger bodies and larger absolute brains.
-Larger species obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
-The patchiness of urban resources seems indeed to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
+Larger species require longer development times and obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
+For example, the large carpenter bees of the genus Xylocopa can live up to two years and are frequent urban dwellers, while small Andrena forest specialists complete their adult life cycle in a few weeks.
+The patchiness of urban resources also seems to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
 There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion.
@@
 All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf.
-In addition, bee trait data and code can be found in the online repositories of Zenodo XXXXXXXKXX and Github https://github.com/ibartomeus/declines_brain.git.
+In addition, bee trait data and code can be found in the online repositories of Zenodo https://doi.org/10.5281/zenodo.8049996 and Github https://github.com/ibartomeus/declines_brain.git.
 
@@
 
-MAC, IB and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC and IB. JBL and IB wrote the manuscript with contributions from all authors.
+IB, MAC and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC and IB. JBL and IB wrote the manuscript with contributions from all authors.
 
@@
 # Acknowledgements
+
+We thank Curro Molina, Ivo Reamakers and Parker Gambino for collecting some of the studies specimens.
 
@@
 :::
-
-\clearpage
-
-# Supplementary material
-
-## Brain and body size measurements
-
-Brain size was measured as the weight of the fixed brain for each specimen.
-First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anaesthetised in cold.
-Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation.
-Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies.
-By doing this, we ensured that our brain measurements mainly accounted for neural tissue.
-We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues.
-Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®.
-Finally, we calculated the average weight for the species with more than one specimen.
-Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.
-
-We considered as a proxy of body size the average intertegular span (ITS) per species.
-ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight (Cane et al., 2006; Kendall et al., 2019).
-All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
-
-Cane, J.
-H., Minckley, R.
-L., Kervin, L.
-J., Roulston, T.
-A.
-H., & Williams, N.
-M.
-(2006).
-Complex responses within a desert bee guild (Hymenoptera: Apiformes) to urban habitat fragmentation.
-_Ecological applications_, _16_(2), 632-644.
-
-Kendall, L.
-K., Rader, R., Gagic, V., Cariveau, D.
-P., Albrecht, M., Baldock, K.
-C., ...
-& Bartomeus, I.
-(2019).
-Pollinator size and its consequences: Robust estimates of body size in pollinating insects.
-_Ecology and Evolution_, _9_(4), 1702-1714.
-
-```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
-dplyr::select(species, cover.names) %>% 
-mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-rename(Cover_names = cover.names) %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-       Discarded = "Barren Land", 
-       Agricultural = "Cultivated Crops", 
-       Natural = "Deciduous Forest", 
-       Urban = "Developed, High Intensity", 
-       Urban = "Developed, Medium Intensity",
-       Urban = "Developed, Low Intensity",
-       Urban = "Developed, Open Space", 
-       Natural = "Emergent Herbaceous Wetlands", 
-       Natural = "Evergreen Forest", 
-       Agricultural = "Hay/Pasture", 
-       Natural = "Herbaceous", 
-       Natural = "Mixed Forest", 
-       Discarded = "Open Water", 
-       Natural = "Shrub/Scrub", 
-       Natural = "Woody Wetlands")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.} Original downloaded cover classes for United States with the respective assigned cover classes and number of bee occurrences found on each of them.",
-booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-\clearpage
-
-```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}
-
-#Create table of occurrences per habitat type (USA)
-#Load libraries
-library(tidyverse)
-library(kableExtra)
-
-#Load extracted data
-pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
-    dplyr::select(Species, Cover_names) %>% 
-    mutate_if(is.character,as.factor)
-
-#Create table with original and new cover categories
-pref %>% 
-select(Cover_names) %>% 
-group_by(Cover_names) %>% 
-summarise(Occurrences = length(Cover_names)) %>% 
-mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
-      Urban = "Airports", #1
-      Discarded = "Bare rocks", #2
-      Natural = "Beaches, dunes, sands", #3
-      Natural = "Broad-leaved forest", #4
-      Natural = "Coastal lagoons", #5
-      Agricultural = "Complex cultivation patterns", #6
-      Natural = "Coniferous forest", #7
-      Urban = "Construction sites", #8
-      Urban = "Continuous urban fabric", #9
-      Urban = "Discontinuous urban fabric", #10
-      Discarded = "Dump sites", #11
-      Discarded = "Estuaries", #12
-      Agricultural = "Fruit trees and berry plantations", #13
-      Urban = "Green urban areas", #14)
-      Urban = "Industrial or commercial units", #15
-      Natural = "Inland marshes", #16
-      Natural = "Intertidal flats", #17
-      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
-      Discarded = "Mineral extraction sites", #19
-      Natural = "Mixed forest", #20
-      Natural = "Moors and heathland", #21
-      Agricultural = "Natural grasslands", #22
-      Agricultural = "Non-irrigated arable land", #23
-      Agricultural = "Pastures", #24
-      Natural = "Peat bogs", #25
-      Urban = "Port areas", #26) 
-      Discarded = "Road and rail networks and associated land", #27
-      Natural = "Salt marshes", #28
-      Discarded = "Sea and ocean", #29
-      Natural = "Sparsely vegetated areas", #30
-      Urban = "Sport and leisure facilities", #31
-      Natural = "Transitional woodland-shrub", #32
-      Agricultural = "Vineyards", #33
-      Discarded = "Water bodies", #34
-      Discarded = "Water courses" #35  
-)) %>% 
-mutate(Cover_names = fct_recode(as.factor(Cover_names),
-      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
-arrange(desc(Occurrences)) %>%  
-kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.} Original downloaded cover classes for Europe with the respective assigned cover classes and number of bee occurrences found on each of them.",booktabs = T,
-align = c("l", "c", "l"), linesep = "") %>% 
-kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
-row_spec(0,bold=TRUE)
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116).
-The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}
-
-#Plot encephalization
-#Load libraries
-library(stringr)
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(elementalist) #devtools::install_github("teunbrand/elementalist")
-
-#Read data
-m = read_csv("../Data/Processing/residuals_model.csv")
-r = read_csv("../Data/Processing/residuals_data.csv")
-d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
-mutate(Species = str_replace_all(query, " ", "_")) %>% 
-select(!c(query, db))
-
-#Add taxonomic rank to dataset
-r = left_join(r,d)
-
-#Add ggplot
-ggplot(m, aes(x = log_it, y = estimate__)) +
-geom_line(size=1) +
-geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
-theme_bw() +
-ylab("log(Brain size)") +
-xlab("log(Body size)") +
-theme(legend.position = c(.95, .60),
-legend.justification = c("right", "top"),
-legend.box.just = "right",
-legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
-legend.key = element_rect(fill = "grey85")) +
-guides(colour=guide_legend(title="Subfamily")) +
-theme(axis.title = element_text(face = "bold"))
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
-
-#Histogram of habitat preferences
-#Load libraries
-library(tidyverse)
-
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
-
-#Merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
-#Convert to long to model everything at the same time
-long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
-
-#Relevel for plotting
-long_data = long_data %>% 
-mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
-
-long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
-
-long_data = long_data %>% 
-mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
-
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
-
-ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
-    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
-    xlab("Habitat preference") +
-    ylab("Frequency") +
-    facet_wrap(~ Habitat) +
-    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    theme_bw() +   
-    theme(axis.title = element_text(face = "bold", size = 14),
-    panel.border=element_rect(size=1),
-    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
-    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
-    panel.grid.minor=element_blank())+
-    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
-    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
-                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
-  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.
-Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
-The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}
-
-plot_all = readRDS("../Data/Processing/plot_all.rds")
-
-plot_all
-
-```

### 2023-06-19 Manuscript/Supplementary_material.Rmd
Manuscript updated to its latest version before submission

---
output: pdf_document
header-includes:
- \usepackage{orcidlink}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{float}
- \floatplacement{figure}{H}
---

\captionsetup[table]{labelformat=empty}

# Supplementary material

## Brain and body size measurements

Brain size was measured as the weight of the fixed brain for each specimen. First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anaesthetised in cold. Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation. Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies. By doing this, we ensured that our brain measurements mainly accounted for neural tissue. We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues. Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®. Finally, we calculated the average weight for the species with more than one specimen. Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.

We considered as a proxy of body size the average intertegular span (ITS) per species. ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight (Cane et al., 2006; Kendall et al., 2019). All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).

Cane, J. H., Minckley, R. L., Kervin, L. J., Roulston, T. A. H., & Williams, N. M. (2006). Complex responses within a desert bee guild (Hymenoptera: Apiformes) to urban habitat fragmentation. _Ecological applications_, _16_(2), 632-644.

Kendall, L. K., Rader, R., Gagic, V., Cariveau, D. P., Albrecht, M., Baldock, K. C., ... & Bartomeus, I. (2019). Pollinator size and its consequences: Robust estimates of body size in pollinating insects. _Ecology and Evolution_, _9_(4), 1702-1714.

\clearpage

```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}

#Create table of occurrences per habitat type (USA)
#Load libraries
library(tidyverse)
library(kableExtra)

#Load extracted data
pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
dplyr::select(species, cover.names) %>% 
mutate_if(is.character,as.factor)

#Create table with original and new cover categories
pref %>% 
rename(Cover_names = cover.names) %>% 
select(Cover_names) %>% 
group_by(Cover_names) %>% 
summarise(Occurrences = length(Cover_names)) %>% 
mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
       Discarded = "Barren Land", 
       Agricultural = "Cultivated Crops", 
       Natural = "Deciduous Forest", 
       Urban = "Developed, High Intensity", 
       Urban = "Developed, Medium Intensity",
       Urban = "Developed, Low Intensity",
       Urban = "Developed, Open Space", 
       Natural = "Emergent Herbaceous Wetlands", 
       Natural = "Evergreen Forest", 
       Agricultural = "Hay/Pasture", 
       Natural = "Herbaceous", 
       Natural = "Mixed Forest", 
       Discarded = "Open Water", 
       Natural = "Shrub/Scrub", 
       Natural = "Woody Wetlands")) %>% 
arrange(desc(Occurrences)) %>%  
kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.} Original downloaded cover classes for United States with the respective assigned cover classes and number of bee occurrences found on each of them.",
booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
row_spec(0,bold=TRUE)

```

\clearpage

```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}

#Create table of occurrences per habitat type (USA)
#Load libraries
library(tidyverse)
library(kableExtra)

#Load extracted data
pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
    dplyr::select(Species, Cover_names) %>% 
    mutate_if(is.character,as.factor)

#Create table with original and new cover categories
pref %>% 
select(Cover_names) %>% 
group_by(Cover_names) %>% 
summarise(Occurrences = length(Cover_names)) %>% 
mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
      Urban = "Airports", #1
      Discarded = "Bare rocks", #2
      Natural = "Beaches, dunes, sands", #3
      Natural = "Broad-leaved forest", #4
      Natural = "Coastal lagoons", #5
      Agricultural = "Complex cultivation patterns", #6
      Natural = "Coniferous forest", #7
      Urban = "Construction sites", #8
      Urban = "Continuous urban fabric", #9
      Urban = "Discontinuous urban fabric", #10
      Discarded = "Dump sites", #11
      Discarded = "Estuaries", #12
      Agricultural = "Fruit trees and berry plantations", #13
      Urban = "Green urban areas", #14)
      Urban = "Industrial or commercial units", #15
      Natural = "Inland marshes", #16
      Natural = "Intertidal flats", #17
      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
      Discarded = "Mineral extraction sites", #19
      Natural = "Mixed forest", #20
      Natural = "Moors and heathland", #21
      Agricultural = "Natural grasslands", #22
      Agricultural = "Non-irrigated arable land", #23
      Agricultural = "Pastures", #24
      Natural = "Peat bogs", #25
      Urban = "Port areas", #26) 
      Discarded = "Road and rail networks and associated land", #27
      Natural = "Salt marshes", #28
      Discarded = "Sea and ocean", #29
      Natural = "Sparsely vegetated areas", #30
      Urban = "Sport and leisure facilities", #31
      Natural = "Transitional woodland-shrub", #32
      Agricultural = "Vineyards", #33
      Discarded = "Water bodies", #34
      Discarded = "Water courses" #35  
)) %>% 
mutate(Cover_names = fct_recode(as.factor(Cover_names),
      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
arrange(desc(Occurrences)) %>%  
kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.} Original downloaded cover classes for Europe with the respective assigned cover classes and number of bee occurrences found on each of them.",booktabs = T,
align = c("l", "c", "l"), linesep = "") %>% 
kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
row_spec(0,bold=TRUE)

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116). The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}

#Plot encephalization
#Load libraries
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(elementalist) #devtools::install_github("teunbrand/elementalist")

#Read data
m = read_csv("../Data/Processing/residuals_model.csv")
r = read_csv("../Data/Processing/residuals_data.csv")
d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
mutate(Species = str_replace_all(query, " ", "_")) %>% 
select(!c(query, db))

#Add taxonomic rank to dataset
r = left_join(r,d)

#Add ggplot
ggplot(m, aes(x = log_it, y = estimate__)) +
geom_line(size=1) +
geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
theme_bw() +
ylab("log(Brain size)") +
xlab("log(Body size)") +
theme(legend.position = c(.95, .60),
legend.justification = c("right", "top"),
legend.box.just = "right",
legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
legend.key = element_rect(fill = "grey85")) +
guides(colour=guide_legend(title="Subfamily")) +
theme(axis.title = element_text(face = "bold"))

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species. A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}

#Plot maps with density of bee occurrences

#Europe----
#Load libraries
library(giscoR) #map of europe
library(ggplot2)
library(dplyr)
library(readr)
library(sf)
library(patchwork)
library(rnaturalearth)

#Load europe map
Europe <- gisco_get_countries(region = "Europe")
#Load countries filtered
euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Load data with records filtered
all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")

#Plot Europe map with counties of interest highlighted
p1 = ggplot(Europe) +
geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white"))

p2 = ggplot(Europe) +
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)

#Insert general plot within specific plot
panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)

#USA-----
#Load libraries

#Load data
usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Load maps
canada <- ne_states(country = "canada", returnclass = "sf") %>%
st_as_sf(crs = 4326)

usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
st_as_sf(crs = 4326)

p1 = ggplot(canada) + 
geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white")) +
geom_map(data = ny_metropolitan_1, 
map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
size = 0.1) +
geom_map(data = usa_states_1, map = usa_states_1,
aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
size = 0.1) +
coord_sf(xlim = c(-86, -62), ylim = c(25, 48))

p2 = ggplot(canada) + 
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))

#Insert general plot within specific plot
panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)

library(patchwork)
panel2  + panel1

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban). The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}

#Histogram of habitat preferences
#Load libraries
library(tidyverse)

#Load preferences
preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")

#Merge datasets
preferences = bind_rows(preferences_europe, preferences_usa)

#Convert to long to model everything at the same time
long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))

#Relevel for plotting
long_data = long_data %>% 
mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))

labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")

long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))

long_data = long_data %>% 
mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))

long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))

ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
    xlab("Habitat preference") +
    ylab("Frequency") +
    facet_wrap(~ Habitat) +
    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    theme_bw() +   
    theme(axis.title = element_text(face = "bold", size = 14),
    panel.border=element_rect(size=1),
    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
    panel.grid.minor=element_blank())+
    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}. Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one). The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}

plot_all = readRDS("../Data/Processing/plot_all.rds")

plot_all

```

### 2023-06-19 Scripts/Google_drive.R

@@ -1,16 +1,17 @@
 #Upload files to google drive
 #install packages
-remotes::install_github("claudiozandonella/trackdown",
-                        build_vignettes = TRUE, force=T)
+#remotes::install_github("claudiozandonella/trackdown",
+#                        build_vignettes = TRUE, force=T)
 
-install.packages("googledrive")
+#install.packages("googledrive")
+
+
+
+library(googledrive)
+library(trackdown)
 
 #set google doc
 googledrive::drive_auth()
-
-#load libraries
-library(googledrive)
-library(trackdown)
 
 #upload file
 

### 2023-06-24 Manuscript/Ms.Rmd
Edits on the ms to submit word doc

@@
 
+Jose B.
+Lanuza$^{1}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
+Collado$^{1,2}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{3}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{3}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
+
 $^{1}$Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
@@
 $^{3}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
+
+**Author for correspondence: barragansljose\@gmail.com**
 
@@
 Yet, not all pollinator species are negatively affected by land use change.
-For instance, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [e.g., _Bombus terrestris_\; @rasmont2008; @theodorou2020].
+For instance, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [e.g., *Bombus terrestris*\; @rasmont2008; @theodorou2020].
 
@@
 The question is then why only some species are able to tolerate urban environments.
-The “cognitive buffer hypothesis” provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on cognition [@sol2009], that is, the processes involved in gathering, storing and reacting to environmental information [@shettleworth2009].
-While the “cognitive buffer hypothesis” receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
+The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on cognition [@sol2009], that is, the processes involved in gathering, storing and reacting to environmental information [@shettleworth2009].
+While the "cognitive buffer hypothesis" receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
 
@@
 Some of these more complex forms of cognition involve the use of simple tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
-In addition, there is also evidence for substantial variation across species in brain size—both in absolute terms and relative to body size— with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
+In addition, there is also evidence for substantial variation across species in brain size---both in absolute terms and relative to body size--- with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
 Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
 
-Here, we report the first test of the “cognitive-buffer hypothesis“ in insects.
-Our test is based on a unique database of brain measures for `r nlevels(factor(unique(d$Species)))` European and North American bee species.
-By means of detailed georeferenced information of species occurrences (GBIF; http://www.gbif.org/), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
+Here, we report the first test of the "cognitive-buffer hypothesis" in insects.
+Our test is based on a unique database of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
+By means of detailed georeferenced information of species occurrences (GBIF; <http://www.gbif.org/>), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
 
@@
 We only considered female specimens because: (i) they are involved in a greater diversity of tasks than males, and hence are expected to experience greater environmental pressures; and (ii), have brains that are structurally and functionally different from those of males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019\; see supplementary text].
+Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; see supplementary text].
 Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
@@
 High values of relative brain size indicate larger brains than expected for their body size and low values denote smaller brains than expected.
-Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* < 0.01).
+Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* \< 0.01).
 However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 1A).
@@
 
-We downloaded occurrence information for all the measured species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe.
+We downloaded occurrence information for all the measured species from the Global Biodiversity Information Facility (GBIF; <http://www.gbif.org/>) for North America and Europe.
 The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
@@
 We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
-Lastly, we considered that a bee species exhibited a “high preference” for any of the three studied habitats when the number of  occurrences observed in this habitat exceeded  the 80th percentile of the values obtained from the simulations.
-On the contrary, the species was considered to exhibit “low preference or avoidance” for the habitat when the observed occurrences were below the 20th percentile.
+Lastly, we considered that a bee species exhibited a "high preference" for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
+On the contrary, the species was considered to exhibit "low preference or avoidance" for the habitat when the observed occurrences were below the 20th percentile.
 To better understand if our findings are affected by the evolutionary histories of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
 
-## (c) Analysis
+## (d) Analysis
 
@@
 Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
-Because we assessed habitat preference as binary (“avoiding” or “preferring”), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
+Because we assessed habitat preference as binary ("avoiding" or "preferring"), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
 Again, we included the phylogenetic covariance matrix as a random factor.
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 1.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="Figure 1.
+Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
 A) Phylogenetic tree at the genus level.
-Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions).
+Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of the United States or from both regions).
 The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted points at the end of the tip branches.
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 2.} Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="Figure 2.
+Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
 The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p"}
@@
 Larger species require longer development times and obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
-For example, the large carpenter bees of the genus Xylocopa can live up to two years and are frequent urban dwellers, while small Andrena forest specialists complete their adult life cycle in a few weeks.
+For example, the large carpenter bees of the genus _Xylocopa_ can live up to two years and are frequent urban dwellers, while small _Andrena_ forest specialists complete their adult life cycle in a few weeks.
 The patchiness of urban resources also seems to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
@@
 
-All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.5s5kuf.
-In addition, bee trait data and code can be found in the online repositories of Zenodo https://doi.org/10.5281/zenodo.8049996 and Github https://github.com/ibartomeus/declines_brain.git.
+All occurrence data used in this study can be downloaded from <https://doi.org/10.15468/dl.5s5kuf>.
+In addition, bee trait data and code can be found in the online repositories of Zenodo <https://doi.org/10.5281/zenodo.8049996> and Github <https://github.com/ibartomeus/declines_brain.git>.
 
@@
 
-Ministerio de Economía y Competitividad, Gobierno de España, Grant/Award Number: CGL2013-47448-P
+Ministerio de Economía y Competitividad, Gobierno de España, Grant/Award Number: CGL2013-47448-P.
 

### 2023-06-27 Manuscript/Ms.Rmd
Minor updates before submission

@@
 output:
-  word_document: default
   pdf_document:
     fig_caption: yes
+  word_document: default
 link-citations: yes
@@
 This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` species of bees, and evaluated the association of their brain sizes with detailed information on habitat preferences.
+To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the association of their brain sizes with detailed information on habitat preferences.
 Our analyses revealed that bees that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.

### 2023-09-01 Scripts/Lecticity.R
Minor update on ms (fig 1 caption)

@@ -1,11 +1,20 @@
+#To answer one of the issues raised by a reviewer
+#We explore how the bee's diet (oligo/polylectic)
+#is associated with habitat preference (1) and brain size (2)
+
+#The data of the diet comes from the paper of Ferran
+#but because is missing for some species
+#I have look the diets for missing ones (~20spp)
+
 #Load libraries
 library(dplyr)
 library(stringr)
+library(ggplot2)
+library(patchwork)
+library(readr)
 
-#Load phylogenetic data from europe
-bee.tree100 = readRDS("Data/Europe_USA/bee.tree100.rds")
-#Load brain data
-brain_weight = readr::read_csv("Data/Processing/brain_weight_data.csv")
+# (1) Lecticity ~ habitat preference ----
+#First we explore this association with habitat preference
 #Load preferences data
 preferences_usa = readr::read_csv("Data/USA_data/preferences_usa.csv") 
 preferences_eu = readr::read_csv("Data/Europe_data/preferences_europe.csv") 
@@ -17,14 +26,12 @@ unique(pref$Species)
 
 #Average preferences between USA and Europe
 preferences = pref %>% 
-  group_by(Species) %>%
-  summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-  ungroup()
-
+group_by(Species) %>%
+summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
+ungroup()
 
 #Read lecticity
-lecticity = readr::read_csv("Data/Bees_Lecticity.csv") 
-
+lecticity = read_csv("Data/Bees_Lecticity.csv") 
 #Delete underscore from tree
 lecticity$species = str_replace(lecticity$species, "_", " ")
 lecticity = lecticity %>% rename(Species = species)
@@ -144,5 +151,74 @@ mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic
 d = d %>% 
 mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))
 
+d %>% 
+ggplot(aes(Lecticity, Agricultural)) +
+geom_violin()
 
 
+#Natural
[…]

### 2023-09-03 Manuscript/Supplementary_material.Rmd
Work in progress for the review of Biol Lett

@@
 ```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure SX}. ", fig.pos="p", out.width="100%"}
+
+#To answer one of the issues raised by a reviewer
+#We explore how the bee's diet (oligo/polylectic)
+#is associated with habitat preference (1) and brain size (2)
+
+#The data of the diet comes from the paper of Ferran
+#but because is missing for some species
+#I have look the diets for missing ones (~20spp)
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(patchwork)
+library(readr)
+
+# (1) Lecticity ~ habitat preference ----
+#First we explore this association with habitat preference
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+
+#Average preferences between USA and Europe
+preferences = pref %>% 
+group_by(Species) %>%
+summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
+ungroup()
+
+#Read lecticity
+lecticity = read_csv("../Data/Bees_Lecticity.csv") 
+#Delete underscore from tree
+lecticity$species = str_replace(lecticity$species, "_", " ")
+lecticity = lecticity %>% rename(Species = species)
+
+#Bind datasets
+d = left_join(preferences, lecticity)
+
+#Check species with lacking info
+#Agapostemon virescens
+#https://lopezuribelab.com/halictidae/agapostemon-virescens/#:~:text=Agapostemon%20virescens%20is%20a%20polylectic,coneflowers%2C%20asters%2C%20and%20goldenrod.
+#As the other Agapostemon, they tend to visit many unrelated plant species
+#but they tend to prefer Asteraceae flowers
+#Polylectic
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Agapostemon virescens", "Polylectic"))
+#Andrena angustior
+#http://dx.doi.org/10.1016/j.biocon.2017.09.009
+#Polylectic
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena angustior", "Polylectic"))
+#Andrena hirticincta
+#https://www.sharpeatmanguides.com/andrena-mining-bees
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena hirticincta", "Oligolectic"))
+#Andrena milwaukeensis
+#https://www.discoverlife.org/20/q?search=Andrena+milwaukeensis
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena milwaukeensis", "Polylectic"))
+#Andrena pilipes
+#https://www.wildbienen.de/eb-apili.htm
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena pilipes", "Polylectic"))
+#Andrena vicina
+#https://www.discoverlife.org/20/q?search=Andrena+vicina
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena vicina", "Polylectic"))
+#Andrena wilkella
+##http://dx.doi.org/10.1016/j.biocon.2017.09.009
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Andrena wilkella", "Oligolectic"))
+
+#Augochlorella aurata
+#https://www.beesofcanada.com/species/augochlorella-aurata
+#https://www.discoverlife.org/20/q?search=Augochlorella+aurata
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Augochlorella aurata", "Polylectic"))
+
+#Augochloropsis metallica
+#https://www.gbif.org/species/1353091
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Augochloropsis metallica", "Polylectic"))
+
+#Bombus griseocollis
+#https://doi.org/10.1002/ecy.2697
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Bombus griseocollis", "Polylectic"))
+
+#Calliopsis andreniformis
+#https://www.sharpeatmanguides.com/calliopsis-andreniformis
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Calliopsis andreniformis", "Polylectic"))
+
+#Ceratina strenua
+#https://www.beesofcanada.com/species/ceratina-strenua
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Ceratina strenua", "Polylectic"))
+
+#Colletes thoracicus
+#https://doi.org/10.1073/pnas.1218503110
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Colletes thoracicus", "Polylectic"))
+
+#Osmia atriventris
+#https://doi.org/10.1073/pnas.1218503110
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Osmia atriventris", "Polylectic"))
+
+#Osmia lignaria
+#https://doi.org/10.1073/pnas.1218503110
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Osmia lignaria", "Polylectic"))
+
+#Osmia pumila
+#https://doi.org/10.1073/pnas.1218503110
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Osmia pumila", "Polylectic"))
+
+#Osmia cornifrons
+#http://dx.doi.org/10.1098/rsos.200225
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Osmia cornifrons", "Polylectic"))
+
+#Lasioglossum malachurum
+#https://www.wildbienen.de/eb-lmala.htm
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Lasioglossum malachurum", "Polylectic"))
+
+#Megachile texana
+#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))
+
+#Megachile texana
+#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))
+
+#Megachile willughbiella
+#https://www.wildbienen.de/eb-mwill.htm
+d = d %>% 
+mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))
+
+#Natural
+p1_1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab("Habitat Preference")+
+xlab(NULL)+
+ggtitle("Natural")+
+theme_bw()
+
+#Agricultural
+p1_2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab(NULL)+
+xlab("Diet")+
+ggtitle("Agricultural")+
+theme_bw()
+
+#Urban
+p1_3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab(NULL)+
+xlab(NULL)+
+ggtitle("Urban")+
+theme_bw()
+
+panel1 = p1_1+p1_2+p1_3
+
+# (2) Lecticity ~ brain size ----
+#Load brain data
+brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
+
+d1 = left_join(brain_weight, d) %>% 
+filter(!is.na(Lecticity))
+
+p2_1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab("Relative brain size")+
+xlab("Diet")+
+theme_bw()
+
+p2_2 = ggplot(d1, aes(x = Lecticity, y = Brain.weight)) + 
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab("Absolute brain size")+
+xlab("Diet")+
+theme_bw()
+
+panel2 =  p2_2 + p2_1
+
+library(cowplot)
+
+plot_grid(plot_grid(panel1, labels = c('A'))+
+  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)),
+
+plot_grid(panel2, labels = c('B'))+
+  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)), nrow = 2)
+
+```

### 2023-09-03 Scripts/Lecticity.R

@@ -157,32 +157,33 @@ geom_violin()
 
 
 #Natural
-p1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
+p1_1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
 geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
 geom_boxplot(colour="black",fill="grey",
 width = .15, 
 outlier.shape = NA) +
 geom_jitter(alpha=0.3,width = 0.1) +
-ylab("Preference")+
+ylab("Habitat Preference")+
 xlab(NULL)+
 ggtitle("Natural")+
-theme_bw()
+theme_bw() 
 
 
 #Agricultural
-p2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
+p1_2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
 geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
 geom_boxplot(colour="black",fill="grey",
 width = .15, 
 outlier.shape = NA) +
 geom_jitter(alpha=0.3,width = 0.1) +
 ylab(NULL)+
+xlab("Diet")+
 ggtitle("Agricultural")+
 theme_bw()
 
 
 #Urban
-p3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
+p1_3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
 geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
 geom_boxplot(colour="black",fill="grey",
 width = .15, 
@@ -193,7 +194,8 @@ xlab(NULL)+
 ggtitle("Urban")+
 theme_bw()
 
-p1+p2+p3
+panel1 = p1_1+p1_2+p1_3 
+
 
 # (2) Lecticity ~ brain size ----
 #Load brain data
@@ -202,23 +204,35 @@ brain_weight = readr::read_csv("Data/Processing/brain_weight_data.csv")
 d1 = left_join(brain_weight, d) %>% 
 filter(!is.na(Lecticity))
 
-p1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
+p2_1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
 geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
 geom_boxplot(colour="black",fill="grey",
 width = .15, 
[…]

### 2023-09-06 Manuscript/Ms.Rmd
Minor update on ms (fig 1 caption)
Response to R

@@
 fontsize: 12pt
+includeline-num: true
 bibliography: references.bib
@@
 - \floatplacement{figure}{H}
+- \usepackage[left]{lineno}
+- \linenumbers
 csl: BiologyLetters.csl
@@
 This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the association of their brain sizes with detailed information on habitat preferences.
-Our analyses revealed that bees that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.
+To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the species-level association between brain size and habitat preferences.
+Our analyses revealed that bee species that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.
 Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
@@
 One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maldaptation [@winfree2011; @brown2009 ; @harrison2019].
+Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maladaptation [@winfree2011; @brown2009 ; @harrison2019].
 Yet, not all pollinator species are negatively affected by land use change.
@@
 
-Human-dominated habitats ---notably cities--- drastically modify the original conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to tolerate urban environments.
+Human-dominated habitats ---notably cities--- drastically modify the original conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+The question is then why only some species are able to tolerate and exploit urban environments.
 The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on cognition [@sol2009], that is, the processes involved in gathering, storing and reacting to environmental information [@shettleworth2009].
@@
 Some of these more complex forms of cognition involve the use of simple tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
-In addition, there is also evidence for substantial variation across species in brain size---both in absolute terms and relative to body size--- with species that have larger brains also exhibiting enhanced cognitive performance [@collado2021].
+In addition, there is also evidence for substantial variation across species in brain size---both in absolute terms and relative to body size--- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
 Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
@@
 Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe (Spain and the Netherlands).
+These specimens were collected opportunistically mostly in semi-natural habitats.
 The dataset includes information of `r nrow(raw)` female individuals from `r nlevels(factor(unique(raw$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
 We only considered female specimens because: (i) they are involved in a greater diversity of tasks than males, and hence are expected to experience greater environmental pressures; and (ii), have brains that are structurally and functionally different from those of males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; see supplementary text].
+Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; see supplementary text S1].
 Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
@@
 
-## (b) Habitat preferences
+## (b) Diet specialisation
+
+To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with both habitat preference and brain size.
+For this, we used the diet information collected in @sayol2020 and complemented the missing information by using the same online resources used in @sayol2020.
+Bee species were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
+We tested for statistical differences in diet specialization per habitat type preference and brain size (i.e., both absolute and relative brain size) by using the Wilcoxon test.
+We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (see details in Figure S2).
+
+## (c) Habitat preferences
 
@@
 We selected the states or countries with the highest density of records for our set of species.
-For North America, we selected states located on the East Coast of the United States (Figure S2A), covering an approximate area of `r usakm2` km$^{2}$.
-For Europe, we selected countries located on the north and centre of the continent (Figure S2B), representing a total area of `r eukm2` km$^{2}$.
+For North America, we selected states located on the East Coast of the United States (Figure S3A), covering an approximate area of `r usakm2` km$^{2}$.
+For Europe, we selected countries located on the north and centre of the continent (Figure S3B), representing a total area of `r eukm2` km$^{2}$.
 To further reduce biases in the data, we only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
@@
 We assigned a habitat type to each GBIF occurrence by merging land cover information with the georeferenced records of species occurrences.
-The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe.
+The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for the United States and the Corine Land Cover (CLC) for Europe.
 After downloading these inventories as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
@@
 
-To evaluate how the association between habitat preference and relative brain size changed by habitat type, we modelled with Bayesian generalised linear models their association.
-For this, we first joined the resulting habitat preference and relative brain size datasets by species.
-The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S3), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+Our macro-ecological approach is justified by: (i) independent data that ensures that any ecological patterns, if present, are likely to be robust and generalizable; (ii) low intraspecific variation of brain and body size in comparison to the interspecific variation (See supplementary text S2 and Figure S4) that provides biological meaning of averaging relative brain sizes per species and the study of the ecological patterns across them.
+
+To evaluate how the association between habitat preference and relative brain size per species changed by habitat type, we used a Bayessian approach to model their association.
+For this, we first joined the resulting habitat preferences per species with their respective average relative brain sizes.
+The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
 Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
@@
 Again, we included the phylogenetic covariance matrix as a random factor.
-Moreover, we also explored how habitat preference changed by brain weight and intertegular span independently.
+Moreover, we also explored how habitat preference changed by average brain weight and intertegular span independently.
 For this, we conducted two analogous models with these two different predictors.
@@
 In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three habitats (Figure 1B).
-The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S3).
-In contrast, most species avoided urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S3).
+The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
+In contrast, most species avoided urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S5).
 
@@
 Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of the United States or from both regions).
-The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted points at the end of the tip branches.
-Larger points indicate larger brains in proportion to their body size and vice versa.
+The deviation of the brain size in relation to the body (i.e., residuals) is represented with filled circles of proportional area to the residuals at the end of the tip branches.
+Larger circles indicate larger brains in proportion to their body size and vice versa.
 B) Heatmap showing the habitat preference for each bee species.
@@
 Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
-Contrarily, bees with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
+Contrarily, bee species with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
 The models of the association between absolute brain size and intertegular span with habitat preference also showed a marked differences between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
 Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
-These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S4).
+These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S6).
 
@@
 
-Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bees with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
+Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bee species with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
 These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to persist and thrive in novel environments.
@@
 There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion.
-Further understanding all these processes represent an important avenue of future research.
+Only a few bee specialists are analyzed in this paper, and further understanding how different life styles facilitate urban exploitation represent an important avenue of future research.
 
@@
 Our analyses focused on brain size mainly for reasons of data availability.
-The use of brain size is justified by the existence of previous evidence that enlarged brains enhance cognitive performance in bees [@collado2021].
+The use of brain size is justified by the existence of previous evidence that enlarged brains enhance some aspects of cognitive performance in bees, such as learning [@collado2021].
 Moreover, brain size is less subject to measurement error or context-dependent biases in comparison with other experimental measures of cognition [@chittka2009; @lihoreau2012].
 Although the use in the future of finer measures like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] is likely to help improve our understanding of bee cognitive abilities.
-Our findings highlight the importance of behavioural responses for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
+Further, there are other alternative explanations beyond the cognitive buffer hypothesis for the enlargement of brains in urban environments.
+Here, we only investigated the role of diet specialisation as a potential confounding factor but other ecological and physiological factors cannot be discarded as potential drivers of this pattern.
+Downscaling our analysis to the individual specimen level where brain size, habitat use and cognitive performance can be tracked through their lifespan would be a challenging but promising next step.
+Overall, our findings highlight the importance of behavioural responses for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
 
@@
 
-We thank Curro Molina, Ivo Reamakers and Parker Gambino for collecting some of the studies specimens.
+We thank Fransico P. Molina, Ivo Reamakers and Parker Gambino for collecting some of the studies specimens.
 

### 2023-09-06 Manuscript/Supplementary_material.Rmd
Response to R

@@
 - \floatplacement{figure}{H}
+
 ---
@@
 
-# Supplementary material
+# Supplementary material for: Bees need larger brains to thrive in urban environments
 
-## Brain and body size measurements
+## Authors: Jose B. Lanuza, Miguel Á. Collado, Ferran Sayol, Daniel Sol and Ignasi Bartomeus
+
+## Supplementary text S1: Brain and body size measurements
 
@@
 All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).
+
+**References**:
 
@@
 _Ecology and Evolution_, _9_(4), 1702-1714.
+
+\clearpage
+
+## Supplementary text S2: Intra- and inter-specific brain body size variation
+
+In average we had 3.31 individuals measured per species with information for both brain and body size.
+These different individuals within a species tended to have similar values of brain and body size (i.e., low variance) in comparison with the variability found across species.
+To show this pattern, we calculated the Intraclass Correlation Coefficient (ICC) that estimates how similar or different are the values within the different groups (i.e., species).
+The ICC values range from 0 to 1.
+For our case, values close to 0 indicate that the different individual measurements per species tend to be very different to each other and values close to 1 indicate that the measurements within a species tend to be very similar or identical.
+In order to compute the ICC, we first fitted an intercept-only linear model with the help of the *lmer* function from the *lme4* library version 1.10-34 (Bates, 2010) where we only included a response variable of interest (i.e., brain or body size) and species as a random effect.
+This allowed us to calculate the variances of interests in order to compute the ICC.
+We calculated the ICC with the help of the *icc* function from the library *performance* version 0.10.3 (Lüdecke et al., 2021).
+This function computes the division between the random effect variance and the total variance (i.e., the sum of the random effect variance and the residual variance).
+We found that for both brain and body size the individual measurements per species tended to be very similar between each other (ICC = 0.947; ICC ICC = 0.931, respectively) but we found high variability across species (see Figure S4).
+Overall, this indicates that the mean values of brain and body size have biological significance and provides support for investigating macro-ecological patterns across species.
+
+**References**:
+
+Bates, D. M. (2010). lme4: Mixed-effects modeling with R.
+
+Lüdecke, D., Ben-Shachar, M.
+S., Patil, I., Waggoner, P., & Makowski, D.
+(2021).
+performance: An R package for assessment, comparison and testing of statistical models.
+Journal of Open Source Software, 6(60).
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
-
-#Histogram of habitat preferences
-#Load libraries
-library(tidyverse)
-
-#Load preferences
-preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
-preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
-
-#Merge datasets
-preferences = bind_rows(preferences_europe, preferences_usa)
-
-#Convert to long to model everything at the same time
-long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
-
-#Relevel for plotting
-long_data = long_data %>% 
-mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
-
-long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
-
-long_data = long_data %>% 
-mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
-
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
-
-ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
-    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
-    xlab("Habitat preference") +
-    ylab("Frequency") +
-    facet_wrap(~ Habitat) +
-    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
-    theme_bw() +   
-    theme(axis.title = element_text(face = "bold", size = 14),
-    panel.border=element_rect(size=1),
-    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
-    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
-    panel.grid.minor=element_blank())+
-    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
-    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
-    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
-    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
-                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
-  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.
-Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
-The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}
-
-plot_all = readRDS("../Data/Processing/plot_all.rds")
-
-plot_all
-
-```
-
-\clearpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure SX}. ", fig.pos="p", out.width="100%"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2}.
+Graphical representation and statistical differences of diet specialisation (i.e., oligolectic and polylectic species) by habitat preference per habitat type (Panel A) and brain size including absolute and relative brain size (Panel B).
+The graphical representation shows boxplots, raw values and a shaded violin plot to allow a better understanding of the spread, density and number of values by each category.
+Note that the raw values are jittered horizontally in order to allow a better visualisation of overlapping points.
+The significance of the statistical differences across values are shown with the resulting P-values of a Wilcoxon test.
+Interestingly, we see that (i) there are few oligolectic species represented in our dataset; (ii) there are no statistical differences between oligolectic and polylectic species in habitat preference (Panel A) and brain size (Panel B); and, (iii) oligolectic species tend to have larger relative brains and, if anything, show lower preferences by urban habitats, indicating that feeding specialisation is unlikely to be driving the observed association between urban preference and brain size.", fig.pos="p", out.width="100%"}
 
@@
 
-#Natural
+#Plotting----
+#Here we load some libraries to include Wilcoxon rank test in the plot
+#for publication
+library(ggpubr)
+library(rstatix)
+
+#First graph Natural habitat
+stat.test1 <- d %>%
+wilcox_test(Natural ~ Lecticity) %>%
+add_significance()
+stat.test1 <- stat.test1 %>% add_xy_position(x = "Lecticity")
+#Build plot
 p1_1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
@@
 geom_jitter(alpha=0.3,width = 0.1) +
-ylab("Habitat Preference")+
+ylab("Habitat preference")+
 xlab(NULL)+
 ggtitle("Natural")+
+stat_pvalue_manual(stat.test1, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
+scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
 
-#Agricultural
+#Second graph Agricultural habitat
+stat.test2 <- d %>%
+wilcox_test(Agricultural ~ Lecticity) %>%
+add_significance()
+stat.test2 <- stat.test2 %>% add_xy_position(x = "Lecticity")
+#Build plot
 p1_2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
@@
 ggtitle("Agricultural")+
+stat_pvalue_manual(stat.test2, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
+scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
 
-#Urban
+#Third graph Urban habitat
+stat.test3 <- d %>%
+wilcox_test(Urban ~ Lecticity) %>%
+add_significance()
+stat.test3 <- stat.test3 %>% add_xy_position(x = "Lecticity")
+#Build plot
 p1_3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
@@
 ggtitle("Urban")+
+stat_pvalue_manual(stat.test3, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
+scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
@@
 
-p2_1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
+#Plotting----
+#First graph brain weight
+stat.test4 <- d1 %>%
+wilcox_test(Brain.weight ~ Lecticity) %>%
+add_significance()
+stat.test4 <- stat.test4 %>% add_xy_position(x = "Lecticity")
+#Build plot
+p2_1 = ggplot(d1, aes(x = Lecticity, y = Brain.weight)) + 
+geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
+geom_boxplot(colour="black",fill="grey",
+width = .15, 
+outlier.shape = NA) +
+geom_jitter(alpha=0.3,width = 0.1) +
+ylab("Absolute brain size")+
+xlab("Diet")+
+stat_pvalue_manual(stat.test4, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
+scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
+theme_bw()
+
+#Second graph residuals
+stat.test5 <- d1 %>%
+wilcox_test(residuals ~ Lecticity) %>%
+add_significance()
+stat.test5 <- stat.test5 %>% add_xy_position(x = "Lecticity")
+#Build plot
+p2_2 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
 geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
@@
 xlab("Diet")+
+stat_pvalue_manual(stat.test5, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
+scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
 
-p2_2 = ggplot(d1, aes(x = Lecticity, y = Brain.weight)) + 
-geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
-geom_boxplot(colour="black",fill="grey",
-width = .15, 
-outlier.shape = NA) +
-geom_jitter(alpha=0.3,width = 0.1) +
-ylab("Absolute brain size")+
-xlab("Diet")+
-theme_bw()
-
-panel2 =  p2_2 + p2_1
+panel2 =  p2_1 + p2_2
 
@@
 plot_grid(plot_grid(panel1, labels = c('A'))+
-  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)),
+  theme(panel.border = element_rect(colour = "black", fill=NA, size=2)),
 
 plot_grid(panel2, labels = c('B'))+
-  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)), nrow = 2)
+  theme(panel.border = element_rect(colour = "black", fill=NA, size=2)), nrow = 2)
 
 ```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S3.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
+A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
+
+#Plot maps with density of bee occurrences
+
+#Europe----
+#Load libraries
+library(giscoR) #map of europe
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(sf)
+library(patchwork)
+library(rnaturalearth)
+
+#Load europe map
+Europe <- gisco_get_countries(region = "Europe")
+#Load countries filtered
+euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
+#Load data with records filtered
+all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+#Plot Europe map with counties of interest highlighted
+p1 = ggplot(Europe) +
+geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
+geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
+aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
+coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white"))
+
+p2 = ggplot(Europe) +
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("B") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(), 
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(40, 60, by = 6)) +
+scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
+
+#Insert general plot within specific plot
+panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
+
+#USA-----
+#Load libraries
+
+#Load data
+usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
+ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
+all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+#Load maps
+canada <- ne_states(country = "canada", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+p1 = ggplot(canada) + 
+geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white")) +
+geom_map(data = ny_metropolitan_1, 
+map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
+size = 0.1) +
+geom_map(data = usa_states_1, map = usa_states_1,
+aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
+size = 0.1) +
+coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
+
+p2 = ggplot(canada) + 
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("A") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(),
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(33, 47, by = 4)) +
+scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
+
+#Insert general plot within specific plot
+panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
+
+library(patchwork)
+panel2  + panel1
+
+```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.
+Mean and standard devaition of brain weight and intertegular distance for all the species included in our analyses (N=89).", fig.pos="p", out.width="100%"}
+
+#Calculate intraclass correlation coefficiet
+#for brain and body size
+#Exploration of the dataset (number of replicates)
+#and show graphically interspecific variation
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(lme4)
+library(performance)
+
+#Read data
+long = read_csv("../Raw_data/all_cleaning.csv")
+
+#Filter out the species that we don't use
+#Load preferences data
+preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+spp = unique(pref$Species)
+
+#Select just the species that we used in the study for calculating preferences
+#after applying our filtering criteria (see mthods for further justification)
+long = long %>% 
+filter(Species %in% spp)
+
+#Show graphically variation in the distribution across species 
+#For both brain and body sizes
+#1st for brain weight
+d = long %>% 
+group_by(Species) %>% 
+summarise(mean = mean(Brain.weight), std = sd(Brain.weight)) %>% 
+arrange(desc(mean))
+#Plotting
+p1 = ggplot(d, aes(x=reorder(Species, mean), y=mean, fill=std)) + 
+geom_point()+
+geom_errorbar(aes(ymin=mean-std, ymax=mean+std), width=.2,
+position=position_dodge(0.05)) +
+theme_bw() +
+coord_flip() +
+theme(legend.position = "none",axis.text.y = element_text(size = 4),
+plot.title = element_text(face = "bold")) +
+xlab("Species") +
+ylab("Brain weight (mg)") +
+ggtitle("A")
+
+#2nd for body size
+d1 = long %>% 
+group_by(Species) %>% 
+summarise(mean = mean(IT), std = sd(IT)) %>% 
+arrange(desc(mean))
+#Plotting
+p2 = ggplot(d1, aes(x=reorder(Species, mean), y=mean, fill=std)) + 
+geom_point()+
+geom_errorbar(aes(ymin=mean-std, ymax=mean+std), width=.2,
+position=position_dodge(0.05)) +
+theme_bw() +
+coord_flip() +
+theme(legend.position = "none",axis.text.y = element_text(size = 4),
+plot.title = element_text(face = "bold")) +
+xlab("Species") +
+ylab("Intertegular distance (mm)")+
+ggtitle("B")
+
+p1 + p2
+
+```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S5.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
+The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
+
+#Histogram of habitat preferences
+#Load libraries
+library(tidyverse)
+
+#Load preferences
+preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
+preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")
+
+#Merge datasets
+preferences = bind_rows(preferences_europe, preferences_usa)
+
+#Convert to long to model everything at the same time
+long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))
+
+#Relevel for plotting
+long_data = long_data %>% 
+mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
+
+labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")
+
+long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))
+
+long_data = long_data %>% 
+mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))
+
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
+
+ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
+    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
+    xlab("Habitat preference") +
+    ylab("Frequency") +
+    facet_wrap(~ Habitat) +
+    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
+    theme_bw() +   
+    theme(axis.title = element_text(face = "bold", size = 14),
+    panel.border=element_rect(size=1),
+    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
+    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
+    panel.grid.minor=element_blank())+
+    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
+    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
+    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
+    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
+                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
+  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))
+
+```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S6}.
+Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
+The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}
+
+plot_all = readRDS("../Data/Processing/plot_all.rds")
+
+plot_all
+
+```

### 2023-09-06 Scripts/ICC_Brain_size.R

@@ -1,36 +1,94 @@
+#Calculate intraclass correlation coefficiet
+#for brain and body size
+#Exploration of the dataset (number of replicates)
+#and show graphically interspecific variation
 
+#Load libraries
 library(dplyr)
 library(stringr)
 library(readr)
-
+library(lme4)
+library(performance)
 
 #Read data
-short = readRDS("Data/Processing/wit.mean.rds")
 long = read_csv("Raw_data/all_cleaning.csv")
 
-
-s = short %>% 
-summarise(mean = mean(Brain.weight), sd = sd(Brain.weight)) %>% 
-pull(sd)
-
-#
-l = long %>% 
-group_by(Species) %>% 
-summarise(std = sd(Brain.weight)) %>% 
+#Check number of replicates per individual
+long %>% 
+group_by(Species) %>%   
+summarise(length = n()) %>% 
 ungroup() %>% 
-summarise(mean = mean(std,  na.rm=TRUE)) %>% 
-pull(mean)
+summarise(mean(length))
 
 
-#Load library
-library(lme4)
+#Filter out the species that we don't use
+#Load preferences data
+preferences_usa = readr::read_csv("Data/USA_data/preferences_usa.csv") 
+preferences_eu = readr::read_csv("Data/Europe_data/preferences_europe.csv") 
+#Bind rows
+pref = bind_rows(preferences_usa, preferences_eu)
+spp = unique(pref$Species)
+
+#Select just the species that we used in the study for calculating preferences
+#after applying our filtering criteria (see mthods for further justification)
+long = long %>% 
+filter(Species %in% spp)
+
+
+#Calculate intraclass correlation coefficient for brain size
 ggplot(long, aes(Brain.weight)) +
 geom_histogram()
 shapiro.test(log(long$Brain.weight))
 qqnorm(log(long$Brain.weight))
 qqline(log(long$Brain.weight))
-
 #Mixed model 
 m1 <- lmer(log(Brain.weight) ~ 1 | Species, data = long)
 performance::icc(m1)
+performance::icc()
+?icc()
 
+#Calculate intraclass correlation coefficient for body size
+ggplot(long, aes(IT)) +
+geom_histogram()
+shapiro.test(log(long$IT))
+qqnorm(log(long$IT))
+qqline(log(long$IT))
+#Mixed model 
[…]

### 2023-09-06 Scripts/Lecticity.R

@@ -21,9 +21,6 @@ preferences_eu = readr::read_csv("Data/Europe_data/preferences_europe.csv")
 #Bind rows
 pref = bind_rows(preferences_usa, preferences_eu)
 
-unique(pref$Species)
-
-
 #Average preferences between USA and Europe
 preferences = pref %>% 
 group_by(Species) %>%
@@ -104,12 +101,6 @@ mutate(Lecticity = replace(Lecticity, Species == "Ceratina strenua", "Polylectic
 d = d %>% 
 mutate(Lecticity = replace(Lecticity, Species == "Colletes thoracicus", "Polylectic"))
 
-#Lasioglossum malachurum
-
-#Megachile texana
-
-#Megachile willughbiella
-
 #Osmia atriventris
 #https://doi.org/10.1073/pnas.1218503110
 d = d %>% 
@@ -151,12 +142,33 @@ mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic
 d = d %>% 
 mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))
 
-d %>% 
-ggplot(aes(Lecticity, Agricultural)) +
-geom_violin()
 
 
-#Natural
+#Explore statistical differences across groups-----
+#1st explore normality
+a = d %>% filter(Lecticity== "Polylectic")
+ggplot(a, aes(Agricultural)) +
+  geom_histogram()
+b = d %>% filter(Lecticity== "Oligolectic")
+ggplot(b, aes(Agricultural)) +
+  geom_histogram()
+#Let's pick a non-parametric option for low sample sizes
+wt1 = wilcox.test(a$Natural, b$Natural)
+wt2 = wilcox.test(a$Agricultural, b$Agricultural)
+wt3 = wilcox.test(a$Urban, b$Urban)
+
+#Plotting----
+#Here we load some libraries to include Wilcoxon rank test in the plot
+#for publication
+library(ggpubr)
+library(rstatix)
+
+#First graph Natural habitat
+stat.test1 <- d %>%
+wilcox_test(Natural ~ Lecticity) %>%
+add_significance()
+stat.test1 <- stat.test1 %>% add_xy_position(x = "Lecticity")
+#Build plot
[…]

### 2023-09-07 Manuscript/Ms.Rmd
updating files

@@
 output:
+  word_document: default
   pdf_document:
     fig_caption: yes
-  word_document: default
 link-citations: yes
@@
 Lanuza$^{1}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
-Collado$^{1,2}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{3}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{3}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
+Collado$^{1,2}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{3}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{3,4}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
 
@@
 $^{3}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
+
+$^{4}$CSIC, Spanish National Research Council, CREAF-UAB, Catalonia, Spain
 
@@
 
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role as plant pollinators.
+The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the negative consequences for their fundamental role as plant pollinators.
 However, not all pollinators are negatively affected by habitat conversion, as certain species find in anthropogenic landscapes appropriate resources to persist and proliferate.
@@
 This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brains for `r nrow(raw)` individuals from `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the species-level association between brain size and habitat preferences.
+To investigate this hypothesis in insects, we measured brains for `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the association between brain size and habitat preferences.
 Our analyses revealed that bee species that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.
@@
 Yet, not all pollinator species are negatively affected by land use change.
-For instance, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [e.g., *Bombus terrestris*\; @rasmont2008; @theodorou2020].
+Indeed, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [@rasmont2008; @theodorou2020].
 
-Human-dominated habitats ---notably cities--- drastically modify the original conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+Human-dominated habitats ---notably cities--- drastically modify the ancestral conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 The question is then why only some species are able to tolerate and exploit urban environments.
-The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on cognition [@sol2009], that is, the processes involved in gathering, storing and reacting to environmental information [@shettleworth2009].
+The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition [@sol2009] to gather, store and plastically react to new information [@shettleworth2009].
 While the "cognitive buffer hypothesis" receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
@@
 Insects have been long associated with the most basic types of learning due to their miniature brain [@giurfa2015; @perry2017].
-However, numerous recent studies have shown that bees and other insects have sophisticated cognition that goes beyond simple associative learning or conditioning [@chittka2009; @avargues2011; @sheehan2011].
-Some of these more complex forms of cognition involve the use of simple tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
+However, recent studies have shown that bees and other insects have sophisticated cognition that goes beyond simple associative learning or conditioning [@chittka2009; @avargues2011; @sheehan2011].
+Some of these more complex forms of cognition involve the use of tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
 In addition, there is also evidence for substantial variation across species in brain size---both in absolute terms and relative to body size--- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
@@
 Our test is based on a unique database of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
-By means of detailed georeferenced information of species occurrences (GBIF; <http://www.gbif.org/>), we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
+By means of detailed georeferenced information of species occurrences, we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
 
@@
 
-Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe (Spain and the Netherlands).
+Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe.
 These specimens were collected opportunistically mostly in semi-natural habitats.
 The dataset includes information of `r nrow(raw)` female individuals from `r nlevels(factor(unique(raw$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
-We only considered female specimens because: (i) they are involved in a greater diversity of tasks than males, and hence are expected to experience greater environmental pressures; and (ii), have brains that are structurally and functionally different from those of males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; see supplementary text S1].
+We considered only female specimens because they engage in a wider variety of tasks, facing greater environmental pressures, and possess distinct brain structures and functions compared to males [@streinzer2013; @roselino2015].
+Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; supplementary text S1].
 Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
@@
 The phylogeny was built with the help of a previously published genus-level phylogeny and it was processed with the help of the packages ape 5.6-2 [@paradis2019], phytools 1.2-0 [@revell2012] and MCMCglmm version 2.33 [@hadfield2010].
-High values of relative brain size indicate larger brains than expected for their body size and low values denote smaller brains than expected.
-Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the evolutionary history of the species (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* \< 0.01).
-However, we found considerable variability in relative brain size within and across the different taxonomic groups (i.e., genera and subfamily level; Figure S1; Figure 1A).
+High values of relative brain size indicate larger brains than expected for their body size, while low values denote smaller brains.
+Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the species’ evolutionary history (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* \< 0.01).
+However, we found considerable variability in relative brain size within and across taxonomic groups (Figure S1; Figure 1A).
 
-## (b) Diet specialisation
-
-To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with both habitat preference and brain size.
-For this, we used the diet information collected in @sayol2020 and complemented the missing information by using the same online resources used in @sayol2020.
-Bee species were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
-We tested for statistical differences in diet specialization per habitat type preference and brain size (i.e., both absolute and relative brain size) by using the Wilcoxon test.
-We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (see details in Figure S2).
-
-## (c) Habitat preferences
+## (b) Habitat preferences
 
@@
 For Europe, we selected countries located on the north and centre of the continent (Figure S3B), representing a total area of `r eukm2` km$^{2}$.
-To further reduce biases in the data, we only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
-In addition, we optimised the match between species occurrence and the land cover data by only using georeferenced records obtained between 1990 and 2022 with a minimum of two decimals of latitude/longitude coordinates.
+We only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
+We optimised the match between species occurrence and the land cover data by only using georeferenced records obtained between 1990 and 2022 with a minimum of two decimals of latitude/longitude coordinates.
 
@@
 Lastly, we considered that a bee species exhibited a "high preference" for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
-On the contrary, the species was considered to exhibit "low preference or avoidance" for the habitat when the observed occurrences were below the 20th percentile.
-To better understand if our findings are affected by the evolutionary histories of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
+On the contrary, we considered a "low preference or avoidance" for the habitat when the observed occurrences were below the 20th percentile.
+To better understand if our findings are affected by the evolutionary history of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
+
+## (c) Diet specialisation
+
+To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with habitat preference and brain size.
+For this, we used existing diet information [see @sayol2020], where bees were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
+We tested for statistical differences in diet across habitat preferences per habitat and brain size by using Wilcoxon test.
+We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (Figure S2).
 
@@
 
-Our macro-ecological approach is justified by: (i) independent data that ensures that any ecological patterns, if present, are likely to be robust and generalizable; (ii) low intraspecific variation of brain and body size in comparison to the interspecific variation (See supplementary text S2 and Figure S4) that provides biological meaning of averaging relative brain sizes per species and the study of the ecological patterns across them.
-
-To evaluate how the association between habitat preference and relative brain size per species changed by habitat type, we used a Bayessian approach to model their association.
+To evaluate how the association between habitat preference and relative brain size per species changed by habitat type, we used a Phylogenetic Bayessian approach to model their association.
 For this, we first joined the resulting habitat preferences per species with their respective average relative brain sizes.
@@
 Because we assessed habitat preference as binary ("avoiding" or "preferring"), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
-Again, we included the phylogenetic covariance matrix as a random factor.
-Moreover, we also explored how habitat preference changed by average brain weight and intertegular span independently.
-For this, we conducted two analogous models with these two different predictors.
-In addition, we also investigated the different trends for the United States and Europe separately.
+To control for the shared evolutionary history of the species, we included the phylogenetic covariance matrix as a random factor.
+Finally, we replicated the analyses for the United States and Europe independently with analogous models.
 
@@
 
-Habitat preference varied substantially across species (Figure 1B) and showed moderate phylogenetic signal (i.e., preference for natural habitat; $\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three habitats (Figure 1B).
+Habitat preference varied substantially across species (Figure 1B) and showed moderate phylogenetic signal ($\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three (Figure 1B).
 The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
@@
 Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of the United States or from both regions).
-The deviation of the brain size in relation to the body (i.e., residuals) is represented with filled circles of proportional area to the residuals at the end of the tip branches.
+The deviation of the brain size in relation to the body (i.e., residuals) is represented with filled circles of proportional area at the end of the tip branches.
 Larger circles indicate larger brains in proportion to their body size and vice versa.
@@
 
-We found that bee species relative brain size is associated with the habitat type preference (Figure 2A; Bayesian R$^2$ = `r r2_main`).
-Specifically, we found that bees with larger brains relative to body size showed a higher preference for urban habitats than bees with smaller relative brain sizes (Figure 2A).
-Contrarily, bee species with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brain sizes (Figure 2A).
+We found that relative brain size is associated with habitat type preference (Figure 2A; Bayesian R$^2$ = `r r2_main`).
+Specifically, we found that bees with larger relative brains showed a higher preference for urban habitats than bees with smaller ones (Figure 2A).
+Contrarily, bee species with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brains (Figure 2A).
 The models of the association between absolute brain size and intertegular span with habitat preference also showed a marked differences between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
-Specifically, we found that bees with larger brains and body sizes tend to appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
-These findings were consistent with the analogous analyses separated by geographical regions (United States and Europe; Figure S6).
+Specifically, we found that bees with larger brains and body sizes appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
+These findings were consistent with the analogous analyses by geographical regions (United States and Europe; Figure S6).
 
@@
 
-Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bee species with a preference for inhabiting urban habitats tended to have larger brains relative to their body while those with forest or agricultural preferences had relatively smaller brains.
-These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to persist and thrive in novel environments.
-In urban environments, individuals are frequently exposed to a variety of new challenges such as novel resources or human disturbances that can change fast in time and space [@wheater2002].
-In such scenarios, a large brain could confer the cognitive flexibility needed to efficiently exploit novel resources while avoiding risks.
-While the exact mechanisms are unclear, we expect cognitive flexibility to be essential in a variety of contexts such as the use of human-made materials for the nest [@allasino2019] or the need to collect pollen and nectar from exotic flowers [@lowenstein2019].
+Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bee species with a preference for inhabiting urban habitats tend to have larger brains relative to their body while those with forest or agricultural preferences have relatively smaller brains.
+These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to tolerate novel environments.
+Since urban environments exhibit dynamic challenges, including novel resources and changing human disturbances [@wheater2002], a large brain may provide the cognitive flexibility to exploit these new resources while avoiding risks.
+Although the exact mechanisms remain unclear, we expect cognitive flexibility to be essential in a variety of contexts such as the use of human-made materials for the nest [@allasino2019] or the use of exotic flowers [@lowenstein2019].
 
-Also in line with the cognitive buffer hypothesis [@ducatez2020], we find that bees that proliferate in urban habitats have bigger bodies and larger absolute brains.
+In line with the cognitive buffer hypothesis [@ducatez2020], urban bees also showed bigger bodies and larger absolute brains.
 Larger species require longer development times and obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
@@
 The patchiness of urban resources also seems to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
-There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion.
-Only a few bee specialists are analyzed in this paper, and further understanding how different life styles facilitate urban exploitation represent an important avenue of future research.
+There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion [@ducatez2015; @sayol2016; @sol2016].
+In our study, we examined only a few bee specialists, highlighting the need for further research on how different diets and life strategies relate to cognition and urban adaptation.
 
-Overall, our findings support and expand previous evidence found in vertebrates that having a large brain size can allow survival in novel environments [@santini2019; @sayol2020b], highlighting that a cognitive buffer is possible even with tiny brains.
+Our findings support and extend upon previous evidence in vertebrates, that having larger brains can facilitate tolerance to urban environments [@santini2019; @sayol2020b], highlighting that a cognitive buffer is possible even with tiny brains.
 The use of the entire brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
-Our analyses focused on brain size mainly for reasons of data availability.
-The use of brain size is justified by the existence of previous evidence that enlarged brains enhance some aspects of cognitive performance in bees, such as learning [@collado2021].
+We primarily analyzed brain size due to data availability, supported by prior findings showing larger brains improving certain cognitive aspects in bees, such as learning [@collado2021].
 Moreover, brain size is less subject to measurement error or context-dependent biases in comparison with other experimental measures of cognition [@chittka2009; @lihoreau2012].
-Although the use in the future of finer measures like neuropils size or mushroom bodies [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995] is likely to help improve our understanding of bee cognitive abilities.
-Further, there are other alternative explanations beyond the cognitive buffer hypothesis for the enlargement of brains in urban environments.
-Here, we only investigated the role of diet specialisation as a potential confounding factor but other ecological and physiological factors cannot be discarded as potential drivers of this pattern.
+However, future finer measures, such as neuropil size or mushroom bodies  [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995], are likely to enhance our understanding of bee cognition.
 Downscaling our analysis to the individual specimen level where brain size, habitat use and cognitive performance can be tracked through their lifespan would be a challenging but promising next step.
-Overall, our findings highlight the importance of behavioural responses for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
+Our findings highlight the importance of cognition for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
 
@@
 
-IB, MAC and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC and IB. JBL and IB wrote the manuscript with contributions from all authors.
+IB, DS, MAC and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC, DS and IB. IB and JBL wrote the manuscript with contributions from all authors.
 
@@
 
-We thank Fransico P. Molina, Ivo Reamakers and Parker Gambino for collecting some of the studies specimens.
+We thank Fransico P. Molina, Ivo Reamakers and Parker Gambino for collecting some of the specimens.
 

### 2023-09-07 Manuscript/Supplementary_material.Rmd
updating files

@@
 This function computes the division between the random effect variance and the total variance (i.e., the sum of the random effect variance and the residual variance).
-We found that for both brain and body size the individual measurements per species tended to be very similar between each other (ICC = 0.947; ICC ICC = 0.931, respectively) but we found high variability across species (see Figure S4).
+We found that for both brain and body size the individual measurements per species tended to be very similar between each other (ICC = 0.947; ICC = 0.931, respectively) but we found high variability across species (see Figure S4).
 Overall, this indicates that the mean values of brain and body size have biological significance and provides support for investigating macro-ecological patterns across species.

### 2023-09-07 Manuscript/Ms.Rmd
updating files before submisison

@@
 Lanuza$^{1}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{1,2}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{3}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+Collado$^{1,2}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3,4}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{3}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
 
@@
 **Author for correspondence: barragansljose\@gmail.com**
-
-\textbf{Author for correspondence: barragansljose@gmail.com}
 
@@
 The question is then why only some species are able to tolerate and exploit urban environments.
-The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition [@sol2009] to gather, store and plastically react to new information [@shettleworth2009].
+The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition to gather, store and plastically react to new information [@shettleworth2009; @sol2009].
 While the "cognitive buffer hypothesis" receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
@@
 We considered only female specimens because they engage in a wider variety of tasks, facing greater environmental pressures, and possess distinct brain structures and functions compared to males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; supplementary text S1].
+Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; Supplementary text S1].
 Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
@@
 We selected the states or countries with the highest density of records for our set of species.
-For North America, we selected states located on the East Coast of the United States (Figure S3A), covering an approximate area of `r usakm2` km$^{2}$.
-For Europe, we selected countries located on the north and centre of the continent (Figure S3B), representing a total area of `r eukm2` km$^{2}$.
+For North America, we selected states located on the East Coast of the United States (Figure S2A), covering an approximate area of `r usakm2` km$^{2}$.
+For Europe, we selected countries located on the north and centre of the continent (Figure S2B), representing a total area of `r eukm2` km$^{2}$.
 We only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
@@
 We tested for statistical differences in diet across habitat preferences per habitat and brain size by using Wilcoxon test.
-We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (Figure S2).
+We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (Figure S3).
 
@@
 For this, we first joined the resulting habitat preferences per species with their respective average relative brain sizes.
+Our macro-ecological approach is justified by: (i) independent data ensuring robust and generalizable ecological patterns and (ii) minimal intraspecific variation in brain and body size compared to interspecific variation (see Supplementary text S2 and Figure S4).
 The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.

### 2023-09-07 Manuscript/Supplementary_material.Rmd
updating files before submisison

@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S2}.
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
+A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
+
+#Plot maps with density of bee occurrences
+
+#Europe----
+#Load libraries
+library(giscoR) #map of europe
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(sf)
+library(patchwork)
+library(rnaturalearth)
+
+#Load europe map
+Europe <- gisco_get_countries(region = "Europe")
+#Load countries filtered
+euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
+#Load data with records filtered
+all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
+
+#Plot Europe map with counties of interest highlighted
+p1 = ggplot(Europe) +
+geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
+geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
+aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
+coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white"))
+
+p2 = ggplot(Europe) +
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("B") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(), 
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(40, 60, by = 6)) +
+scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
+
+#Insert general plot within specific plot
+panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
+
+#USA-----
+#Load libraries
+
+#Load data
+usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
+ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
+all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
+
+#Load maps
+canada <- ne_states(country = "canada", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
+st_as_sf(crs = 4326)
+
+p1 = ggplot(canada) + 
+geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
+microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
+panel.background = element_rect(fill = "white")) +
+geom_map(data = ny_metropolitan_1, 
+map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
+size = 0.1) +
+geom_map(data = usa_states_1, map = usa_states_1,
+aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
+size = 0.1) +
+coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
+
+p2 = ggplot(canada) + 
+geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
+geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
+geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
+viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
+coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
+theme(panel.background = element_rect(fill = "white"),
+panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+ylab("Latitude") +
+xlab("Longitude") +
+ggtitle("A") +
+theme(legend.position = c(0, 1),legend.justification = c(0, 1),
+legend.background = element_blank(),
+legend.box.background = element_blank(),
+legend.key = element_blank(),
+legend.title = element_text(face = "bold"),
+axis.title = element_text(face="bold", size = 14),
+plot.title = element_text(face="bold", size = 18),
+legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
+scale_y_continuous(breaks = seq(33, 47, by = 4)) +
+scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
+guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
+ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
+
+#Insert general plot within specific plot
+panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
+
+library(patchwork)
+panel2  + panel1
+
+```
+
+\clearpage
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S3}.
 Graphical representation and statistical differences of diet specialisation (i.e., oligolectic and polylectic species) by habitat preference per habitat type (Panel A) and brain size including absolute and relative brain size (Panel B).
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S3.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
-B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
-
-#Plot maps with density of bee occurrences
-
-#Europe----
-#Load libraries
-library(giscoR) #map of europe
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(sf)
-library(patchwork)
-library(rnaturalearth)
-
-#Load europe map
-Europe <- gisco_get_countries(region = "Europe")
-#Load countries filtered
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Load data with records filtered
-all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-
-#Plot Europe map with counties of interest highlighted
-p1 = ggplot(Europe) +
-geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
-geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
-aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
-coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white"))
-
-p2 = ggplot(Europe) +
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("B") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(), 
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(40, 60, by = 6)) +
-scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)
-
-#Insert general plot within specific plot
-panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)
-
-#USA-----
-#Load libraries
-
-#Load data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
-all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Load maps
-canada <- ne_states(country = "canada", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
-st_as_sf(crs = 4326)
-
-p1 = ggplot(canada) + 
-geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
-microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
-panel.background = element_rect(fill = "white")) +
-geom_map(data = ny_metropolitan_1, 
-map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
-size = 0.1) +
-geom_map(data = usa_states_1, map = usa_states_1,
-aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
-size = 0.1) +
-coord_sf(xlim = c(-86, -62), ylim = c(25, 48))
-
-p2 = ggplot(canada) + 
-geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
-geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
-geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
-viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
-coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
-theme(panel.background = element_rect(fill = "white"),
-panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-ylab("Latitude") +
-xlab("Longitude") +
-ggtitle("A") +
-theme(legend.position = c(0, 1),legend.justification = c(0, 1),
-legend.background = element_blank(),
-legend.box.background = element_blank(),
-legend.key = element_blank(),
-legend.title = element_text(face = "bold"),
-axis.title = element_text(face="bold", size = 14),
-plot.title = element_text(face="bold", size = 18),
-legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
-scale_y_continuous(breaks = seq(33, 47, by = 4)) +
-scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
-guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
-ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))
-
-#Insert general plot within specific plot
-panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)
-
-library(patchwork)
-panel2  + panel1
-
-```
-
-\clearpage
-
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}.

### 2023-09-08 Manuscript/Ms.Rmd
updates befor saving repo

@@
 \textbf{Jose B.
-Lanuza$^{1}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{1,2}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{3,4}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{3}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
+Lanuza$^{1,2}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
+Collado$^{1,3}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{4}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{4,5}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
 
 Jose B.
-Lanuza$^{1}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
-Collado$^{1,2}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{3}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{3,4}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
+Lanuza$^{1,2}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
+Collado$^{1,3}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{4}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{4,5}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
 
@@
 
-$^{2}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
+$^{2}$German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
 
-$^{3}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
+$^{3}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
 
-$^{4}$CSIC, Spanish National Research Council, CREAF-UAB, Catalonia, Spain
+$^{4}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
+
+$^{5}$CSIC, Spanish National Research Council, CREAF-UAB, Catalonia, Spain
 

### 2023-09-08 Scripts/0_Processing/0_Preprocessing_new.R

@@ -11,12 +11,12 @@ library(ggplot2)
 
 #Read datasets
 #1st Sayol 
-sayol = read_csv("Raw_data/BeeTraits_individual.csv") %>% 
+sayol = read_csv("Raw_data/sayol_data.csv") %>% 
 dplyr::select(!c(head.weight.g, no.optic.lobes.weight.mg,body.mass.g)) %>% 
 mutate(species = str_replace(species, "_", " "))
 #snd Collado (is similar but Collado has some extra species)
 #Note Sayol data is clean but Collado still needs some processing
-collado <- read.csv("Raw_data/brains.it.csv", dec = ",") %>% 
+collado <- read.csv("Raw_data/collado_data.csv", dec = ",") %>% 
 dplyr::select(!c(Place, No.optic.lobes.weight, 
 Notes, Head.weight..g., Plant, Date, Weigth..g., ID, Genus)) %>% 
 rename(Brain.weight = Brain.Weight..mg.)
@@ -131,7 +131,7 @@ wit.mean <- merge(weights.mean, IT.mean)
 
 #Save data ----
 #Save data to load raw values in methods
-write_csv(all_cleaning, "Raw_data/all_cleaning.csv")
+write_csv(all_cleaning, "Data/Processing/all_cleaning.csv")
 #Save data to calculate residuals
 saveRDS(wit.mean, "Data/Processing/wit.mean.rds")
 

### 2023-09-08 Scripts/Lecticity.R

@@ -28,7 +28,7 @@ summarise(across(everything(), list(mean),.names = "{.col}")) %>%
 ungroup()
 
 #Read lecticity
-lecticity = read_csv("Data/Bees_Lecticity.csv") 
+lecticity = read_csv("Data/Lecticity/Bees_Lecticity.csv") 
 #Delete underscore from tree
 lecticity$species = str_replace(lecticity$species, "_", " ")
 lecticity = lecticity %>% rename(Species = species)
@@ -142,7 +142,7 @@ mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic
 d = d %>% 
 mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))
 
-
+write_csv(d, "Data/Lecticity/Lecticity_all.csv")
 
 #Explore statistical differences across groups-----
 #1st explore normality

### 2023-09-08 Manuscript/Ms.Rmd
Data availability statement update

@@
 #Load final dataset 
-raw = read_csv("../Raw_data/all_cleaning.csv")
+raw = read_csv("../Data/Processing/all_cleaning.csv")
 
@@
 
-All occurrence data used in this study can be downloaded from <https://doi.org/10.15468/dl.5s5kuf>.
-In addition, bee trait data and code can be found in the online repositories of Zenodo <https://doi.org/10.5281/zenodo.8049996> and Github <https://github.com/ibartomeus/declines_brain.git>.
+All data and code can be accessed at Dryad Digital Repository: <https://doi.org/10.5061/dryad.zw3r228dr> [@Lanuza2023].
 

### 2023-09-08 Scripts/0_Processing/0_Preprocessing_new.R

@@ -134,5 +134,6 @@ wit.mean <- merge(weights.mean, IT.mean)
 write_csv(all_cleaning, "Data/Processing/all_cleaning.csv")
 #Save data to calculate residuals
 saveRDS(wit.mean, "Data/Processing/wit.mean.rds")
+write_csv(wit.mean, "Data/Processing/wit.mean.csv")
 
 

### 2023-10-23 Manuscript/Ms.Rmd
Minor revision Biol Lett

@@
 ---
-title: "Bees need larger brains to thrive in urban environments"
+title: "Brain size predicts bees' tolerance to urban environments"
 output:
-  word_document: default
   pdf_document:
     fig_caption: yes
+  word_document: default
 link-citations: yes
@@
 However, not all pollinators are negatively affected by habitat conversion, as certain species find in anthropogenic landscapes appropriate resources to persist and proliferate.
-The reason why some species thrive in anthropogenic environments while most find them inhospitable remains poorly understood.
+The reason why some species tolerate anthropogenic environments while most find them inhospitable remains poorly understood.
 The cognitive buffer hypothesis, widely supported in vertebrates but untested in insects, offers a potential explanation.
 This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brains for `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated the association between brain size and habitat preferences.
-Our analyses revealed that bee species that prefer urban habitats had larger brains relative to their body size than those who prefer forested or agricultural habitats.
+To investigate this hypothesis in insects, we measured brain size for `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated their association with the degree of habitat occupancy.
+Our analyses revealed that bee species mainly found in urban habitats had larger brains relative to their body size than those that tend to occur in forested or agricultural habitats.
 Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
@@
 
-Keywords: relative brain size, habitat preference, Apoidea, urbanisation, pollinators
+Keywords: relative brain size, habitat occupancy, Apoidea, urbanisation, pollinators
 
@@
 One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maladaptation [@winfree2011; @brown2009 ; @harrison2019].
+Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maladaptation [@winfree2011; @brown2009; @harrison2019].
 Yet, not all pollinator species are negatively affected by land use change.
@@
 
-Human-dominated habitats ---notably cities--- drastically modify the ancestral conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
+Human-dominated habitats --- notably cities --- drastically modify the ancestral conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
 The question is then why only some species are able to tolerate and exploit urban environments.
-The "cognitive buffer hypothesis" provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition to gather, store and plastically react to new information [@shettleworth2009; @sol2009].
-While the "cognitive buffer hypothesis" receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
+The 'cognitive buffer hypothesis' provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition to gather, store and plastically react to new information [@shettleworth2009; @sol2009].
+While the 'cognitive buffer hypothesis' receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
 
@@
 Some of these more complex forms of cognition involve the use of tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
-In addition, there is also evidence for substantial variation across species in brain size---both in absolute terms and relative to body size--- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
+In addition, there is also evidence for substantial variation across species in brain size --- both in absolute terms and relative to body size --- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
 Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
 
-Here, we report the first test of the "cognitive-buffer hypothesis" in insects.
-Our test is based on a unique database of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
-By means of detailed georeferenced information of species occurrences, we characterise habitat preferences for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
+Here, we report the first test of the 'cognitive-buffer hypothesis' in insects.
+Our test is based on a unique dataset of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
+By means of detailed georeferenced information of species occurrences, we characterise the degree of habitat occupancy for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
 
@@
 
-Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Europe.
+Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Northern Central Europe.
 These specimens were collected opportunistically mostly in semi-natural habitats.
@@
 
-## (b) Habitat preferences
+## (b) Habitat occupancy
 
@@
 
-Habitat preferences were estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
-For this, we generated 10,000 randomised matrices based on the occurrence matrix with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
+The degree of habitat occupancy was estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
+Although previous work has referred to this metric as ‘habitat preference’ [@collado2019], we have opted to use ‘degree of habitat occupancy’ as it avoids the connotation of active choice by the different species.
+To calculate habitat occupancy, we generated 10,000 randomised matrices based on the occurrence matrix with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
 We used the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
@@
 We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
-Lastly, we considered that a bee species exhibited a "high preference" for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
-On the contrary, we considered a "low preference or avoidance" for the habitat when the observed occurrences were below the 20th percentile.
-To better understand if our findings are affected by the evolutionary history of the species, we estimated the phylogenetic signal of relative brain size and habitat preference across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
+Lastly, we considered that a bee species exhibited a ‘high habitat occupancy’ for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
+On the contrary, we considered a ‘low habitat occupancy’ for the habitat when the observed occurrences were below the 20th percentile.
+To better understand if our findings are affected by the evolutionary history of the species, we estimated the phylogenetic signal of relative brain size and the degree of habitat occupancy across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
 
@@
 
-To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with habitat preference and brain size.
+To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with the degree of habitat occupancy and brain size.
 For this, we used existing diet information [see @sayol2020], where bees were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
-We tested for statistical differences in diet across habitat preferences per habitat and brain size by using Wilcoxon test.
-We found a low number of specialist species in our dataset with no clear associations with habitat preference or brain size that can explain the relationship between habitat preference and brain size (Figure S3).
+We tested for statistical differences in diet across the degree of habitat occupancy per habitat type and brain size by using Wilcoxon test.
+We found a low number of specialist species in our dataset with no clear associations with the degree of habitat occupancy or brain size that can explain the observed relationship between the two (Figure S3).
 
@@
 
-To evaluate how the association between habitat preference and relative brain size per species changed by habitat type, we used a Phylogenetic Bayessian approach to model their association.
-For this, we first joined the resulting habitat preferences per species with their respective average relative brain sizes.
+To evaluate how the association between the degree of habitat occupancy and relative brain size per species changed by habitat type, we used a Phylogenetic Bayessian approach to model their association.
+For this, we first joined the resulting values of the degree of habitat occupancy per species with their respective average relative brain sizes.
 Our macro-ecological approach is justified by: (i) independent data ensuring robust and generalizable ecological patterns and (ii) minimal intraspecific variation in brain and body size compared to interspecific variation (see Supplementary text S2 and Figure S4).
-The resulting distribution of habitat preferences for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat preferences close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
-Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as avoiding or preferring a given habitat).
-Because we assessed habitat preference as binary ("avoiding" or "preferring"), we specified a Bernouilli distribution where habitat preference was the response variable and relative brain size the predictor.
+The resulting distribution of the degree of habitat occupancy for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat occupancy close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
+In other words, simulated occurrences were substantially lower or higher than the observed ones.
+Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as high or low habitat occupancy).
+Because we assessed the degree of habitat occupancy as binary ('low' or 'high'), we specified a Bernouilli distribution where the degree of habitat occupancy was the response variable and relative brain size the predictor.
 To control for the shared evolutionary history of the species, we included the phylogenetic covariance matrix as a random factor.
@@
 
-Habitat preference varied substantially across species (Figure 1B) and showed moderate phylogenetic signal ($\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species preferred one or two habitat types but more rarely occurred indistinctly in the three (Figure 1B).
-The most preferred habitats were the agricultural and natural ones with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
-In contrast, most species avoided urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high preference for this habitat type (Figure 1B and Figure S5).
+The degree of habitat occupancy varied substantially across species (Figure 1B) and showed moderate phylogenetic signal ($\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
+In general, most bee species have high occupancy in one or two habitat types but more rarely occurred indistinctly in the three (Figure 1B).
+The habitats with the highest occurrences in comparison with random expectations were the agricultural and natural ones, with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
+In contrast, most species showed low degree of occupancy in urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high occupancy in this habitat type (Figure 1B and Figure S5).
 
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="Figure 1.
-Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=89).
+Phylogenetic relationship and degree of habitat occupancy for the selected bee species with brain weight and intertegular distance information (N=89).
 A) Phylogenetic tree at the genus level.
@@
 Larger circles indicate larger brains in proportion to their body size and vice versa.
-B) Heatmap showing the habitat preference for each bee species.
+B) Heatmap showing the degree of habitat occupancy for each bee species.
 The columns delimit the habitat type (i.e., natural, agricultural and urban) and the rows the different bee species."}
@@
    # scale_fill_gradient(high="black",  low="white") +
-    scale_fill_viridis_c(name = "Habitat\npreference") + 
+  scale_fill_continuous("Habitat occupancy",type = "viridis", 
+                        breaks = c(0, 0.5,  1), 
+                        labels = c("0 (Low)", "0.5","1 (High)")) +
+
+  #  scale_fill_viridis_c(name = "Habitat\noccupancy") + 
     theme_minimal() +
@@
 
-We found that relative brain size is associated with habitat type preference (Figure 2A; Bayesian R$^2$ = `r r2_main`).
-Specifically, we found that bees with larger relative brains showed a higher preference for urban habitats than bees with smaller ones (Figure 2A).
-Contrarily, bee species with smaller relative brains showed higher preference in natural and agricultural habitats than bees with larger relative brains (Figure 2A).
-The models of the association between absolute brain size and intertegular span with habitat preference also showed a marked differences between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
+We found that relative brain size was associated with the degree of occupancy across the different habitat types (Figure 2A; Bayesian R$^2$ = `r r2_main`).
+Specifically, we found that bees with larger relative brains showed a higher occupancy in urban habitats than bees with smaller ones (Figure 2A).
+Contrarily, bee species with smaller relative brains showed higher occupancy in natural and agricultural habitats than bees with larger relative brains (Figure 2A).
+The models of the association between absolute brain size and intertegular span with habitat occupancy also showed a marked difference between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
 Specifically, we found that bees with larger brains and body sizes appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="Figure 2.
-Association between relative brain size (A), brain weight (B) and intertegular distance (C) with habitat preference by habitat type (i.e., natural, agricultural and urban).
+Association between relative brain size (A), brain weight (B) and intertegular distance (C) with the degree of habitat occupancy by habitat type (i.e., natural, agricultural and urban).
 The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p"}
@@
 theme_bw() +
-ylab("Habitat preference") +
+ylab("Habitat occupancy") +
 xlab("Relative brain size") +
@@
 
-Using data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we find that bee species with a preference for inhabiting urban habitats tend to have larger brains relative to their body while those with forest or agricultural preferences have relatively smaller brains.
+By analysing data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we observed that bee species inhabiting urban habitats tend to have larger brains relative to their body while those found in forest or agricultural habitats exhibit relatively smaller brains.
 These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to tolerate novel environments.

### 2023-10-23 Manuscript/Supplementary_material.Rmd
Minor revision Biol Lett

@@
 
-# Supplementary material for: Bees need larger brains to thrive in urban environments
+# Supplementary material for: Brain size predicts bees' tolerance to urban environments
 
@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S3}.
-Graphical representation and statistical differences of diet specialisation (i.e., oligolectic and polylectic species) by habitat preference per habitat type (Panel A) and brain size including absolute and relative brain size (Panel B).
+Graphical representation and statistical differences of diet specialisation (i.e., oligolectic and polylectic species) by the degree of habitat occupancy per habitat type (Panel A) and brain size including absolute and relative brain size (Panel B).
 The graphical representation shows boxplots, raw values and a shaded violin plot to allow a better understanding of the spread, density and number of values by each category.
@@
 The significance of the statistical differences across values are shown with the resulting P-values of a Wilcoxon test.
-Interestingly, we see that (i) there are few oligolectic species represented in our dataset; (ii) there are no statistical differences between oligolectic and polylectic species in habitat preference (Panel A) and brain size (Panel B); and, (iii) oligolectic species tend to have larger relative brains and, if anything, show lower preferences by urban habitats, indicating that feeding specialisation is unlikely to be driving the observed association between urban preference and brain size.", fig.pos="p", out.width="100%"}
+Interestingly, we see that (i) there are few oligolectic species represented in our dataset; (ii) there are no statistical differences between oligolectic and polylectic species in habitat occupancy (Panel A) and brain size (Panel B); and, (iii) oligolectic species tend to have larger relative brains and, if anything, show lower occupancy in urban habitats, indicating that feeding specialisation is unlikely to be driving the observed association between urban occupancy and brain size.", fig.pos="p", out.width="100%"}
 
@@
 #Read lecticity
-lecticity = read_csv("../Data/Bees_Lecticity.csv") 
+lecticity = read_csv("../Data/Lecticity/Bees_Lecticity.csv") 
 #Delete underscore from tree
@@
 geom_jitter(alpha=0.3,width = 0.1) +
-ylab("Habitat preference")+
+ylab("Habitat occupancy")+
 xlab(NULL)+
@@
 stat_pvalue_manual(stat.test1, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
-scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
+scale_y_continuous(breaks = c(0, 0.5,1),expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
@@
 stat_pvalue_manual(stat.test2, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
-scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
+scale_y_continuous(breaks = c(0, 0.5,1),expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
@@
 stat_pvalue_manual(stat.test3, label =  "Wilcoxon test, p = {p}",vjust = -0.05, bracket.nudge.y = 0.05) +
-scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
+scale_y_continuous(breaks = c(0, 0.5,1), expand = expansion(mult = c(0.05, 0.2))) +
 theme_bw()
@@
 #Read data
-long = read_csv("../Raw_data/all_cleaning.csv")
+long = read_csv("../Data/Processing/all_cleaning.csv")
 
@@
 
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S5.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban).
-The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S5.} Frequency distribution of the degree of habitat occupancy compared to null expectations for each habitat type (i.e., natural, agricultural and urban).
+The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for the United States.
+The categories of 'low' and 'high' indicate that the observed occurrences are consistently under- or over-represented when compared to random expectations.", fig.pos="p"}
 
@@
 
-long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))
+long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low", "No difference", "High"))
 
@@
     geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
-    xlab("Habitat preference") +
+    xlab("Degree of habitat occupancy") +
     ylab("Frequency") +
@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S6}.
-Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
+Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with the degree of habitat occupancy separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one).
 The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}

### 2023-10-23 Scripts/Figure_4_Preference_Models/Figure_4_Plotting_all.R

@@ -39,7 +39,7 @@ geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, si
 geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
-ylab("Habitat preference") +
+ylab("Habitat occupancy") +
 xlab("Relative brain size") +
 scale_colour_viridis_d() +
 scale_fill_viridis_d() +
@@ -117,7 +117,7 @@ geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, si
 geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
-ylab("Habitat preference") +
+ylab("Habitat occupancy") +
 xlab("Relative brain size") +
 ggtitle("D") +  
 scale_colour_viridis_d() +
@@ -201,7 +201,7 @@ geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, si
 geom_line(aes(color=Habitat),size=0.7) +
 geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
 theme_bw() +
-ylab("Habitat preference") +
+ylab("Habitat occupancy") +
 xlab("Relative brain size") +
 ggtitle("A") +  
 scale_colour_viridis_d() +

### 2023-10-23 Scripts/Lecticity.R

@@ -144,6 +144,9 @@ mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Pol
 
 write_csv(d, "Data/Lecticity/Lecticity_all.csv")
 
+d1 = d %>% select(Species, Lecticity)
+write_csv(d1, "Data/Lecticity/Lecticity.csv")
+
 #Explore statistical differences across groups-----
 #1st explore normality
 a = d %>% filter(Lecticity== "Polylectic")

### 2023-10-25 Manuscript/Figure_1.Rmd
Minor revision

--- 2023-10-23 Manuscript/Ms.Rmd
@@
 ---
-title: "Brain size predicts bees' tolerance to urban environments"
-output:
-  pdf_document:
-    fig_caption: yes
-  word_document: default
-link-citations: yes
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
-fontsize: 12pt
-includeline-num: true
-bibliography: references.bib
-always_allow_html: true
+output: pdf_document
 header-includes:
-- \usepackage{orcidlink}
 - \usepackage{caption,setspace}
 - \captionsetup[figure]{labelformat=empty}
-- \usepackage{float}
-- \floatplacement{figure}{H}
-- \usepackage[left]{lineno}
-- \linenumbers
-csl: BiologyLetters.csl
 ---
@@
 \captionsetup[table]{labelformat=empty}
-
-```{r echo=FALSE,warning=FALSE}
-
-knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
-```
-
-\textbf{Jose B.
-Lanuza$^{1,2}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{1,3}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{4}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{4,5}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
-
-Jose B.
-Lanuza$^{1,2}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
-Collado$^{1,3}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{4}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{4,5}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
-
-$^{1}$Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
-
-$^{2}$German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
-
-$^{3}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
-
-$^{4}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
-
-$^{5}$CSIC, Spanish National Research Council, CREAF-UAB, Catalonia, Spain
-
-**Author for correspondence: barragansljose\@gmail.com**
-
-```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-library(readr)
-library(dplyr)
-
-#Load final dataset 
-raw = read_csv("../Data/Processing/all_cleaning.csv")
-
-d = raw %>% 
-group_by(Species) %>% 
-summarise(individuals = n())
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-s = unique(pref$Species)
-
-raw = raw %>% filter(Species %in% s)
-
-#Load taxonomic rank script
-tax = read_csv("../Data/Processing/taxonomic_rank.csv")
-
-```
-
-# Abstract
-
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the negative consequences for their fundamental role as plant pollinators.
-However, not all pollinators are negatively affected by habitat conversion, as certain species find in anthropogenic landscapes appropriate resources to persist and proliferate.
-The reason why some species tolerate anthropogenic environments while most find them inhospitable remains poorly understood.
-The cognitive buffer hypothesis, widely supported in vertebrates but untested in insects, offers a potential explanation.
-This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brain size for `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated their association with the degree of habitat occupancy.
-Our analyses revealed that bee species mainly found in urban habitats had larger brains relative to their body size than those that tend to occur in forested or agricultural habitats.
-Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
-Our results provide the first empirical support for the cognitive buffer hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
-
-Keywords: relative brain size, habitat occupancy, Apoidea, urbanisation, pollinators
-
-\clearpage
-
-# 1. Introduction
-
-Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
-Regrettably, there is increasing evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
-One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maladaptation [@winfree2011; @brown2009; @harrison2019].
-Yet, not all pollinator species are negatively affected by land use change.
-Indeed, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [@rasmont2008; @theodorou2020].
-
-Human-dominated habitats --- notably cities --- drastically modify the ancestral conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to tolerate and exploit urban environments.
-The 'cognitive buffer hypothesis' provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition to gather, store and plastically react to new information [@shettleworth2009; @sol2009].
-While the 'cognitive buffer hypothesis' receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
-
-Insects have been long associated with the most basic types of learning due to their miniature brain [@giurfa2015; @perry2017].
-However, recent studies have shown that bees and other insects have sophisticated cognition that goes beyond simple associative learning or conditioning [@chittka2009; @avargues2011; @sheehan2011].
-Some of these more complex forms of cognition involve the use of tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
-In addition, there is also evidence for substantial variation across species in brain size --- both in absolute terms and relative to body size --- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
-Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
-
-Here, we report the first test of the 'cognitive-buffer hypothesis' in insects.
-Our test is based on a unique dataset of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
-By means of detailed georeferenced information of species occurrences, we characterise the degree of habitat occupancy for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
-
-\clearpage
-
-# 2. Methods
-
-```{r Trait info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr)
-#Load brain data
-brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Select unique species
-sp = pref %>% distinct(Species) %>% pull()
-
-#Filter species that are in the phylo
-brain_weight = brain_weight %>% filter(Species %in% sp)
-largest_encephalisation = brain_weight %>% filter(residuals == max(residuals)) %>% select(Species) %>% pull()
-lowest_encephalisation = brain_weight %>% filter(residuals == min(residuals)) %>% select(Species) %>% pull()
-
-#Read model output
-r2 = read_rds("../Data/Processing/bayessianr2.rds")
-r2 = round(r2[1],2)
-
-#Read phylo signal
-phylo = read_rds("../Data/Processing/phylo_signal.rds")
-lambda = round(phylo$lambda, 2)
-phylo_nat = read_rds("../Data/Processing/phylo_signal_natural.rds")
-lambda_nat = round(phylo_nat$lambda, 2)
-p_nat = round(phylo_nat$P,2)
-
-#Read summary of habitat preference
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Average preferences between USA and Europe
-preferences = pref %>% 
-group_by(Species) %>%
-summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-ungroup()
-#Agricultural
-pref_agr = pref %>% 
-dplyr::select(Agricultural) %>%  
-filter(Agricultural > 0.80) %>% 
-nrow()
-#Natural
-pref_nat = pref %>% 
-dplyr::select(Natural) %>%  
-filter(Natural > 0.80) %>% 
-nrow()
-#Urban
-pref_urb1 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban < 0.20) %>% 
-nrow()
-#Urban
-pref_urb2 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban > 0.80) %>% 
-nrow()
-
-#Read R2 MAIN MODEL
-r2_main = readRDS("../Data/Processing/r2_main.rds")
-r2_main = round(r2_main[[1]],2)
-#Read R2 MODEL BRAIN WEIGHT
-r2_brain = readRDS("../Data/Processing/r2_brain.rds")
-r2_brain = round(r2_brain[[1]],2)
-#Read R2 MODEL ITS
-r2_its = readRDS("../Data/Processing/r2_its.rds")
-r2_its = round(r2_its[[1]],2)
-
-```
-
-## (a) Brain measurements
-
-Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Northern Central Europe.
-These specimens were collected opportunistically mostly in semi-natural habitats.
-The dataset includes information of `r nrow(raw)` female individuals from `r nlevels(factor(unique(raw$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
-We considered only female specimens because they engage in a wider variety of tasks, facing greater environmental pressures, and possess distinct brain structures and functions compared to males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; Supplementary text S1].
-Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
-
-Following @sayol2020, we estimated relative brain size as the residuals of a log-log phylogenetic linear model of brain mass against body size, using the Bayesian approximation implemented in the function *brm* from the package *brms* version 2.18 [@burkner2017].
-The phylogeny was built with the help of a previously published genus-level phylogeny and it was processed with the help of the packages ape 5.6-2 [@paradis2019], phytools 1.2-0 [@revell2012] and MCMCglmm version 2.33 [@hadfield2010].
-High values of relative brain size indicate larger brains than expected for their body size, while low values denote smaller brains.
-Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the species’ evolutionary history (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* \< 0.01).
-However, we found considerable variability in relative brain size within and across taxonomic groups (Figure S1; Figure 1A).
-
-## (b) Habitat occupancy
-
-```{r Loading study site areas, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Calculate study site areas
-#Load libraries
-library(sf)
-library(sfheaders)
-
-#EUROPE:
-#Read data
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Add area col
-euro_map$area <- st_area(euro_map)
-#Create dataframe with the correct units
-d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Calculate area
-eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")
-
-#USA:
-#Read data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-#Not an sf object, convert into multypolygon
-usa = sfheaders::sf_multipolygon(
-  obj = usa_states_1
-  , multipolygon_id = "group"
-  , x = "long"
-  , y = "lat"
-) %>% st_set_crs(4326)
-#Calculate area
-usa$area <- st_area(usa)
-#Create dataframe with the correct units
-d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Sum areas
-usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")
-
-```
-
-```{r Species and occurrences info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read occurrence records and extract relevant info
-#Load libraries
-library(data.table)
-#Load clean data with bee occurrences
-europe <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-usa <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Europe
-eu_occurrences = prettyNum(nrow(europe), big.mark=",")
-records_eu = europe %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#USA
-usa_occurrences = prettyNum(nrow(usa), big.mark=",")
-records_usa = usa %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#Number of species
-usa_spp = nrow(records_usa)
-eu_spp = nrow(records_eu)
-spp_number = nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
-spp_shared = nrow(bind_rows(records_eu, records_usa)) - nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
-
-#Median number of records per dataset
-eu_median = median(records_eu$no_rows)
-usa_median = median(records_usa$no_rows)
-
-#Load now dataset with full set of species
-all_spp = readRDS("../Data/Processing/wit.mean.rds")
-spp_number_nofilter = nrow(all_spp) #total number of species
-
-```
-
-We downloaded occurrence information for all the measured species from the Global Biodiversity Information Facility (GBIF; <http://www.gbif.org/>) for North America and Europe.
-The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
-We selected the states or countries with the highest density of records for our set of species.
-For North America, we selected states located on the East Coast of the United States (Figure S2A), covering an approximate area of `r usakm2` km$^{2}$.
-For Europe, we selected countries located on the north and centre of the continent (Figure S2B), representing a total area of `r eukm2` km$^{2}$.
-We only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
-We optimised the match between species occurrence and the land cover data by only using georeferenced records obtained between 1990 and 2022 with a minimum of two decimals of latitude/longitude coordinates.
-
-We assigned a habitat type to each GBIF occurrence by merging land cover information with the georeferenced records of species occurrences.
-The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for the United States and the Corine Land Cover (CLC) for Europe.
-After downloading these inventories as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
-To simplify the interpretation and conduct a joint analysis for both regions, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
-With this information we can build an occurrence matrix with species in rows, habitats in columns, and cells depicting the number of occurrences per species-habitat combination.
-
-The degree of habitat occupancy was estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
-Although previous work has referred to this metric as ‘habitat preference’ [@collado2019], we have opted to use ‘degree of habitat occupancy’ as it avoids the connotation of active choice by the different species.
-To calculate habitat occupancy, we generated 10,000 randomised matrices based on the occurrence matrix with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
-We used the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
-This maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
-Lastly, we considered that a bee species exhibited a ‘high habitat occupancy’ for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
-On the contrary, we considered a ‘low habitat occupancy’ for the habitat when the observed occurrences were below the 20th percentile.
-To better understand if our findings are affected by the evolutionary history of the species, we estimated the phylogenetic signal of relative brain size and the degree of habitat occupancy across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
-
-## (c) Diet specialisation
-
-To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with the degree of habitat occupancy and brain size.
-For this, we used existing diet information [see @sayol2020], where bees were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
-We tested for statistical differences in diet across the degree of habitat occupancy per habitat type and brain size by using Wilcoxon test.
-We found a low number of specialist species in our dataset with no clear associations with the degree of habitat occupancy or brain size that can explain the observed relationship between the two (Figure S3).
-
-## (d) Analysis
-
-To evaluate how the association between the degree of habitat occupancy and relative brain size per species changed by habitat type, we used a Phylogenetic Bayessian approach to model their association.
-For this, we first joined the resulting values of the degree of habitat occupancy per species with their respective average relative brain sizes.
-Our macro-ecological approach is justified by: (i) independent data ensuring robust and generalizable ecological patterns and (ii) minimal intraspecific variation in brain and body size compared to interspecific variation (see Supplementary text S2 and Figure S4).
-The resulting distribution of the degree of habitat occupancy for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat occupancy close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
-In other words, simulated occurrences were substantially lower or higher than the observed ones.
-Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as high or low habitat occupancy).
-Because we assessed the degree of habitat occupancy as binary ('low' or 'high'), we specified a Bernouilli distribution where the degree of habitat occupancy was the response variable and relative brain size the predictor.
-To control for the shared evolutionary history of the species, we included the phylogenetic covariance matrix as a random factor.
-Finally, we replicated the analyses for the United States and Europe independently with analogous models.
-
-All our models were run with 4,000 iterations with previous 1,000 warm up iterations, using non-informative or weakly informative priors [@burkner2017].
-Further, the different posterior predictive checks were conducted with the function *pp_check*, also from the *brms* package.
-All our analyses were undertaken in R version 4.0.5 [@r2021], and all data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
-
-# 3. Results
-
-The degree of habitat occupancy varied substantially across species (Figure 1B) and showed moderate phylogenetic signal ($\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species have high occupancy in one or two habitat types but more rarely occurred indistinctly in the three (Figure 1B).
-The habitats with the highest occurrences in comparison with random expectations were the agricultural and natural ones, with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
-In contrast, most species showed low degree of occupancy in urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high occupancy in this habitat type (Figure 1B and Figure S5).
 
@@
 ```
-
-\clearpage
-
-We found that relative brain size was associated with the degree of occupancy across the different habitat types (Figure 2A; Bayesian R$^2$ = `r r2_main`).
-Specifically, we found that bees with larger relative brains showed a higher occupancy in urban habitats than bees with smaller ones (Figure 2A).
-Contrarily, bee species with smaller relative brains showed higher occupancy in natural and agricultural habitats than bees with larger relative brains (Figure 2A).
-The models of the association between absolute brain size and intertegular span with habitat occupancy also showed a marked difference between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
-Specifically, we found that bees with larger brains and body sizes appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
-These findings were consistent with the analogous analyses by geographical regions (United States and Europe; Figure S6).
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="Figure 2.
-Association between relative brain size (A), brain weight (B) and intertegular distance (C) with the degree of habitat occupancy by habitat type (i.e., natural, agricultural and urban).
-The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p"}
-
-#Load libraries
-library(readr)
-library(dplyr)
-library(ggplot2)
-
-#Load data for plotting
-long_data <- read_csv("../Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")
-
-model1 = read_csv("../Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")
-
-model2 = read_csv("../Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")
-
-model3 = read_csv("../Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")
-
-#Relevel for plotting
-long_data = long_data %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model1 = model1 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model2 = model2 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
-#Relevel for plotting
-model3 = model3 %>% 
-    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))
-
-#colorBlindness::displayAllColors(viridis::viridis(4))
-
-set.seed(3)
-
-all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
-geom_jitter(width = 0.05, height = 0,data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab("Habitat occupancy") +
-xlab("Relative brain size") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("A") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-all_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab(NULL) +
-xlab("Brain weight (mg)") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("B") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-all_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
-geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
-geom_line(aes(color=Habitat),size=0.7) +
-geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
-theme_bw() +
-ylab(NULL) +
-xlab("Intertegular distance (mm)") +
-scale_colour_viridis_d() +
-scale_fill_viridis_d() +
-scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
-ggtitle("C") +
-theme(plot.title = element_text(face = "bold", size=18))
-
-library(patchwork)
-
-all_1 + (all_2/all_3) + plot_layout(guides = "collect") &
-theme(legend.position = 'bottom', panel.border=element_rect(size=1),
-      axis.title=element_text(face="bold"),
-      legend.box.background = element_rect(colour = "black",size=1))
-
-```
-
-\clearpage
-
-# 4. Discussion
-
-By analysing data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we observed that bee species inhabiting urban habitats tend to have larger brains relative to their body while those found in forest or agricultural habitats exhibit relatively smaller brains.
-These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to tolerate novel environments.
-Since urban environments exhibit dynamic challenges, including novel resources and changing human disturbances [@wheater2002], a large brain may provide the cognitive flexibility to exploit these new resources while avoiding risks.
-Although the exact mechanisms remain unclear, we expect cognitive flexibility to be essential in a variety of contexts such as the use of human-made materials for the nest [@allasino2019] or the use of exotic flowers [@lowenstein2019].
-
-In line with the cognitive buffer hypothesis [@ducatez2020], urban bees also showed bigger bodies and larger absolute brains.
-Larger species require longer development times and obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
-For example, the large carpenter bees of the genus _Xylocopa_ can live up to two years and are frequent urban dwellers, while small _Andrena_ forest specialists complete their adult life cycle in a few weeks.
-The patchiness of urban resources also seems to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
-There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion [@ducatez2015; @sayol2016; @sol2016].
-In our study, we examined only a few bee specialists, highlighting the need for further research on how different diets and life strategies relate to cognition and urban adaptation.
-
-Our findings support and extend upon previous evidence in vertebrates, that having larger brains can facilitate tolerance to urban environments [@santini2019; @sayol2020b], highlighting that a cognitive buffer is possible even with tiny brains.
-The use of the entire brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
-We primarily analyzed brain size due to data availability, supported by prior findings showing larger brains improving certain cognitive aspects in bees, such as learning [@collado2021].
-Moreover, brain size is less subject to measurement error or context-dependent biases in comparison with other experimental measures of cognition [@chittka2009; @lihoreau2012].
-However, future finer measures, such as neuropil size or mushroom bodies  [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995], are likely to enhance our understanding of bee cognition.
-Downscaling our analysis to the individual specimen level where brain size, habitat use and cognitive performance can be tracked through their lifespan would be a challenging but promising next step.
-Our findings highlight the importance of cognition for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
-
-# Data accessibility
-
-All data and code can be accessed at Dryad Digital Repository: <https://doi.org/10.5061/dryad.zw3r228dr> [@Lanuza2023].
-
-# Authors' contributions
-
-IB, DS, MAC and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC, DS and IB. IB and JBL wrote the manuscript with contributions from all authors.
-
-# Funding
-
-Ministerio de Economía y Competitividad, Gobierno de España, Grant/Award Number: CGL2013-47448-P.
-
-# Acknowledgements
-
-We thank Fransico P. Molina, Ivo Reamakers and Parker Gambino for collecting some of the specimens.
-
-# Conflict of interest declaration
-
-We declare we have no competing interests.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::

### 2023-10-25 Manuscript/Figure_2.Rmd
Minor revision

--- 2023-10-23 Manuscript/Ms.Rmd
@@
 ---
-title: "Brain size predicts bees' tolerance to urban environments"
-output:
-  pdf_document:
-    fig_caption: yes
-  word_document: default
-link-citations: yes
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
-fontsize: 12pt
-includeline-num: true
-bibliography: references.bib
-always_allow_html: true
+output: pdf_document
 header-includes:
-- \usepackage{orcidlink}
 - \usepackage{caption,setspace}
 - \captionsetup[figure]{labelformat=empty}
-- \usepackage{float}
-- \floatplacement{figure}{H}
-- \usepackage[left]{lineno}
-- \linenumbers
-csl: BiologyLetters.csl
 ---
-
-\captionsetup[table]{labelformat=empty}
-
-```{r echo=FALSE,warning=FALSE}
-
-knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
-```
-
-\textbf{Jose B.
-Lanuza$^{1,2}$} \orcidlink{0000-0002-0287-409X} \textbf{, Miguel Á.
-Collado$^{1,3}$} \orcidlink{0000-0002-4216-317X} \textbf{, Ferran Sayol$^{4}$} \orcidlink{0000-0003-3540-7487} \textbf{, Daniel Sol$^{4,5}$} \orcidlink{0000-0001-6346-6949} \textbf{and Ignasi Bartomeus$^{1}$} \orcidlink{0000-0001-7893-4389}
-
-Jose B.
-Lanuza$^{1,2}$ https://orcid.org/0000-0002-0287-409X, Miguel Á.
-Collado$^{1,3}$ https://orcid.org/0000-0002-4216-317X, Ferran Sayol$^{4}$ https://orcid.org/0000-0003-3540-7487, Daniel Sol$^{4,5}$ https://orcid.org/0000-0001-6346-6949 and Ignasi Bartomeus$^{1}$ https://orcid.org/0000-0001-7893-4389
-
-$^{1}$Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain
-
-$^{2}$German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
-
-$^{3}$Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain
-
-$^{4}$Centre for Ecological Research and Forestry Applications (CREAF), Bellaterra, Catalonia, Spain
-
-$^{5}$CSIC, Spanish National Research Council, CREAF-UAB, Catalonia, Spain
-
-**Author for correspondence: barragansljose\@gmail.com**
-
-```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-library(readr)
-library(dplyr)
-
-#Load final dataset 
-raw = read_csv("../Data/Processing/all_cleaning.csv")
-
-d = raw %>% 
-group_by(Species) %>% 
-summarise(individuals = n())
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-s = unique(pref$Species)
-
-raw = raw %>% filter(Species %in% s)
-
-#Load taxonomic rank script
-tax = read_csv("../Data/Processing/taxonomic_rank.csv")
-
-```
-
-# Abstract
-
-The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the negative consequences for their fundamental role as plant pollinators.
-However, not all pollinators are negatively affected by habitat conversion, as certain species find in anthropogenic landscapes appropriate resources to persist and proliferate.
-The reason why some species tolerate anthropogenic environments while most find them inhospitable remains poorly understood.
-The cognitive buffer hypothesis, widely supported in vertebrates but untested in insects, offers a potential explanation.
-This theory suggests that species with larger brains have enhanced behavioural plasticity, enabling them to confront and adapt to novel challenges.
-To investigate this hypothesis in insects, we measured brain size for `r nlevels(factor(unique(raw$Species)))` bee species, and evaluated their association with the degree of habitat occupancy.
-Our analyses revealed that bee species mainly found in urban habitats had larger brains relative to their body size than those that tend to occur in forested or agricultural habitats.
-Additionally, urban bees exhibited larger body sizes and, consequently, larger absolute brain sizes.
-Our results provide the first empirical support for the cognitive buffer hypothesis in invertebrates, suggesting that a large brain in bees could confer behavioural advantages to tolerate urban environments.
-
-Keywords: relative brain size, habitat occupancy, Apoidea, urbanisation, pollinators
-
-\clearpage
-
-# 1. Introduction
-
-Pollinators deliver a fundamental ecosystem service on which the earth's vegetation and human economy depend [@ollerton2017].
-Regrettably, there is increasing evidence of recent declines in pollinator populations [@bartomeus2019; @potts2010; @scheper2014].
-One of the main contributing factors to the current pollinator declines is the alteration and loss of their habitat due to human activity [@harrison2015; @harrison2018].
-Anthropogenic landscapes present new challenges for the survival and reproduction of organisms, increasing their risk of extinction by maladaptation [@winfree2011; @brown2009; @harrison2019].
-Yet, not all pollinator species are negatively affected by land use change.
-Indeed, some bee species are able to tolerate human-altered environments [@biesmeijer2006; @bartomeus2013; @collado2019] or even to thrive [@rasmont2008; @theodorou2020].
-
-Human-dominated habitats --- notably cities --- drastically modify the ancestral conditions where pollinators evolved, but can also offer unique ecological opportunities in the form of new nesting spots, shelter from phytosanitary products, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019].
-The question is then why only some species are able to tolerate and exploit urban environments.
-The 'cognitive buffer hypothesis' provides an explanation for this conundrum, suggesting that in novel environments the chances to successfully survive and reproduce depends on enhanced cognition to gather, store and plastically react to new information [@shettleworth2009; @sol2009].
-While the 'cognitive buffer hypothesis' receives ample support from studies of vertebrates [@sol2008; @amiel2011; @sayol2016; @howell2023], similar evidence is lacking for insects.
-
-Insects have been long associated with the most basic types of learning due to their miniature brain [@giurfa2015; @perry2017].
-However, recent studies have shown that bees and other insects have sophisticated cognition that goes beyond simple associative learning or conditioning [@chittka2009; @avargues2011; @sheehan2011].
-Some of these more complex forms of cognition involve the use of tools, attention, social learning or metacognitive processes [@giurfa2015; @perry2017].
-In addition, there is also evidence for substantial variation across species in brain size --- both in absolute terms and relative to body size --- with species that have larger brains also exhibiting enhanced cognitive performance, at least for some tasks such as conditional learning [@collado2021].
-Therefore, we can ask whether the varying success of insects in human-altered habitats might be explained by variation in brain size.
-
-Here, we report the first test of the 'cognitive-buffer hypothesis' in insects.
-Our test is based on a unique dataset of brain measures for `r nlevels(factor(unique(raw$Species)))` European and North American bee species.
-By means of detailed georeferenced information of species occurrences, we characterise the degree of habitat occupancy for all the species, and use a phylogenetically-informed comparative analysis to assess whether bees that proliferate in human-altered habitats have enlarged brains compared to those that avoid them.
-
-\clearpage
-
-# 2. Methods
-
-```{r Trait info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr)
-#Load brain data
-brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Select unique species
-sp = pref %>% distinct(Species) %>% pull()
-
-#Filter species that are in the phylo
-brain_weight = brain_weight %>% filter(Species %in% sp)
-largest_encephalisation = brain_weight %>% filter(residuals == max(residuals)) %>% select(Species) %>% pull()
-lowest_encephalisation = brain_weight %>% filter(residuals == min(residuals)) %>% select(Species) %>% pull()
-
-#Read model output
-r2 = read_rds("../Data/Processing/bayessianr2.rds")
-r2 = round(r2[1],2)
-
-#Read phylo signal
-phylo = read_rds("../Data/Processing/phylo_signal.rds")
-lambda = round(phylo$lambda, 2)
-phylo_nat = read_rds("../Data/Processing/phylo_signal_natural.rds")
-lambda_nat = round(phylo_nat$lambda, 2)
-p_nat = round(phylo_nat$P,2)
-
-#Read summary of habitat preference
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-#Average preferences between USA and Europe
-preferences = pref %>% 
-group_by(Species) %>%
-summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-ungroup()
-#Agricultural
-pref_agr = pref %>% 
-dplyr::select(Agricultural) %>%  
-filter(Agricultural > 0.80) %>% 
-nrow()
-#Natural
-pref_nat = pref %>% 
-dplyr::select(Natural) %>%  
-filter(Natural > 0.80) %>% 
-nrow()
-#Urban
-pref_urb1 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban < 0.20) %>% 
-nrow()
-#Urban
-pref_urb2 = pref %>% 
-dplyr::select(Urban) %>%  
-filter(Urban > 0.80) %>% 
-nrow()
-
-#Read R2 MAIN MODEL
-r2_main = readRDS("../Data/Processing/r2_main.rds")
-r2_main = round(r2_main[[1]],2)
-#Read R2 MODEL BRAIN WEIGHT
-r2_brain = readRDS("../Data/Processing/r2_brain.rds")
-r2_brain = round(r2_brain[[1]],2)
-#Read R2 MODEL ITS
-r2_its = readRDS("../Data/Processing/r2_its.rds")
-r2_its = round(r2_its[[1]],2)
-
-```
-
-## (a) Brain measurements
-
-Our dataset contains measurements of brain and body size for bee specimens captured on flowers by hand netting in different areas of the East Coast of the United States and Northern Central Europe.
-These specimens were collected opportunistically mostly in semi-natural habitats.
-The dataset includes information of `r nrow(raw)` female individuals from `r nlevels(factor(unique(raw$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` families and `r nlevels(factor(unique(tax$genus)))` genera.
-We considered only female specimens because they engage in a wider variety of tasks, facing greater environmental pressures, and possess distinct brain structures and functions compared to males [@streinzer2013; @roselino2015].
-Brain size was measured as the weight of fixed brains [@sayol2020] and body size as intertegular span [@kendall2019; Supplementary text S1].
-Given that brain size scales allometrically with body size [@sayol2020], we considered for our analysis the size of the brain relative to the body.
-
-Following @sayol2020, we estimated relative brain size as the residuals of a log-log phylogenetic linear model of brain mass against body size, using the Bayesian approximation implemented in the function *brm* from the package *brms* version 2.18 [@burkner2017].
-The phylogeny was built with the help of a previously published genus-level phylogeny and it was processed with the help of the packages ape 5.6-2 [@paradis2019], phytools 1.2-0 [@revell2012] and MCMCglmm version 2.33 [@hadfield2010].
-High values of relative brain size indicate larger brains than expected for their body size, while low values denote smaller brains.
-Our bee dataset showed a strong allometric relationship between brain and body size (Bayesian R$^2$ = `r r2`) that was constrained by the species’ evolutionary history (phylogenetic signal of relative brain size, $\lambda$ = `r lambda`; *P* \< 0.01).
-However, we found considerable variability in relative brain size within and across taxonomic groups (Figure S1; Figure 1A).
-
-## (b) Habitat occupancy
-
-```{r Loading study site areas, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Calculate study site areas
-#Load libraries
-library(sf)
-library(sfheaders)
-
-#EUROPE:
-#Read data
-euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
-#Add area col
-euro_map$area <- st_area(euro_map)
-#Create dataframe with the correct units
-d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Calculate area
-eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")
-
-#USA:
-#Read data
-usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
-#Not an sf object, convert into multypolygon
-usa = sfheaders::sf_multipolygon(
-  obj = usa_states_1
-  , multipolygon_id = "group"
-  , x = "long"
-  , y = "lat"
-) %>% st_set_crs(4326)
-#Calculate area
-usa$area <- st_area(usa)
-#Create dataframe with the correct units
-d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
-mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
-#Sum areas
-usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")
-
-```
-
-```{r Species and occurrences info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read occurrence records and extract relevant info
-#Load libraries
-library(data.table)
-#Load clean data with bee occurrences
-europe <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
-usa <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")
-
-#Europe
-eu_occurrences = prettyNum(nrow(europe), big.mark=",")
-records_eu = europe %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#USA
-usa_occurrences = prettyNum(nrow(usa), big.mark=",")
-records_usa = usa %>% 
-group_by(species) %>% 
-summarize(no_rows= length(species))
-
-#Number of species
-usa_spp = nrow(records_usa)
-eu_spp = nrow(records_eu)
-spp_number = nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
-spp_shared = nrow(bind_rows(records_eu, records_usa)) - nrow(bind_rows(records_eu, records_usa) %>% distinct(species))
-
-#Median number of records per dataset
-eu_median = median(records_eu$no_rows)
-usa_median = median(records_usa$no_rows)
-
-#Load now dataset with full set of species
-all_spp = readRDS("../Data/Processing/wit.mean.rds")
-spp_number_nofilter = nrow(all_spp) #total number of species
-
-```
-
-We downloaded occurrence information for all the measured species from the Global Biodiversity Information Facility (GBIF; <http://www.gbif.org/>) for North America and Europe.
-The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017].
-We selected the states or countries with the highest density of records for our set of species.
-For North America, we selected states located on the East Coast of the United States (Figure S2A), covering an approximate area of `r usakm2` km$^{2}$.
-For Europe, we selected countries located on the north and centre of the continent (Figure S2B), representing a total area of `r eukm2` km$^{2}$.
-We only included species with a minimum number of 50 records and whose geographic distribution was larger than the sampled area (i.e., excluding species at the edge of their distributions).
-We optimised the match between species occurrence and the land cover data by only using georeferenced records obtained between 1990 and 2022 with a minimum of two decimals of latitude/longitude coordinates.
-
-We assigned a habitat type to each GBIF occurrence by merging land cover information with the georeferenced records of species occurrences.
-The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for the United States and the Corine Land Cover (CLC) for Europe.
-After downloading these inventories as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively.
-To simplify the interpretation and conduct a joint analysis for both regions, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban (see Tables S1 and S2 for details).
-With this information we can build an occurrence matrix with species in rows, habitats in columns, and cells depicting the number of occurrences per species-habitat combination.
-
-The degree of habitat occupancy was estimated by assessing whether the occurrences of species in a given habitat were more frequent than expected by chance.
-Although previous work has referred to this metric as ‘habitat preference’ [@collado2019], we have opted to use ‘degree of habitat occupancy’ as it avoids the connotation of active choice by the different species.
-To calculate habitat occupancy, we generated 10,000 randomised matrices based on the occurrence matrix with the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008].
-We used the method 'r2dtable', which maintains row and column sums constant by using Patefield's algorithm [@patefield1981].
-This maintains the proportional dominance of the species and habitats constant but reshuffles their associations.
-We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones (i.e., percentile).
-Lastly, we considered that a bee species exhibited a ‘high habitat occupancy’ for any of the three studied habitats when the number of occurrences observed in this habitat exceeded the 80th percentile of the values obtained from the simulations.
-On the contrary, we considered a ‘low habitat occupancy’ for the habitat when the observed occurrences were below the 20th percentile.
-To better understand if our findings are affected by the evolutionary history of the species, we estimated the phylogenetic signal of relative brain size and the degree of habitat occupancy across habitats for our set of species with the help of the function *phylosig* from the *phytools* package.
-
-## (c) Diet specialisation
-
-To ensure diet is not acting as a confounding factor we investigated how diet specialisation is associated with the degree of habitat occupancy and brain size.
-For this, we used existing diet information [see @sayol2020], where bees were classified as oligolectic when they use a single plant family to feed their brood or polylectic when they use several.
-We tested for statistical differences in diet across the degree of habitat occupancy per habitat type and brain size by using Wilcoxon test.
-We found a low number of specialist species in our dataset with no clear associations with the degree of habitat occupancy or brain size that can explain the observed relationship between the two (Figure S3).
-
-## (d) Analysis
-
-To evaluate how the association between the degree of habitat occupancy and relative brain size per species changed by habitat type, we used a Phylogenetic Bayessian approach to model their association.
-For this, we first joined the resulting values of the degree of habitat occupancy per species with their respective average relative brain sizes.
-Our macro-ecological approach is justified by: (i) independent data ensuring robust and generalizable ecological patterns and (ii) minimal intraspecific variation in brain and body size compared to interspecific variation (see Supplementary text S2 and Figure S4).
-The resulting distribution of the degree of habitat occupancy for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat occupancy close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
-In other words, simulated occurrences were substantially lower or higher than the observed ones.
-Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as high or low habitat occupancy).
-Because we assessed the degree of habitat occupancy as binary ('low' or 'high'), we specified a Bernouilli distribution where the degree of habitat occupancy was the response variable and relative brain size the predictor.
-To control for the shared evolutionary history of the species, we included the phylogenetic covariance matrix as a random factor.
-Finally, we replicated the analyses for the United States and Europe independently with analogous models.
-
-All our models were run with 4,000 iterations with previous 1,000 warm up iterations, using non-informative or weakly informative priors [@burkner2017].
-Further, the different posterior predictive checks were conducted with the function *pp_check*, also from the *brms* package.
-All our analyses were undertaken in R version 4.0.5 [@r2021], and all data processing and graphics were done with the set of packages from the tidyverse version 1.3.0 [@tidyverse19].
-
-# 3. Results
-
-The degree of habitat occupancy varied substantially across species (Figure 1B) and showed moderate phylogenetic signal ($\lambda$ = `r lambda_nat`; *P* = `r p_nat`).
-In general, most bee species have high occupancy in one or two habitat types but more rarely occurred indistinctly in the three (Figure 1B).
-The habitats with the highest occurrences in comparison with random expectations were the agricultural and natural ones, with `r pref_agr` and `r pref_nat` species over 80% of the values from null models, respectively (Figure 1B and Figure S5).
-In contrast, most species showed low degree of occupancy in urban habitats (`r pref_urb1` species under the 20th percentile) and just `r pref_urb2` species showed high occupancy in this habitat type (Figure 1B and Figure S5).
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="Figure 1.
-Phylogenetic relationship and degree of habitat occupancy for the selected bee species with brain weight and intertegular distance information (N=89).
-A) Phylogenetic tree at the genus level.
-Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of the United States or from both regions).
-The deviation of the brain size in relation to the body (i.e., residuals) is represented with filled circles of proportional area at the end of the tip branches.
-Larger circles indicate larger brains in proportion to their body size and vice versa.
-B) Heatmap showing the degree of habitat occupancy for each bee species.
-The columns delimit the habitat type (i.e., natural, agricultural and urban) and the rows the different bee species."}
-
-library(ggtree) #v3.5.3.002
-library(ggplot2)
-library(stringr)
-library(tidyverse)
-library(tibble)
-
-#Load phylogenetic data from europe
-bee.tree100 = readRDS("../Data/Europe_USA/bee.tree100.rds")
-#Load brain data
-brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
-#Load preferences data
-preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
-preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
-#Bind rows
-pref = bind_rows(preferences_usa, preferences_eu)
-
-#Average preferences between USA and Europe
-preferences = pref %>% 
-group_by(Species) %>%
-summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
-ungroup()
-
-#Delete underscore from tree
-bee.tree100$tip.label = str_replace(bee.tree100$tip.label, "_", " ")
-
-#Generate vector with species
-v = bee.tree100$tip.label
-#Filter species with that are in the phylo
-brain_weight = brain_weight %>% filter(Species %in% v)
-
-#Fill a column based on condition
-brain_weight = brain_weight %>% 
-    mutate(Country = case_when((brain_weight$Species %in% preferences_eu$Species & 
-    brain_weight$Species %in% preferences_usa$Species ~ "USA and Europe"),
-    (brain_weight$Species %in% preferences_usa$Species ~ "USA"),
-    (brain_weight$Species %in% preferences_eu$Species ~ "Europe")))
-
-#Plot tree
-p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country)+
-    geom_tippoint(aes(size = residuals, color=Country))+
-    theme_tree2()+scale_size_continuous(range = c(-1, 4),name="Relative\nbrain size")+
-    scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
-                       labels = c("Europe", "USA", "USA and Europe"),
-                       breaks = c("Europe", "USA", "USA and Europe")) +
-  ggtitle("A") + theme(plot.title = element_text(size=18, face="bold"))
-
-#Convert wide format to long for plotting heatmap
-d=gather(preferences, condition, measurement, Agricultural:Urban, factor_key=TRUE)
-
-#Relevel for plotting
-d = d %>% 
-mutate(condition=fct_relevel(condition,c("Natural","Agricultural", "Urban")))
-
-p2 = ggplot(d, aes(x=condition, y=Species)) + 
-    geom_tile(aes(fill=measurement),color = "white") +
-   # scale_fill_gradient(high="black",  low="white") +
-  scale_fill_continuous("Habitat occupancy",type = "viridis", 
-                        breaks = c(0, 0.5,  1), 
-                        labels = c("0 (Low)", "0.5","1 (High)")) +
-
-  #  scale_fill_viridis_c(name = "Habitat\noccupancy") + 
-    theme_minimal() +
-    xlab(NULL) + 
-    ylab(NULL) + 
-    theme(axis.text.x = element_text(face="bold", color="black", 
-    size=7.5,  vjust=1.15),
-    axis.text.y = element_text(face="italic", color="black", size=7.5, hjust=0),
-    plot.title = element_text(size=18, face="bold")) +
-   ggtitle("B")
-
-#Nice way to insert ordered heatmap
-library(aplot)
-p2 %>% insert_left(p) 
-```
-
-\clearpage
-
-We found that relative brain size was associated with the degree of occupancy across the different habitat types (Figure 2A; Bayesian R$^2$ = `r r2_main`).
-Specifically, we found that bees with larger relative brains showed a higher occupancy in urban habitats than bees with smaller ones (Figure 2A).
-Contrarily, bee species with smaller relative brains showed higher occupancy in natural and agricultural habitats than bees with larger relative brains (Figure 2A).
-The models of the association between absolute brain size and intertegular span with habitat occupancy also showed a marked difference between habitat types (Figures 2B and 2C; Bayesian R$^2$ = `r r2_brain`; Bayesian R$^2$ = `r r2_its`, respectively).
-Specifically, we found that bees with larger brains and body sizes appear more in urban habitats and bees with smaller brains and body sizes are more common in natural and agricultural ones (Figures 2B and 2C).
-These findings were consistent with the analogous analyses by geographical regions (United States and Europe; Figure S6).
 
@@
 library(ggplot2)
+library(forcats)
 
@@
 ```
-
-\clearpage
-
-# 4. Discussion
-
-By analysing data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we observed that bee species inhabiting urban habitats tend to have larger brains relative to their body while those found in forest or agricultural habitats exhibit relatively smaller brains.
-These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to tolerate novel environments.
-Since urban environments exhibit dynamic challenges, including novel resources and changing human disturbances [@wheater2002], a large brain may provide the cognitive flexibility to exploit these new resources while avoiding risks.
-Although the exact mechanisms remain unclear, we expect cognitive flexibility to be essential in a variety of contexts such as the use of human-made materials for the nest [@allasino2019] or the use of exotic flowers [@lowenstein2019].
-
-In line with the cognitive buffer hypothesis [@ducatez2020], urban bees also showed bigger bodies and larger absolute brains.
-Larger species require longer development times and obtain greater net benefits from exploring and learning, especially in heterogeneous environments [@bateman2012; @theodorou2021].
-For example, the large carpenter bees of the genus _Xylocopa_ can live up to two years and are frequent urban dwellers, while small _Andrena_ forest specialists complete their adult life cycle in a few weeks.
-The patchiness of urban resources also seems to favour larger body sizes, as suggests the observation that foraging distance is positively associated with body size in bees [@greenleaf2007; @kendall2019].
-There are also shreds of evidence that associate increased tolerance to urbanisation with wider ecological niches [@geslin2013; @buchholz2020; @prendergast2022], and both larger bodies and relative brains are thought to be features that can facilitate broader diets and niche expansion [@ducatez2015; @sayol2016; @sol2016].
-In our study, we examined only a few bee specialists, highlighting the need for further research on how different diets and life strategies relate to cognition and urban adaptation.
-
-Our findings support and extend upon previous evidence in vertebrates, that having larger brains can facilitate tolerance to urban environments [@santini2019; @sayol2020b], highlighting that a cognitive buffer is possible even with tiny brains.
-The use of the entire brain size as a proxy for cognitive performance is not exempt from criticism [@chittka2009; @lihoreau2012].
-We primarily analyzed brain size due to data availability, supported by prior findings showing larger brains improving certain cognitive aspects in bees, such as learning [@collado2021].
-Moreover, brain size is less subject to measurement error or context-dependent biases in comparison with other experimental measures of cognition [@chittka2009; @lihoreau2012].
-However, future finer measures, such as neuropil size or mushroom bodies  [the suspected centres of cognitive processes in bees\; @barth1997; @withers1995; @fahrbach1995], are likely to enhance our understanding of bee cognition.
-Downscaling our analysis to the individual specimen level where brain size, habitat use and cognitive performance can be tracked through their lifespan would be a challenging but promising next step.
-Our findings highlight the importance of cognition for understanding the dynamics of insect populations in altered environments and stresses the need to avoid viewing them as passive agents of external pressures.
-
-# Data accessibility
-
-All data and code can be accessed at Dryad Digital Repository: <https://doi.org/10.5061/dryad.zw3r228dr> [@Lanuza2023].
-
-# Authors' contributions
-
-IB, DS, MAC and JBL designed the study. MAC and FS collected the data. JBL led the analysis with help from MAC, DS and IB. IB and JBL wrote the manuscript with contributions from all authors.
-
-# Funding
-
-Ministerio de Economía y Competitividad, Gobierno de España, Grant/Award Number: CGL2013-47448-P.
-
-# Acknowledgements
-
-We thank Fransico P. Molina, Ivo Reamakers and Parker Gambino for collecting some of the specimens.
-
-# Conflict of interest declaration
-
-We declare we have no competing interests.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::

### 2023-10-25 Manuscript/Ms.Rmd
Minor revision

@@
 The resulting distribution of the degree of habitat occupancy for each of the habitats analysed followed a zero-one inflated beta distribution (Figure S5), indicating that there were high frequencies of habitat occupancy close to 0 or 1 but low frequencies of intermediate values between 0 and 1.
-In other words, simulated occurrences were substantially lower or higher than the observed ones.
+In other words, simulated occurrences per species and habitat were consistently lower or higher than the observed ones.
 Hence, in our analyses we take a conservative approach and only modelled the extremes of the distribution (i.e., species classified as high or low habitat occupancy).
@@
 
-By analysing data from two different continents and `r nlevels(factor(unique(raw$Species)))` bee species, we observed that bee species inhabiting urban habitats tend to have larger brains relative to their body while those found in forest or agricultural habitats exhibit relatively smaller brains.
+By analysing `r nlevels(factor(unique(raw$Species)))` bee species from two different continents, we observed that bee species inhabiting urban habitats tend to have larger brains relative to their body while those found in forest or agricultural habitats exhibit relatively smaller brains.
 These results are in line with the cognitive buffer hypothesis [@sol2009], which predicts that a large relative brain should provide enhanced behavioural plasticity to tolerate novel environments.

### 2023-10-25 Manuscript/Supplementary_material.Rmd
Minor revision

@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species.
-A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
+A) East Coast of the United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.
 B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}
